### 2014-06-27 Detectors.tex
First commit

\section{Detectors}\label{sec:detectors}
\subsection{Operating principle of an interferometric detector}\label{sec:principles}
All of the man-made detectors discussed in this section work on the principle of interferometry. Such detectors work by taking a beam of monochromatic light and splitting it into two beams travelling at some angle to each other. Each beam is passed in to a Fabry-P\'{e}rot cavity where it undergoes a number of round trips before being recombined to form an interference pattern. The ends of the cavity are, in the ideal case, freely floating test masses which move in response to a passing GW, this effect is measured by observing the changing interference pattern.

The response of a detector to an incident plane fronted GW depends upon the relative orientations of the detector and the incoming wave. Choosing the origin of our coordinate system to be the beam-splitter of the interferometer and let $l_{1}^{i}$ and $l_{2}^{i}$ be unit 3-vectors pointing along the two arms. In the absence of noise the output of the detector is the difference in strain between the two arms \citep{MTW}
\begin{equation}\label{eq:hoftxx} h(t)=\frac{1}{2}h_{ij}\left( l_{1}^{i}l_{1}^{j}-l_{2}^{i}l_{2}^{j} \right)\; . \end{equation}
where $h_{ij}$ are the spatial components of the GW metric perturbation. Let $n^{i}$ be the unit 3-vector pointing towards the source of the GWs, with spherical polar angles $(\theta,\phi)$ relative to some axes fixed to the detector, also let $p^{i}$ and $q^{i}$ be unit vectors orthogonal to $n^{i}$. There remains a freedom in the coordinates described, a rotation of $p^{i}$ and $q^{i}$ through an angle $\psi$ about $n^{i}$. Define the basis tensors,
\begin{eqnarray}
H^{+}_{ij}&=p_{i}p_{j}-q_{i}q_{j} \nonumber \\
H^{\times}_{ij}&=p_{i}q_{j}+q_{i}p_{j} \; .
\end{eqnarray}
Consider a single frequency component of the incident GW, the strain induced by the incident GW may be written as
\begin{equation}\label{eq:hijxx} h_{ij}=A_{+}H^{+}_{ij}\cos\left(2\pi ft\right)+A_{\times}H^{\times}_{ij}\cos\left(2\pi ft+\Delta \phi\right) \; ,\end{equation}
where $A_{+}$ and $A_{\times}$ are the amplitudes of the two polarisation states. Combining (\ref{eq:hoftxx}) and (\ref{eq:hijxx}) allows the detector output to be written as
\begin{equation} h(t)=F^{+}(\theta,\phi,\psi)A_{+}\cos\left(2\pi ft\right)+F^{\times}(\theta,\phi,\psi)A_{\times}\cos\left(2\pi f t + \Delta\phi \right)\; , \end{equation}
where the response functions inherit their angular dependence from the choice of coordinates,
\begin{eqnarray}
F^{+}(\theta,\phi,\psi)&=\frac{1}{2}H^{+}_{ij}\left(l_{1}^{i}l_{1}^{j}-l_{2}^{i}l_{2}^{j}\right) \nonumber \\
F^{\times}(\theta,\phi,\psi)&=\frac{1}{2}H^{\times}_{ij}\left(l_{1}^{i}l_{1}^{j}-l_{2}^{i}l_{2}^{j}\right) \; . 
\end{eqnarray}
The response function of a two-arm interferometric detector is quadrapolar, an example is plotted in figure \ref{fig:LIGO}.

\begin{figure}[h!]
 \centering
 \includegraphics[trim=0cm 0cm 0cm 0cm, width=0.99\textwidth]{LIGO_detectorframe.pdf}
 \caption{The angular response function of an interferometric detector shown both as a surface plot and in an Aitoff-Hammer projection. The two detector arms lie in the $xy$ plane either side of one of the zeros in the response.}
 \label{fig:LIGO}
\end{figure}

Throughout this paper detector sensitivity refers to the polarisation and sky averaged sensitivity, $F$, where
\begin{equation}\label{eq:skyav} F^{2}=\int_{0}^{2\pi}\frac{\textrm{d}\psi}{2\pi}\; \int_{0}^{\pi}\int_{0}^{2\pi}\frac{\sin\theta\,\textrm{d}\theta\textrm{d}\phi}{4\pi}\;\left(\frac{F^{+}\left(\theta,\phi,\psi\right)^{2}+F^{\times}\left(\theta,\phi,\psi\right)^{2}}{2}\right)\; .\end{equation}
For a single $90\degree$ interferometer, such as LIGO (Laser Interferometer GW Observatory), the sky and polarisation averaged response is $F=\sqrt{1/5}\approx 0.447$.

A detector may consist of several interferometers. Let $F_{a}$ be the averaged response of the $a^{\textrm{th}}$ interferometer. The average response of the detector network is obtained by adding in quadrature,
\begin{equation} F_{\textrm{Total}}^{2}=\frac{1}{\left(\textrm{Number of interferometers}\right)}\sum_{x=1}^{N}F_{x}^{2} \; .\end{equation}

The averaging in (\ref{eq:skyav}) assumes a uniform distribution of polarisation angles $\psi$. This will be the case for a stochastic background, however for a non-inspiralling circular binary the polarisation will be a function of the two spherical polar angles ($\iota,\xi$) specifying the orientation of the binary's orbital angular momentum. Here $\iota$ is the polar angle between the orbital angular momentum and the line joining the source to the detector while $\xi$ is the azimuthal angle around the same line. In this case the correct average to characterise the detector sensitivity would be an an average over all four angles ($\theta,\phi,\iota,\xi$). If the binary is inspiralling then the polarisation will depend on still more parameters which need to be averaged over. These more complicated averages all have the property that they depend on both the detector and the source, hence they are unhelpful for our present purpose. In addition the different averages don't work out to be so different from each other; \cite{1993PhRvD..47.2198F} calculated the sensitivity for a detector with the LIGO geometry averaged over the four angles ($\theta,\phi,\iota,\xi$) as $\sqrt{4/25}=0.4$ times peak sensitivity, which should be compared with the value $\sqrt{1/5}\approx 0.447$ above. For the remainder of this paper the 3 angle average defined in (\ref{eq:skyav}) will be used.

\subsection{Ground based detectors}\label{sec:ground}
On Earth approximations to the free floating test masses described above are achieved by suspending the mass from a pendulum system with natural frequency much greater than that of the GW. The interferometric detectors listed in Table \ref{table:t} are all sensitive to GWs in the frequency range ${\cal{O}}(10-10^{3})\,\textrm{Hz}$. They all have multiple narrow lines in their sensitivity curves which arise from noise sources in the instrument including resonances in the suspension system and electronic noise at multiples of $60\,\textrm{Hz}$: these have been removed in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega} for clarity. The detectors fall broadly into three categories: 1\textsuperscript{st} generation detectors which have already operated, 2\textsuperscript{nd} generation detectors currently under construction and 3\textsuperscript{rd} generation detectors at the planning stage.

\begin{table}[h!]
\caption{\label{table:t} For the aVIRGO sensitivity curve an interpolation to the data published on \url{https://wwwcascina.virgo.infn.it/advirgo/} (2013) was used. For KAGRA (Kamioka Gravitational Wave Detector) an interpolation to the data for version D of the detector published on \url{http://gwcenter.icrr.u-tokyo.ac.jp/en/researcher/parameter} (2013) was used. For the remaining detectors analytic fits to the sensitivity curves due to \cite{Sathyaprakash} were used.}
\begin{indented}
\item[]\begin{tabular}{ l l l l l }
\br
{\bf Detector} & {\bf Country} & {\bf Arm length} & {\bf  Approximate date} & {\bf Generation} \\
\mr
  GEO600 	&	Germany 	& $600\,\textrm{m}$ 	& 2001-present 	   & 1$^{\textrm{st}}$\\
  TAMA300 	& 	Japan		& $300\,\textrm{m}$ 	& 1995-present     & 1$^{\textrm{st}}$\\
  iLIGO		&	US		& $4\,\textrm{km}$ 	& 2004-2010 	   & 1$^{\textrm{st}}$\\
  iVIRGO	& 	Italy		& $3\,\textrm{km}$ 	& 2007-2010 	   & 1$^{\textrm{st}}$\\
  aLIGO 	&	US		& $4\,\textrm{km}$ 	& \emph{est.} 2016 & 2$^{\textrm{nd}}$\\
  KAGRA		&	Japan		& $3\,\textrm{km}$ 	& \emph{est.} 2018 & 2$^{\textrm{nd}}$\\
  aVIRGO	&	Italy	 	& $3\,\textrm{km}$ 	& \emph{est.} 2017 & 2$^{\textrm{nd}}$\\
  ET (Einstein Telescope) &	Italy		& $10\,\textrm{km}$ 	& \emph{est.} 2025 & 3$^{\textrm{rd}}$\\
\br
\end{tabular}
\end{indented}
\end{table}

\subsection{Space based detectors}\label{sec:space}
Space based detectors works on similar principles to ground based detectors but with the test masses residing inside of independent, widely separated satellites. Space based detectors are sensitive to lower frequency gravitational waves than their ground based counterparts, this is partly because space based detectors can have much longer arms and partly because they are unaffected by seismic noise which limits the low frequency performance of ground based detectors.

The canonical design for a space based detector is LISA (Laser Interferometer Space Antenna), which would be sensitive to milli-Hz GWs. LISA would consist of three satellites flying in a triangular constellation with arms of length $5\times 10^{9}\,\textrm{m}$ in a $1\,\textrm{AU}$ orbit around the sun, trailing the Earth by $20\degree$. The laser arms in a LISA like detector are not a Fabry-P\'{e}rot cavity, the light only travels once along each arm. All of the sources discussed here lie within LISA's sensitivity reach. eLISA (evolved LISA) is a simpler version of LISA designed to probe the same frequency range, while proposals such as ALIA (Advanced Laser Interferometer Antenna), BBO (Big Bang Observer) and DECIGO (DECI-Hertz Interferometer GW Observatory) are designed to probe deci-Hz GWs.

\subsubsection{LISA and eLISA}
We use an analytic fit to the instrumental noise curve given by \cite{Sathyaprakash}. When observing individual sources with LISA there is an additional contribution to the noise from a background of unresolvable binaries. This is not included here as we consider the background as a source of GWs, see section \ref{sec:GB}. eLISA is a re-scoped version of the classic LISA mission, the main differences are shorter arms ($10^{9}\,\textrm{m}$ instead of $5\times 10^{9}\,\textrm{m}$), two laser arms instead of three and a different orbit (drifting away from Earth instead of $20\degree$ Earth trailing). The effect of these changes is a slightly reduced peak sensitivity and a shift to higher frequencies. We use an analytic fit to the instrumental noise curve given by \cite{Amaro-Seoane-et-al}.

\subsubsection{ALIA, DECIGO and BBO}
These missions are designed to probe the deci-Hz region of the GW spectrum, they are considerably more ambitious than the LISA or eLISA mission and their launches will be further into the future. For ALIA a simple analytic fit to the sensitivity curve is used, while for DECIGO and BBO fits to the sensitivity curves given by \cite{2011PhRvD..83d4011Y} are used.

\subsection{Pulsar timing arrays (PTAs)}\label{sec:PTAgeneralproperties}
\cjm{I am going to rewrite this section and give a similar level of detail as was given for interferometric detectors in section \ref{sec:principles}.}

PTAs can be thought of as naturally occurring interferometers with galactic scale arm lengths, hence they are sensitive to much lower frequencies than the detectors considered so far. Each pulsar is a very regular clock and the measured arrival time can be compared against a prediction leaving a residual which includes the effects of passing GWs. \cite{SesanaVecchioColancino} describe the shape of the sensitivity curve obtained by correlating timing residuals from each of $N_{\textrm{p}}$ pulsars in an array. The total timing residual is the sum of the residuals due to noise and due to GWs,
\begin{equation} \delta t= \delta t_{n} + \delta t _{h} \; .\end{equation}
The maximum timing residual caused by a passing GW is the light travel time of along wavelength times the strain amplitude of the GW, $\delta t_{h}\approx h/f$. An upper limit to any background may be placed by assuming the residuals are due entirely to GWs. First consider the case of two pulsars; the minimum instantaneous detectable timing residual due to GWs is proportional to the root mean square error in the measurement of the timing residual, $\delta t_{h}\approx\delta t_{\textrm{r.m.s}}$. The minimum detectable signal decreases with the number of measured timing residuals (or data points). Over an observation time $T$, with individual observations taken at intervals of $\Delta T$ the minimum detectable signal is given by
\begin{equation}\label{eq:GWtimingresidual} \delta t_{h} \propto \frac{\delta t_{\textrm{r.m.s}}}{\sqrt{T\Delta T ^{-1}}} \; . \end{equation}
The PTA is sensitive to a minimum GW frequency of $T^{-1}$ and a maximum frequency of $\Delta T^{-1}$, hence the bandwidth is $\Delta f \approx \Delta T^{-1}-T^{-1}$. Assuming $\Delta T \ll T$ gives $T\Delta T^{-1} \approx T \Delta f$. Therefore the minimum characteristic strain the two-pulsar PTA is sensitive to is given by
\begin{equation} h_{c}=\frac{\delta t_{\textrm{r.m.s.}}f}{\left(T\Delta f\right)^{1/4}} \; . \end{equation}
When using more than two pulsars each separate pair of pulsars forms an independent detector, so the number of detectors scales as $N_{\textrm{p}}^{2}$. The optimum SNR is given by adding the individual SNRs in quadrature. Hence the total sensitivity is given by
\begin{equation}\label{eq:PTA} h_{c}(f) \propto \frac{\delta t_{\textrm{r.m.s.}}f}{N_{\textrm{p}}^{1/2}\left( T\Delta f \right)^{1/4}} \; . \end{equation}
The characteristic strain that the PTA is sensitive to scales linearly with $f$ between the minimum and maximum frequencies. This gives the wedge shaped curves plotted in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega}. The absolute value of the sensitivity is fixed by normalising (\ref{eq:PTA}) to agree with a calculated limit at a given frequency for each PTA.

There is a discrepancy between the treatment of PTA sensitivity curves here and the higher frequency detectors discussed in sections \ref{sec:ground} and \ref{sec:space}. When observing a long lived source, such as an inspiral, with a high frequency detector the convention was to define a \emph{characteristic} strain to satisfy (\ref{eq:hc}). Here the convention is to leave the strain untouched and instead adjust the PTA sensitivity curve with observation time, again to satisfy (\ref{eq:hc}). This discrepancy is an unfortunate result of the conventions in use by the different GW communities; however it is also natural given the sources under observation. When observing a transient source, such as a burst or inspiral, which changes within the lifetime of the detector it is natural to consider the detector as performing constantly while the signal changes. However when observing a monochromatic source or a stochastic background which is unchanging over the mission lifetime it is more natural to consider the source as being fixed and the sensitivity of the detector gradually improving. All that is required by the definition in (\ref{eq:hc}) is that the ratio $h_{c}(f)/h_{n}(f)$ is constant.

\subsubsection{EPTA/PPTA/NANOGrav}
The PTAs currently in operation are the EPTA (European Pulsar Timing Array), PPTA (Parkes Pulsar Timing Array, based in  Australia) and NANOGrav (North American Nanohertz Observatory for GWs). There are published limits on the amplitude of the stochastic background from all three detectors: currently the lowest is from the EPTA, \cite{Haasteren}. The EPTA curves in the figures show the  published limit based on an analysis of 5 pulsars over approximately 10 years.

\subsubsection{IPTA}
Combining the existing arrays would yield a single PTA using approximately 4 times as many pulsars pulsars. The IPTA curves plotted in the figures are based on 20 pulsars timed for twice as long as the current EPTA.

\subsubsection{SKA}
The sensitivity curve plotted in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega} for the SKA (Square Kilometer Array) assume a values of $\delta t^{2}_{\textrm{r.m.s.}}$ a factor of 10 better than current PTAs and 20 pulsars timed for twice as long as the current EPTA.

### 2014-06-27 Introduction.tex
First commit

\section{Introduction}

The next few years promises to deliver the first direct detection of gravitational waves (GWs). This will most likely be achieved by the advanced LIGO \citep{2010CQGra..27h4006H} detector operating in the frequency range $(10-10^{3})\,\textrm{Hz}$. By the end of the decade it is expected that pulsar timing arrays \citep{1990ApJ361300F} will detect very low frequency GWs around $10^{-8}\,\textrm{Hz}$. Further into the future space based detectors, such as eLISA \citep{2012CQGra..29l4016A}, will probe GWs in the millihertz regime. This will herald the beginning of multi-wavelength GW astronomy. There already exists an extensive literature assessing the potential of all of these detectors to probe the astrophysics of various potential sources. There are several different methods commonly used to describe the sensitivity of a GW detector and the strength of a GW source. It is common practice in GW astronomy to summarise this information on a sensitivity curve plot. When producing these plots it is desirable to have a consistent convention between detectors and sources which allows information about both detectors and sources to be plotted on the same graph. Another desirable feature of such a plot is that the detectors and sources are represented in such a way that the relative detectability of the sources is immediately apparent from the plot.

In this work we discuss the differing conventions commonly used in sensitivity curve plots. The amplitude of a GW is a strain, a dimensionless quantity $h$. This gives a fractional change in length, or equivalently light travel time, across a detector. The strain is small, making it a challenge to measure. To calibrate our expectations for detection it is necessary to quantify the sensitivity of our instruments and the strength of the target signals. There are three commonly used descriptions: the characteristic strain, the power spectral density and the spectral energy density. We aim to disambiguate these three and give a concrete comparison of different detectors. It is hoped that this will provide a useful reference for new and experienced researchers in this field alike.

We begin by expounding the various conventions and the relationships between them in sections \ref{sec:conventions} and \ref{sec:voc}. A review of GW detectors (both current and proposed for the future) is given in section \ref{sec:detectors} and a review of GW sources is given in section \ref{sec:sources}. In \ref{app:a} several example sensitivity curves are presented; there is also a website at \url{http://www.ast.cam.ac.uk/~rhc26/sources/} which can be used to generate similar figures with user specified detectors and sources included.

\section{Signal analysis}\label{sec:conventions}

Gravitational radiation has two independent polarization states denoted $+$ and $\times$; a general signal can be described as a linear combination the two polarization states, $h \nobreak=\nobreak A_{+} h_{+}\nobreak +\nobreak A_{\times} h_{\times}$. The sensitivity of a detector to each of these depends upon the relative orientations of the source and detector. The output of a gravitational wave detector, $s(t)$, will contain a superposition of noise, $n(t)$, and possibly a signal, $h(t)$,
\begin{equation} s(t)=n(t)+h(t) \; . \end{equation}
It will be necessary to work with the Fourier transform of the signal, the most common convention used in the GW community is
\begin{eqnarray} x(t)&={\cal{F}}^{-1}\left\{\tilde{x}(f)\right\}&=\int_{-\infty}^{\infty}\textrm{d}f\;\tilde{x}(f)\exp (2\pi \rmi f t )  \nonumber \\
\tilde{x}(f) &=\,\;{\cal{F}}\left\{x(t)\right\}&=\int_{-\infty}^{\infty}\textrm{d}t\;x(t)\exp (-2\pi \rmi f t )\; .\end{eqnarray}
For simplicity it is assumed that the noise in the GW detector is stationary and Gaussian (with zero mean), under these assumptions the noise is fully characterised via the \emph{one-sided} noise power spectral density, $S_{n}(f)$,\footnote{An alternative convention is to use the two-sided power spectral density $S^{(2)}_{n}(f)=S_{n}(f)/2$.}, 
\begin{equation}\label{eq:psd} \left<\tilde{n}(f)\tilde{n}^{*}(f')\right>=\frac{1}{2}\delta (f-f')S_{n}(f) \; .\end{equation}
Angled brackets $\left<\ldots\right>$ denotes an ensemble average over many realisations of the noise. In reality we have only a single system to work with, so the ensemble average is replaced by a time average. The procedure is to measure the noise over a sufficiently long duration, $T$, and then compute the Fourier transform, $\tilde{n}(f)$, with a frequency resolution $\Delta f = T^{-1}$. This is then repeated many times to give the ensemble average; this procedure assumes the system is ergodic. Since the GW signal and detector output are real it follows that $\tilde{h}(-f)=\tilde{h}^{*}(f)$ and $\tilde{n}(-f)=\tilde{n}^{*}(f)$; therefore, $S_{n}(f)=S_{n}(-f)$. The fact the $S_{n}(f)$ is an even function means that Fourier integrals over all frequencies can instead be written as integrals over positive frequencies only (see, for example, equations (\ref{eq:meansquare1}) and (\ref{eq:snrinnerprod})), for this reason $S_{n}(f)$ is called the \emph{one-sided} power spectral density. It is straightforward to show that the power spectral density has the property that integrated over all positive frequencies it gives the mean square noise amplitude in the detector.
\begin{eqnarray}\label{eq:meansquare}
\overline{n(t)^{2}}&=\lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T}\textrm{d}t\;\left| n(t) \right|^{2} \nonumber \\
&=\lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T}\textrm{d}t\; \int_{-\infty}^{\infty}\textrm{d}f\;\int_{-\infty}^{\infty}\textrm{d}f'\;\tilde{n}(f)\tilde{n}^{*}(f')\exp\left(2\pi \rmi (f-f')t \right)
\end{eqnarray}
Using the property of Fourier transforms, ${\cal{F}}\left\{ n(t+\tau) \right\}(f)=\tilde{n}(f)\exp (-2\pi \rmi f\tau )$, the exponential factors in equation (\ref{eq:meansquare}) may be absorbed as a time translation of the noise by an amount $t$. The noise is stationary and so it's long term time average is invariant under this time translation and is given by Equation (\ref{eq:psd}).
\begin{eqnarray}\label{eq:meansquare1}
\overline{n(t)^{2}}& = \lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T}\int_{-\infty}^{\infty}\textrm{d}f\;\int_{-\infty}^{\infty}\textrm{d}f'\;{\cal{F}}\left\{ n(\tau-t) \right\}(f){\cal{F}}\left\{ n(\tau-t) \right\}(f')^{*} \nonumber \\
&=\int_{-\infty}^{\infty}\textrm{d}f\;\int_{-\infty}^{\infty}\textrm{d}f'\;\left<\tilde{n}(f)\tilde{n}^{*}(f')\right>\nonumber \\
& = \int_{0}^{\infty}\textrm{d}f\; S_{n}(f)
\end{eqnarray}

Using the property of Fourier transforms, ${\cal{F}}\left\{ n(t+\tau) \right\}(f)=\tilde{n}(f)\exp (-2\pi \rmi f\tau )$, the time average can be turned into an ensemble average and hence the definition of power spectral density in (\ref{eq:psd}) may be used.
\begin{eqnarray}\label{eq:meansquare1}
%\left<n(t)^{2}\right>&=\lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-\infty}^{\infty}\textrm{d}f\;\int_{-\infty}^{\infty}\textrm{d}f'\;\int_{-T}^{T}\textrm{d}t\; {\cal{F}}\left\{n(\tau+t)\right\}\left[{\cal{F}}\left\{n(\tau+t)\right\}\right]^{*}\nonumber \\
\overline{n(t)^{2}}& = \lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-\infty}^{\infty}\textrm{d}f\;\int_{-\infty}^{\infty}\textrm{d}f'\;\int_{-T}^{T}\textrm{d}t\; \tilde{n}(f)\tilde{n}^{*}(f') \nonumber \\
&=\int_{-\infty}^{\infty}\textrm{d}f\;\int_{-\infty}^{\infty}\textrm{d}f'\;\left<\tilde{n}(f)\tilde{n}^{*}(f')\right>\nonumber \\
%&= \int_{-\infty}^{\infty}\textrm{d}f\;\int_{-\infty}^{\infty}\textrm{d}f'\;\frac{1}{2}S_{n}(f)\delta (f-f')\nonumber\\
& = \int_{0}^{\infty}\textrm{d}f\; S_{n}(f)
\end{eqnarray}
Given a detector output the challenge is to extract the signal. There is a well known solution to this problem which involves constructing a Wiener optimal filter. Let $K(t)$ be the filter function with Fourier transform $\tilde{K}(f)$.
Convolving this with the detector output gives a contribution from the signal and a contribution from the noise
\begin{equation}\label{eq:conv} \left(s*K\right)(\tau)=\int_{-\infty}^{\infty}\textrm{d}t\;(n(t)+h(t))K(t-\tau)=S+N\; .\end{equation}
The signal contribution, $S$, is defined as the expectation of the convolution in (\ref{eq:conv}) when a signal is present, maximised by varying the offset $\tau$. Since the expectation of pure noise is zero it follows that
\begin{equation} S = \int_{-\infty}^{\infty}\textrm{d}t\;h(t)K(t)=\int_{-\infty}^{\infty}\textrm{d}t\;h(t)K^{*}(t)=\int_{-\infty}^{\infty}\textrm{d}f\; \tilde{h}(f)\tilde{K}^{*}(f) \; .\end{equation}
The squared contribution from noise, $N^{2}$, is defined as the mean square of the convolution in (\ref{eq:conv}) when no signal is present,
\begin{eqnarray} 
N^{2}&=\int_{-\infty}^{\infty}\textrm{d}t\;\int_{-\infty}^{\infty}\textrm{d}t'\;K(t)K(t')\left<n(t)n(t')\right> \nonumber \\
&=\int_{-\infty}^{\infty}\textrm{d}t\;\int_{-\infty}^{\infty}\textrm{d}t'\;K(t)K^{*}(t')\int_{-\infty}^{\infty}\textrm{d}f\;\int_{-\infty}^{\infty}\textrm{d}f'\;\left<\tilde{n}(f)\tilde{n}^{*}(f')\right>\exp\left(2\pi\rmi(ft-f't')\right)\nonumber \\
%&=\int_{-\infty}^{\infty}\textrm{d}f\;\int_{-\infty}^{\infty}\textrm{d}t\;\int_{-\infty}^{\infty}\textrm{d}t'\;\frac{1}{2}S_{n}(f)K(t)K^{*}(t')\exp\left(2\pi\rmi f (t-t')\right)\nonumber \\
&=\int_{-\infty}^{\infty}\textrm{d}f \; \frac{1}{2}S_{n}(f)\tilde{K}(f)\tilde{K}^{*}(f) \; .
 \end{eqnarray}
Where the definition of $S_{n}(f)$ from (\ref{eq:psd}) has been used. Hence the optimum signal-to-noise ratio (SNR), $\varrho$, is given by
\begin{equation}\label{eq:SNRinnerprod} 
\varrho^{2}= \frac{(S)^{2}}{N^{2}}= \frac{\left( \frac{1}{2}S_{n}(f)\tilde{K}(f) \Big| \tilde{h} \right)^{2}}{\left( \frac{1}{2}S_{n}(f)\tilde{K}(f) \Big| \frac{1}{2}S_{n}(f)\tilde{K}(f) \right)} \;,
\end{equation}
where the inner product between a pair of signals $A$ and $B$ has been defined
\begin{equation}\label{eq:snrinnerprod} \left( A | B \right) = 4\Re\left\{\int_{0}^{\infty}\textrm{d}f\;\frac{\tilde{A}^{*}(f)\tilde{B}(f)}{S_{n}(f)}\right\} \; .\end{equation}
The optimum filter is that which maximises the SNR in (\ref{eq:SNRinnerprod}). From the Cauchy-Schwarz inequality it follows that the optimum filter is
\begin{equation} \tilde{K}(f)=\frac{\tilde{h}(f)}{S_{n}(f)} \; . \end{equation}
This is the Wiener filter, it may be multiplied by an arbitrary constant since this will not change the SNR. Notice that in order to construct the Wiener filter it is necessary to know \emph{a priori} the form of the signal, $\tilde{h}(f)$. It is common to define the noise amplitude and characteristic strain as
\begin{eqnarray}\label{eq:strains} h_{n}(f)^{2}&=fS_{n}(f) \\
\label{eq:strains1} \label{eq:strainsagain}               h_{c}(f)^{2}&=4f^{2}\left| \tilde{h}(f) \right|^{2} \; , \end{eqnarray}
so that the SNR in (\ref{eq:SNRinnerprod}) may be written compactly as
\begin{equation}\label{eq:hc} 
\varrho^{2}=\int_{0}^{\infty} \textrm{d}\left(\log f\right)\; \left(\frac{h_{c}(f)}{h_{n}(f)}\right)^{2} \;.
\end{equation}
The strain amplitudes $h_{n}(f)$ and $h_{c}(f)$ are dimensionless, while $S_{n}(f)$ has units of inverse frequency.

Whilst the magnitude of the Fourier transform of the signal, $h(f)=\left|\tilde{h}(f)\right|$, provides a simple quantification of the GW amplitude as a function of frequency it has one main deficiency. For an inspiralling source the instantaneous amplitude can be orders of magnitude below the noise level in the detector; however, as the signal continues over many orbits the SNR can be integrated up to a detectable level. It is desirable to have a quantification of the GW amplitude that accounts for this effect; three such methods are described in the following sections.

\subsection{Characteristic strain}
One sensible choice of quantities to plot on a sensitivity curve is $h_{n}$ for the detector and $h_{c}$ for the source (see figure \ref{fig:hc}). Using this convention, when plotting on a log-log scale, the area between the source and detector curves is related to the SNR via (\ref{eq:hc}). This convention allows the reader to ``integrate by eye" for a given detector to assess the detectability of a given source. Also, when using this convention the reader does not need to worry about the observation time for each source because this is already accounted for in the definition of characteristic strain. An additional advantage of this convention is that the values on the strain axis for the detector curve have a simple physical interpretation, they correspond to the root-mean-square noise in a bandwidth $f$. One downside to plotting characteristic strain is that the values on the strain axis do not directly relate to the amplitude of the waves from the source. Another disadvantage is that the characteristic strain for a monochromatic source is infinite!

\subsection{Power spectral density}\label{sec:psd}
Another common quantity to plot on sensitivity curves is the square root of the power spectral density (see figure \ref{fig:S}), which from (\ref{eq:strains}) is given by
\begin{eqnarray}\label{eq:temp1} \sqrt{S_{n}(f)}&=h_{n}(f)f^{-1/2} \quad \textrm{for detectors, and by analogy} \\
\sqrt{S_{h}(f)}&=h_{c}(f)f^{-1/2} \quad \textrm{for sources.}\label{eq:ShforSources}\end{eqnarray}
The power spectral density is the most commonly plotted quantity in sensitivity curves in the literature and it has the nice property, proven in equation (\ref{eq:meansquare1}), that integrated over all positive frequencies it gives the mean square amplitude of the signal in the detector.

\subsection{Energy density}
A third way of describing the detectability of a GW is through the energy carried by the waves, this quantity has the obvious advantage of having a clear physical significance. This quantity is most commonly used in sensitivity curves showing stochastic backgrounds of GWs, see section \ref{sec:stoch}. The energy in GWs is described by the Isaacson stress tensor \citep{MTW}
\begin{equation} T_{\mu\nu}=\frac{c^{2}}{32\pi G}\left<\partial_{\mu}\bar{h}_{\alpha\beta}\partial_{\nu}\bar{h}^{\alpha\beta}\right> \;.\end{equation}
The energy density, $\rho c^{2}$, is given by the $T_{00}$ component of this tensor. Using this together with (\ref{eq:psd}) to calculate the energy density in the GW signal gives
\begin{eqnarray}\label{eq:specNRGdensity}
\rho c^{2}&=\frac{c^{2}}{16\pi G}\int_{-\infty}^{\infty}\textrm{d}f\;\left(2\pi f\right)^{2}\tilde{h}(f)\tilde{h}^{*}(f)=\int_{0}^{\infty}\textrm{d}f\;\frac{\pi c^{2}}{4G}f^{2}S_{h}(f)\; ,
\end{eqnarray} 
where the definitions (\ref{eq:ShforSources}) and (\ref{eq:strainsagain}) have been used. Equation (\ref{eq:specNRGdensity}) motivates the definition of the spectral energy density in GWs, $S_{\textrm{E}}$, the energy per unit volume of space, per unit frequency, \cite{HellingsDowns}
\begin{equation}\label{eq:spectralenergydensity} S_{\textrm{E}}(f)=\frac{\pi c^{2}}{4G} f^{2}S_{h}(f) \; ,\end{equation}
and a corresponding expression for the noise with $S_{n}(f)$. It is usual to define the dimensionless quantity $\Omega_{\textrm{GW}}$ as the energy density per logarithmic frequency interval normalised to the critical density of the universe
\begin{equation}\label{eq:omega} \Omega_{\textrm{GW}}(f)=\frac{fS_{\textrm{E}}(f)}{\rho_{\textrm{c}}c^{2}}\; . \end{equation}
The critical density of the universe is $\rho_{\textrm{c}}=3H_{0}^{2}/8\pi G$, where $H_{0}=h_{100}\times 100\, \textrm{km}\,\textrm{s}^{-1}\,\textrm{Mpc}^{-1}$ is the Hubble constant (this equation serves as a definition of $h_{100}$, which here has nothing to do with strain). The most common quantity related to energy density to be plotted on sensitivity curves is $\Omega_{\textrm{GW}}h_{100}^{2}$ (figure \ref{fig:omega}). This quantity has one aesthetic advantage over the others: it accounts for the fact that there is less energy in low frequency waves of the same amplitude and does not place the sensitivity curves of very low frequency detectors much higher than most ground based detectors.

\subsection{Relating the different descriptions}
The dimensionless energy density in GWs, $\Omega_{\textrm{GW}}$, the spectral energy density in GWs, $S_{\textrm{E}}$, the one-sided power spectral density, $S_{h}$, and the characteristic strain are all related via
\begin{equation}\label{eq:omega} 
\Omega_{\textrm{GW}}(f)=\frac{fS_{\textrm{E}}(f)}{\rho_{\textrm{c}}c^{2}}=\frac{\pi}{4G\rho_{\textrm{c}}}f^{2}h_{c}(f)^{2}=\frac{\pi}{4G\rho_{\textrm{c}}}f^{3}S_{h}(f)  \; ,
\end{equation}
and corresponding expressions for the noise with $S_{n}(f)$ and $h_{n}(f)$.

### 2014-06-27 Plots.tex
First commit

\appendix
\section{Sensitivity curves}\label{app:a}
The plots in this section show all of the detectors and sources described in the text. Interactive versions of these plots may be viewed on-line, \url{http://www.ast.cam.ac.uk/~rhc26/sources/}. The detector noise curves all have their resonance spikes removed for clarity. 
\begin{figure}[h!]
 \centering
 \includegraphics[trim=0cm 0cm 0cm 0cm, width=0.85\textwidth]{figure1.pdf}
 \caption{A plot of characteristic strain against frequency for a variety of detectors and sources.}
 \label{fig:hc}
\end{figure}
\begin{figure}[h!]
 \centering
 \includegraphics[trim=0cm 0cm 0cm 0cm, width=0.85\textwidth]{figure2.pdf}
 \caption{A plot of the square root of power spectral density against frequency for a variety of detectors and sources.}
 \label{fig:S}
\end{figure}
\begin{figure}[h!]
 \centering
 \includegraphics[trim=0cm 0cm 0cm 0cm, width=0.85\textwidth]{figure3.pdf}
 \caption{A plot of the dimensionless energy density in GWs against frequency for a variety of detectors and sources.}
 \label{fig:omega}
\end{figure}

### 2014-06-27 Sources.tex
First commit

\section{Astrophysical sources}\label{sec:sources}
All the sources described here are represented by boxes in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega}. The boxes are drawn in such a way that there is a reasonable event rate for sources at a detectable SNR. Descriptions of each source are given in the following sections. Sources with short durations (i.e., burst sources) and sources which evolve in time over much longer timescales than our observations are drawn with flat topped boxes. Inspiralling sources which change their frequency over observable timescales are drawn with sloping tops. The slope for the top of the box is chosen to be $-2/3$ for all sources, which is also the appropriate gradient for a stochastic backgrounds of binaries. The width of the box gives the range of frequencies sources of a given type can have a while remaining at a detectable amplitude.

\subsection{Sources for ground based detectors}

\subsubsection{Neutron star binaries}
The inspiral and merger of a pair of neutron stars is the primary target for ground based detectors. The expected event rate for this type of source is uncertain, but estimates centre around $\rho_{\textrm{NS-NS}}=1.3\times 10^{-4}\;\textrm{Mpc}^{-3}\textrm{yr}^{-1}$, \cite{CBC}. Plotted in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega} is a box with an amplitude such that it produces a ratio $h_{c}/h_{n}=16$ for advanced LIGO at peak sensitivity and a width between $(3-300)\,\textrm{Hz}$.

\subsubsection{Supernovae}
Simulations of core collapse supernova events show that GWs between $(10^{2}-10^{3})\,\textrm{Hz}$ can be produced. The GW signal undergoes ${\cal{O}}(1)$ oscillation and hence is burst-like. \cite{2002A&A...393..523D} calculate the average maximum amplitude of GWs for a supernova at distance $r$ as
\begin{equation} h=8.9\times 10^{-21}\left( \frac{10 \,\textrm{kpc}}{r} \right) \; .\end{equation}
The event rate for supernova is approximately $\rho_{\textrm{SN}}=5\times10^{-4}\,\textrm{Mpc}^{-3}\textrm{yr}^{-1}$. The box plotted in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega} corresponds to a distance $r=3 \,\textrm{Mpc}$ with the frequency range quoted above. The LIGO and VIRGO detectors have already placed bounds on the event rate for these sources, \cite{Bursts}.

\subsubsection{Continuous waves from rotating neutron stars}

{\cjm{If someone is feeling keen they can write a very short section on these, maybe using this reference \cite{Einstein@Home}}}.

\subsection{Sources for space based detectors}
For a review of the GW sources for space based missions see, for example, \cite{Amaro-Seoane-et-al}, \cite{Gairetal} of the \cite{eLISAyellowbook}.

\subsubsection{Massive black hole binaries}
Space based detectors will be sensitive to equal mass mergers in the range $(10^{4}-10^{7})\,\Msun$. Predictions of the event rate for these mergers range from ${\cal{O}}(3-300)\,\textrm{yr}^{-1}$ for LISA with SNRs of up to 10000. The uncertainty in the rate reflects our uncertainty in the growth mechanisms of the supermassive black hole population. Plotted in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega} is a box with a ratio $h_{c}/h_{n}=100$ for eLISA at it's peak sensitivity. The range of frequencies plotted is $(3\times 10^{-4}-3\times 10^{-1})\,\textrm{Hz}$, this corresponds to circular binaries in the mass range quoted above.

\subsubsection{Galactic white dwarf binaries} \label{sec:GB}

\cjm{This section should probably be larger and more detailed. These are by far the most numerous sources and at the moment I don't even mention verification binaries.}

These divide into two classes, the unresolvable and the resolvable galactic binaries. The distinction between resolvable and unresolvable is detector specific; here we choose LISA. For the unresolvable binaries the box plotted corresponds to the estimate of the background due to \cite{Nelemans},
\begin{equation} h_{c}(f)= 5\times 10^{-21} \left(\frac{f}{10^{-3}\,\textrm{Hz}}\right)^{-2/3} \; . \end{equation}
This background is unchanging over the mission lifetime, therefore, as discussed in section \ref{sec:PTAgeneralproperties}, when plotting the amplitude on a sensitivity curve plot an observation time must be specified; here an observation time of one year is used. Estimates for the event rate of resolvable binaries centre around ${\cal{O}}(10^{3})$ events for eLISA. The box plotted in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega} has an ratio $h_{c}/h_{n}=50$ for eLISA at its peak sensitivity, the frequency range of the box is $\left(3\times10^{-4}-10^{-2}\right)\,\textrm{Hz}$, based on visual inspection of Monte Carlo population simulation results presented in \cite{Amaro-Seoane-et-al}.

\subsubsection{Extreme mass ratio inspirals}
EMRI events occur when a compact stellar mass object inspirals into a supermassive black hole. There is extreme uncertainty in the event rate for EMRIs due to the poorly constrained astrophysics in galactic centres; the best guess estimate is ${\cal{O}}(10)$ events per year with eLISA with SNR $\ge 20$. The box plotted in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega} has a characteristic strain of $h_{c}=3\times 10^{-20}$ at $10^{-2}\;\textrm{Hz}$ which corresponds to a $10\Msun$ BH inspiralling into a $10^{6}\Msun$ black hole at a luminosity distance of $1\,\textrm{Gpc}$. The frequency width of the box is somewhat unknown, EMRI events can occur into a black hole of any mass, and hence EMRIs can occur at any frequency. However there is substantial uncertainty in the black hole mass function, so the box in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega} is drawn with a width comparable to that of the LISA sensitivity curve.

\subsection{Sources for PTAs}

\subsubsection{Supermassive black hole binaries}
The main target for PTAs is a stochastic background of GWs produced by a population of supermassive black hole binaries at cosmological distances. Supermassive black holes are known to lie at the centres of most galaxies and the black hole mergers are associated with the mergers of the host galaxies. The current best published limit for the amplitude of the stochastic background is $h_{c}=6\times 10^{-15}$ at a frequency of $f_{0}=1\,\textrm{yr}^{-1}$, \cite{Haasteren}. There is strong theoretical evidence that the actual background lies close to the current limit, \cite{imminentdetectionofgravitationalwaves} and \cite{Sesana-2012}.

Sources at higher frequencies are inspiralling faster and hence there are fewer of them per frequency bin. At a certain frequency, these sources will cease to be a background and become individually resolvable. The exact distinction between resolvable and unresolvable is detector specific. Plotted in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega} for the unresolvable background is a third of the current limit with a cut off frequency of $f=1\,\textrm{yr}^{-1}$ which is suggested by Monte Carlo population studies, \cite{SesanaVecchioColancino}. For the resolvable sources the amplitude of the current limit is plotted between $\left(3\times 10^{-9} - 3\times 10^{-7}\right)\,\textrm{Hz}$.

\subsection{Cosmological sources}
\cjm{Christopher, would you be happy to write a section on very high redshift sources of stochastic backgrounds. Stressing that these in principal can be at any frequency. And also saying that because of the lack of certainty here we don't include these sources in the figures. Maybe reference recent excitement caused by BICEP2 results.}

### 2014-06-27 Typesofsource.tex
First commit

\section{Types of source}\label{sec:voc}

Sources of GWs can be broadly split into three catagories: well modelled sources, unmodelled (or poorly-modelled) transient sources and unmodelled stochastic backgrounds. The classic example of a well modelled source is the inspiral of two compact objects, this is discussed in section \ref{sec:insp}. Stochastic backgrounds can either be formed from many overlapping sources which are each individually susceptible to being modelled, or from some intrinsically random process, these are discussed in section \ref{sec:stoch}. The classic example of a unmodelled (or poorly-modelled) transient source is a supernova, searches of this type are often called burst searched and are discussed in section \ref{sec:bursts}.

\subsection{Inspirals}\label{sec:insp}
Inspiralling binaries spend a variable amount of time in each frequency band. If $\phi$ is the orbital phase then the number of cycles generated at frequency $f$ can be estimated as
\begin{equation}\label{eq:inspiral} {\cal{N}}_{\textrm{cycles}} = \frac{f}{2\pi}\frac{\textrm{d}\phi}{\textrm{d}f}=\frac{f^{2}}{\dot{f}} \; ,\end{equation}
where an overdot represents the time derivative and $f=\dot{\phi}$. The squared SNR will scale with ${\cal{N}}_{\textrm{cycles}}$, so it would be expected that $h_{c}(f)\approx \sqrt{{\cal{N}}_{\textrm{cycles}}}\tilde{h}(f)$. The exact conversion can be derived by considering the Fourier transform in the stationary phase approximation. Consider a source signal with approximately constant amplitude, $h$, and central frequency $f'$.
\begin{eqnarray}&h(t)= h\exp\left(\rmi \phi(t)\right) \\
\Rightarrow &\tilde{h}(f)=\int_{-\infty}^{\infty}d\textrm{t}\;h\exp \left(2\pi \rmi \left(\frac{\phi(t)}{2\pi t}-f\right)t\right)\end{eqnarray}
Without loss of generality assume an initial phase of zero and expand the exponent in powers of $t$,
\begin{eqnarray} \label{eq:FTofEMRI} \tilde{h}(f) &\approx \int_{-\infty}^{\infty}\rmd t\; h \exp\left[2\pi\rmi \dot{f'} \left(t^{2}+\frac{f'-f}{\dot{f'}}t\right)\right]\nonumber \\
&\approx h\exp\left[-\pi\rmi\frac{\left(f'-f\right)^{2}}{2\dot{f'}}\right]\int_{-\infty}^{\infty}\rmd t\;\exp\left[2\pi\rmi\dot{f'}\left(t+\frac{f'-f}{2\dot{f'}}\right)^{2}\right] \nonumber \\
&\approx\frac{h}{\sqrt{2i\dot{f'}}}\exp\left[-\pi\rmi\frac{\left(f'-f\right)^{2}}{2\dot{f'}}\right]\; .
\end{eqnarray}
From (\ref{eq:strains}) and (\ref{eq:FTofEMRI}) the characteristic strain and the Fourier transform can be related for inspiralling sources \citep{FinnThorne}
\begin{equation}\label{eq:insphc}h_{c}(f) = \sqrt{\frac{2f^{2}}{\dot{f}}}h\;.\end{equation}
Equation (\ref{eq:insphc}) is the relation between $h_{c}(f)$ and the instantaneous amplitude $h$ for an inspiralling source; for other types of source a new relation which satisfies (\ref{eq:hc}) will have to be found. Equation (\ref{eq:hc}) should be considered as the definition of characteristic strain and (\ref{eq:insphc}) a consequence of it for inspirals.

\subsection{Stochastic backgrounds}\label{sec:stoch}
Another important source of GWs are stochastic backgrounds due to large populations of individually unresolvable sources. The population of sources will in general be at cosmological distances and it is necessary to distinguish the frequency in the source rest frame, $f_{\textrm{r}}$, from the measured frequency, $f$; these are related through the redshift, $z$, via $f_{\textrm{r}}=(1+z)f$. The comoving number density of sources, $n$, producing the background will also be a function of redshift; if the sources producing the stochastic background are all in the local universe then simply set $n(z)=\delta (z)$ and replace $d_{\textrm{L}}(z)$ with $d$ in all that follows (where $d_{\textrm{L}}(z)$ and $d$ are respectively the luminosity and comoving distances to the source, $d_{\textrm{L}}(z)=(1+z)d_{m}$). Equation (\ref{eq:omega}) gives an expression for the energy density in GWs per logarithmic frequency interval,
\begin{equation}\label{eq:stoch} fS_{\textrm{E}}(f)=\frac{\pi c^{2}}{4G}f^{2}h_{c}(f)^{2} \; . \end{equation}
The total energy emitted in the logarithmic frequency interval $\textrm{d}\left(\log (f_{\textrm{r}})\right)$ by a single binary in the population is $\left(\textrm{d}E_{\textrm{GW}}/\textrm{d}(\log f_{\textrm{r}})\right)\textrm{d}(\log f_{\textrm{r}})$; the energy density may be written as
\begin{equation}\label{eq:Phinney} fS_{\textrm{E}}(f)=\int_{0}^{\infty}dz\; \frac{\textrm{d}n}{\textrm{d}z}\frac{1}{(1+z)}\frac{1}{d_{\textrm{L}}(z)^{2}}\frac{\textrm{d}E_{\textrm{GW}}}{\textrm{d}\left(\log f_{\textrm{r}} \right)} \; , \end{equation}
where the factor of $\left( 1+z \right)^{-1}$ accounts for the redshift of the energy.

For simplicity consider all the binaries comprising the background to be in circular orbits with frequencies $\nu=f_{\textrm{r}}/2$, and to be far from their last stable orbit; the energy radiated may then be calculated using the quadrupole approximation, \cite{petersmathews1963}. The chirp mass is defined as ${\cal{M}}=\mu^{3/5}M^{2/5}$, where $\mu$ is the reduced mass and $M$ is the total mass of the binary. The number density of sources will also be a function of chirp mass, $n(z,{\cal{M}})$. The energy in GWs from a single binary per logarithmic frequency interval is, 
\begin{equation}\label{eq:Thorne} \frac{\textrm{d}E_{\textrm{GW}}}{\textrm{d}\left(\log f_{\textrm{r}} \right)}=\frac{G^{2/3}\pi^{2/3}}{3}{\cal{M}}^{5/3}f_{\textrm{r}}^{2/3} \; , \end{equation}
between an minimum and maximum frequency set by the initial and final radius of the binary orbit. Here we assume that the maximum and minimum frequencies are outside of the range of frequencies probed by our detector and hence may be neglected. Using (\ref{eq:stoch}), (\ref{eq:Phinney}) and (\ref{eq:Thorne}) an expression for characteristic strain can now be found, \cite{SesanaVecchioColancino}
\begin{equation}\label{eq:bigint}
h_{c}(f)^{2}=\frac{4G^{5/3}}{3\pi^{1/3}c^{2}}f^{-4/3}\int_{0}^{\infty}\textrm{d}z\;\int_{0}^{\infty}\textrm{d}{\cal{M}}\;\frac{\textrm{d}^{2}n}{\textrm{d}z\,\textrm{d}{\cal{M}}}\frac{1}{d_{\textrm{L}}(z)^{2}}\left( \frac{{\cal{M}}^{5}}{1+z} \right)^{1/3}\; .
\end{equation}
From (\ref{eq:bigint}) it can be seen that the characteristic strain due to a stochastic background of binaries is a power law in frequency with spectral index $\alpha=-2/3$. The amplitude of the background depends on the population statistics of the binaries under consideration via $n(z,{\cal{M}})$. The power law is often parametrised as
\begin{equation}\label{eq:power} h_{c}(f)= A\left(\frac{f}{f_{0}}\right)^{\alpha}\; , \end{equation}
and constraints are then placed on $A$. In practice this power law will also have upper and lower frequency cut-offs related to the population of objects creating the GW spectrum. The stochastic background due to other sources, such as cosmic strings or relic GWs from the big bang, can also be written in the same form as (\ref{eq:power}) but with different spectral indices: $\alpha=-7/6$ for cosmic strings or $\alpha$ in the range $-1$ to $-0.8$ for relic GWs, \cite{Jenet}.

\subsection{Burst sources}\label{sec:bursts}

\cjm{This section is currently much shorter than both \ref{sec:insp} and \ref{sec:stoch}, Rob said he was going to add more detail on searching for unmodelled phenomenon using time-frequency plots.}

A signal is burst-like if it's duration at a detectable amplitude is only a few times the wave period. If this is the case then the signal does not have time to accumulate SNR in each frequency band in the way inspirals do. If the source generates ${\cal{N}}_{\textrm{cycles}}$ of radiation then the relation between $h_{c}(f)$ and $h$ is simply
\begin{equation}\label{eq:simple} h_{c}(f)=h\sqrt{{\cal{N}}_{\textrm{cycles}}} \; . \end{equation}

### 2014-06-27 noise_curve.tex
First commit

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    INSTITUTE OF PHYSICS PUBLISHING                                   %
%                                                                      %
%   `Preparing an article for publication in an Institute of Physics   %
%    Publishing journal using LaTeX'                                   %
%                                                                      %
%    LaTeX source code `ioplau2e.tex' used to generate `author         %
%    guidelines', the documentation explaining and demonstrating use   %
%    of the Institute of Physics Publishing LaTeX preprint files       %
%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
%                                                                      %
%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[fleqn,12pt]{iopart}
\newcommand{\gguide}{{\it Preparing graphics for IOP journals}}
%Uncomment next line if AMS fonts required
\usepackage{iopams}  
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{url}
\usepackage{natbib}
\usepackage{appendix}
\usepackage[margin=1cm]{caption}
\usepackage{fullpage}
\usepackage{hyperref}
\usepackage{color}

\def\newblock{\hskip .11em plus .33em minus .07em}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% 	definition of a solar mass symbol for use in equations.      %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\Msun}{\ensuremath{M_{ \odot }}}
\newcommand{\degree}{\ensuremath{^\circ}}
\newcommand{\cplb}[1]{{\color{red} #1}}
\newcommand{\rhc}[1]{{\color{green} #1}}
\newcommand{\cjm}[1]{{\color{blue} #1}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\title{Gravitational wave sensitivity curves}
\author{C J Moore}
\address{Institute of Astronomy, Madingley Road, Cambridge, CB3 0HA, United Kingdom}
\ead{cjm96@ast.cam.ac.uk}

\begin{abstract}
There are several common conventions in use by the gravitational wave community to describe the amplitude of sources and the sensitivity of detectors. These are frequently confused. We outline the merits and differences between the various quantities used for parametrising noise curves and characterising gravitational wave amplitudes. Sensitivity curves plots for each of the conventions are presented and interactive versions have been made available on-line at \url{http://www.ast.cam.ac.uk/~rhc26/sources/}.
\end{abstract}

\input{Introduction}
\input{Typesofsource}
\input{Detectors}
\input{Sources}

\section{Concluding remarks}\label{sec:discussion}
When quantifying the sensitivity of a GW detector and the loudness of a GW source there are three commonly used quantities: the characteristic strain, the power spectral density, and the spectral energy density. Here we have carefully defined these quantities and derived the relationships between them. In addition example sensitivity curve plots using each of these quantities have been presented for a wide range of ground based detectors, space based detectors and pulsar timing arrays. Also plotted are predicted amplitudes for a wide range of sources, these predictions are based on astrophysical estimates of the event rates for a range of processes and are subject to varying degrees of uncertainty. There also exists the exciting possibility of sources nobody has yet considered. Interactive versions of these plots with user specified detectors and sources are made available on-line at \url{http://www.ast.cam.ac.uk/~rhc26/sources/}. Trying to summarise an entire field of astronomy on one plot is an impossible task; however we hope that the figures and analysis presented here provide useful insight.

\ack{Christopher Moore is supported by the STFC. I would like to thank Dr Jonathan Gair, Christopher Berry and Robert Cole for help producing this work. I would also like to thank Robert Cole for help producing the figures on the website. }

\bibliographystyle{apj}
\bibliography{bibliography,apj-jour}    %% includes the journal abbrevs
%\bibliography{bibliography}

\input{Plots}

\end{document}

### 2014-06-30 Sources.tex
Cosmological sources

@@
 Estimates for the event rate of resolvable binaries centre around ${\cal{O}}(10^{3})$ events for eLISA.
-The box plotted in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega} has an ratio $h_{c}/h_{n}=50$ for eLISA at its peak sensitivity, the frequency range of the box is $\left(3\times10^{-4}-10^{-2}\right)\,\textrm{Hz}$, based on visual inspection of Monte Carlo population simulation results presented in \cite{Amaro-Seoane-et-al}.
+The box plotted in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega} has an ratio $h_{c}/h_{n}=50$ for eLISA at its peak sensitivity, the frequency range of the box is $\left(3\times10^{-4}\right.$--$\left.10^{-2}\right)~\mathrm{Hz}$, based on visual inspection of Monte Carlo population simulation results presented in \cite{Amaro-Seoane-et-al}.
 
@@
 EMRI events occur when a compact stellar mass object inspirals into a supermassive black hole.
-There is extreme uncertainty in the event rate for EMRIs due to the poorly constrained astrophysics in galactic centres; the best guess estimate is ${\cal{O}}(10)$ events per year with eLISA with SNR $\ge 20$.
-The box plotted in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega} has a characteristic strain of $h_{c}=3\times 10^{-20}$ at $10^{-2}\;\textrm{Hz}$ which corresponds to a $10\Msun$ BH inspiralling into a $10^{6}\Msun$ black hole at a luminosity distance of $1\,\textrm{Gpc}$.
+There is extreme uncertainty in the event rate for EMRIs due to the poorly constrained astrophysics in galactic centres; the best guess estimate is ${\mathcal{O}}(10)$ events per year with eLISA with SNR $\ge 20$.
+The box plotted in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega} has a characteristic strain of $h_{c}=3\times 10^{-20}$ at $10^{-2}~\mathrm{Hz}$ which corresponds to a $10\Msun$ BH inspiralling into a $10^{6}\Msun$ black hole at a luminosity distance of $1~\mathrm{Gpc}$.
 The frequency width of the box is somewhat unknown, EMRI events can occur into a black hole of any mass, and hence EMRIs can occur at any frequency.
@@
 Supermassive black holes are known to lie at the centres of most galaxies and the black hole mergers are associated with the mergers of the host galaxies.
-The current best published limit for the amplitude of the stochastic background is $h_{c}=6\times 10^{-15}$ at a frequency of $f_{0}=1\,\textrm{yr}^{-1}$, \cite{Haasteren}.
+The current best published limit for the amplitude of the stochastic background is $h_{c} = 6\times 10^{-15}$ at a frequency of $f_{0}=1~\mathrm{yr}^{-1}$, \cite{Haasteren}.
 There is strong theoretical evidence that the actual background lies close to the current limit, \cite{imminentdetectionofgravitationalwaves} and \cite{Sesana-2012}.
@@
 The exact distinction between resolvable and unresolvable is detector specific.
-Plotted in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega} for the unresolvable background is a third of the current limit with a cut off frequency of $f=1\,\textrm{yr}^{-1}$ which is suggested by Monte Carlo population studies, \cite{SesanaVecchioColancino}.
-For the resolvable sources the amplitude of the current limit is plotted between $\left(3\times 10^{-9} - 3\times 10^{-7}\right)\,\textrm{Hz}$.
+Plotted in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega} for the unresolvable background is a third of the current limit with a cut off frequency of $f=1~\mathrm{yr}^{-1}$ which is suggested by Monte Carlo population studies, \cite{SesanaVecchioColancino}.
+For the resolvable sources the amplitude of the current limit is plotted between $\left(3\times 10^{-9}\right.$--$\left.3\times 10^{-7}\right)~\mathrm{Hz}$.
 
 \subsection{Cosmological sources}
-\cjm{Christopher, would you be happy to write a section on very high redshift sources of stochastic backgrounds.
-Stressing that these in principal can be at any frequency.
-And also saying that because of the lack of certainty here we don't include these sources in the figures.
-Maybe reference recent excitement caused by BICEP2 results.}
+
+In addition to the sources above, early Universe processes, such as inflation \citep{Grishchuk2005} or a first-order phase transition \citep{Binetruy2012}, could have created GWs.
+More speculatively, it has been hypothesised that cosmic strings could also be a potential source \citep{Damour2005,Binetruy2012}.
+These relic GWs allow us to explore energy scales far beyond those accessible by other means, providing insight into new and exotic physics.
+The excitement surrounding the tentative discovery by BICEP2 of the imprint of primordial GWs (generated during inflation) in the cosmic microwave background \citep{Ade2014}, and the subsequent flurry of activity, has shown the scientific potential of such cosmological GWs.
+These GW signals are so alluring because they probe unknown physics; this also makes them difficult to predict.
+Cosmological stochastic backgrounds have been predicted across a range of frequencies with considerable variation in amplitude.
+As a consequence of this uncertainty, although we could learn much from measuring these signals, we have not included them amongst the sources shown in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega}.

### 2014-06-30 Typesofsource.tex
New branch and typo corrections

@@
 
-Sources of GWs can be broadly split into three catagories: well modelled sources, unmodelled (or poorly-modelled) transient sources and unmodelled stochastic backgrounds.
+Sources of GWs can be broadly split into three categories: well modelled sources, for which we have a description of the expected waveform; stochastic backgrounds, for which we can describe the statistical behaviour; and unmodelled (or poorly-modelled) transient sources.
 The classic example of a well modelled source is the inspiral of two compact objects, this is discussed in section \ref{sec:insp}.
-Stochastic backgrounds can either be formed from many overlapping sources which are each individually susceptible to being modelled, or from some intrinsically random process, these are discussed in section \ref{sec:stoch}.
-The classic example of a unmodelled (or poorly-modelled) transient source is a supernova, searches of this type are often called burst searched and are discussed in section \ref{sec:bursts}.
+Stochastic backgrounds can either be formed from many overlapping sources, which are each individually susceptible to being modelled, or from some intrinsically random process, these are discussed in section \ref{sec:stoch}.
+An example of an unmodelled (or poorly-modelled) transient source is a supernova; searches of this type are often called burst searches and are discussed in section \ref{sec:bursts}.
 
@@
 \subsection{Stochastic backgrounds}\label{sec:stoch}
-Another important source of GWs are stochastic backgrounds due to large populations of individually unresolvable sources.
+Another important source of GWs arises from stochastic backgrounds due to large populations of individually unresolvable sources.
 The population of sources will in general be at cosmological distances and it is necessary to distinguish the frequency in the source rest frame, $f_{\textrm{r}}$, from the measured frequency, $f$; these are related through the redshift, $z$, via $f_{\textrm{r}}=(1+z)f$.
@@
 In practice this power law will also have upper and lower frequency cut-offs related to the population of objects creating the GW spectrum.
-The stochastic background due to other sources, such as cosmic strings or relic GWs from the big bang, can also be written in the same form as (\ref{eq:power}) but with different spectral indices: $\alpha=-7/6$ for cosmic strings or $\alpha$ in the range $-1$ to $-0.8$ for relic GWs, \cite{Jenet}.
+The stochastic background due to other sources, such as cosmic strings or relic GWs from the early Universe, can also be written in the same form as (\ref{eq:power}) but with different spectral indices: $\alpha=-7/6$ for cosmic strings or $\alpha$ in the range $-1$ to $-0.8$ for relic GWs, \cite{Jenet}.
 

### 2014-06-30 Typesofsource.tex
Burst sources

@@
 
-\cjm{This section is currently much shorter than both \ref{sec:insp} and \ref{sec:stoch}, Rob said he was going to add more detail on searching for unmodelled phenomenon using time-frequency plots.}
+Some sources of GWs can produce signals with large amplitudes, greater than the detector noise.
+The duration of such a signal is typically very short, of the order of a few wave periods, and so there is not time to accumulate SNR in each frequency band in the way inspirals do.
+As a consequence, waveform models are not required to produce a detection; we simply rely on the "excess power" produced by these so-called burst sources.
+Typically, we may be looking for signals from core collapse supernovae \citep{Ott2009}, the late stages of merging compact binaries, or more generally, signals from any unexpected or poorly modelled sources.
 
-A signal is burst-like if it's duration at a detectable amplitude is only a few times the wave period.
-If this is the case then the signal does not have time to accumulate SNR in each frequency band in the way inspirals do.
-If the source generates ${\cal{N}}_{\textrm{cycles}}$ of radiation then the relation between $h_{c}(f)$ and $h$ is simply
-\begin{equation}\label{eq:simple} h_{c}(f)=h\sqrt{{\cal{N}}_{\textrm{cycles}}} \; . \end{equation}
+Burst searches are often carried out using time-frequency techniques.
+The data stream from a detector is temporally split into segments, the length of which can be tuned to give greater sensitivity to particular sources.
+Each segment is then transformed into the frequency domain, whitened and normalised to the noise spectrum to produce a time-frequency plot.
+Potential GW signals are identified by searching for clusters of pixels that contain an excess of power (see e.g.
+\cite{Bursts}).
+
+The presence of excess power across a number of pixels eliminates modelled noise sources, but such a cluster may also be caused by atypical noise within a detector.
+We can improve our confidence of a GW signal by making use of information obtained from other GW detectors.
+Signals across a network of detectors should have compatible arrival times (given the sky direction) as well as consistent amplitudes, frequencies and shapes of the waveform.
+Different pipelines are currently in use that analyse the signal consistency in different ways: both coincidence searches \citep{Chatterji2004} and fully coherent methods \citep{Klimenko2008} are used.
+
+An important aspect of burst search algorithms is to accurately estimate the noise properties within each time segment.
+To this extent, null data streams can be constructed that should be insensitive to real GW signals.
+In order to estimate the false alarm rate, the data from different detectors can be shifted in time to remove any genuine coincident GW signals.
+These time-shifts are then analysed to simulate the potential occurence of coincident noise events.
+The algorithms are tuned using time-shifted data to ensure there is no bias in the final search.
+
+As discussed in \ref{sec:insp}, the expected relation between $h_{c}(f)$ and a typical waveform $\tilde{h}(f)$ is simply
+\begin{equation}\label{eq:simple} h_{c}(f)=\tilde{h}(f)\sqrt{{\cal{N}}_{\textrm{cycles}}} \; , \end{equation}
+where ${\cal{N}}_{\textrm{cycles}}$ is the number of cycles of radiation generated by the source.
+Here, ${\cal{N}}_{\textrm{cycles}}$ is of order unity and $\tilde{h}(f)$ is strongly peaked around some central frequency.

### 2014-06-30 Introduction.tex

@@
 
+%CPLB: alternate openning I put together a while ago.
+%Even the best of us can be frustrated by the use of different conventions or units when describing a single quantity.
+It is a common occurrence in physics that  there is a particular parameterization that is perfect for one experiment or calculation, but this is not of universal applicability.
+When comparing results across studies it is then necessary to trawl through the literature to check that $x$ is really the same parameter.
+Astronomy and astrophysics are rife with specialized units, from the jansky (a spectral flux density of $10^{-26}$ watts per square metre per hertz) to the solar neutrino unit (a neutrino flux that produces an interaction rate of $10^{-36}$ per second per target atom).
+Inevitably, things get lost in translation.
+The mismatch of conventions can be confusing to those new to the field, especially students.
+
 The next few years promises to deliver the first direct detection of gravitational waves (GWs).
-This will most likely be achieved by the advanced LIGO \citep{2010CQGra..27h4006H} detector operating in the frequency range $(10-10^{3})\,\textrm{Hz}$.
-By the end of the decade it is expected that pulsar timing arrays \citep{1990ApJ361300F} will detect very low frequency GWs around $10^{-8}\,\textrm{Hz}$.
-Further into the future space based detectors, such as eLISA \citep{2012CQGra..29l4016A}, will probe GWs in the millihertz regime.
-This will herald the beginning of multi-wavelength GW astronomy.
-There already exists an extensive literature assessing the potential of all of these detectors to probe the astrophysics of various potential sources.
+This will most likely be achieved by the advanced LIGO \citep{2010CQGra..27h4006H} and Virgo \citep{Accadia2011} detectors operating in the frequency range $(10$--$10^{3})~\mathrm{Hz}$.
+By the end of the decade, it is expected that pulsar timing arrays \citep{1990ApJ361300F} shall detect very low frequency GWs around $10^{-8}~\mathrm{Hz}$.
+Further into the future, space-based detectors, such as eLISA \citep{2012CQGra..29l4016A}, will probe GWs in the millihertz regime.
+This shall herald the beginning of multi-wavelength GW astronomy.
+
+There already exists an extensive literature assessing the potential of all of these detectors to probe the astrophysics of various sources.
 There are several different methods commonly used to describe the sensitivity of a GW detector and the strength of a GW source.
 It is common practice in GW astronomy to summarise this information on a sensitivity curve plot.
-When producing these plots it is desirable to have a consistent convention between detectors and sources which allows information about both detectors and sources to be plotted on the same graph.
-Another desirable feature of such a plot is that the detectors and sources are represented in such a way that the relative detectability of the sources is immediately apparent from the plot.
+When producing these plots, it is desirable to have a consistent convention between detectors and sources which allows information about both to be plotted on the same graph.
+\cplb{Ideally,} the detectors and sources are represented in such a way that the relative detectability of the signals is immediately apparent from the plot.
 
@@
 
-We begin by expounding the various conventions and the relationships between them in sections \ref{sec:conventions} and \ref{sec:voc}.
-A review of GW detectors (both current and proposed for the future) is given in section \ref{sec:detectors} and a review of GW sources is given in section \ref{sec:sources}.
-In \ref{app:a} several example sensitivity curves are presented; there is also a website at \url{http://www.ast.cam.ac.uk/~rhc26/sources/} which can be used to generate similar figures with user specified detectors and sources included.
+We begin by expounding the various conventions and the relationships between the conventions in sections \ref{sec:conventions} and \ref{sec:voc}.
+A review of GW detectors (both current and proposed) is given in section \ref{sec:detectors} and a review of GW sources is given in section \ref{sec:sources}.
+In \ref{app:a} several example sensitivity curves are presented.
+\cplb{A website where similar figures can be generated is available at \url{www.ast.cam.ac.uk/~rhc26/sources/}.
+Here the user may select which sources and detectors to include to tailor the figure to their specific requirements.}
 
@@
 It will be necessary to work with the Fourier transform of the signal, the most common convention used in the GW community is
-\begin{eqnarray} x(t)&={\cal{F}}^{-1}\left\{\tilde{x}(f)\right\}&=\int_{-\infty}^{\infty}\textrm{d}f\;\tilde{x}(f)\exp (2\pi \rmi f t )  \nonumber \\
-\tilde{x}(f) &=\,\;{\cal{F}}\left\{x(t)\right\}&=\int_{-\infty}^{\infty}\textrm{d}t\;x(t)\exp (-2\pi \rmi f t )\; .\end{eqnarray}
+\begin{eqnarray} x(t)&={\cal{F}}^{-1}\left\{\tilde{x}(f)\right\}&=\int_{-\infty}^{\infty}\mathrm{d}f\;\tilde{x}(f)\exp (2\pi \rmi f t )  \nonumber \\
+\tilde{x}(f) &=\,\;{\cal{F}}\left\{x(t)\right\}&=\int_{-\infty}^{\infty}\mathrm{d}t\;x(t)\exp (-2\pi \rmi f t )\; .\end{eqnarray}
 For simplicity it is assumed that the noise in the GW detector is stationary and Gaussian (with zero mean), under these assumptions the noise is fully characterised via the \emph{one-sided} noise power spectral density, $S_{n}(f)$,\footnote{An alternative convention is to use the two-sided power spectral density $S^{(2)}_{n}(f)=S_{n}(f)/2$.},
@@
 \begin{eqnarray}\label{eq:meansquare}
-\overline{n(t)^{2}}&=\lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T}\textrm{d}t\;\left| n(t) \right|^{2} \nonumber \\
-&=\lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T}\textrm{d}t\; \int_{-\infty}^{\infty}\textrm{d}f\;\int_{-\infty}^{\infty}\textrm{d}f'\;\tilde{n}(f)\tilde{n}^{*}(f')\exp\left(2\pi \rmi (f-f')t \right)
+\overline{n(t)^{2}}&=\lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T}\mathrm{d}t\;\left| n(t) \right|^{2} \nonumber \\
+&=\lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T}\mathrm{d}t\; \int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\tilde{n}(f)\tilde{n}^{*}(f')\exp\left(2\pi \rmi (f-f')t \right)
 \end{eqnarray}
@@
 \begin{eqnarray}\label{eq:meansquare1}
-\overline{n(t)^{2}}& = \lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T}\int_{-\infty}^{\infty}\textrm{d}f\;\int_{-\infty}^{\infty}\textrm{d}f'\;{\cal{F}}\left\{ n(\tau-t) \right\}(f){\cal{F}}\left\{ n(\tau-t) \right\}(f')^{*} \nonumber \\
-&=\int_{-\infty}^{\infty}\textrm{d}f\;\int_{-\infty}^{\infty}\textrm{d}f'\;\left<\tilde{n}(f)\tilde{n}^{*}(f')\right>\nonumber \\
-& = \int_{0}^{\infty}\textrm{d}f\; S_{n}(f)
+\overline{n(t)^{2}}& = \lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T}\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;{\cal{F}}\left\{ n(\tau-t) \right\}(f){\cal{F}}\left\{ n(\tau-t) \right\}(f')^{*} \nonumber \\
+&=\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\left<\tilde{n}(f)\tilde{n}^{*}(f')\right>\nonumber \\
+& = \int_{0}^{\infty}\mathrm{d}f\; S_{n}(f)
 \end{eqnarray}
@@
 \begin{eqnarray}\label{eq:meansquare1}
-%\left<n(t)^{2}\right>&=\lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-\infty}^{\infty}\textrm{d}f\;\int_{-\infty}^{\infty}\textrm{d}f'\;\int_{-T}^{T}\textrm{d}t\; {\cal{F}}\left\{n(\tau+t)\right\}\left[{\cal{F}}\left\{n(\tau+t)\right\}\right]^{*}\nonumber \\
-\overline{n(t)^{2}}& = \lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-\infty}^{\infty}\textrm{d}f\;\int_{-\infty}^{\infty}\textrm{d}f'\;\int_{-T}^{T}\textrm{d}t\; \tilde{n}(f)\tilde{n}^{*}(f') \nonumber \\
-&=\int_{-\infty}^{\infty}\textrm{d}f\;\int_{-\infty}^{\infty}\textrm{d}f'\;\left<\tilde{n}(f)\tilde{n}^{*}(f')\right>\nonumber \\
-%&= \int_{-\infty}^{\infty}\textrm{d}f\;\int_{-\infty}^{\infty}\textrm{d}f'\;\frac{1}{2}S_{n}(f)\delta (f-f')\nonumber\\
-& = \int_{0}^{\infty}\textrm{d}f\; S_{n}(f)
+%\left<n(t)^{2}\right>&=\lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\int_{-T}^{T}\mathrm{d}t\; {\cal{F}}\left\{n(\tau+t)\right\}\left[{\cal{F}}\left\{n(\tau+t)\right\}\right]^{*}\nonumber \\
+\overline{n(t)^{2}}& = \lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\int_{-T}^{T}\mathrm{d}t\; \tilde{n}(f)\tilde{n}^{*}(f') \nonumber \\
+&=\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\left<\tilde{n}(f)\tilde{n}^{*}(f')\right>\nonumber \\
+%&= \int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\frac{1}{2}S_{n}(f)\delta (f-f')\nonumber\\
+& = \int_{0}^{\infty}\mathrm{d}f\; S_{n}(f)
 \end{eqnarray}
@@
 Convolving this with the detector output gives a contribution from the signal and a contribution from the noise
-\begin{equation}\label{eq:conv} \left(s*K\right)(\tau)=\int_{-\infty}^{\infty}\textrm{d}t\;(n(t)+h(t))K(t-\tau)=S+N\; .\end{equation}
+\begin{equation}\label{eq:conv} \left(s*K\right)(\tau)=\int_{-\infty}^{\infty}\mathrm{d}t\;(n(t)+h(t))K(t-\tau)=S+N\; .\end{equation}
 The signal contribution, $S$, is defined as the expectation of the convolution in (\ref{eq:conv}) when a signal is present, maximised by varying the offset $\tau$.
 Since the expectation of pure noise is zero it follows that
-\begin{equation} S = \int_{-\infty}^{\infty}\textrm{d}t\;h(t)K(t)=\int_{-\infty}^{\infty}\textrm{d}t\;h(t)K^{*}(t)=\int_{-\infty}^{\infty}\textrm{d}f\; \tilde{h}(f)\tilde{K}^{*}(f) \; .\end{equation}
+\begin{equation} S = \int_{-\infty}^{\infty}\mathrm{d}t\;h(t)K(t)=\int_{-\infty}^{\infty}\mathrm{d}t\;h(t)K^{*}(t)=\int_{-\infty}^{\infty}\mathrm{d}f\; \tilde{h}(f)\tilde{K}^{*}(f) \; .\end{equation}
 The squared contribution from noise, $N^{2}$, is defined as the mean square of the convolution in (\ref{eq:conv}) when no signal is present,
 \begin{eqnarray} 
-N^{2}&=\int_{-\infty}^{\infty}\textrm{d}t\;\int_{-\infty}^{\infty}\textrm{d}t'\;K(t)K(t')\left<n(t)n(t')\right> \nonumber \\
-&=\int_{-\infty}^{\infty}\textrm{d}t\;\int_{-\infty}^{\infty}\textrm{d}t'\;K(t)K^{*}(t')\int_{-\infty}^{\infty}\textrm{d}f\;\int_{-\infty}^{\infty}\textrm{d}f'\;\left<\tilde{n}(f)\tilde{n}^{*}(f')\right>\exp\left(2\pi\rmi(ft-f't')\right)\nonumber \\
-%&=\int_{-\infty}^{\infty}\textrm{d}f\;\int_{-\infty}^{\infty}\textrm{d}t\;\int_{-\infty}^{\infty}\textrm{d}t'\;\frac{1}{2}S_{n}(f)K(t)K^{*}(t')\exp\left(2\pi\rmi f (t-t')\right)\nonumber \\
-&=\int_{-\infty}^{\infty}\textrm{d}f \; \frac{1}{2}S_{n}(f)\tilde{K}(f)\tilde{K}^{*}(f) \; .
+N^{2}&=\int_{-\infty}^{\infty}\mathrm{d}t\;\int_{-\infty}^{\infty}\mathrm{d}t'\;K(t)K(t')\left<n(t)n(t')\right> \nonumber \\
+&=\int_{-\infty}^{\infty}\mathrm{d}t\;\int_{-\infty}^{\infty}\mathrm{d}t'\;K(t)K^{*}(t')\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\left<\tilde{n}(f)\tilde{n}^{*}(f')\right>\exp\left(2\pi\rmi(ft-f't')\right)\nonumber \\
+%&=\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}t\;\int_{-\infty}^{\infty}\mathrm{d}t'\;\frac{1}{2}S_{n}(f)K(t)K^{*}(t')\exp\left(2\pi\rmi f (t-t')\right)\nonumber \\
+&=\int_{-\infty}^{\infty}\mathrm{d}f \; \frac{1}{2}S_{n}(f)\tilde{K}(f)\tilde{K}^{*}(f) \; .
  \end{eqnarray}
@@
 where the inner product between a pair of signals $A$ and $B$ has been defined
-\begin{equation}\label{eq:snrinnerprod} \left( A | B \right) = 4\Re\left\{\int_{0}^{\infty}\textrm{d}f\;\frac{\tilde{A}^{*}(f)\tilde{B}(f)}{S_{n}(f)}\right\} \; .\end{equation}
+\begin{equation}\label{eq:snrinnerprod} \left( A | B \right) = 4\Re\left\{\int_{0}^{\infty}\mathrm{d}f\;\frac{\tilde{A}^{*}(f)\tilde{B}(f)}{S_{n}(f)}\right\} \; .\end{equation}
 The optimum filter is that which maximises the SNR in (\ref{eq:SNRinnerprod}). From the Cauchy-Schwarz inequality it follows that the optimum filter is
@@
 \begin{equation}\label{eq:hc} 
-\varrho^{2}=\int_{0}^{\infty} \textrm{d}\left(\log f\right)\; \left(\frac{h_{c}(f)}{h_{n}(f)}\right)^{2} \;.
+\varrho^{2}=\int_{0}^{\infty} \mathrm{d}\left(\log f\right)\; \left(\frac{h_{c}(f)}{h_{n}(f)}\right)^{2} \;.
 \end{equation}
@@
 Another common quantity to plot on sensitivity curves is the square root of the power spectral density (see figure \ref{fig:S}), which from (\ref{eq:strains}) is given by
-\begin{eqnarray}\label{eq:temp1} \sqrt{S_{n}(f)}&=h_{n}(f)f^{-1/2} \quad \textrm{for detectors, and by analogy} \\
-\sqrt{S_{h}(f)}&=h_{c}(f)f^{-1/2} \quad \textrm{for sources.}\label{eq:ShforSources}\end{eqnarray}
+\begin{eqnarray}\label{eq:temp1} \sqrt{S_{n}(f)}&=h_{n}(f)f^{-1/2} \quad \mathrm{for detectors, and by analogy} \\
+\sqrt{S_{h}(f)}&=h_{c}(f)f^{-1/2} \quad \mathrm{for sources.}\label{eq:ShforSources}\end{eqnarray}
 The power spectral density is the most commonly plotted quantity in sensitivity curves in the literature and it has the nice property, proven in equation (\ref{eq:meansquare1}), that integrated over all positive frequencies it gives the mean square amplitude of the signal in the detector.
@@
 \begin{eqnarray}\label{eq:specNRGdensity}
-\rho c^{2}&=\frac{c^{2}}{16\pi G}\int_{-\infty}^{\infty}\textrm{d}f\;\left(2\pi f\right)^{2}\tilde{h}(f)\tilde{h}^{*}(f)=\int_{0}^{\infty}\textrm{d}f\;\frac{\pi c^{2}}{4G}f^{2}S_{h}(f)\; ,
+\rho c^{2}&=\frac{c^{2}}{16\pi G}\int_{-\infty}^{\infty}\mathrm{d}f\;\left(2\pi f\right)^{2}\tilde{h}(f)\tilde{h}^{*}(f)=\int_{0}^{\infty}\mathrm{d}f\;\frac{\pi c^{2}}{4G}f^{2}S_{h}(f)\; ,
 \end{eqnarray} 
 where the definitions (\ref{eq:ShforSources}) and (\ref{eq:strainsagain}) have been used.
-Equation (\ref{eq:specNRGdensity}) motivates the definition of the spectral energy density in GWs, $S_{\textrm{E}}$, the energy per unit volume of space, per unit frequency, \cite{HellingsDowns}
-\begin{equation}\label{eq:spectralenergydensity} S_{\textrm{E}}(f)=\frac{\pi c^{2}}{4G} f^{2}S_{h}(f) \; ,\end{equation}
+Equation (\ref{eq:specNRGdensity}) motivates the definition of the spectral energy density in GWs, $S_{\mathrm{E}}$, the energy per unit volume of space, per unit frequency, \cite{HellingsDowns}
+\begin{equation}\label{eq:spectralenergydensity} S_{\mathrm{E}}(f)=\frac{\pi c^{2}}{4G} f^{2}S_{h}(f) \; ,\end{equation}
 and a corresponding expression for the noise with $S_{n}(f)$.
-It is usual to define the dimensionless quantity $\Omega_{\textrm{GW}}$ as the energy density per logarithmic frequency interval normalised to the critical density of the universe
-\begin{equation}\label{eq:omega} \Omega_{\textrm{GW}}(f)=\frac{fS_{\textrm{E}}(f)}{\rho_{\textrm{c}}c^{2}}\; . \end{equation}
-The critical density of the universe is $\rho_{\textrm{c}}=3H_{0}^{2}/8\pi G$, where $H_{0}=h_{100}\times 100\, \textrm{km}\,\textrm{s}^{-1}\,\textrm{Mpc}^{-1}$ is the Hubble constant (this equation serves as a definition of $h_{100}$, which here has nothing to do with strain).
-The most common quantity related to energy density to be plotted on sensitivity curves is $\Omega_{\textrm{GW}}h_{100}^{2}$ (figure \ref{fig:omega}).
+It is usual to define the dimensionless quantity $\Omega_{\mathrm{GW}}$ as the energy density per logarithmic frequency interval normalised to the critical density of the universe
+\begin{equation}\label{eq:omega} \Omega_{\mathrm{GW}}(f)=\frac{fS_{\mathrm{E}}(f)}{\rho_{\mathrm{c}}c^{2}}\; . \end{equation}
+The critical density of the universe is $\rho_{\mathrm{c}}=3H_{0}^{2}/8\pi G$, where $H_{0}=h_{100}\times 100\, \mathrm{km}\,\mathrm{s}^{-1}\,\mathrm{Mpc}^{-1}$ is the Hubble constant (this equation serves as a definition of $h_{100}$, which here has nothing to do with strain).
+The most common quantity related to energy density to be plotted on sensitivity curves is $\Omega_{\mathrm{GW}}h_{100}^{2}$ (figure \ref{fig:omega}).
 This quantity has one aesthetic advantage over the others: it accounts for the fact that there is less energy in low frequency waves of the same amplitude and does not place the sensitivity curves of very low frequency detectors much higher than most ground based detectors.
@@
 \subsection{Relating the different descriptions}
-The dimensionless energy density in GWs, $\Omega_{\textrm{GW}}$, the spectral energy density in GWs, $S_{\textrm{E}}$, the one-sided power spectral density, $S_{h}$, and the characteristic strain are all related via
+The dimensionless energy density in GWs, $\Omega_{\mathrm{GW}}$, the spectral energy density in GWs, $S_{\mathrm{E}}$, the one-sided power spectral density, $S_{h}$, and the characteristic strain are all related via
 \begin{equation}\label{eq:omega} 
-\Omega_{\textrm{GW}}(f)=\frac{fS_{\textrm{E}}(f)}{\rho_{\textrm{c}}c^{2}}=\frac{\pi}{4G\rho_{\textrm{c}}}f^{2}h_{c}(f)^{2}=\frac{\pi}{4G\rho_{\textrm{c}}}f^{3}S_{h}(f)  \; ,
+\Omega_{\mathrm{GW}}(f)=\frac{fS_{\mathrm{E}}(f)}{\rho_{\mathrm{c}}c^{2}}=\frac{\pi}{4G\rho_{\mathrm{c}}}f^{2}h_{c}(f)^{2}=\frac{\pi}{4G\rho_{\mathrm{c}}}f^{3}S_{h}(f)  \; ,
 \end{equation}

### 2014-06-30 noise_curve.tex

@@
 
-\title{Gravitational wave sensitivity curves}
-\author{C J Moore}
-\address{Institute of Astronomy, Madingley Road, Cambridge, CB3 0HA, United Kingdom}
+\title{Gravitational-wave sensitivity curves}
+\author{C J Moore$^1$, C P L Berry$^{1,\,2}$ \& R H Cole$^1$}
+\address{$^1$ Institute of Astronomy, Madingley Road, Cambridge, CB3 0HA, United Kingdom}
+\address{$^2$ School of Physics and Astronomy, University of Birmingham, Edgbaston, Birmingham B15 2TT, United Kingdom}
 \ead{cjm96@ast.cam.ac.uk}
@@
 \begin{abstract}
-There are several common conventions in use by the gravitational wave community to describe the amplitude of sources and the sensitivity of detectors.
+There are several common conventions use by the gravitational wave community to describe the amplitude of sources and the sensitivity of detectors.
 These are frequently confused.
-We outline the merits and differences between the various quantities used for parametrising noise curves and characterising gravitational wave amplitudes.
-Sensitivity curves plots for each of the conventions are presented and interactive versions have been made available on-line at \url{http://www.ast.cam.ac.uk/~rhc26/sources/}.
+We outline the merits of and differences between the various quantities used for parameterizing noise curves and characterizing gravitational wave amplitudes.
+We conclude by producing plots that consistently compare different detectors.
+Similar figures can be generated on-line for general use at \url{http://www.ast.cam.ac.uk/~rhc26/sources/}.
+
+\cplb{Slightly tweaked and added PACS}
 \end{abstract}
+
+\pacs{04.30.--w, 04.30.Db, 04.80.Nn, 95.55.Ym}
 

### 2014-06-30 Detectors.tex
Typo corrections

@@
 The response of a detector to an incident plane fronted GW depends upon the relative orientations of the detector and the incoming wave.
-Choosing the origin of our coordinate system to be the beam-splitter of the interferometer and let $l_{1}^{i}$ and $l_{2}^{i}$ be unit 3-vectors pointing along the two arms.
+Choose the origin of our coordinate system to be the beam-splitter of the interferometer and let $l_{1}^{i}$ and $l_{2}^{i}$ be unit 3-vectors pointing along the two arms.
 In the absence of noise the output of the detector is the difference in strain between the two arms \citep{MTW}
-\begin{equation}\label{eq:hoftxx} h(t)=\frac{1}{2}h_{ij}\left( l_{1}^{i}l_{1}^{j}-l_{2}^{i}l_{2}^{j} \right)\; . \end{equation}
+\begin{equation}\label{eq:hoftxx} h(t)=\frac{1}{2}h_{ij}\left( l_{1}^{i}l_{1}^{j}-l_{2}^{i}l_{2}^{j} \right)\; , \end{equation}
 where $h_{ij}$ are the spatial components of the GW metric perturbation.
@@
 \end{eqnarray}
-The response function of a two-arm interferometric detector is quadrapolar, an example is plotted in figure \ref{fig:LIGO}.
+The response function of a two-arm interferometric detector is quadrupolar, an example is plotted in figure \ref{fig:LIGO}.
 
@@
 The interferometric detectors listed in Table \ref{table:t} are all sensitive to GWs in the frequency range ${\cal{O}}(10-10^{3})\,\textrm{Hz}$.
-They all have multiple narrow lines in their sensitivity curves which arise from noise sources in the instrument including resonances in the suspension system and electronic noise at multiples of $60\,\textrm{Hz}$: these have been removed in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega} for clarity.
-The detectors fall broadly into three categories: 1\textsuperscript{st} generation detectors which have already operated, 2\textsuperscript{nd} generation detectors currently under construction and 3\textsuperscript{rd} generation detectors at the planning stage.
+They all have multiple narrow lines in their sensitivity curves, which arise from noise sources in the instrument including resonances in the suspension system and electronic noise at multiples of $60\,\textrm{Hz}$: these have been removed in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega} for clarity.
+The detectors fall broadly into three categories: 1\textsuperscript{st} generation detectors, which have already operated; 2\textsuperscript{nd} generation detectors currently under construction; and 3\textsuperscript{rd} generation detectors at the planning stage.
 
@@
 \subsection{Space based detectors}\label{sec:space}
-Space based detectors works on similar principles to ground based detectors but with the test masses residing inside of independent, widely separated satellites.
-Space based detectors are sensitive to lower frequency gravitational waves than their ground based counterparts, this is partly because space based detectors can have much longer arms and partly because they are unaffected by seismic noise which limits the low frequency performance of ground based detectors.
+Space based detectors work on similar principles to ground based detectors but with the test masses residing inside of independent, widely separated satellites.
+Space based detectors are sensitive to lower frequency GWs than their ground based counterparts, this is partly because space based detectors can have much longer arms and partly because they are unaffected by seismic noise which limits the low frequency performance of ground based detectors.
 

### 2014-06-30 Introduction.tex
Typo corrections

@@
 
-Gravitational radiation has two independent polarization states denoted $+$ and $\times$; a general signal can be described as a linear combination the two polarization states, $h \nobreak=\nobreak A_{+} h_{+}\nobreak +\nobreak A_{\times} h_{\times}$.
+Gravitational radiation has two independent polarization states denoted $+$ and $\times$; a general signal can be described as a linear combination of the two polarization states, $h \nobreak=\nobreak A_{+} h_{+}\nobreak +\nobreak A_{\times} h_{\times}$.
 The sensitivity of a detector to each of these depends upon the relative orientations of the source and detector.
@@
 \begin{equation}\label{eq:psd} \left<\tilde{n}(f)\tilde{n}^{*}(f')\right>=\frac{1}{2}\delta (f-f')S_{n}(f) \; .\end{equation}
-Angled brackets $\left<\ldots\right>$ denotes an ensemble average over many realisations of the noise.
+Angled brackets $\left<\ldots\right>$ denote an ensemble average over many realisations of the noise.
 In reality we have only a single system to work with, so the ensemble average is replaced by a time average.
@@
 Since the GW signal and detector output are real it follows that $\tilde{h}(-f)=\tilde{h}^{*}(f)$ and $\tilde{n}(-f)=\tilde{n}^{*}(f)$; therefore, $S_{n}(f)=S_{n}(-f)$.
-The fact the $S_{n}(f)$ is an even function means that Fourier integrals over all frequencies can instead be written as integrals over positive frequencies only (see, for example, equations (\ref{eq:meansquare1}) and (\ref{eq:snrinnerprod})), for this reason $S_{n}(f)$ is called the \emph{one-sided} power spectral density.
+The fact $S_{n}(f)$ is an even function means that Fourier integrals over all frequencies can instead be written as integrals over positive frequencies only (see, for example, equations (\ref{eq:meansquare1}) and (\ref{eq:snrinnerprod})), it is for this reason that $S_{n}(f)$ is called the \emph{one-sided} power spectral density.
 It is straightforward to show that the power spectral density has the property that integrated over all positive frequencies it gives the mean square noise amplitude in the detector.
@@
 Using the property of Fourier transforms, ${\cal{F}}\left\{ n(t+\tau) \right\}(f)=\tilde{n}(f)\exp (-2\pi \rmi f\tau )$, the exponential factors in equation (\ref{eq:meansquare}) may be absorbed as a time translation of the noise by an amount $t$.
-The noise is stationary and so it's long term time average is invariant under this time translation and is given by Equation (\ref{eq:psd}).
+The noise is stationary and so its long term time average is invariant under this time translation and is given by Equation (\ref{eq:psd}).
 \begin{eqnarray}\label{eq:meansquare1}
@@
 %&=\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}t\;\int_{-\infty}^{\infty}\mathrm{d}t'\;\frac{1}{2}S_{n}(f)K(t)K^{*}(t')\exp\left(2\pi\rmi f (t-t')\right)\nonumber \\
-&=\int_{-\infty}^{\infty}\mathrm{d}f \; \frac{1}{2}S_{n}(f)\tilde{K}(f)\tilde{K}^{*}(f) \; .
+&=\int_{-\infty}^{\infty}\mathrm{d}f \; \frac{1}{2}S_{n}(f)\tilde{K}(f)\tilde{K}^{*}(f) \; ,
  \end{eqnarray}
-Where the definition of $S_{n}(f)$ from (\ref{eq:psd}) has been used. Hence the optimum signal-to-noise ratio (SNR), $\varrho$, is given by
+where the definition of $S_{n}(f)$ from (\ref{eq:psd}) has been used. Hence the optimum signal-to-noise ratio (SNR), $\varrho$, is given by
 \begin{equation}\label{eq:SNRinnerprod} 

### 2014-06-30 Sources.tex
Typo corrections

@@
 \subsection{Sources for space based detectors}
-For a review of the GW sources for space based missions see, for example, \cite{Amaro-Seoane-et-al}, \cite{Gairetal} of the \cite{eLISAyellowbook}.
+For a review of the GW sources for space based missions see, for example, \cite{Amaro-Seoane-et-al}, \cite{Gairetal} or the \cite{eLISAyellowbook}.
 
@@
 The uncertainty in the rate reflects our uncertainty in the growth mechanisms of the supermassive black hole population.
-Plotted in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega} is a box with a ratio $h_{c}/h_{n}=100$ for eLISA at it's peak sensitivity.
+Plotted in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega} is a box with a ratio $h_{c}/h_{n}=100$ for eLISA at its peak sensitivity.
 The range of frequencies plotted is $(3\times 10^{-4}-3\times 10^{-1})\,\textrm{Hz}$, this corresponds to circular binaries in the mass range quoted above.

### 2014-07-01 Detectors.tex
random tiny change for test purposes
completely rewritten sec4.4 on PTAs

@@
 \section{Detectors}\label{sec:detectors}
+
 \subsection{Operating principle of an interferometric detector}\label{sec:principles}
@@
 where the response functions inherit their angular dependence from the choice of coordinates,
-\begin{eqnarray}
+\begin{eqnarray}\label{eq:responsefuncs}
 F^{+}(\theta,\phi,\psi)&=\frac{1}{2}H^{+}_{ij}\left(l_{1}^{i}l_{1}^{j}-l_{2}^{i}l_{2}^{j}\right) \nonumber \\
@@
 \subsection{Pulsar timing arrays (PTAs)}\label{sec:PTAgeneralproperties}
-\cjm{I am going to rewrite this section and give a similar level of detail as was given for interferometric detectors in section \ref{sec:principles}.}
+\cjm{
+PTAs can be thought of as naturally occurring interferometers with galactic scale arm lengths, hence they are sensitive to much lower frequencies than the detectors considered so far.
+Each pulsar is a very regular clock and the measured pulse arrival time (in practice, arrival times are corrected to arrival times at the solar-system  barycenter, SSB) can be compared against a prediction leaving a residual which includes the effects of passing GWs.
+Using an array of these pulsars spread across the sky allows us to correlate residuals between different pulsars and exploit the fact that the GWs influence all pulsars where as intrinsic pulsar noise clearly does not.
+The correlation between different pulsars depends only on their angular separation on the sky and has a distinctive shape, known as the Hellings and Downs curve \citep{HellingsDowns}, which we will now derive.
 
-PTAs can be thought of as naturally occurring interferometers with galactic scale arm lengths, hence they are sensitive to much lower frequencies than the detectors considered so far.
-Each pulsar is a very regular clock and the measured arrival time can be compared against a prediction leaving a residual which includes the effects of passing GWs.
-\cite{SesanaVecchioColancino} describe the shape of the sensitivity curve obtained by correlating timing residuals from each of $N_{\textrm{p}}$ pulsars in an array.
-The total timing residual is the sum of the residuals due to noise and due to GWs,
-\begin{equation} \delta t= \delta t_{n} + \delta t _{h} \; .\end{equation}
-The maximum timing residual caused by a passing GW is the light travel time of along wavelength times the strain amplitude of the GW, $\delta t_{h}\approx h/f$.
-An upper limit to any background may be placed by assuming the residuals are due entirely to GWs.
-First consider the case of two pulsars; the minimum instantaneous detectable timing residual due to GWs is proportional to the root mean square error in the measurement of the timing residual, $\delta t_{h}\approx\delta t_{\textrm{r.m.s}}$.
-The minimum detectable signal decreases with the number of measured timing residuals (or data points).
-Over an observation time $T$, with individual observations taken at intervals of $\Delta T$ the minimum detectable signal is given by
-\begin{equation}\label{eq:GWtimingresidual} \delta t_{h} \propto \frac{\delta t_{\textrm{r.m.s}}}{\sqrt{T\Delta T ^{-1}}} \; . \end{equation}
-The PTA is sensitive to a minimum GW frequency of $T^{-1}$ and a maximum frequency of $\Delta T^{-1}$, hence the bandwidth is $\Delta f \approx \Delta T^{-1}-T^{-1}$.
-Assuming $\Delta T \ll T$ gives $T\Delta T^{-1} \approx T \Delta f$.
-Therefore the minimum characteristic strain the two-pulsar PTA is sensitive to is given by
-\begin{equation} h_{c}=\frac{\delta t_{\textrm{r.m.s.}}f}{\left(T\Delta f\right)^{1/4}} \; . \end{equation}
-When using more than two pulsars each separate pair of pulsars forms an independent detector, so the number of detectors scales as $N_{\textrm{p}}^{2}$.
-The optimum SNR is given by adding the individual SNRs in quadrature.
-Hence the total sensitivity is given by
-\begin{equation}\label{eq:PTA} h_{c}(f) \propto \frac{\delta t_{\textrm{r.m.s.}}f}{N_{\textrm{p}}^{1/2}\left( T\Delta f \right)^{1/4}} \; . \end{equation}
-The characteristic strain that the PTA is sensitive to scales linearly with $f$ between the minimum and maximum frequencies.
+The redshift of the rate of arrival of pulses for a pulsar at a distance $L$ from Earth, in the direction of the unit spatial vector $\vec{p}$ induced by a GW travelling in the direction of the unit vector $\hat{\Omega}$ is given by \citep{anholm-2009}
+\begin{equation}\label{eq:pulsarredshift} z(t,\hat{\Omega})=\frac{1}{2}\frac{\hat{p}^{j}\hat{p}^{i}} {1+\hat{\Omega}\cdot\hat{p}}\left(h_{ij}^{\textrm{Pulsar}}(t-L,\hat{\Omega} )-h_{ij}^{\textrm{Earth}}(t,\hat{\Omega} )\right)=\frac{1}{2}\frac{\hat{p}^{j}\hat{p}^{i}} {1+\hat{\Omega}\cdot\hat{p}}\Delta h_{ij}(t,\hat{\Omega})\; .
+\end{equation}
+The redshift includes two terms, the pulsar term and the Earth term.
+The pulsar term is often neglected in PTA analysis as it can be considered as an extra noise term which averages to zero.
+The experimentally measured quantity is not the redshift but the timing residual, the two are related via
+\begin{equation}\label{eq:restored}  R(t,\hat{\Omega})=\int_{0}^{t}\textrm{d}t'\;z(t',\hat{\Omega}) \; . \end{equation}
+All of the pulsars, and the Earth, reside in the same metric perturbation field which may be expressed in terms of its Fourier transform
+\begin{equation} h_{ij}(t,\hat{x})=\sum_{A=+,\times}\int\textrm{d}f\;\iint_{S_{2}}\textrm{d}\hat{\Omega}\; \tilde{h}_{A}(f,\hat{\Omega})e_{ij}^{A}(\hat{\Omega})e^{ 2\pi i f (t-\hat{\Omega}\cdot\vec{x}) } \; ,\end{equation}
+where $e^{A}_{ij}(\hat{\Omega})$ is the $A$ polarisation basis tensor for direction $\hat{\Omega}$.
+Choosing the SSB as the origin of our coordinate system, so the pulsar is at position $L\hat{p}$, gives
+\begin{equation}\label{eq:deltahij} \Delta h_{ij}(t,\hat{\Omega})=\sum_{A=+,\times}\int\textrm{d}f\; \tilde{h}_{A}(f,\hat{\Omega})e_{ij}^{A}(\hat{\Omega})e^{ 2\pi i f t }\left(e^{-2\pi\rmi fL (1+\hat{p}\cdot\hat{\Omega})}-1\right) \; .
+\end{equation}
+From (\ref{eq:pulsarredshift}) and (\ref{eq:deltahij}) the Fourier transform of the redshift, $\tilde{z}(f,\hat{\Omega})$, may be read off as
+\begin{equation}\label{eq:45} \tilde{z}(f,\hat{\Omega})=\left(e^{-2\pi\rmi fL (1+\hat{p}\cdot\hat{\Omega})}-1\right)\sum_{A=+,\times} \tilde{h}_{A}(f,\hat{\Omega})F^{A}(\hat{\Omega})\;, \textrm{where}\; F^{A}(\hat{\Omega})=\frac{e_{ij}^{A}(\hat{\Omega})\hat{p}^{j}\hat{p}^{i}} {2(1+\hat{\Omega}\cdot\hat{p})} \, .\end{equation}
+The function $F^{A}(\hat{\Omega})$ may be regarded as the PTA equivalent of the detector response functions in (\ref{eq:responsefuncs}).
+The stochastic background of GWs is fully characterised by the one-sided power spectral density via the following expectation value
+\begin{equation}\label{eq:stochback} \left<h_{A}^{*}(f,\hat{\Omega})h_{A'}(f',\hat{\Omega}')\right>=\frac{1}{2}S_{h}(f)\delta^{(2)}(\hat{\Omega},\hat{\Omega}')\delta_{AA'}\delta(f-f') \; ,\end{equation}
+where $\delta^{(2)}(\hat{\Omega},\hat{\Omega}')$ is the delta function on the sphere.
+From (\ref{eq:stochback}) and (\ref{eq:45}) the expectation of the product of signals from two different pulsars in directions $\hat{p}_{1}$ and $\hat{p}_{2}$ may be evaluated
+\begin{eqnarray} & \left<z_{1}(f)z_{2}^{*}(f')\right>=\frac{1}{2}S_{h}(f)\delta (f-f')\Gamma(f)\quad \textrm{where,} \\
+& \Gamma(f)=\sum_{A=+,\times}\iint_{{\cal{S}}^{2}}\textrm{d}\hat{\Omega}\,\left(e^{2\pi\rmi fL_{1}(1+\hat{\Omega}\cdot\hat{p}_{1})}-1\right)\left(e^{-2\pi\rmi fL_{2}(1+\hat{\Omega}\cdot\hat{p}_{2})}-1\right)F_{1}^{A}(\hat{\Omega})F_{2}^{A}(\hat{\Omega}) \; .
+\nonumber\end{eqnarray}
+The overlap function, $\Gamma(f)$, tends to a constant value in the limit that the distances to the pulsars are large compared to the wavelength of GWs, PTAs operate in this limit, so the overlap may be approximated as a constant
+\begin{equation} \Gamma(f)\approx\Gamma_{0}= \sum_{A=+,\times}\iint_{{\cal{S}}^{2}}\textrm{d}\hat{\Omega}\;F_{1}^{A}(\hat{\Omega})F_{2}^{A}(\hat{\Omega})\;.\end{equation}
+Neglecting the exponential terms in the overlap is the frequency domain equivalent of neglecting the pulsar term in (\ref{eq:pulsarredshift}).
+The integral may be evaluated to give an expression depending only on the angle, $\theta$, between the two pulsars; this is the famous Hellings and Downs curve, see figure \ref{fig:HnD},
+\begin{equation} \Gamma_{0}=\frac{1}{2}+\frac{3x}{2}\left(\log (x)-\frac{1}{6}\right) \quad \textrm{where}\quad x=\frac{1-\cos\theta}{2}\; .\end{equation}
+
+\begin{figure}[h!]
+ \centering
+ \includegraphics[trim=0cm 0cm 0cm 0cm, width=0.6\textwidth]{HnDcurve.pdf}
+ \caption{The Hellings and Downs curve, the correlation between two pulsars seperated on the sky by an angle $\theta$.}
+ \label{fig:HnD}
+\end{figure}
+
+The sensitivity bandwidth of a PTA is set by the sampling properties of the data set.
+If measurements are spaced in time by $\delta t$ and taken for a total length of time $T$ then the PTA is sensitive to frequencies in the range $(1/T)<f<(1/\delta t)$.
+The characteristic strain that the PTA is sensitive to scales linearly with $f$ in this range.
 This gives the wedge shaped curves plotted in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega}.
-The absolute value of the sensitivity is fixed by normalising (\ref{eq:PTA}) to agree with a calculated limit at a given frequency for each PTA.
+The absolute value of the sensitivity is fixed by normalising to a calculated limit at a given frequency for each PTA.
+For a discussion of the sensitivities of PTA to both individual sources and stochastic background see \cite{MooreTaylorGair}.
 
@@
 When observing a transient source, such as a burst or inspiral, which changes within the lifetime of the detector it is natural to consider the detector as performing constantly while the signal changes.
-However when observing a monochromatic source or a stochastic background which is unchanging over the mission lifetime it is more natural to consider the source as being fixed and the sensitivity of the detector gradually improving.
+However when observing a monochromatic source or a stochastic background which is unchanging over the detector lifetime it is more natural to consider the source as being fixed and the sensitivity of the detector gradually improving.
 All that is required by the definition in (\ref{eq:hc}) is that the ratio $h_{c}(f)/h_{n}(f)$ is constant.
+
+}
 

### 2014-07-02 Sources.tex
small changes to section 5

@@
 \section{Astrophysical sources}\label{sec:sources}
-All the sources described here are represented by boxes in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega}.
+All the sources described here are represented by shaded boxes in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega}.
 The boxes are drawn in such a way that there is a reasonable event rate for sources at a detectable SNR.
 Descriptions of each source are given in the following sections.
-Sources with short durations (i.e., burst sources) and sources which evolve in time over much longer timescales than our observations are drawn with flat topped boxes.
-Inspiralling sources which change their frequency over observable timescales are drawn with sloping tops.
-The slope for the top of the box is chosen to be $-2/3$ for all sources, which is also the appropriate gradient for a stochastic backgrounds of binaries.
+Sources with short durations and sources which evolve in time over much longer timescales than our observations are drawn with flat topped boxes.
+Inspiralling binaries, or stochastic backgrounds of binaries, are drawn with a sloping top proportional to $f^{-2/3}$, which is result derived in section \ref{sec:stoch}.
 The width of the box gives the range of frequencies sources of a given type can have a while remaining at a detectable amplitude.
@@
 \subsubsection{Galactic white dwarf binaries} \label{sec:GB}
+These are the most numerous GW sources, they are also the only source where several guaranteed sources have already been indentified by electromagnetic observations, these are known as the verification binaries, \cite{2006CQGra..23S.809S}.
 
-\cjm{This section should probably be larger and more detailed. These are by far the most numerous sources and at the moment I don't even mention verification binaries.}
-
-These divide into two classes, the unresolvable and the resolvable galactic binaries.
-The distinction between resolvable and unresolvable is detector specific; here we choose LISA.
-For the unresolvable binaries the box plotted corresponds to the estimate of the background due to \cite{Nelemans},
+Galactic binaries divide into two classes; the unresolvable and the resolvable galactic binaries.
+The unresolvable binaries overlap to form a stochastic background as discussed in section \ref{sec:stoch}.
+Plotted in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega} is the estimate of this background
+due to \cite{Nelemans} where an observation time of one year has been assumed
 \begin{equation} h_{c}(f)= 5\times 10^{-21} \left(\frac{f}{10^{-3}\,\textrm{Hz}}\right)^{-2/3} \; . \end{equation}
-This background is unchanging over the mission lifetime, therefore, as discussed in section \ref{sec:PTAgeneralproperties}, when plotting the amplitude on a sensitivity curve plot an observation time must be specified; here an observation time of one year is used.
 Estimates for the event rate of resolvable binaries centre around ${\cal{O}}(10^{3})$ events for eLISA.

### 2014-07-02 noise_curve.tex
tweaks
new acknowledgements

@@
 \title{Gravitational-wave sensitivity curves}
-\author{C J Moore$^1$, C P L Berry$^{1,\,2}$ \& R H Cole$^1$}
+\author{C J Moore$^1$, R H Cole$^1$ \& C P L Berry$^{1,\,2}$}
 \address{$^1$ Institute of Astronomy, Madingley Road, Cambridge, CB3 0HA, United Kingdom}
@@
 
-\ack{Christopher Moore is supported by the STFC.
-I would like to thank Dr Jonathan Gair, Christopher Berry and Robert Cole for help producing this work.
-I would also like to thank Robert Cole for help producing the figures on the website.
-}
+\ack{C Moore and R Cole are supported by the STFC. C Berry is supported by ... We are indebted to Dr Jonathan Gair, for his useful advice and for proof-reading the manuscript.}
 

### 2014-07-02 Detectors.tex

@@
 \subsection{Pulsar timing arrays (PTAs)}\label{sec:PTAgeneralproperties}
-\cjm{
 PTAs can be thought of as naturally occurring interferometers with galactic scale arm lengths, hence they are sensitive to much lower frequencies than the detectors considered so far.
@@
 
-}
-
 \subsubsection{EPTA/PPTA/NANOGrav}

### 2014-07-02 Introduction.tex

@@
 When producing these plots, it is desirable to have a consistent convention between detectors and sources which allows information about both to be plotted on the same graph.
-\cplb{Ideally,} the detectors and sources are represented in such a way that the relative detectability of the signals is immediately apparent from the plot.
+Ideally, the detectors and sources are represented in such a way that the relative detectability of the signals is immediately apparent from the plot.
 
@@
 In \ref{app:a} several example sensitivity curves are presented.
-\cplb{A website where similar figures can be generated is available at \url{www.ast.cam.ac.uk/~rhc26/sources/}.
-Here the user may select which sources and detectors to include to tailor the figure to their specific requirements.}
+A website where similar figures can be generated is available at \url{www.ast.cam.ac.uk/~rhc26/sources/}.
+Here the user may select which sources and detectors to include to tailor the figure to their specific requirements.
 

### 2014-07-02 Sources.tex

@@
 
-{\cjm{If someone is feeling keen they can write a very short section on these, maybe using this reference \cite{Einstein@Home}}}.
+{\cjm{If someone is feeling keen they can write a very short section on these, maybe using this reference \cite{Einstein@Home}.
+Can whoever writes this section say that in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega} the rotating neutron stars share the same box as supernova}.
 
@@
 The current best published limit for the amplitude of the stochastic background is $h_{c} = 6\times 10^{-15}$ at a frequency of $f_{0}=1~\mathrm{yr}^{-1}$, \cite{Haasteren}.
-There is strong theoretical evidence that the actual background lies close to the current limit, \cite{imminentdetectionofgravitationalwaves} and \cite{Sesana-2012}.
+There is strong theoretical evidence that the actual background lies close to the current limit, \cite{Sesana-2012}.
 
-Sources at higher frequencies are inspiralling faster and hence there are fewer of them per frequency bin.
+Supermassive black hole binaries at higher frequencies are inspiralling faster and hence there are fewer of them per frequency bin.
 At a certain frequency, these sources will cease to be a background and become individually resolvable.
-The exact distinction between resolvable and unresolvable is detector specific.
+It is currently unclear whether PTAs will detect an individual binary or a stochastic background first.
 Plotted in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega} for the unresolvable background is a third of the current limit with a cut off frequency of $f=1~\mathrm{yr}^{-1}$ which is suggested by Monte Carlo population studies, \cite{SesanaVecchioColancino}.

### 2014-07-02 Detectors.tex
changed refs to Appendix A

@@
 The interferometric detectors listed in Table \ref{table:t} are all sensitive to GWs in the frequency range ${\cal{O}}(10-10^{3})\,\textrm{Hz}$.
-They all have multiple narrow lines in their sensitivity curves, which arise from noise sources in the instrument including resonances in the suspension system and electronic noise at multiples of $60\,\textrm{Hz}$: these have been removed in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega} for clarity.
+They all have multiple narrow lines in their sensitivity curves, which arise from noise sources in the instrument including resonances in the suspension system and electronic noise at multiples of $60\,\textrm{Hz}$: these have been removed in the \ref{app:a} figures for clarity.
 The detectors fall broadly into three categories: 1\textsuperscript{st} generation detectors, which have already operated; 2\textsuperscript{nd} generation detectors currently under construction; and 3\textsuperscript{rd} generation detectors at the planning stage.
@@
 The characteristic strain that the PTA is sensitive to scales linearly with $f$ in this range.
-This gives the wedge shaped curves plotted in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega}.
+This gives the wedge shaped curves plotted in \ref{app:a}.
 The absolute value of the sensitivity is fixed by normalising to a calculated limit at a given frequency for each PTA.
@@
 \subsubsection{SKA}
-The sensitivity curve plotted in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega} for the SKA (Square Kilometer Array) assume a values of $\delta t^{2}_{\textrm{r.m.s.}}$ a factor of 10 better than current PTAs and 20 pulsars timed for twice as long as the current EPTA.
+The sensitivity curves plotted in \ref{app:a} for the SKA (Square Kilometer Array) assume values of $\delta t^{2}_{\textrm{r.m.s.}}$ a factor of 10 better than current PTAs and 20 pulsars timed for twice as long as the current EPTA.

### 2014-07-02 Detectors.tex
added the detector references Christopher suggested to table 1

--- 2014-07-02 Detectors.tex
@@
 \mr
-  GEO600 	&	Germany 	& $600\,\textrm{m}$ 	& 2001-present 	   & 1$^{\textrm{st}}$\\
-  TAMA300 	& 	Japan		& $300\,\textrm{m}$ 	& 1995-present     & 1$^{\textrm{st}}$\\
-  iLIGO		&	US		& $4\,\textrm{km}$ 	& 2004-2010 	   & 1$^{\textrm{st}}$\\
-  iVIRGO	& 	Italy		& $3\,\textrm{km}$ 	& 2007-2010 	   & 1$^{\textrm{st}}$\\
-  aLIGO 	&	US		& $4\,\textrm{km}$ 	& \emph{est.} 2016 & 2$^{\textrm{nd}}$\\
-  KAGRA		&	Japan		& $3\,\textrm{km}$ 	& \emph{est.} 2018 & 2$^{\textrm{nd}}$\\
-  aVIRGO	&	Italy	 	& $3\,\textrm{km}$ 	& \emph{est.} 2017 & 2$^{\textrm{nd}}$\\
-  ET (Einstein Telescope) &	Italy		& $10\,\textrm{km}$ 	& \emph{est.} 2025 & 3$^{\textrm{rd}}$\\
+  GEO600 $^{a}$	&	Germany 	& $600\,\textrm{m}$ 	& 2001-present 	   & 1$^{\textrm{st}}$\\
+  TAMA300 $^{b}$& 	Japan		& $300\,\textrm{m}$ 	& 1995-present     & 1$^{\textrm{st}}$\\
+  iLIGO	$^{c}$	&	US		& $4\,\textrm{km}$ 	& 2004-2010 	   & 1$^{\textrm{st}}$\\
+  iVIRGO $^{d}$	& 	Italy		& $3\,\textrm{km}$ 	& 2007-2010 	   & 1$^{\textrm{st}}$\\
+  aLIGO $^{e}$	&	US		& $4\,\textrm{km}$ 	& \emph{est.} 2016 & 2$^{\textrm{nd}}$\\
+  KAGRA	$^{f}$	&	Japan		& $3\,\textrm{km}$ 	& \emph{est.} 2018 & 2$^{\textrm{nd}}$\\
+  aVIRGO $^{g}$	&	Italy	 	& $3\,\textrm{km}$ 	& \emph{est.} 2017 & 2$^{\textrm{nd}}$\\
+  ET (Einstein Telescope)$^{h}$ &Italy		& $10\,\textrm{km}$ 	& \emph{est.} 2025 & 3$^{\textrm{rd}}$\\
+\br
+\multicolumn{5}{l}{$^{a}$\cite{2008CQGra..25k4043G}, $^{b}$\cite{2002CQGra..19.1409A}, }\\
+\multicolumn{5}{l}{$^{c}$\cite{2009RPPh...72g6901A}, $^{d}$\cite{accadia_virgo:_2012}, $^{e}$\cite{2010CQGra..27h4006H}, }\\
+\multicolumn{5}{l}{$^{f}$\cite{2012CQGra..29l4007S}, $^{g}$\cite{Acernese2009}, $^{h}$\cite{2011CQGra..28i4013H}.}\\
 \br

### 2014-07-02 Sources.tex

@@
 \section{Astrophysical sources}\label{sec:sources}
-All the sources described here are represented by shaded boxes in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega}.
+All the sources described here are represented by shaded boxes in \ref{app:a}.
 The boxes are drawn in such a way that there is a reasonable event rate for sources at a detectable SNR.
@@
 The expected event rate for this type of source is uncertain, but estimates centre around $\rho_{\textrm{NS-NS}}=1.3\times 10^{-4}\;\textrm{Mpc}^{-3}\textrm{yr}^{-1}$, \cite{CBC}.
-Plotted in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega} is a box with an amplitude such that it produces a ratio $h_{c}/h_{n}=16$ for advanced LIGO at peak sensitivity and a width between $(3-300)\,\textrm{Hz}$.
+Plotted in \ref{app:a} is a box with an amplitude such that it produces a ratio $h_{c}/h_{n}=16$ for advanced LIGO at peak sensitivity and a width between $(3-300)\,\textrm{Hz}$.
 
@@
 The event rate for supernova is approximately $\rho_{\textrm{SN}}=5\times10^{-4}\,\textrm{Mpc}^{-3}\textrm{yr}^{-1}$.
-The box plotted in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega} corresponds to a distance $r=3 \,\textrm{Mpc}$ with the frequency range quoted above.
+The boxes plotted in \ref{app:a} correspond to a distance $r=3 \,\textrm{Mpc}$ with the frequency range quoted above.
 The LIGO and VIRGO detectors have already placed bounds on the event rate for these sources, \cite{Bursts}.
@@
 The uncertainty in the rate reflects our uncertainty in the growth mechanisms of the supermassive black hole population.
-Plotted in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega} is a box with a ratio $h_{c}/h_{n}=100$ for eLISA at its peak sensitivity.
+Plotted in \ref{app:a} are boxes with a ratio $h_{c}/h_{n}=100$ for eLISA at its peak sensitivity.
 The range of frequencies plotted is $(3\times 10^{-4}-3\times 10^{-1})\,\textrm{Hz}$, this corresponds to circular binaries in the mass range quoted above.
@@
 
-Galactic binaries divide into two classes; the unresolvable and the resolvable galactic binaries.
+Galactic binaries divide into two classes: the unresolvable and the resolvable galactic binaries.
 The unresolvable binaries overlap to form a stochastic background as discussed in section \ref{sec:stoch}.
-Plotted in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega} is the estimate of this background
-due to \cite{Nelemans} where an observation time of one year has been assumed
+Plotted in \ref{app:a} is the estimate of this background due to \cite{Nelemans} where an observation time of one year has been assumed
 \begin{equation} h_{c}(f)= 5\times 10^{-21} \left(\frac{f}{10^{-3}\,\textrm{Hz}}\right)^{-2/3} \; . \end{equation}
 Estimates for the event rate of resolvable binaries centre around ${\cal{O}}(10^{3})$ events for eLISA.
-The box plotted in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega} has an ratio $h_{c}/h_{n}=50$ for eLISA at its peak sensitivity, the frequency range of the box is $\left(3\times10^{-4}\right.$--$\left.10^{-2}\right)~\mathrm{Hz}$, based on visual inspection of Monte Carlo population simulation results presented in \cite{Amaro-Seoane-et-al}.
+The boxes plotted in \ref{app:a} have a ratio $h_{c}/h_{n}=50$ for eLISA at its peak sensitivity.
+The frequency range of the box is $\left(3\times10^{-4}\right.$--$\left.10^{-2}\right)~\mathrm{Hz}$, based on visual inspection of Monte Carlo population simulation results presented in \cite{Amaro-Seoane-et-al}.
 
@@
 There is extreme uncertainty in the event rate for EMRIs due to the poorly constrained astrophysics in galactic centres; the best guess estimate is ${\mathcal{O}}(10)$ events per year with eLISA with SNR $\ge 20$.
-The box plotted in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega} has a characteristic strain of $h_{c}=3\times 10^{-20}$ at $10^{-2}~\mathrm{Hz}$ which corresponds to a $10\Msun$ BH inspiralling into a $10^{6}\Msun$ black hole at a luminosity distance of $1~\mathrm{Gpc}$.
-The frequency width of the box is somewhat unknown, EMRI events can occur into a black hole of any mass, and hence EMRIs can occur at any frequency.
-However there is substantial uncertainty in the black hole mass function, so the box in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega} is drawn with a width comparable to that of the LISA sensitivity curve.
+The boxes plotted in \ref{app:a} have a characteristic strain of $h_{c}=3\times 10^{-20}$ at $10^{-2}~\mathrm{Hz}$, which corresponds to a $10\Msun$ BH inspiralling into a $10^{6}\Msun$ black hole at a luminosity distance of $1~\mathrm{Gpc}$.
+The frequency width of the box is somewhat unknown; EMRI events can occur into a black hole of any mass, and hence EMRIs can occur at any frequency.
+However there is substantial uncertainty in the black hole mass function, so the boxes in \ref{app:a} are drawn with a width comparable to that of the LISA sensitivity curve.
 
@@
 It is currently unclear whether PTAs will detect an individual binary or a stochastic background first.
-Plotted in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega} for the unresolvable background is a third of the current limit with a cut off frequency of $f=1~\mathrm{yr}^{-1}$ which is suggested by Monte Carlo population studies, \cite{SesanaVecchioColancino}.
+Plotted in \ref{app:a} for the unresolvable background is a third of the current limit with a cut off frequency of $f=1~\mathrm{yr}^{-1}$ which is suggested by Monte Carlo population studies, \cite{SesanaVecchioColancino}.
 For the resolvable sources the amplitude of the current limit is plotted between $\left(3\times 10^{-9}\right.$--$\left.3\times 10^{-7}\right)~\mathrm{Hz}$.
@@
 Cosmological stochastic backgrounds have been predicted across a range of frequencies with considerable variation in amplitude.
-As a consequence of this uncertainty, although we could learn much from measuring these signals, we have not included them amongst the sources shown in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega}.
+As a consequence of this uncertainty, although we could learn much from measuring these signals, we have not included them amongst the sources shown in \ref{app:a}.

### 2014-07-03 Introduction.tex
Introduction

@@
 The next few years promises to deliver the first direct detection of gravitational waves (GWs).
-This will most likely be achieved by the advanced LIGO \citep{2010CQGra..27h4006H} and Virgo \citep{Accadia2011} detectors operating in the frequency range $(10$--$10^{3})~\mathrm{Hz}$.
-By the end of the decade, it is expected that pulsar timing arrays \citep{1990ApJ361300F} shall detect very low frequency GWs around $10^{-8}~\mathrm{Hz}$.
-Further into the future, space-based detectors, such as eLISA \citep{2012CQGra..29l4016A}, will probe GWs in the millihertz regime.
-This shall herald the beginning of multi-wavelength GW astronomy.
+This will most likely be achieved by the advanced Laser Interferometer Gravitational-wave Observatory \citep[LIGO;][]{2010CQGra..27h4006H} and Virgo \citep{Accadia2011} detectors operating in the frequency range $(10$--$10^{3})~\mathrm{Hz}$.
+By the end of the decade, it is expected that pulsar timing arrays \citep[PTAs;][]{1990ApJ361300F} shall detect very low frequency GWs around $10^{-8}~\mathrm{Hz}$.
+Further into the future, space-based detectors, such as the evolved Laser Interferometer Space Antenna \citep[eLISA;][]{2012CQGra..29l4016A}, will probe GWs in the millihertz regime.
+These advances shall herald the beginning of multi-wavelength GW astronomy as a means of observing the Universe.
 
@@
 There are several different methods commonly used to describe the sensitivity of a GW detector and the strength of a GW source.
-It is common practice in GW astronomy to summarise this information on a sensitivity curve plot.
+It is common practice to summarise this information on a sensitivity-curve plot.
 When producing these plots, it is desirable to have a consistent convention between detectors and sources which allows information about both to be plotted on the same graph.
@@
 
-In this work we discuss the differing conventions commonly used in sensitivity curve plots.
+In this work, we discuss the differing conventions commonly used in GW astronomy.
 The amplitude of a GW is a strain, a dimensionless quantity $h$.
 This gives a fractional change in length, or equivalently light travel time, across a detector.
-The strain is small, making it a challenge to measure.
-To calibrate our expectations for detection it is necessary to quantify the sensitivity of our instruments and the strength of the target signals.
-There are three commonly used descriptions: the characteristic strain, the power spectral density and the spectral energy density.
+The strain is small, making it a challenge to measure: we are yet to obtain a direct detection
+of a GW. To calibrate our expectations for detection, it is necessary to quantify the sensitivity of our instruments and the strength of their target signals.
+
+When discussing the loudness of sources and the sensitivity of detectors there are three commonly used parametrizations based upon the strain: the characteristic strain, the power spectral density and the spectral energy density.
 We aim to disambiguate these three and give a concrete comparison of different detectors.

### 2014-07-03 noise_curve.tex
added the detector references Christopher suggested to table 1
Introduction

@@
 \usepackage{graphicx}
+\usepackage{multirow}
 \usepackage{amssymb}
@@
 \begin{abstract}
-There are several common conventions use by the gravitational wave community to describe the amplitude of sources and the sensitivity of detectors.
+There are several common conventions use by the gravitational-wave community to describe the amplitude of sources and the sensitivity of detectors.
 These are frequently confused.
-We outline the merits of and differences between the various quantities used for parameterizing noise curves and characterizing gravitational wave amplitudes.
+We outline the merits of and differences between the various quantities used for parameterizing noise curves and characterizing gravitational-wave amplitudes.
 We conclude by producing plots that consistently compare different detectors.
 Similar figures can be generated on-line for general use at \url{http://www.ast.cam.ac.uk/~rhc26/sources/}.
-
-\cplb{Slightly tweaked and added PACS}
 \end{abstract}
@@
 
-\ack{C Moore and R Cole are supported by the STFC. C Berry is supported by ... We are indebted to Dr Jonathan Gair, for his useful advice and for proof-reading the manuscript.}
+\ack{CJM, RHC and CPLB are supported by the Science \& Technology Facilities Council.
+We are indebted to Jonathan Gair, for useful advice and proof-reading the manuscript.
+We also thank Stephen Taylor and Justin Ellis for useful conversations about pulsar timing.}
 

### 2014-07-03 Introduction.tex
Section 2

@@
 
-\section{Signal analysis}\label{sec:conventions}
+\section{Signal parametrization}\label{sec:conventions}
+
+\subsection{Signal analysis preliminaries}
 
 Gravitational radiation has two independent polarization states denoted $+$ and $\times$; a general signal can be described as a linear combination of the two polarization states, $h \nobreak=\nobreak A_{+} h_{+}\nobreak +\nobreak A_{\times} h_{\times}$.
-The sensitivity of a detector to each of these depends upon the relative orientations of the source and detector.
-The output of a gravitational wave detector, $s(t)$, will contain a superposition of noise, $n(t)$, and possibly a signal, $h(t)$,
-\begin{equation} s(t)=n(t)+h(t) \; . \end{equation}
-It will be necessary to work with the Fourier transform of the signal, the most common convention used in the GW community is
-\begin{eqnarray} x(t)&={\cal{F}}^{-1}\left\{\tilde{x}(f)\right\}&=\int_{-\infty}^{\infty}\mathrm{d}f\;\tilde{x}(f)\exp (2\pi \rmi f t )  \nonumber \\
-\tilde{x}(f) &=\,\;{\cal{F}}\left\{x(t)\right\}&=\int_{-\infty}^{\infty}\mathrm{d}t\;x(t)\exp (-2\pi \rmi f t )\; .\end{eqnarray}
-For simplicity it is assumed that the noise in the GW detector is stationary and Gaussian (with zero mean), under these assumptions the noise is fully characterised via the \emph{one-sided} noise power spectral density, $S_{n}(f)$,\footnote{An alternative convention is to use the two-sided power spectral density $S^{(2)}_{n}(f)=S_{n}(f)/2$.},
-\begin{equation}\label{eq:psd} \left<\tilde{n}(f)\tilde{n}^{*}(f')\right>=\frac{1}{2}\delta (f-f')S_{n}(f) \; .\end{equation}
-Angled brackets $\left<\ldots\right>$ denote an ensemble average over many realisations of the noise.
-In reality we have only a single system to work with, so the ensemble average is replaced by a time average.
-The procedure is to measure the noise over a sufficiently long duration, $T$, and then compute the Fourier transform, $\tilde{n}(f)$, with a frequency resolution $\Delta f = T^{-1}$.
-This is then repeated many times to give the ensemble average; this procedure assumes the system is ergodic.
-Since the GW signal and detector output are real it follows that $\tilde{h}(-f)=\tilde{h}^{*}(f)$ and $\tilde{n}(-f)=\tilde{n}^{*}(f)$; therefore, $S_{n}(f)=S_{n}(-f)$.
-The fact $S_{n}(f)$ is an even function means that Fourier integrals over all frequencies can instead be written as integrals over positive frequencies only (see, for example, equations (\ref{eq:meansquare1}) and (\ref{eq:snrinnerprod})), it is for this reason that $S_{n}(f)$ is called the \emph{one-sided} power spectral density.
-It is straightforward to show that the power spectral density has the property that integrated over all positive frequencies it gives the mean square noise amplitude in the detector.
+The sensitivity of a detector to these depends upon the relative orientations of the source and detector.
+The output of a gravitational wave detector $s(t)$ contains a superposition of noise $n(t)$ and (possibly) the signal $h(t)$,
+\begin{equation}
+s(t) = n(t)+h(t) \; .
+\end{equation}
+We will work with the Fourier transform of the signal, using the conventions that
+\begin{eqnarray} 
+\tilde{x}(f) &= {\mathcal{F}}\left\{x(t)\right\} &= \int_{-\infty}^{\infty}\mathrm{d}t\;x(t)\exp (-2\pi \rmi f t ) \; , \\
+x(t) &= {\mathcal{F}}^{-1}\left\{\tilde{x}(f)\right\} &= \int_{-\infty}^{\infty}\mathrm{d}f\;\tilde{x}(f)\exp (2\pi \rmi f t ) \; .
+\end{eqnarray}
+
+For simplicity, it is assumed that the noise in the GW detector is stationary and Gaussian (with zero mean); under these assumptions the noise is fully characterised via the one-sided noise power spectral density $S_{n}(f)$,
+\begin{equation}\label{eq:psd}
+\left<\tilde{n}(f)\tilde{n}^{*}(f')\right>=\frac{1}{2}\delta (f-f')S_{n}(f) \; ,
+\end{equation}
+where angle brackets $\left<\ldots\right>$ denote an ensemble average over many noise realisations \citep{Cutler1994}.
+In reality, we have only a realisation to work with, however the ensemble average can replaced by a time average for stationary stochastic noise.
+The procedure is to measure the noise over a sufficiently long duration $T$ and then compute the Fourier transform $\tilde{n}(f)$ with a frequency resolution $\Delta f = T^{-1}$; this is repeated many times to give an average.
+The noise power spectral density $S_{n}(f)$ has units of inverse frequency.
+
+Since the GW signal and detector output are both real it follows that $\tilde{h}(-f)=\tilde{h}^{*}(f)$ and $\tilde{n}(-f)=\tilde{n}^{*}(f)$; therefore, $S_{n}(f)=S_{n}(-f)$.
+The fact that $S_{n}(f)$ is an even function means that Fourier integrals over all frequencies can instead be written as integrals over positive frequencies only, e.g., (\ref{eq:meansquare1}) and (\ref{eq:snrinnerprod})); it is for this reason that $S_{n}(f)$ is called the \emph{one-sided} power spectral density.\footnote{An alternative convention is to use the two-sided power spectral density $S^{(2)}_{n}(f) = S_{n}(f)/2$.}
+
+When integrated over all positive frequencies the power spectral density gives the mean square noise amplitude. Starting with taking the average of the square of the detector noise:
 \begin{eqnarray}\label{eq:meansquare}
-\overline{n(t)^{2}}&=\lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T}\mathrm{d}t\;\left| n(t) \right|^{2} \nonumber \\
-&=\lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T}\mathrm{d}t\; \int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\tilde{n}(f)\tilde{n}^{*}(f')\exp\left(2\pi \rmi (f-f')t \right)
+\overline{n(t)^{2}} &= \lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T}\mathrm{d}t\;\left| n(t) \right|^{2} \nonumber \\
+ &= \lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T}\mathrm{d}t\; \int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\tilde{n}(f)\tilde{n}^{*}(f')\exp\left(2\pi \rmi ft \right)\exp\left(-2\pi \rmi f't \right)
 \end{eqnarray}
-Using the property of Fourier transforms, ${\cal{F}}\left\{ n(t+\tau) \right\}(f)=\tilde{n}(f)\exp (-2\pi \rmi f\tau )$, the exponential factors in equation (\ref{eq:meansquare}) may be absorbed as a time translation of the noise by an amount $t$.
-The noise is stationary and so its long term time average is invariant under this time translation and is given by Equation (\ref{eq:psd}).
+Using the Fourier property ${\mathcal{F}}\left\{ n(t+\tau) \right\} = \tilde{n}(f)\exp (-2\pi \rmi f\tau )$, the exponential factors may be absorbed as a time-translation of the noise by an amount $t$.
+The noise is stationary and is invariant under this time translation.
+We are thus taking the time average...
+\cplb{I give up here. This derivation makes no sense in it's current state.}
 \begin{eqnarray}\label{eq:meansquare1}
-\overline{n(t)^{2}}& = \lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T}\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;{\cal{F}}\left\{ n(\tau-t) \right\}(f){\cal{F}}\left\{ n(\tau-t) \right\}(f')^{*} \nonumber \\
+\overline{n(t)^{2}} & = \lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T}\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;{\cal{F}}\left\{ n(\tau-t) \right\}(f){\cal{F}}\left\{ n(\tau-t) \right\}(f')^{*} \nonumber \\
 &=\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\left<\tilde{n}(f)\tilde{n}^{*}(f')\right>\nonumber \\
@@
 \end{eqnarray}
-
 Using the property of Fourier transforms, ${\cal{F}}\left\{ n(t+\tau) \right\}(f)=\tilde{n}(f)\exp (-2\pi \rmi f\tau )$, the time average can be turned into an ensemble average and hence the definition of power spectral density in (\ref{eq:psd}) may be used.
@@
 \end{eqnarray}
-Given a detector output the challenge is to extract the signal.
-There is a well known solution to this problem which involves constructing a Wiener optimal filter.
+\cplb{Try someting based on Wiener--Khinchin theorem?}
+
+Given a detector output, the challenge is to extract the signal.
+There is a well known solution to this problem which involves constructing a Wiener optimal filter \cplb{[citation needed]}.
 Let $K(t)$ be the filter function with Fourier transform $\tilde{K}(f)$.
 Convolving this with the detector output gives a contribution from the signal and a contribution from the noise
-\begin{equation}\label{eq:conv} \left(s*K\right)(\tau)=\int_{-\infty}^{\infty}\mathrm{d}t\;(n(t)+h(t))K(t-\tau)=S+N\; .\end{equation}
-The signal contribution, $S$, is defined as the expectation of the convolution in (\ref{eq:conv}) when a signal is present, maximised by varying the offset $\tau$.
-Since the expectation of pure noise is zero it follows that
-\begin{equation} S = \int_{-\infty}^{\infty}\mathrm{d}t\;h(t)K(t)=\int_{-\infty}^{\infty}\mathrm{d}t\;h(t)K^{*}(t)=\int_{-\infty}^{\infty}\mathrm{d}f\; \tilde{h}(f)\tilde{K}^{*}(f) \; .\end{equation}
-The squared contribution from noise, $N^{2}$, is defined as the mean square of the convolution in (\ref{eq:conv}) when no signal is present,
+\begin{equation}\label{eq:conv}
+\left(s*K\right)(\tau) = \int_{-\infty}^{\infty}\mathrm{d}t\;(n(t)+h(t))K(t-\tau) = \mathcal{S} + \mathcal{N} \; .
+\end{equation}
+The signal contribution $\mathcal{S}$ is defined as the expectation of the convolution in (\ref{eq:conv}) when a signal is present, maximised by varying the offset $\tau$.
+\cplb{I think this needs expanding.} Since the expectation of pure noise is zero it, follows that
+\begin{equation}
+\mathcal{S} = \int_{-\infty}^{\infty}\mathrm{d}t\;h(t)K(t)=\int_{-\infty}^{\infty}\mathrm{d}t\;h(t)K^{*}(t)=\int_{-\infty}^{\infty}\mathrm{d}f\; \tilde{h}(f)\tilde{K}^{*}(f) \; .
+\end{equation}
+The squared contribution from noise $\mathcal{N}^{2}$ is defined as the mean square of the convolution in (\ref{eq:conv}) when no signal is present,
 \begin{eqnarray} 
-N^{2}&=\int_{-\infty}^{\infty}\mathrm{d}t\;\int_{-\infty}^{\infty}\mathrm{d}t'\;K(t)K(t')\left<n(t)n(t')\right> \nonumber \\
-&=\int_{-\infty}^{\infty}\mathrm{d}t\;\int_{-\infty}^{\infty}\mathrm{d}t'\;K(t)K^{*}(t')\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\left<\tilde{n}(f)\tilde{n}^{*}(f')\right>\exp\left(2\pi\rmi(ft-f't')\right)\nonumber \\
+\mathcal{N}^{2} &= \int_{-\infty}^{\infty}\mathrm{d}t\;\int_{-\infty}^{\infty}\mathrm{d}t'\;K(t)K(t')\left<n(t)n(t')\right> \nonumber \\
+ &= \int_{-\infty}^{\infty}\mathrm{d}t\;\int_{-\infty}^{\infty}\mathrm{d}t'\;K(t)K^{*}(t')\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\left<\tilde{n}(f)\tilde{n}^{*}(f')\right>\exp\left[2\pi\rmi(ft-f't')\right]\nonumber \\
 %&=\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}t\;\int_{-\infty}^{\infty}\mathrm{d}t'\;\frac{1}{2}S_{n}(f)K(t)K^{*}(t')\exp\left(2\pi\rmi f (t-t')\right)\nonumber \\
-&=\int_{-\infty}^{\infty}\mathrm{d}f \; \frac{1}{2}S_{n}(f)\tilde{K}(f)\tilde{K}^{*}(f) \; ,
+ &= \int_{-\infty}^{\infty}\mathrm{d}f \; \frac{1}{2}S_{n}(f)\tilde{K}(f)\tilde{K}^{*}(f) \; ,
  \end{eqnarray}
-where the definition of $S_{n}(f)$ from (\ref{eq:psd}) has been used. Hence the optimum signal-to-noise ratio (SNR), $\varrho$, is given by
+using the definition of $S_{n}(f)$ from (\ref{eq:psd}). Hence the optimum signal-to-noise ratio (SNR) $\varrho$ is given by \cplb{Why is this optimal?}
 \begin{equation}\label{eq:SNRinnerprod} 
-\varrho^{2}= \frac{(S)^{2}}{N^{2}}= \frac{\left( \frac{1}{2}S_{n}(f)\tilde{K}(f) \Big| \tilde{h} \right)^{2}}{\left( \frac{1}{2}S_{n}(f)\tilde{K}(f) \Big| \frac{1}{2}S_{n}(f)\tilde{K}(f) \right)} \;,
+\varrho^{2} = \frac{\mathcal{S}^{2}}{\mathcal{N}^{2}}= \frac{\innerprod{\frac{1}{2}S_{n}(f)\tilde{K}(f)}{\tilde{h}}^{2}}{\innerprod{\frac{1}{2}S_{n}(f)\tilde{K}(f)}{\frac{1}{2}S_{n}(f)\tilde{K}(f)}} \;,
 \end{equation}
-where the inner product between a pair of signals $A$ and $B$ has been defined
+\cplb{Inner product should be between signals: you don't specify the Fourier transform} where we have introduced the inner product between signal $A$ and $B$ as \citep{Finn1992}
 \begin{equation}\label{eq:snrinnerprod} \left( A | B \right) = 4\Re\left\{\int_{0}^{\infty}\mathrm{d}f\;\frac{\tilde{A}^{*}(f)\tilde{B}(f)}{S_{n}(f)}\right\} \; .\end{equation}
-The optimum filter is that which maximises the SNR in (\ref{eq:SNRinnerprod}). From the Cauchy-Schwarz inequality it follows that the optimum filter is
-\begin{equation} \tilde{K}(f)=\frac{\tilde{h}(f)}{S_{n}(f)} \; . \end{equation}
-This is the Wiener filter, it may be multiplied by an arbitrary constant since this will not change the SNR.
-Notice that in order to construct the Wiener filter it is necessary to know \emph{a priori} the form of the signal, $\tilde{h}(f)$.
-It is common to define the noise amplitude and characteristic strain as
-\begin{eqnarray}\label{eq:strains} h_{n}(f)^{2}&=fS_{n}(f) \\
-\label{eq:strains1} \label{eq:strainsagain}               h_{c}(f)^{2}&=4f^{2}\left| \tilde{h}(f) \right|^{2} \; , \end{eqnarray}
-so that the SNR in (\ref{eq:SNRinnerprod}) may be written compactly as
-\begin{equation}\label{eq:hc} 
-\varrho^{2}=\int_{0}^{\infty} \mathrm{d}\left(\log f\right)\; \left(\frac{h_{c}(f)}{h_{n}(f)}\right)^{2} \;.
+The optimum filter maximises the SNR in (\ref{eq:SNRinnerprod}). From the Cauchy--Schwarz inequality, it follows that the optimum filter is
+\begin{equation}
+\tilde{K}(f)=\frac{\tilde{h}(f)}{S_{n}(f)} \; .
 \end{equation}
-The strain amplitudes $h_{n}(f)$ and $h_{c}(f)$ are dimensionless, while $S_{n}(f)$ has units of inverse frequency.
+This is the Wiener filter, it may be multiplied by an arbitrary constant since this does not change the SNR. Using this form for $\tilde{K}(f)$, the squared SNR is
+\begin{equation}
+\varrho^2 = \int_0^\infty\mathrm{d}f \frac{4| \tilde{h}(f)|^{2}}{S_n(f)} = \innerprod{h}{h}.
+\label{eq:traditionalSNR} 
+\end{equation}
+In order to construct the Wiener filter, it is necessary to know \emph{a priori} the form of the signal, $\tilde{h}(f)$.
 
-Whilst the magnitude of the Fourier transform of the signal, $h(f)=\left|\tilde{h}(f)\right|$, provides a simple quantification of the GW amplitude as a function of frequency it has one main deficiency.
-For an inspiralling source the instantaneous amplitude can be orders of magnitude below the noise level in the detector; however, as the signal continues over many orbits the SNR can be integrated up to a detectable level.
-It is desirable to have a quantification of the GW amplitude that accounts for this effect; three such methods are described in the following sections.
+Whilst the magnitude of the Fourier transform of the signal $|\tilde{h}(f)|$ provides a simple quantification of the GW amplitude as a function of frequency, it has one main deficiency.
+For an inspiralling source the instantaneous amplitude can be orders of magnitude below the noise level in the detector; however, as the signal continues over many orbits, the SNR can be integrated up to a detectable level.
+It is useful to have a quantification of the GW amplitude that accounts for this effect; three such methods are described in the following subsections.
 
 \subsection{Characteristic strain}
-One sensible choice of quantities to plot on a sensitivity curve is $h_{n}$ for the detector and $h_{c}$ for the source (see figure \ref{fig:hc}).
-Using this convention, when plotting on a log-log scale, the area between the source and detector curves is related to the SNR via (\ref{eq:hc}).
-This convention allows the reader to ``integrate by eye" for a given detector to assess the detectability of a given source.
-Also, when using this convention the reader does not need to worry about the observation time for each source because this is already accounted for in the definition of characteristic strain.
-An additional advantage of this convention is that the values on the strain axis for the detector curve have a simple physical interpretation, they correspond to the root-mean-square noise in a bandwidth $f$.
-One downside to plotting characteristic strain is that the values on the strain axis do not directly relate to the amplitude of the waves from the source.
-Another disadvantage is that the characteristic strain for a monochromatic source is infinite!
+
+The characteristic strain $h_c$ is designed to include the effect of integrating an inspiralling signal. Its counterpart for describing noise is the noise amplitude $h_n$. These are defined as
+\begin{eqnarray}\label{eq:strain-hc} 
+h_\mathrm{c}(f)^{2} &= 4f^{2}\left| \tilde{h}(f) \right|^{2} \; ,\\
+h_{n}(f)^{2} &= fS_{n}(f),
+\label{eq:strain-hn}
+\end{eqnarray}
+such that the SNR in (\ref{eq:traditionalSNR}) may be written
+\begin{equation}\label{eq:hc} 
+\varrho^{2} = \int_{0}^{\infty} \mathrm{d}\left(\log f\right)\; \left[\frac{h_\mathrm{c}(f)}{h_{n}(f)}\right]^{2} \;.
+\end{equation}
+The strain amplitudes $h_\mathrm{c}(f)$ and $h_{c}(f)$ are dimensionless.
+Using this convention, when plotting on a log--log scale, the area between the source and detector curves is related to the SNR via (\ref{eq:hc}).
+This convention allows the reader to integrate by eye to assess the detectability of a given source (see figure \ref{fig:hc}).
+
+An additional advantage of this convention is that the values on the strain axis for the detector curve $h_n(F)$ have a simple physical interpretation: they correspond to the root-mean-square noise in a bandwidth $f$.
+One downside to plotting characteristic strain is that the values on the strain axis $h_\mathrm{c}(f)$ do not directly relate to the amplitude of the waves from the source.
+Another disadvantage is that the characteristic strain for a monochromatic source is infinite.
 
 \subsection{Power spectral density}\label{sec:psd}
-Another common quantity to plot on sensitivity curves is the square root of the power spectral density (see figure \ref{fig:S}), which from (\ref{eq:strains}) is given by
-\begin{eqnarray}\label{eq:temp1} \sqrt{S_{n}(f)}&=h_{n}(f)f^{-1/2} \quad \mathrm{for detectors, and by analogy} \\
-\sqrt{S_{h}(f)}&=h_{c}(f)f^{-1/2} \quad \mathrm{for sources.}\label{eq:ShforSources}\end{eqnarray}
-The power spectral density is the most commonly plotted quantity in sensitivity curves in the literature and it has the nice property, proven in equation (\ref{eq:meansquare1}), that integrated over all positive frequencies it gives the mean square amplitude of the signal in the detector.
+
+\cplb{Can we abbreviate to PSD?}
+
+A second commonly used quantity for sensitivity curves is the square root of the power spectral density (see figure \ref{fig:S}). When discussing a detector this is
+\begin{equation}\label{eq:temp1}
+\sqrt{S_{n}(f)} = h_{n}(f)f^{-1/2} \; ,
+\end{equation}
+where we have used (\ref{eq:strain-hn}); by analogy, we can define an equivalent for for source amplitudes
+\begin{equation}
+\sqrt{S_{h}(f)} = h_\mathrm{c}(f)f^{-1/2} = 2 f^{1/2} \left| \tilde{h}(f) \right| \; ,
+\label{eq:ShforSources}
+\end{equation}
+where we have used (\ref{eq:strain-hc}).
+Both $\sqrt{S_{n}(f)}$ and $\sqrt{S_{h}(f)}$ have units of $\mathrm{Hz^{-1/2}}$.
+The root power spectral density is the most frequently plotted quantity in the literature.
+
+The power spectral density has the nice property, demonstrated in (\ref{eq:meansquare1}), that integrated over all positive frequencies it gives the mean square amplitude of the signal in the detector.
+However, in one important regard it is less appealing than characteristic strain: the height of the source above the detector curve is no longer is directly related to the SNR.
 
 \subsection{Energy density}
-A third way of describing the detectability of a GW is through the energy carried by the waves, this quantity has the obvious advantage of having a clear physical significance.
-This quantity is most commonly used in sensitivity curves showing stochastic backgrounds of GWs, see section \ref{sec:stoch}.
-The energy in GWs is described by the Isaacson stress tensor \citep{MTW}
-\begin{equation} T_{\mu\nu}=\frac{c^{2}}{32\pi G}\left<\partial_{\mu}\bar{h}_{\alpha\beta}\partial_{\nu}\bar{h}^{\alpha\beta}\right> \;.\end{equation}
-The energy density, $\rho c^{2}$, is given by the $T_{00}$ component of this tensor. Using this together with (\ref{eq:psd}) to calculate the energy density in the GW signal gives
-\begin{eqnarray}\label{eq:specNRGdensity}
-\rho c^{2}&=\frac{c^{2}}{16\pi G}\int_{-\infty}^{\infty}\mathrm{d}f\;\left(2\pi f\right)^{2}\tilde{h}(f)\tilde{h}^{*}(f)=\int_{0}^{\infty}\mathrm{d}f\;\frac{\pi c^{2}}{4G}f^{2}S_{h}(f)\; ,
+
+A third way of describing the amplitude of a GW is through the energy carried by the waves.
+This has the advantage of having a clear physical significance.
+The energy density is most commonly used in sensitivity curves showing stochastic backgrounds of GWs (see section \ref{sec:stoch}).
+
+The energy in GWs is described by the Isaacson stress tensor \citep[section 35.15]{MTW}
+\begin{equation}
+T_{\mu\nu}=\frac{c^{2}}{32\pi G}\left<\partial_{\mu}\bar{h}_{\alpha\beta}\partial_{\nu}\bar{h}^{\alpha\beta}\right> \;,
+\end{equation}
+where the anle brackets denote averaging over several wavelengths or periods. The energy density $\rho c^{2}$ is given by the $T_{00}$ component of this tensor. Consequently
+\begin{eqnarray}
+\label{eq:specNRGdensity}
+\rho c^{2} &=& \frac{c^{2}}{16\pi G}\int_{-\infty}^{\infty}\mathrm{d}f\;\left(2\pi f\right)^{2}\tilde{h}(f)\tilde{h}^{*}(f) \\*
+ &=& \int_{0}^{\infty}\mathrm{d}f\;\frac{\pi c^{2}}{4G}f^{2}S_{h}(f)\; ,
 \end{eqnarray} 
-where the definitions (\ref{eq:ShforSources}) and (\ref{eq:strainsagain}) have been used.
-Equation (\ref{eq:specNRGdensity}) motivates the definition of the spectral energy density in GWs, $S_{\mathrm{E}}$, the energy per unit volume of space, per unit frequency, \cite{HellingsDowns}
-\begin{equation}\label{eq:spectralenergydensity} S_{\mathrm{E}}(f)=\frac{\pi c^{2}}{4G} f^{2}S_{h}(f) \; ,\end{equation}
-and a corresponding expression for the noise with $S_{n}(f)$.
-It is usual to define the dimensionless quantity $\Omega_{\mathrm{GW}}$ as the energy density per logarithmic frequency interval normalised to the critical density of the universe
-\begin{equation}\label{eq:omega} \Omega_{\mathrm{GW}}(f)=\frac{fS_{\mathrm{E}}(f)}{\rho_{\mathrm{c}}c^{2}}\; . \end{equation}
-The critical density of the universe is $\rho_{\mathrm{c}}=3H_{0}^{2}/8\pi G$, where $H_{0}=h_{100}\times 100\, \mathrm{km}\,\mathrm{s}^{-1}\,\mathrm{Mpc}^{-1}$ is the Hubble constant (this equation serves as a definition of $h_{100}$, which here has nothing to do with strain).
-The most common quantity related to energy density to be plotted on sensitivity curves is $\Omega_{\mathrm{GW}}h_{100}^{2}$ (figure \ref{fig:omega}).
+where the definition (\ref{eq:ShforSources}) have been used.
+Equation (\ref{eq:specNRGdensity}) gives the definition of the spectral energy density, the energy per unit volume of space, per unit frequency \citep{HellingsDowns}
+\begin{equation}\label{eq:spectralenergydensity}
+S_{\mathrm{E}}(f)=\frac{\pi c^{2}}{4G} f^{2}S_{h}(f) \; ;
+\end{equation}
+a corresponding expression for the noise can be formulated by replacing $S_h(F)$ with $S_{n}(f)$.
+
+Cosmological studies often work in terms of the dimensionless quantity $\Omega_{\mathrm{GW}}$, the energy density per logarithmic frequency interval normalised to the critical density of the universe:
+\begin{equation}\label{eq:omega}
+\Omega_{\mathrm{GW}}(f)= \frac{fS_{\mathrm{E}}(f)}{\rho_{\mathrm{c}}c^{2}} = \frac{\pi}{4G\rho_{c}}f^{3}S_{h}(f) = \frac{\pi}{4G\rho_{c}}f^{2}h_{c}(f)^{2}\; .
+\end{equation}
+The critical density of the universe is
+\begin{equation}
+\rho_{\mathrm{c}}=3H_{0}^{2}/8\pi G \;,
+\label{eq:crit-density}
+\end{equation}
+ where $H_{0}$ is the Hubble constant which is commonly parametrized as
+\begin{equation}
+H_0 =h_{100}\times 100~\mathrm{km\,s^{-1}\,Mpc^{-1}}.
+\end{equation}
+The reduced Hubble parameter $h_{100}$ has nothing to do with strain.
+The most common quantity related to energy density to be plotted on sensitivity curves is $\Omega_{\mathrm{GW}}h_{100}^{2}$ (figure \ref{fig:omega}) as this removes sensitivity to the measured value of the Hubble constant.
+
 This quantity has one aesthetic advantage over the others: it accounts for the fact that there is less energy in low frequency waves of the same amplitude and does not place the sensitivity curves of very low frequency detectors much higher than most ground based detectors.
@@
 \subsection{Relating the different descriptions}
-The dimensionless energy density in GWs, $\Omega_{\mathrm{GW}}$, the spectral energy density in GWs, $S_{\mathrm{E}}$, the one-sided power spectral density, $S_{h}$, and the characteristic strain are all related via
+The dimensionless energy density in GWs $\Omega_{\mathrm{GW}}$, spectral energy density $S_{\mathrm{E}}$, one-sided power spectral density $S_{h}$, characteristic strain $h_\mathrm{c}$ and frequency-domain strain $\tilde{h}(f)$ are related via
 \begin{equation}\label{eq:omega} 
-\Omega_{\mathrm{GW}}(f)=\frac{fS_{\mathrm{E}}(f)}{\rho_{\mathrm{c}}c^{2}}=\frac{\pi}{4G\rho_{\mathrm{c}}}f^{2}h_{c}(f)^{2}=\frac{\pi}{4G\rho_{\mathrm{c}}}f^{3}S_{h}(f)  \; ,
+H_0^2\Omega_{\mathrm{GW}}(f)= \frac{8 \pi G}{3 c^{2}} fS_{\mathrm{E}}(f) = \frac{2\pi^2}{3} f^3 S_h(f) = \frac{2\pi^2}{3} f^2 h_\mathrm{c}^2(f) = \frac{2\pi^2}{3} f^4 \left|\tilde{h}(f)\right|^2\; ,
 \end{equation}
-and corresponding expressions for the noise with $S_{n}(f)$ and $h_{n}(f)$.
+using (\ref{eq:strain-hc}), (\ref{eq:ShforSources}), (\ref{eq:spectralenergydensity}), (\ref{eq:omega}) and (\ref{eq:crit-density}).
+Corresponding expressions for the noise are obtained by substituting $S_{n}(f)$ for $S_h(f)$, $h_{n}(f)$ for $h_\mathrm{c}(f)$ and $\tilde{n}(f)$ for $\tilde{h}(f)$.

### 2014-07-03 Introduction.tex
Section 2

@@
 Since the GW signal and detector output are both real it follows that $\tilde{h}(-f)=\tilde{h}^{*}(f)$ and $\tilde{n}(-f)=\tilde{n}^{*}(f)$; therefore, $S_{n}(f)=S_{n}(-f)$.
-The fact that $S_{n}(f)$ is an even function means that Fourier integrals over all frequencies can instead be written as integrals over positive frequencies only, e.g., (\ref{eq:meansquare1}) and (\ref{eq:snrinnerprod})); it is for this reason that $S_{n}(f)$ is called the \emph{one-sided} power spectral density.\footnote{An alternative convention is to use the two-sided power spectral density $S^{(2)}_{n}(f) = S_{n}(f)/2$.}
+The fact that $S_{n}(f)$ is an even function means that Fourier integrals over all frequencies can instead be written as integrals over positive frequencies only, e.g., (\ref{eq:meansquare1}) and (\ref{eq:snrinnerprod}); it is for this reason that $S_{n}(f)$ is called the \emph{one-sided} power spectral density.\footnote{An alternative convention is to use the two-sided power spectral density $S^{(2)}_{n}(f) = S_{n}(f)/2$.}
 
@@
 Cosmological studies often work in terms of the dimensionless quantity $\Omega_{\mathrm{GW}}$, the energy density per logarithmic frequency interval normalised to the critical density of the universe:
-\begin{equation}\label{eq:omega}
+\begin{equation}
+\label{eq:omega}
 \Omega_{\mathrm{GW}}(f)= \frac{fS_{\mathrm{E}}(f)}{\rho_{\mathrm{c}}c^{2}} = \frac{\pi}{4G\rho_{c}}f^{3}S_{h}(f) = \frac{\pi}{4G\rho_{c}}f^{2}h_{c}(f)^{2}\; .

### 2014-07-03 Introduction.tex
Section 2

@@
 \label{eq:omega}
-\Omega_{\mathrm{GW}}(f)= \frac{fS_{\mathrm{E}}(f)}{\rho_{\mathrm{c}}c^{2}} = \frac{\pi}{4G\rho_{c}}f^{3}S_{h}(f) = \frac{\pi}{4G\rho_{c}}f^{2}h_{c}(f)^{2}\; .
+\Omega_{\mathrm{GW}}(f) = \frac{fS_{\mathrm{E}}(f)}{\rho_{\mathrm{c}}c^{2}} \; .
 \end{equation}
 The critical density of the universe is
-\begin{equation}
+\begin{equation}\label{eq:crit-density}
 \rho_{\mathrm{c}}=3H_{0}^{2}/8\pi G \;,
-\label{eq:crit-density}
 \end{equation}
@@
 
-This quantity has one aesthetic advantage over the others: it accounts for the fact that there is less energy in low frequency waves of the same amplitude and does not place the sensitivity curves of very low frequency detectors much higher than most ground based detectors.
+This quantity has one aesthetic advantage over the others: it automatically accounts for the fact that there is less energy in low frequency waves of the same amplitude.
+However, unlike characteristic strain, the area between the source and detector curves is no longer simply related to the SNR.
 
@@
 The dimensionless energy density in GWs $\Omega_{\mathrm{GW}}$, spectral energy density $S_{\mathrm{E}}$, one-sided power spectral density $S_{h}$, characteristic strain $h_\mathrm{c}$ and frequency-domain strain $\tilde{h}(f)$ are related via
-\begin{equation}\label{eq:omega} 
-H_0^2\Omega_{\mathrm{GW}}(f)= \frac{8 \pi G}{3 c^{2}} fS_{\mathrm{E}}(f) = \frac{2\pi^2}{3} f^3 S_h(f) = \frac{2\pi^2}{3} f^2 h_\mathrm{c}^2(f) = \frac{2\pi^2}{3} f^4 \left|\tilde{h}(f)\right|^2\; ,
+\begin{equation}
+H_0^2\Omega_{\mathrm{GW}}(f)= \frac{8 \pi G}{3 c^{2}} fS_{\mathrm{E}}(f) = \frac{2\pi^2}{3} f^3 S_h(f) = \frac{2\pi^2}{3} f^2 h_\mathrm{c}^2(f) = \frac{8\pi^2}{3} f^4 \left|\tilde{h}(f)\right|^2\; ,
 \end{equation}

### 2014-07-04 Detectors.tex
Attempt at adding stuff to astrophysical sources

@@
 \begin{table}[h!]
-\caption{\label{table:t} For the aVIRGO sensitivity curve an interpolation to the data published on \url{https://wwwcascina.virgo.infn.it/advirgo/} (2013) was used.
+\caption{\label{table:t} For the AdV sensitivity curve an interpolation to the data published on \url{https://wwwcascina.virgo.infn.it/advirgo/} (2013) was used.
 For KAGRA (Kamioka Gravitational Wave Detector) an interpolation to the data for version D of the detector published on \url{http://gwcenter.icrr.u-tokyo.ac.jp/en/researcher/parameter} (2013) was used.
@@
   iLIGO	$^{c}$	&	US		& $4\,\textrm{km}$ 	& 2004-2010 	   & 1$^{\textrm{st}}$\\
-  iVIRGO $^{d}$	& 	Italy		& $3\,\textrm{km}$ 	& 2007-2010 	   & 1$^{\textrm{st}}$\\
-  aLIGO $^{e}$	&	US		& $4\,\textrm{km}$ 	& \emph{est.} 2016 & 2$^{\textrm{nd}}$\\
+  Virgo $^{d}$	& 	Italy		& $3\,\textrm{km}$ 	& 2007-2010 	   & 1$^{\textrm{st}}$\\
+  aLIGO $^{e}$	&	US		& $4\,\textrm{km}$ 	& \emph{est.} 2015 & 2$^{\textrm{nd}}$\\
   KAGRA	$^{f}$	&	Japan		& $3\,\textrm{km}$ 	& \emph{est.} 2018 & 2$^{\textrm{nd}}$\\
-  aVIRGO $^{g}$	&	Italy	 	& $3\,\textrm{km}$ 	& \emph{est.} 2017 & 2$^{\textrm{nd}}$\\
-  ET (Einstein Telescope)$^{h}$ &Italy		& $10\,\textrm{km}$ 	& \emph{est.} 2025 & 3$^{\textrm{rd}}$\\
+  AdV $^{g}$	&	Italy	 	& $3\,\textrm{km}$ 	& \emph{est.} 2016 & 2$^{\textrm{nd}}$\\
+  ET (Einstein Telescope)$^{h}$ & Italy		& $10\,\textrm{km}$ 	& \emph{est.} 2025 & 3$^{\textrm{rd}}$\\
 \br
@@
 
-\subsection{Pulsar timing arrays (PTAs)}\label{sec:PTAgeneralproperties}
+\subsection{Pulsar timing arrays}\label{sec:PTAgeneralproperties}
 PTAs can be thought of as naturally occurring interferometers with galactic scale arm lengths, hence they are sensitive to much lower frequencies than the detectors considered so far.
@@
 
-\begin{figure}[h!]
+\begin{figure}
  \centering

### 2014-07-04 Sources.tex
Attempt at adding stuff to astrophysical sources

@@
 \section{Astrophysical sources}\label{sec:sources}
+
 All the sources described here are represented by shaded boxes in \ref{app:a}.
@@
 Descriptions of each source are given in the following sections.
-Sources with short durations and sources which evolve in time over much longer timescales than our observations are drawn with flat topped boxes.
-Inspiralling binaries, or stochastic backgrounds of binaries, are drawn with a sloping top proportional to $f^{-2/3}$, which is result derived in section \ref{sec:stoch}.
+Sources with short durations (i.e., burst sources) and sources which evolve in time over much longer timescales than our observations are drawn with flat-topped boxes \cplb{for $h_\mathrm{c}(f)$}.
+Inspiralling binaries, or stochastic backgrounds of binaries, are drawn with a sloping top proportional to $f^{-2/3}$ \cplb{for $h_\mathrm{c}(f)$}, which is result derived in section \ref{sec:stoch}.
 The width of the box gives the range of frequencies sources of a given type can have a while remaining at a detectable amplitude.
@@
 \subsubsection{Neutron star binaries}
+
 The inspiral and merger of a pair of neutron stars is the primary target for ground based detectors.
-The expected event rate for this type of source is uncertain, but estimates centre around $\rho_{\textrm{NS-NS}}=1.3\times 10^{-4}\;\textrm{Mpc}^{-3}\textrm{yr}^{-1}$, \cite{CBC}.
-Plotted in \ref{app:a} is a box with an amplitude such that it produces a ratio $h_{c}/h_{n}=16$ for advanced LIGO at peak sensitivity and a width between $(3-300)\,\textrm{Hz}$.
+The expected event rate for this type of source is uncertain, but estimates centre around $\rho_{\mathrm{NS-NS}}=1.3\times 10^{-4}~\mathrm{Mpc^{-3}\,yr^{-1}}$ \citep{CBC}.
+Plotted in \ref{app:a} is a box with an amplitude such that it produces a ratio $h_\mathrm{c}/h_{n}=16$ for Advanced LIGO at peak sensitivity and a width between $(3$--$300)~\mathrm{Hz}$.
 
 \subsubsection{Supernovae}
-Simulations of core collapse supernova events show that GWs between $(10^{2}-10^{3})\,\textrm{Hz}$ can be produced.
-The GW signal undergoes ${\cal{O}}(1)$ oscillation and hence is burst-like.
-\cite{2002A&A...393..523D} calculate the average maximum amplitude of GWs for a supernova at distance $r$ as
-\begin{equation} h=8.9\times 10^{-21}\left( \frac{10 \,\textrm{kpc}}{r} \right) \; .\end{equation}
-The event rate for supernova is approximately $\rho_{\textrm{SN}}=5\times10^{-4}\,\textrm{Mpc}^{-3}\textrm{yr}^{-1}$.
-The boxes plotted in \ref{app:a} correspond to a distance $r=3 \,\textrm{Mpc}$ with the frequency range quoted above.
-The LIGO and VIRGO detectors have already placed bounds on the event rate for these sources, \cite{Bursts}.
+
+Simulations of core-collapse supernovae show that GWs between $(10^{2}$--$10^{3})~\mathrm{Hz}$ can be produced \citep{Kotake2006}.
+The GW signal undergoes ${\mathcal{O}}(1)$ oscillation and is hence burst-like.
+\citet{2002A&A...393..523D} calculate the average maximum amplitude of GWs for a supernova at distance $r$ as
+\begin{equation}
+h=8.9\times 10^{-21}\left( \frac{10~\mathrm{kpc}}{r} \right) \; .
+\end{equation}
+The event rate for supernovae is approximately $\rho_{\mathrm{SN}} = 5\times10^{-4}~\mathrm{Mpc^{-3}\,yr^{-1}}$.
+The boxes plotted in \ref{app:a} correspond to a distance $r=3~\mathrm{Mpc}$ with the frequency range quoted above.
+The LIGO and Virgo detectors have already placed bounds on the event rate for these sources \citep{Bursts}.
 
@@
 
-{\cjm{If someone is feeling keen they can write a very short section on these, maybe using this reference \cite{Einstein@Home}.
-Can whoever writes this section say that in figures \ref{fig:hc}, \ref{fig:S} and \ref{fig:omega} the rotating neutron stars share the same box as supernova}.
+Rotating neutron stars are a source of continuous GWs if they possess some degree of axial asymmetry \citep{Abbott2007, Prix2009, Einstein@Home}.
+The signals are near monochromatic with a frequency twice the rotation frequency of the neutron star, and are a potential source for ground-based detectors.
+The amplitude of the GWs depends upon the deformation of the neutron star, which itself depends upon the neutron star equation of state.
+This is currently uncertain \citep{Lattimer2012}.
+Several known pulsars could be sources for the advanced detectors and upper limits from the initial detectors help to constrain the deformations \citep{Aasi2014a}.
+The boxes plotted in \ref{app:a} is the same as for supernovae \cplb{for reasons I do not entirely understand}.
 
 \subsection{Sources for space based detectors}
-For a review of the GW sources for space based missions see, for example, \cite{Amaro-Seoane-et-al}, \cite{Gairetal} or the \cite{eLISAyellowbook}.
+
+For a review of the GW sources for space based missions see, for example, \citet{Amaro-Seoane-et-al}, \citet{Gairetal} or \citet{eLISAyellowbook}.
 
 \subsubsection{Massive black hole binaries}
-Space based detectors will be sensitive to equal mass mergers in the range $(10^{4}-10^{7})\,\Msun$.
-Predictions of the event rate for these mergers range from ${\cal{O}}(3-300)\,\textrm{yr}^{-1}$ for LISA with SNRs of up to 10000.
-The uncertainty in the rate reflects our uncertainty in the growth mechanisms of the supermassive black hole population.
-Plotted in \ref{app:a} are boxes with a ratio $h_{c}/h_{n}=100$ for eLISA at its peak sensitivity.
-The range of frequencies plotted is $(3\times 10^{-4}-3\times 10^{-1})\,\textrm{Hz}$, this corresponds to circular binaries in the mass range quoted above.
+
+Space-based detectors shall be sensitive to equal-mass mergers in the range $(10^{4}$--$10^{7})\,\Msun$.
+Predictions of the event rate for these mergers range from ${\mathcal{O}}(3$--$300)~\mathrm{yr}^{-1}$ for LISA with SNRs of up to $10^4$ \cplb{[citation needed]}.
+The large range in the rate reflects our uncertainty in the growth mechanisms of the supermassive black hole population \citep{Volonteri2010}.
+Plotted in \ref{app:a} are boxes with a ratio $h_\mathrm{c}/h_{n}=100$ for eLISA at its peak sensitivity.
+The range of frequencies plotted is $(3\times 10^{-4}$--$3\times 10^{-1})~\mathrm{Hz}$, this corresponds to circular binaries in the mass range quoted above.
 
 \subsubsection{Galactic white dwarf binaries} \label{sec:GB}
-These are the most numerous GW sources, they are also the only source where several guaranteed sources have already been indentified by electromagnetic observations, these are known as the verification binaries, \cite{2006CQGra..23S.809S}.
+
+For space-based detectors, these are the most numerous GW sources; they are also the only guaranteed source since several detectable systems (known as verification binaries) have already been identified by electromagnetic observations \citep{2006CQGra..23S.809S}.
 
@@
 The unresolvable binaries overlap to form a stochastic background as discussed in section \ref{sec:stoch}.
-Plotted in \ref{app:a} is the estimate of this background due to \cite{Nelemans} where an observation time of one year has been assumed
-\begin{equation} h_{c}(f)= 5\times 10^{-21} \left(\frac{f}{10^{-3}\,\textrm{Hz}}\right)^{-2/3} \; . \end{equation}
-Estimates for the event rate of resolvable binaries centre around ${\cal{O}}(10^{3})$ events for eLISA.
-The boxes plotted in \ref{app:a} have a ratio $h_{c}/h_{n}=50$ for eLISA at its peak sensitivity.
-The frequency range of the box is $\left(3\times10^{-4}\right.$--$\left.10^{-2}\right)~\mathrm{Hz}$, based on visual inspection of Monte Carlo population simulation results presented in \cite{Amaro-Seoane-et-al}.
+\cplb{The distinction between resolvable and unresolvable is detector specific; here we choose LISA.
+This boundary will not be too different for eLISA but would move substantially for either of the decihertz detectors.} Plotted in \ref{app:a} is the estimate of this background due to \citet{Nelemans} where an observation time of one year has been assumed,
+\begin{equation}
+h_\mathrm{c}(f)= 5\times 10^{-21} \left(\frac{f}{10^{-3}~\mathrm{Hz}}\right)^{-2/3} \; .
+ \end{equation}
+Estimates for the event rate of resolvable binaries centre around ${\mathcal{O}}(10^{3})$ events for eLISA.
+The boxes plotted in \ref{app:a} have a ratio $h_\mathrm{c}/h_{n}=50$ for eLISA at its peak sensitivity.
+The frequency range of the box is $\left(3\times10^{-4}\right.$--$\left.10^{-2}\right)~\mathrm{Hz}$, estimated from Monte Carlo population simulation results presented in \citet{Amaro-Seoane-et-al}.
 
-\subsubsection{Extreme mass ratio inspirals}
-EMRI events occur when a compact stellar mass object inspirals into a supermassive black hole.
-There is extreme uncertainty in the event rate for EMRIs due to the poorly constrained astrophysics in galactic centres; the best guess estimate is ${\mathcal{O}}(10)$ events per year with eLISA with SNR $\ge 20$.
+\subsubsection{Extreme mass-ratio inspirals}
+
+EMRIs occur when a compact stellar mass object inspirals into a supermassive black hole.
+There is extreme uncertainty in the event rate for EMRIs due to the poorly constrained astrophysics in galactic centres \citep[e.g.,][]{Merritt2011}; the best guess estimate is ${\mathcal{O}}(10)$ events per year with eLISA with SNR $\ge 20$ \cplb{[citation needed]}.
 The boxes plotted in \ref{app:a} have a characteristic strain of $h_{c}=3\times 10^{-20}$ at $10^{-2}~\mathrm{Hz}$, which corresponds to a $10\Msun$ BH inspiralling into a $10^{6}\Msun$ black hole at a luminosity distance of $1~\mathrm{Gpc}$.
 The frequency width of the box is somewhat unknown; EMRI events can occur into a black hole of any mass, and hence EMRIs can occur at any frequency.
-However there is substantial uncertainty in the black hole mass function, so the boxes in \ref{app:a} are drawn with a width comparable to that of the LISA sensitivity curve.
+However, there is substantial uncertainty in the black hole mass function \cplb{[citation needed]}, so the boxes in \ref{app:a} are drawn with a width comparable to that of the LISA sensitivity curve.
 
@@
 \subsubsection{Supermassive black hole binaries}
-The main target for PTAs is a stochastic background of GWs produced by a population of supermassive black hole binaries at cosmological distances.
-Supermassive black holes are known to lie at the centres of most galaxies and the black hole mergers are associated with the mergers of the host galaxies.
-The current best published limit for the amplitude of the stochastic background is $h_{c} = 6\times 10^{-15}$ at a frequency of $f_{0}=1~\mathrm{yr}^{-1}$, \cite{Haasteren}.
-There is strong theoretical evidence that the actual background lies close to the current limit, \cite{Sesana-2012}.
+
+The main target for PTAs is a stochastic background of GWs produced by a population of supermassive black hole binaries at cosmological distances \citep{SesanaVecchioColancino}.
+Supermassive black holes are known to lie at the centres of most galaxies and the black hole mergers are associated with the mergers of the host galaxies \citep{Volonteri2003,Ferrarese2005}.
+The current best published limit for the amplitude of the stochastic background is $h_\mathrm{c} = 6\times 10^{-15}$ at a frequency of $f_{0}=1~\mathrm{yr}^{-1}$ \citep{Haasteren}.
+There is strong theoretical evidence that the actual background lies close to the current limit \citep{Sesana-2012}.
 
@@
 It is currently unclear whether PTAs will detect an individual binary or a stochastic background first.
-Plotted in \ref{app:a} for the unresolvable background is a third of the current limit with a cut off frequency of $f=1~\mathrm{yr}^{-1}$ which is suggested by Monte Carlo population studies, \cite{SesanaVecchioColancino}.
+Plotted in \ref{app:a} for the unresolvable background is a third of the current limit with a cut off frequency of $f=1~\mathrm{yr}^{-1}$ which is suggested by Monte Carlo population studies \citep{SesanaVecchioColancino}.
 For the resolvable sources the amplitude of the current limit is plotted between $\left(3\times 10^{-9}\right.$--$\left.3\times 10^{-7}\right)~\mathrm{Hz}$.
@@
 In addition to the sources above, early Universe processes, such as inflation \citep{Grishchuk2005} or a first-order phase transition \citep{Binetruy2012}, could have created GWs.
-More speculatively, it has been hypothesised that cosmic strings could also be a potential source \citep{Damour2005,Binetruy2012}.
+More speculatively, it has been hypothesised that cosmic strings could also be a potential source \citep{Damour2005,Binetruy2012,Aasi2014}.
 These relic GWs allow us to explore energy scales far beyond those accessible by other means, providing insight into new and exotic physics.

### 2014-07-04 noise_curve.tex
Section 2
Ending

@@
 \usepackage[margin=1cm]{caption}
+\usepackage{braket}
 \usepackage{fullpage}
@@
 \newcommand{\degree}{\ensuremath{^\circ}}
+\newcommand{\innerprod}[2]{\ensuremath{\left({#1}\middle|{#2}\right)}}
 \newcommand{\cplb}[1]{{\color{red} #1}}
@@
 \section{Concluding remarks}\label{sec:discussion}
-When quantifying the sensitivity of a GW detector and the loudness of a GW source there are three commonly used quantities: the characteristic strain, the power spectral density, and the spectral energy density.
-Here we have carefully defined these quantities and derived the relationships between them.
-In addition example sensitivity curve plots using each of these quantities have been presented for a wide range of ground based detectors, space based detectors and pulsar timing arrays.
-Also plotted are predicted amplitudes for a wide range of sources, these predictions are based on astrophysical estimates of the event rates for a range of processes and are subject to varying degrees of uncertainty.
-There also exists the exciting possibility of sources nobody has yet considered.
-Interactive versions of these plots with user specified detectors and sources are made available on-line at \url{http://www.ast.cam.ac.uk/~rhc26/sources/}.
-Trying to summarise an entire field of astronomy on one plot is an impossible task; however we hope that the figures and analysis presented here provide useful insight.
+
+When quantifying the sensitivity of a GW detector and the loudness of a GW source, there are three commonly used quantities: the characteristic strain, the power spectral density, and the spectral energy density.
+We have carefully defined these quantities and derived the relationships between them.
+The characteristic strain (section \ref{sec:character-strain}) is most directly related to the SNR, the power spectral density is mostly closely related to the mean square amplitude in the detector, and the energy density has a clear physical interpretation.
+We have produced example plots using each of these quantities for a wide range of detectors and sources.
+The predicted source amplitudes are based on astrophysical estimates of the event rates and are subject to varying degrees of uncertainty.\footnote{There is also the exciting possibility of sources yet to be considered.} Interactive versions of these plots, with user-specified detectors and sources, are available on-line at \url{http://www.ast.cam.ac.uk/~rhc26/sources/}.
+Trying to summarise an entire field of astronomy on one plot is an impossible task; however, we hope that the figures and analysis presented here provide useful insight.
 

### 2014-07-04 Detectors.tex
Detectors to PTAs

@@
 \subsection{Operating principle of an interferometric detector}\label{sec:principles}
+
 All of the man-made detectors discussed in this section work on the principle of interferometry.
 Such detectors work by taking a beam of monochromatic light and splitting it into two beams travelling at some angle to each other.
-Each beam is passed in to a Fabry-P\'{e}rot cavity where it undergoes a number of round trips before being recombined to form an interference pattern.
+Each beam is passed in to an optical cavity where it undergoes a number of round trips before being recombined to form an interference pattern.
 The ends of the cavity are, in the ideal case, freely floating test masses which move in response to a passing GW, this effect is measured by observing the changing interference pattern.
+% LIGO isn't Fabry-P\'{e}rot
 
-The response of a detector to an incident plane fronted GW depends upon the relative orientations of the detector and the incoming wave.
-Choose the origin of our coordinate system to be the beam-splitter of the interferometer and let $l_{1}^{i}$ and $l_{2}^{i}$ be unit 3-vectors pointing along the two arms.
-In the absence of noise the output of the detector is the difference in strain between the two arms \citep{MTW}
-\begin{equation}\label{eq:hoftxx} h(t)=\frac{1}{2}h_{ij}\left( l_{1}^{i}l_{1}^{j}-l_{2}^{i}l_{2}^{j} \right)\; , \end{equation}
+The response of a detector to an incident plane-fronted GW depends upon the relative orientations of the detector and the incoming wave.
+Let us choose the origin of our coordinate system to be the beam-splitter of the interferometer, and $l_{1}^{i}$ and $l_{2}^{i}$ to be unit 3-vectors pointing along the two arms.
+In the absence of noise the output of the detector is the difference in strain between the two arms \citep{Thorne1987} \cplb{I think you mean this rather than MTW}
+\begin{equation}\label{eq:hoftxx}
+h(t)=\frac{1}{2}h_{ij}\left( l_{1}^{i}l_{1}^{j}-l_{2}^{i}l_{2}^{j} \right)\; ,
+\end{equation}
 where $h_{ij}$ are the spatial components of the GW metric perturbation.
-Let $n^{i}$ be the unit 3-vector pointing towards the source of the GWs, with spherical polar angles $(\theta,\phi)$ relative to some axes fixed to the detector, also let $p^{i}$ and $q^{i}$ be unit vectors orthogonal to $n^{i}$.
-There remains a freedom in the coordinates described, a rotation of $p^{i}$ and $q^{i}$ through an angle $\psi$ about $n^{i}$.
-Define the basis tensors,
+Let $\hat{r}^{i}$ be the unit 3-vector pointing towards the source of the GWs, with spherical polar angles $(\theta,\phi)$ relative to some axes fixed to the detector, and let $p^{i}$ and $q^{i}$ be unit vectors orthogonal to $\hat{r}^{i}$.
+We can now define the basis tensors
 \begin{eqnarray}
@@
 \end{eqnarray}
-Consider a single frequency component of the incident GW, the strain induced by the incident GW may be written as
-\begin{equation}\label{eq:hijxx} h_{ij}=A_{+}H^{+}_{ij}\cos\left(2\pi ft\right)+A_{\times}H^{\times}_{ij}\cos\left(2\pi ft+\Delta \phi\right) \; ,\end{equation}
+There remains a freedom in the coordinates described, a rotation of $p^{i}$ and $q^{i}$ through an angle $\psi$ about $\hat{r}^{i}$ known as the polarization angle.
+For a single frequency component, the strain induced by GW may be written as
+\begin{equation}\label{eq:hijxx}
+h_{ij}=A_{+}H^{+}_{ij}\cos\left(2\pi ft\right)+A_{\times}H^{\times}_{ij}\cos\left(2\pi ft+\Delta \phi\right) \; ,
+\end{equation}
 where $A_{+}$ and $A_{\times}$ are the amplitudes of the two polarisation states. Combining (\ref{eq:hoftxx}) and (\ref{eq:hijxx}) allows the detector output to be written as
-\begin{equation} h(t)=F^{+}(\theta,\phi,\psi)A_{+}\cos\left(2\pi ft\right)+F^{\times}(\theta,\phi,\psi)A_{\times}\cos\left(2\pi f t + \Delta\phi \right)\; , \end{equation}
+\begin{equation}
+h(t)=F^{+}(\theta,\phi,\psi)A_{+}\cos\left(2\pi ft\right)+F^{\times}(\theta,\phi,\psi)A_{\times}\cos\left(2\pi f t + \Delta\phi \right)\; ,
+\end{equation}
 where the response functions inherit their angular dependence from the choice of coordinates,
 \begin{eqnarray}\label{eq:responsefuncs}
-F^{+}(\theta,\phi,\psi)&=\frac{1}{2}H^{+}_{ij}\left(l_{1}^{i}l_{1}^{j}-l_{2}^{i}l_{2}^{j}\right) \nonumber \\
+F^{+}(\theta,\phi,\psi)&=\frac{1}{2}H^{+}_{ij}\left(l_{1}^{i}l_{1}^{j}-l_{2}^{i}l_{2}^{j}\right) \; , \\
 F^{\times}(\theta,\phi,\psi)&=\frac{1}{2}H^{\times}_{ij}\left(l_{1}^{i}l_{1}^{j}-l_{2}^{i}l_{2}^{j}\right) \; . 
@@
 
-\begin{figure}[h!]
+\begin{figure}
  \centering
  \includegraphics[trim=0cm 0cm 0cm 0cm, width=0.99\textwidth]{LIGO_detectorframe.pdf}
- \caption{The angular response function of an interferometric detector shown both as a surface plot and in an Aitoff-Hammer projection.
-The two detector arms lie in the $xy$ plane either side of one of the zeros in the response.}
+ \caption{The angular response function of an interferometric detector shown both as a surface plot and in an Aitoff--Hammer projection \cplb{[citation needed]}.
+The two detector arms lie in the $x$--$y$ plane either side of one of the zeros in the response.
+\cplb{Can you label the colour bar/explain what the surface indicates?}}
  \label{fig:LIGO}
@@
 
-Throughout this paper detector sensitivity refers to the polarisation and sky averaged sensitivity, $F$, where
-\begin{equation}\label{eq:skyav} F^{2}=\int_{0}^{2\pi}\frac{\textrm{d}\psi}{2\pi}\; \int_{0}^{\pi}\int_{0}^{2\pi}\frac{\sin\theta\,\textrm{d}\theta\textrm{d}\phi}{4\pi}\;\left(\frac{F^{+}\left(\theta,\phi,\psi\right)^{2}+F^{\times}\left(\theta,\phi,\psi\right)^{2}}{2}\right)\; .\end{equation}
-For a single $90\degree$ interferometer, such as LIGO (Laser Interferometer GW Observatory), the sky and polarisation averaged response is $F=\sqrt{1/5}\approx 0.447$.
+Throughout this paper detector sensitivity refers to the polarisation and sky averaged sensitivity $F$, where
+\begin{equation}\label{eq:skyav}
+F^{2}=\int_{0}^{2\pi}\frac{\mathrm{d}\psi}{2\pi}\; \int_{0}^{2\pi} \frac{\mathrm{d}\phi}{2\pi}\; \int_{0}^{\pi}\frac{\sin\theta\,\mathrm{d}\theta}{2}\;\left[\frac{F^{+}\left(\theta,\phi,\psi\right)^{2}+F^{\times}\left(\theta,\phi,\psi\right)^{2}}{2}\right]\; .
+\end{equation}
+For a single $90\degree$-interferometer, such as LIGO, the sky and polarisation averaged response is $F=\sqrt{1/5}\approx 0.447$.
 
 A detector may consist of several interferometers.
-Let $F_{a}$ be the averaged response of the $a^{\textrm{th}}$ interferometer.
-The average response of the detector network is obtained by adding in quadrature,
-\begin{equation} F_{\textrm{Total}}^{2}=\frac{1}{\left(\textrm{Number of interferometers}\right)}\sum_{x=1}^{N}F_{x}^{2} \; .\end{equation}
+Let $F_{a}$ be the averaged response of the $a$-th interferometer, the average response of a network of $k$ detectors is obtained by adding in quadrature,
+\begin{equation} F_{\mathrm{Total}}^{2}=\frac{1}{k)}\sum_{a\,=\,1}^{k}F_{a}^{2} \; .\end{equation}
 
 The averaging in (\ref{eq:skyav}) assumes a uniform distribution of polarisation angles $\psi$.
-This will be the case for a stochastic background, however for a non-inspiralling circular binary the polarisation will be a function of the two spherical polar angles ($\iota,\xi$) specifying the orientation of the binary's orbital angular momentum.
-Here $\iota$ is the polar angle between the orbital angular momentum and the line joining the source to the detector while $\xi$ is the azimuthal angle around the same line.
-In this case the correct average to characterise the detector sensitivity would be an an average over all four angles ($\theta,\phi,\iota,\xi$).
-If the binary is inspiralling then the polarisation will depend on still more parameters which need to be averaged over.
-These more complicated averages all have the property that they depend on both the detector and the source, hence they are unhelpful for our present purpose.
-In addition the different averages don't work out to be so different from each other; \cite{1993PhRvD..47.2198F} calculated the sensitivity for a detector with the LIGO geometry averaged over the four angles ($\theta,\phi,\iota,\xi$) as $\sqrt{4/25}=0.4$ times peak sensitivity, which should be compared with the value $\sqrt{1/5}\approx 0.447$ above.
-For the remainder of this paper the 3 angle average defined in (\ref{eq:skyav}) will be used.
+This is the case for a stochastic background; however, for a non-inspiralling circular binary, the polarisation is a function of the two spherical polar angles ($\iota,\xi$) specifying the orientation of the binary's orbital angular momentum.
+Here, $\iota$ is the inclination angle, the polar angle between the orbital angular momentum and the line joining the source to the detector \cplb{($-\hat{r}$)} and $\xi$ is the azimuthal angle around the same line.
+In this case, to characterise the detector sensitivity we average over all four angles ($\theta,\phi,\iota,\xi$).
+If the binary is inspiralling, then the polarisation depends on still more parameters which need to be averaged over.
+These more complicated averages all have the property that they depend on both the detector and the source, hence they are unhelpful for our present purpose \cplb{of \ldots}.
+Additionally, the different averages do not work out to be so different from each other: \citet{1993PhRvD..47.2198F} calculated the sensitivity for a detector with the LIGO geometry averaged over the four angles ($\theta,\phi,\iota,\xi$) as $\sqrt{4/25}=0.4$ times peak sensitivity, which should be compared with the value $\sqrt{1/5}\approx 0.447$ above.
+For the remainder of this paper the three-angle average defined in (\ref{eq:skyav}) is used.
 
-\subsection{Ground based detectors}\label{sec:ground}
-On Earth approximations to the free floating test masses described above are achieved by suspending the mass from a pendulum system with natural frequency much greater than that of the GW.
-The interferometric detectors listed in Table \ref{table:t} are all sensitive to GWs in the frequency range ${\cal{O}}(10-10^{3})\,\textrm{Hz}$.
-They all have multiple narrow lines in their sensitivity curves, which arise from noise sources in the instrument including resonances in the suspension system and electronic noise at multiples of $60\,\textrm{Hz}$: these have been removed in the \ref{app:a} figures for clarity.
-The detectors fall broadly into three categories: 1\textsuperscript{st} generation detectors, which have already operated; 2\textsuperscript{nd} generation detectors currently under construction; and 3\textsuperscript{rd} generation detectors at the planning stage.
+\subsection{Ground-based detectors}\label{sec:ground}
 
-\begin{table}[h!]
-\caption{\label{table:t} For the AdV sensitivity curve an interpolation to the data published on \url{https://wwwcascina.virgo.infn.it/advirgo/} (2013) was used.
-For KAGRA (Kamioka Gravitational Wave Detector) an interpolation to the data for version D of the detector published on \url{http://gwcenter.icrr.u-tokyo.ac.jp/en/researcher/parameter} (2013) was used.
-For the remaining detectors analytic fits to the sensitivity curves due to \cite{Sathyaprakash} were used.}
+Ground-based detectors are the most numerous.
+A collection of interferometric detectors are listed in table \ref{table:t}, these are sensitive to GWs in the frequency range ${\mathcal{O}}(10$--$10^{3})~\mathrm{Hz}$.
+They all simulate free-floating test masses by suspending a mass from a pendulum system with natural frequency much greater than that of the GW.
+Their sensitivity curves include narrow lines, which arise from noise sources in the instrument including resonances in the suspension system and electrical noise at multiples of $60~\mathrm{Hz}$: these have been removed in the \ref{app:a} figures for clarity.
+The detectors fall broadly into three categories: first-generation detectors, which have already operated; second-generation detectors currently under construction, and third-generation detectors at the planning stage.
+
+The most notable ground-based detectors are LIGO and Virgo \citep{Abramovici1992,Acernese2008a}.
+These work in collaboration, supported by GEO600 \citep{Willke2002}.
+LIGO and Virgo have completed science runs as first-generation detectors, now referred to as Initial LIGO (iLIGO) and Initial Virgo (iVirgo).
+They are now being upgraded to second-generation detectors, advanced LIGO (aLIGO) and advanced Virgo (AdV).
+\cplb{I think AdV is the official acronym, although it does break the pattern.} LIGO has two observatories, one at Hanford, Washington, which has two detectors, and another at Livingston, Louisiana.
+There is an agreement to move one of the upgraded Hanford detector systems to a location in India \citep{LIGO-India,Unnikrishnan2013}.
+The advanced detectors should start operation in the next couple of years, with LIGO-India following further in the future.
+
+TAMA300 is a Japanese first-generation detector.
+Its successor is the Kamioka Gravitational Wave Detector (KAGRA), formerly the Large-scale Cryogenic Gravitational wave Telescope (LCGT), which is located underground in the Kamioka mine.
+It employs more sophisticated noise-reduction techniques than LIGO or Virgo, such as cryogenic cooling \citep{2012CQGra..29l4007S}.
+
+The Einstein Telescope (ET) is an ambitious proposal to construct an underground third-generation detector \citep{Punturo2010,Sathyaprakash2012}.
+Its location would provide shielding from seismic noise, allowing it to observe frequencies of $10$--$10^4~\mathrm{Hz}$.
+
+\begin{table}
+\caption{\label{table:t} Summary of ground-based laser interferometers.}
 \begin{indented}
@@
 \mr
-  GEO600 $^{a}$	&	Germany 	& $600\,\textrm{m}$ 	& 2001-present 	   & 1$^{\textrm{st}}$\\
-  TAMA300 $^{b}$& 	Japan		& $300\,\textrm{m}$ 	& 1995-present     & 1$^{\textrm{st}}$\\
-  iLIGO	$^{c}$	&	US		& $4\,\textrm{km}$ 	& 2004-2010 	   & 1$^{\textrm{st}}$\\
-  Virgo $^{d}$	& 	Italy		& $3\,\textrm{km}$ 	& 2007-2010 	   & 1$^{\textrm{st}}$\\
-  aLIGO $^{e}$	&	US		& $4\,\textrm{km}$ 	& \emph{est.} 2015 & 2$^{\textrm{nd}}$\\
-  KAGRA	$^{f}$	&	Japan		& $3\,\textrm{km}$ 	& \emph{est.} 2018 & 2$^{\textrm{nd}}$\\
-  AdV $^{g}$	&	Italy	 	& $3\,\textrm{km}$ 	& \emph{est.} 2016 & 2$^{\textrm{nd}}$\\
-  ET (Einstein Telescope)$^{h}$ & Italy		& $10\,\textrm{km}$ 	& \emph{est.} 2025 & 3$^{\textrm{rd}}$\\
+  GEO600$^{a}$	&	Germany 	& $600~\mathrm{m}$ 	& 2001--present	   & First \\
+  TAMA300$^{b}$ & 	Japan		& $300~\mathrm{m}$ 	& 1995--present    & First \\
+  iLIGO$^{c}$	&	USA		& $4~\mathrm{km}$ 	& 2004--2010 	   & First \\
+  Virgo$^{d}$	& 	Italy		& $3~\mathrm{km}$ 	& 2007--2010 	   & First \\
+  aLIGO$^{e}$	&	USA		& $4~\mathrm{km}$ 	& \emph{est.} 2015 & Second \\
+  AdV$^{f}$	&	Italy	 	& $3~\mathrm{km}$ 	& \emph{est.} 2016 & Second \\
+  KAGRA$^{g}$	&	Japan		& $3~\mathrm{km}$ 	& \emph{est.} 2018 & Second \\
+  ET$^{h}$      & 	---		& $10~\mathrm{km}$ 	& \emph{est.} 2025 & Third \\
 \br
-\multicolumn{5}{l}{$^{a}$\cite{2008CQGra..25k4043G}, $^{b}$\cite{2002CQGra..19.1409A}, }\\
-\multicolumn{5}{l}{$^{c}$\cite{2009RPPh...72g6901A}, $^{d}$\cite{accadia_virgo:_2012}, $^{e}$\cite{2010CQGra..27h4006H}, }\\
-\multicolumn{5}{l}{$^{f}$\cite{2012CQGra..29l4007S}, $^{g}$\cite{Acernese2009}, $^{h}$\cite{2011CQGra..28i4013H}.}\\
+\multicolumn{5}{l}{$^{a}$\citet{2008CQGra..25k4043G}, $^{b}$\citet{2002CQGra..19.1409A}, $^{c}$\citet{2009RPPh...72g6901A}, $^{d}$\citet{accadia_virgo:2012}, }\\
+\multicolumn{5}{l}{$^{e}$\citet{2010CQGra..27h4006H}, $^{f}$\citet{Acernese2009}, $^{g}$\citet{2012CQGra..29l4007S}, $^{h}$\citet{2011CQGra..28i4013H}.}\\
 \br
@@
 
-\subsection{Space based detectors}\label{sec:space}
-Space based detectors work on similar principles to ground based detectors but with the test masses residing inside of independent, widely separated satellites.
-Space based detectors are sensitive to lower frequency GWs than their ground based counterparts, this is partly because space based detectors can have much longer arms and partly because they are unaffected by seismic noise which limits the low frequency performance of ground based detectors.
+In \ref{app:a}, we use an interpolation to the data published on \url{https://wwwcascina.virgo.infn.it/advirgo/} (2013) for the AdV sensitivity curve, an interpolation to the data for version D of the detector published on \url{http://gwcenter.icrr.u-tokyo.ac.jp/en/researcher/parameter} (2013) for KAGRA and analytic fits to the sensitivity curves from \citet{Sathyaprakash} for the remaining detectors.
 
-The canonical design for a space based detector is LISA (Laser Interferometer Space Antenna), which would be sensitive to milli-Hz GWs.
-LISA would consist of three satellites flying in a triangular constellation with arms of length $5\times 10^{9}\,\textrm{m}$ in a $1\,\textrm{AU}$ orbit around the sun, trailing the Earth by $20\degree$.
-The laser arms in a LISA like detector are not a Fabry-P\'{e}rot cavity, the light only travels once along each arm.
-All of the sources discussed here lie within LISA's sensitivity reach.
-eLISA (evolved LISA) is a simpler version of LISA designed to probe the same frequency range, while proposals such as ALIA (Advanced Laser Interferometer Antenna), BBO (Big Bang Observer) and DECIGO (DECI-Hertz Interferometer GW Observatory) are designed to probe deci-Hz GWs.
+\subsection{Space-based detectors}\label{sec:space}
+
+Space-based detectors work on similar principles to ground based detectors, but with the test masses residing inside of independent, widely separated satellites.
+Space-based detectors are sensitive to lower frequency GWs than their ground-based counterparts; this is partly because space based detectors can have much longer arms, and partly because they are unaffected by seismic noise which limits the low frequency performance of ground-based detectors.
+
+The canonical design for a space-based detector is the aser Interferometer Space Antenna (LISA), which is sensitive to millihertz GWs.
+LISA would consist of three satellites flying in a triangular constellation with arms of length $5\times 10^{9}~\mathrm{m}$ in a $1~\mathrm{AU}$ orbit around the sun, trailing the Earth by $20\degree$.
+The laser arms in a LISA like detector are not a cavity, the light only travels once along each arm.
+eLISA is a simpler version of LISA designed to probe the same frequency range, while proposals such as the Advanced Laser Interferometer Antenna (ALIA), Big Bang Observer (BBO) and Deci-hertz Interferometer GW Observatory (DECIGO) are designed to probe decihertz GWs.
 
 \subsubsection{LISA and eLISA}
-We use an analytic fit to the instrumental noise curve given by \cite{Sathyaprakash}.
+
+The instrumental noise curves for LISA by the analytic fit given by \citet{Sathyaprakash}, which we use for the plots in \ref{app:a}.
 When observing individual sources with LISA there is an additional contribution to the noise from a background of unresolvable binaries.
-This is not included here as we consider the background as a source of GWs, see section \ref{sec:GB}.
-eLISA is a re-scoped version of the classic LISA mission, the main differences are shorter arms ($10^{9}\,\textrm{m}$ instead of $5\times 10^{9}\,\textrm{m}$), two laser arms instead of three and a different orbit (drifting away from Earth instead of $20\degree$ Earth trailing).
+This is not included here as we consider the background as a source of GWs (see section \ref{sec:GB}).
+eLISA is a re-scoped version of the classic LISA mission, the main differences are shorter arms ($10^{9}~\mathrm{m}$ instead of $5\times 10^{9}~\mathrm{m}$), two laser arms instead of three, and a different orbit (drifting away from Earth instead of $20\degree$ Earth trailing).
 The effect of these changes is a slightly reduced peak sensitivity and a shift to higher frequencies.
-We use an analytic fit to the instrumental noise curve given by \cite{Amaro-Seoane-et-al}.
+In \ref{app:a}, we use an analytic fit to the instrumental noise curve given by \citet{Amaro-Seoane-et-al}.
 
-\subsubsection{ALIA, DECIGO and BBO}
-These missions are designed to probe the deci-Hz region of the GW spectrum, they are considerably more ambitious than the LISA or eLISA mission and their launches will be further into the future.
-For ALIA a simple analytic fit to the sensitivity curve is used, while for DECIGO and BBO fits to the sensitivity curves given by \cite{2011PhRvD..83d4011Y} are used.
+\subsubsection{DECIGO, ALIA and BBO}
+
+These missions are designed to probe the decihertz region of the GW spectrum; they are considerably more ambitious than the LISA or eLISA mission and their launches will be further into the future.
+In \ref{app:a} we use a simple analytic fit to the sensitivity curve for ALIA \cplb{[citation needed]}, while for DECIGO and BBO, fits to the sensitivity curves given by \citet{2011PhRvD..83d4011Y} are used.
 
@@
 The redshift of the rate of arrival of pulses for a pulsar at a distance $L$ from Earth, in the direction of the unit spatial vector $\vec{p}$ induced by a GW travelling in the direction of the unit vector $\hat{\Omega}$ is given by \citep{anholm-2009}
-\begin{equation}\label{eq:pulsarredshift} z(t,\hat{\Omega})=\frac{1}{2}\frac{\hat{p}^{j}\hat{p}^{i}} {1+\hat{\Omega}\cdot\hat{p}}\left(h_{ij}^{\textrm{Pulsar}}(t-L,\hat{\Omega} )-h_{ij}^{\textrm{Earth}}(t,\hat{\Omega} )\right)=\frac{1}{2}\frac{\hat{p}^{j}\hat{p}^{i}} {1+\hat{\Omega}\cdot\hat{p}}\Delta h_{ij}(t,\hat{\Omega})\; .
+\begin{equation}\label{eq:pulsarredshift} z(t,\hat{\Omega})=\frac{1}{2}\frac{\hat{p}^{j}\hat{p}^{i}} {1+\hat{\Omega}\cdot\hat{p}}\left(h_{ij}^{\mathrm{Pulsar}}(t-L,\hat{\Omega} )-h_{ij}^{\mathrm{Earth}}(t,\hat{\Omega} )\right)=\frac{1}{2}\frac{\hat{p}^{j}\hat{p}^{i}} {1+\hat{\Omega}\cdot\hat{p}}\Delta h_{ij}(t,\hat{\Omega})\; .
 \end{equation}
@@
 The experimentally measured quantity is not the redshift but the timing residual, the two are related via
-\begin{equation}\label{eq:restored}  R(t,\hat{\Omega})=\int_{0}^{t}\textrm{d}t'\;z(t',\hat{\Omega}) \; . \end{equation}
+\begin{equation}\label{eq:restored}  R(t,\hat{\Omega})=\int_{0}^{t}\mathrm{d}t'\;z(t',\hat{\Omega}) \; . \end{equation}
 All of the pulsars, and the Earth, reside in the same metric perturbation field which may be expressed in terms of its Fourier transform
-\begin{equation} h_{ij}(t,\hat{x})=\sum_{A=+,\times}\int\textrm{d}f\;\iint_{S_{2}}\textrm{d}\hat{\Omega}\; \tilde{h}_{A}(f,\hat{\Omega})e_{ij}^{A}(\hat{\Omega})e^{ 2\pi i f (t-\hat{\Omega}\cdot\vec{x}) } \; ,\end{equation}
+\begin{equation} h_{ij}(t,\hat{x})=\sum_{A=+,\times}\int\mathrm{d}f\;\iint_{S_{2}}\mathrm{d}\hat{\Omega}\; \tilde{h}_{A}(f,\hat{\Omega})e_{ij}^{A}(\hat{\Omega})e^{ 2\pi i f (t-\hat{\Omega}\cdot\vec{x}) } \; ,\end{equation}
 where $e^{A}_{ij}(\hat{\Omega})$ is the $A$ polarisation basis tensor for direction $\hat{\Omega}$.
 Choosing the SSB as the origin of our coordinate system, so the pulsar is at position $L\hat{p}$, gives
-\begin{equation}\label{eq:deltahij} \Delta h_{ij}(t,\hat{\Omega})=\sum_{A=+,\times}\int\textrm{d}f\; \tilde{h}_{A}(f,\hat{\Omega})e_{ij}^{A}(\hat{\Omega})e^{ 2\pi i f t }\left(e^{-2\pi\rmi fL (1+\hat{p}\cdot\hat{\Omega})}-1\right) \; .
+\begin{equation}\label{eq:deltahij} \Delta h_{ij}(t,\hat{\Omega})=\sum_{A=+,\times}\int\mathrm{d}f\; \tilde{h}_{A}(f,\hat{\Omega})e_{ij}^{A}(\hat{\Omega})e^{ 2\pi i f t }\left(e^{-2\pi\rmi fL (1+\hat{p}\cdot\hat{\Omega})}-1\right) \; .
 \end{equation}
 From (\ref{eq:pulsarredshift}) and (\ref{eq:deltahij}) the Fourier transform of the redshift, $\tilde{z}(f,\hat{\Omega})$, may be read off as
-\begin{equation}\label{eq:45} \tilde{z}(f,\hat{\Omega})=\left(e^{-2\pi\rmi fL (1+\hat{p}\cdot\hat{\Omega})}-1\right)\sum_{A=+,\times} \tilde{h}_{A}(f,\hat{\Omega})F^{A}(\hat{\Omega})\;, \textrm{where}\; F^{A}(\hat{\Omega})=\frac{e_{ij}^{A}(\hat{\Omega})\hat{p}^{j}\hat{p}^{i}} {2(1+\hat{\Omega}\cdot\hat{p})} \, .\end{equation}
+\begin{equation}\label{eq:45} \tilde{z}(f,\hat{\Omega})=\left(e^{-2\pi\rmi fL (1+\hat{p}\cdot\hat{\Omega})}-1\right)\sum_{A=+,\times} \tilde{h}_{A}(f,\hat{\Omega})F^{A}(\hat{\Omega})\;, \mathrm{where}\; F^{A}(\hat{\Omega})=\frac{e_{ij}^{A}(\hat{\Omega})\hat{p}^{j}\hat{p}^{i}} {2(1+\hat{\Omega}\cdot\hat{p})} \, .\end{equation}
 The function $F^{A}(\hat{\Omega})$ may be regarded as the PTA equivalent of the detector response functions in (\ref{eq:responsefuncs}).
@@
 From (\ref{eq:stochback}) and (\ref{eq:45}) the expectation of the product of signals from two different pulsars in directions $\hat{p}_{1}$ and $\hat{p}_{2}$ may be evaluated
-\begin{eqnarray} & \left<z_{1}(f)z_{2}^{*}(f')\right>=\frac{1}{2}S_{h}(f)\delta (f-f')\Gamma(f)\quad \textrm{where,} \\
-& \Gamma(f)=\sum_{A=+,\times}\iint_{{\cal{S}}^{2}}\textrm{d}\hat{\Omega}\,\left(e^{2\pi\rmi fL_{1}(1+\hat{\Omega}\cdot\hat{p}_{1})}-1\right)\left(e^{-2\pi\rmi fL_{2}(1+\hat{\Omega}\cdot\hat{p}_{2})}-1\right)F_{1}^{A}(\hat{\Omega})F_{2}^{A}(\hat{\Omega}) \; .
+\begin{eqnarray} & \left<z_{1}(f)z_{2}^{*}(f')\right>=\frac{1}{2}S_{h}(f)\delta (f-f')\Gamma(f)\quad \mathrm{where,} \\
+& \Gamma(f)=\sum_{A=+,\times}\iint_{{\cal{S}}^{2}}\mathrm{d}\hat{\Omega}\,\left(e^{2\pi\rmi fL_{1}(1+\hat{\Omega}\cdot\hat{p}_{1})}-1\right)\left(e^{-2\pi\rmi fL_{2}(1+\hat{\Omega}\cdot\hat{p}_{2})}-1\right)F_{1}^{A}(\hat{\Omega})F_{2}^{A}(\hat{\Omega}) \; .
 \nonumber\end{eqnarray}
 The overlap function, $\Gamma(f)$, tends to a constant value in the limit that the distances to the pulsars are large compared to the wavelength of GWs, PTAs operate in this limit, so the overlap may be approximated as a constant
-\begin{equation} \Gamma(f)\approx\Gamma_{0}= \sum_{A=+,\times}\iint_{{\cal{S}}^{2}}\textrm{d}\hat{\Omega}\;F_{1}^{A}(\hat{\Omega})F_{2}^{A}(\hat{\Omega})\;.\end{equation}
+\begin{equation} \Gamma(f)\approx\Gamma_{0}= \sum_{A=+,\times}\iint_{{\cal{S}}^{2}}\mathrm{d}\hat{\Omega}\;F_{1}^{A}(\hat{\Omega})F_{2}^{A}(\hat{\Omega})\;.\end{equation}
 Neglecting the exponential terms in the overlap is the frequency domain equivalent of neglecting the pulsar term in (\ref{eq:pulsarredshift}).
 The integral may be evaluated to give an expression depending only on the angle, $\theta$, between the two pulsars; this is the famous Hellings and Downs curve, see figure \ref{fig:HnD},
-\begin{equation} \Gamma_{0}=\frac{1}{2}+\frac{3x}{2}\left(\log (x)-\frac{1}{6}\right) \quad \textrm{where}\quad x=\frac{1-\cos\theta}{2}\; .\end{equation}
+\begin{equation} \Gamma_{0}=\frac{1}{2}+\frac{3x}{2}\left(\log (x)-\frac{1}{6}\right) \quad \mathrm{where}\quad x=\frac{1-\cos\theta}{2}\; .\end{equation}
 
@@
 \subsubsection{SKA}
-The sensitivity curves plotted in \ref{app:a} for the SKA (Square Kilometer Array) assume values of $\delta t^{2}_{\textrm{r.m.s.}}$ a factor of 10 better than current PTAs and 20 pulsars timed for twice as long as the current EPTA.
+The sensitivity curves plotted in \ref{app:a} for the SKA (Square Kilometer Array) assume values of $\delta t^{2}_{\mathrm{r.m.s.}}$ a factor of 10 better than current PTAs and 20 pulsars timed for twice as long as the current EPTA.

### 2014-07-04 Detectors.tex
Up to PTAs

@@
 \section{Detectors}\label{sec:detectors}
+
+In this section we introduce the detector noise curves used in \ref{app:a}.
+We begin with with a description of the basic operation of detectors.
+We then discuss ground-based detectors, space-based detectors and PTAs.
+The latter function somewhat different than conventional interferometers, so we include a brief introduction to PTA analysis.
+References for the noise curve used for individual detectors can be found in the relevant subsections, further information about the detectors can be in discovered these.
 
@@
 
-In \ref{app:a}, we use an interpolation to the data published on \url{https://wwwcascina.virgo.infn.it/advirgo/} (2013) for the AdV sensitivity curve, an interpolation to the data for version D of the detector published on \url{http://gwcenter.icrr.u-tokyo.ac.jp/en/researcher/parameter} (2013) for KAGRA and analytic fits to the sensitivity curves from \citet{Sathyaprakash} for the remaining detectors.
+We use an interpolation to the data published on \url{https://wwwcascina.virgo.infn.it/advirgo/} (2013) for the AdV sensitivity curve, an interpolation to the data for version D of the detector published on \url{http://gwcenter.icrr.u-tokyo.ac.jp/en/researcher/parameter} (2013) for KAGRA and analytic fits to the sensitivity curves from \citet{Sathyaprakash} for the remaining detectors.
 
@@
 
-The canonical design for a space-based detector is the aser Interferometer Space Antenna (LISA), which is sensitive to millihertz GWs.
+The canonical design for a space-based detector is the Laser Interferometer Space Antenna (LISA), which is sensitive to millihertz GWs.
 LISA would consist of three satellites flying in a triangular constellation with arms of length $5\times 10^{9}~\mathrm{m}$ in a $1~\mathrm{AU}$ orbit around the sun, trailing the Earth by $20\degree$.
 The laser arms in a LISA like detector are not a cavity, the light only travels once along each arm.
-eLISA is a simpler version of LISA designed to probe the same frequency range, while proposals such as the Advanced Laser Interferometer Antenna (ALIA), Big Bang Observer (BBO) and Deci-hertz Interferometer GW Observatory (DECIGO) are designed to probe decihertz GWs.
+eLISA is a reduced version of LISA designed to probe the same frequency range, while proposals such as the Advanced Laser Interferometer Antenna (ALIA), Big Bang Observer (BBO) and Deci-hertz Interferometer GW Observatory (DECIGO) are designed to probe decihertz GWs.
 
@@
 The effect of these changes is a slightly reduced peak sensitivity and a shift to higher frequencies.
-In \ref{app:a}, we use an analytic fit to the instrumental noise curve given by \citet{Amaro-Seoane-et-al}.
+We use an analytic fit to the instrumental noise curve given by \citet{Amaro-Seoane-et-al}.
 
@@
 These missions are designed to probe the decihertz region of the GW spectrum; they are considerably more ambitious than the LISA or eLISA mission and their launches will be further into the future.
-In \ref{app:a} we use a simple analytic fit to the sensitivity curve for ALIA \cplb{[citation needed]}, while for DECIGO and BBO, fits to the sensitivity curves given by \citet{2011PhRvD..83d4011Y} are used.
+We use a simple analytic fit to the sensitivity curve for ALIA \cplb{[citation needed]}, while for DECIGO and BBO, fits to the sensitivity curves given by \citet{2011PhRvD..83d4011Y} are used.
 
 \subsection{Pulsar timing arrays}\label{sec:PTAgeneralproperties}
-PTAs can be thought of as naturally occurring interferometers with galactic scale arm lengths, hence they are sensitive to much lower frequencies than the detectors considered so far.
-Each pulsar is a very regular clock and the measured pulse arrival time (in practice, arrival times are corrected to arrival times at the solar-system  barycenter, SSB) can be compared against a prediction leaving a residual which includes the effects of passing GWs.
-Using an array of these pulsars spread across the sky allows us to correlate residuals between different pulsars and exploit the fact that the GWs influence all pulsars where as intrinsic pulsar noise clearly does not.
-The correlation between different pulsars depends only on their angular separation on the sky and has a distinctive shape, known as the Hellings and Downs curve \citep{HellingsDowns}, which we will now derive.
 
-The redshift of the rate of arrival of pulses for a pulsar at a distance $L$ from Earth, in the direction of the unit spatial vector $\vec{p}$ induced by a GW travelling in the direction of the unit vector $\hat{\Omega}$ is given by \citep{anholm-2009}
-\begin{equation}\label{eq:pulsarredshift} z(t,\hat{\Omega})=\frac{1}{2}\frac{\hat{p}^{j}\hat{p}^{i}} {1+\hat{\Omega}\cdot\hat{p}}\left(h_{ij}^{\mathrm{Pulsar}}(t-L,\hat{\Omega} )-h_{ij}^{\mathrm{Earth}}(t,\hat{\Omega} )\right)=\frac{1}{2}\frac{\hat{p}^{j}\hat{p}^{i}} {1+\hat{\Omega}\cdot\hat{p}}\Delta h_{ij}(t,\hat{\Omega})\; .
+PTAs can be thought of as naturally occurring interferometers with galactic-scale arm lengths.
+Accordingly, they are sensitive to much lower frequencies than the detectors previously discussed.
+Each pulsar is a regular clock and the measured pulse arrival time can be compared against a prediction, leaving a residual which includes the effects of passing GWs.
+Using an array of these pulsars spread across the sky allows us to correlate residuals between different pulsars, to exploit the fact that GWs influence all pulsars whereas intrinsic pulsar noise does not.
+The correlation between different pulsars depends only on their angular separation on the sky, and has a distinctive shape, known as the Hellings and Downs curve \citep{HellingsDowns}.
+
+The redshift of the rate of arrival of pulses for a pulsar at a distance $L$ from the Solar-System barycentre (SSB), in the direction of the unit spatial vector $\hat{p}$ induced by a GW travelling in direction of the unit vector $\hat{\Omega}$ is \citep{anholm-2009}
+\begin{equation}\label{eq:pulsarredshift}
+z(t,\hat{\Omega})= \frac{1}{2} \frac{\hat{p}^{j}\hat{p}^{i}} {1+\hat{\Omega}\cdot\hat{p}}\left[h_{ij}^{\mathrm{Pulsar}}\left(t-\frac{L}/c,\hat{\Omega} \right)-h_{ij}^{\mathrm{Earth}}(t,\hat{\Omega} )\right] = \frac{1}{2}\frac{\hat{p}^{j}\hat{p}^{i}}{1+\hat{\Omega}\cdot\hat{p}} \Delta h_{ij}(t,\hat{\Omega})\; .
 \end{equation}
-The redshift includes two terms, the pulsar term and the Earth term.
-The pulsar term is often neglected in PTA analysis as it can be considered as an extra noise term which averages to zero.
+The redshift includes two terms: the pulsar term and the Earth term.
+The pulsar term is often neglected in PTA analysis as it can be considered as an extra noise term which averages to zero across the array.
 The experimentally measured quantity is not the redshift but the timing residual, the two are related via
-\begin{equation}\label{eq:restored}  R(t,\hat{\Omega})=\int_{0}^{t}\mathrm{d}t'\;z(t',\hat{\Omega}) \; . \end{equation}
-All of the pulsars, and the Earth, reside in the same metric perturbation field which may be expressed in terms of its Fourier transform
-\begin{equation} h_{ij}(t,\hat{x})=\sum_{A=+,\times}\int\mathrm{d}f\;\iint_{S_{2}}\mathrm{d}\hat{\Omega}\; \tilde{h}_{A}(f,\hat{\Omega})e_{ij}^{A}(\hat{\Omega})e^{ 2\pi i f (t-\hat{\Omega}\cdot\vec{x}) } \; ,\end{equation}
-where $e^{A}_{ij}(\hat{\Omega})$ is the $A$ polarisation basis tensor for direction $\hat{\Omega}$.
+\begin{equation}\label{eq:restored}
+R(t,\hat{\Omega}) = \int_{0}^{t}\mathrm{d}t'\;z(t',\hat{\Omega}) \; .
+\end{equation}
+All of the pulsars, and the Earth, are subject to the same metric perturbation field. This may be expressed in terms of its Fourier transform
+\begin{equation}
+h_{ij}(t,\hat{x}) = \sum_{A\,=\,+,\,\times}\int\mathrm{d}f\;\iint_{\mathbb{S}_{2}}\mathrm{d}\hat{\Omega}\; \tilde{h}_{A}(f,\hat{\Omega})e_{ij}^{A}(\hat{\Omega}) \exp\left[2\pi i f (t-\hat{\Omega}\cdot\vec{x}\right] \; ,
+\end{equation}
+where $e^{A}_{ij}(\hat{\Omega})$ is the $A$ polarisation basis tensor for direction $\hat{\Omega}$ \cplb{What is $x$?}.
 Choosing the SSB as the origin of our coordinate system, so the pulsar is at position $L\hat{p}$, gives
-\begin{equation}\label{eq:deltahij} \Delta h_{ij}(t,\hat{\Omega})=\sum_{A=+,\times}\int\mathrm{d}f\; \tilde{h}_{A}(f,\hat{\Omega})e_{ij}^{A}(\hat{\Omega})e^{ 2\pi i f t }\left(e^{-2\pi\rmi fL (1+\hat{p}\cdot\hat{\Omega})}-1\right) \; .
+\begin{equation}\label{eq:deltahij}
+\Delta h_{ij}(t,\hat{\Omega}) = \sum_{A\,=\,+,\,\times}\int\mathrm{d}f\; \tilde{h}_{A}(f,\hat{\Omega})e_{ij}^{A}(\hat{\Omega})\exp\left(2\pi \rmi f t\right)\left\{\exp\left[-2\pi\rmi fL \left(1+\hat{p}\cdot\hat{\Omega}\right)\right]-1\right\} \; .
 \end{equation}
-From (\ref{eq:pulsarredshift}) and (\ref{eq:deltahij}) the Fourier transform of the redshift, $\tilde{z}(f,\hat{\Omega})$, may be read off as
-\begin{equation}\label{eq:45} \tilde{z}(f,\hat{\Omega})=\left(e^{-2\pi\rmi fL (1+\hat{p}\cdot\hat{\Omega})}-1\right)\sum_{A=+,\times} \tilde{h}_{A}(f,\hat{\Omega})F^{A}(\hat{\Omega})\;, \mathrm{where}\; F^{A}(\hat{\Omega})=\frac{e_{ij}^{A}(\hat{\Omega})\hat{p}^{j}\hat{p}^{i}} {2(1+\hat{\Omega}\cdot\hat{p})} \, .\end{equation}
+From (\ref{eq:pulsarredshift}) and (\ref{eq:deltahij}), the Fourier transform of the redshift $\tilde{z}(f,\hat{\Omega})$ can be identified as
+\begin{equation}\label{eq:45}
+\tilde{z}(f,\hat{\Omega}) = \left\{\exp\left[-2\pi\rmi fL \left(1+\hat{p}\cdot\hat{\Omega}\right)\right]-1\right\} \sum_{A\,=\,+,\,\times} \tilde{h}_{A}(f,\hat{\Omega})F^{A}(\hat{\Omega}) \; ,
+\end{equation}
+where
+\begin{equation}
+F^{A}(\hat{\Omega}) = \frac{e_{ij}^{A}(\hat{\Omega})\hat{p}^{j}\hat{p}^{i}} {2\left(1+\hat{\Omega}\cdot\hat{p}\right)} \, .
+\end{equation}
 The function $F^{A}(\hat{\Omega})$ may be regarded as the PTA equivalent of the detector response functions in (\ref{eq:responsefuncs}).
-The stochastic background of GWs is fully characterised by the one-sided power spectral density via the following expectation value
-\begin{equation}\label{eq:stochback} \left<h_{A}^{*}(f,\hat{\Omega})h_{A'}(f',\hat{\Omega}')\right>=\frac{1}{2}S_{h}(f)\delta^{(2)}(\hat{\Omega},\hat{\Omega}')\delta_{AA'}\delta(f-f') \; ,\end{equation}
-where $\delta^{(2)}(\hat{\Omega},\hat{\Omega}')$ is the delta function on the sphere.
-From (\ref{eq:stochback}) and (\ref{eq:45}) the expectation of the product of signals from two different pulsars in directions $\hat{p}_{1}$ and $\hat{p}_{2}$ may be evaluated
-\begin{eqnarray} & \left<z_{1}(f)z_{2}^{*}(f')\right>=\frac{1}{2}S_{h}(f)\delta (f-f')\Gamma(f)\quad \mathrm{where,} \\
-& \Gamma(f)=\sum_{A=+,\times}\iint_{{\cal{S}}^{2}}\mathrm{d}\hat{\Omega}\,\left(e^{2\pi\rmi fL_{1}(1+\hat{\Omega}\cdot\hat{p}_{1})}-1\right)\left(e^{-2\pi\rmi fL_{2}(1+\hat{\Omega}\cdot\hat{p}_{2})}-1\right)F_{1}^{A}(\hat{\Omega})F_{2}^{A}(\hat{\Omega}) \; .
-\nonumber\end{eqnarray}
-The overlap function, $\Gamma(f)$, tends to a constant value in the limit that the distances to the pulsars are large compared to the wavelength of GWs, PTAs operate in this limit, so the overlap may be approximated as a constant
-\begin{equation} \Gamma(f)\approx\Gamma_{0}= \sum_{A=+,\times}\iint_{{\cal{S}}^{2}}\mathrm{d}\hat{\Omega}\;F_{1}^{A}(\hat{\Omega})F_{2}^{A}(\hat{\Omega})\;.\end{equation}
-Neglecting the exponential terms in the overlap is the frequency domain equivalent of neglecting the pulsar term in (\ref{eq:pulsarredshift}).
-The integral may be evaluated to give an expression depending only on the angle, $\theta$, between the two pulsars; this is the famous Hellings and Downs curve, see figure \ref{fig:HnD},
-\begin{equation} \Gamma_{0}=\frac{1}{2}+\frac{3x}{2}\left(\log (x)-\frac{1}{6}\right) \quad \mathrm{where}\quad x=\frac{1-\cos\theta}{2}\; .\end{equation}
-
+The stochastic background of GWs is fully characterised by the one-sided power spectral density via the expectation value
+\begin{equation}\label{eq:stochback}
+\left<h_{A}^{*}(f,\hat{\Omega})h_{A'}(f',\hat{\Omega}')\right> = \frac{1}{2}S_{h}(f)\delta^{(2)}(\hat{\Omega}, \hat{\Omega}')\delta_{AA'}\delta(f-f') \; ,
+\end{equation}
+where $\delta^{(2)}(\hat{\Omega},\hat{\Omega}')$ is the delta-function on the sphere.
+From (\ref{eq:45}) and (\ref{eq:stochback}), the expectation of the product of signals from two different pulsars in directions $\hat{p}_{1}$ and $\hat{p}_{2}$ may be evaluated AS
+\begin{equation} 
+\left<z_{1}(f)z_{2}^{*}(f')\right> = \frac{1}{2}S_{h}(f)\delta (f-f')\Gamma(f)
+\end{equation}
+where,
+\begin{eqnarray}
+\Gamma(f) &= \sum_{A\,=\,+,\,\times} & \iint_{{\mathbb{S}}^{2}}\mathrm{d}\hat{\Omega}\,\left\{\exp\left[2\pi\rmi fL_{1}\left(1+\hat{\Omega}\cdot\hat{p}_{1}\right)\right]-1\right\} \nonumber \\*
+ & & \times \left\{\exp\left[-2\pi\rmi fL_{2}\left(1+\hat{\Omega}\cdot\hat{p}_{2}\right)\right]-1\right\} F_{1}^{A}(\hat{\Omega})F_{2}^{A}(\hat{\Omega}) \; .
+\end{eqnarray}
+The overlap function $\Gamma(f)$ tends to a constant value in the limit that the distances to the pulsars are large compared to the wavelength of GWs; PTAs operate in this limit, so the overlap may be approximated as a constant,
+\begin{equation}
+\Gamma(f) \approx \Gamma_{0} = \sum_{A\,=\,+,\,\times}\iint_{{\mathbb{S}}^{2}}\mathrm{d}\hat{\Omega}\;F_{1}^{A}(\hat{\Omega})F_{2}^{A}(\hat{\Omega})\;.
+\end{equation}
+Neglecting the exponential terms in the overlap is the frequency-domain equivalent of neglecting the pulsar term in (\ref{eq:pulsarredshift}).
+The integral may be evaluated to give an expression depending only on the angle $\theta$ between the two pulsars; this is the famous Hellings and Downs curve, shown in figure \ref{fig:HnD},
+\begin{equation}
+\Gamma_{0} = \frac{1}{2}+\frac{3x}{2}\left(\ln x -\frac{1}{6}\right)
+\end{equation}
+where
+\begin{equation}
+x = \frac{1-\cos\theta}{2}\; .
+\end{equation}
 \begin{figure}
@@
  \includegraphics[trim=0cm 0cm 0cm 0cm, width=0.6\textwidth]{HnDcurve.pdf}
- \caption{The Hellings and Downs curve, the correlation between two pulsars seperated on the sky by an angle $\theta$.}
+ \caption{The Hellings and Downs curve, the correlation between two pulsars separated on the sky by an angle $\theta$ \citep{HellingsDowns}.}
  \label{fig:HnD}
@@
 The sensitivity bandwidth of a PTA is set by the sampling properties of the data set.
-If measurements are spaced in time by $\delta t$ and taken for a total length of time $T$ then the PTA is sensitive to frequencies in the range $(1/T)<f<(1/\delta t)$.
+If measurements are spaced in time by $\delta t$ and taken for a total length of time $T$, then the PTA is sensitive to frequencies in the range $(1/T) < f < (1/\delta t)$.
 The characteristic strain that the PTA is sensitive to scales linearly with $f$ in this range.
@@
 The absolute value of the sensitivity is fixed by normalising to a calculated limit at a given frequency for each PTA.
-For a discussion of the sensitivities of PTA to both individual sources and stochastic background see \cite{MooreTaylorGair}.
+For a discussion of the sensitivities of PTA to both individual sources and stochastic background, see \citet{MooreTaylorGair}.
 
 There is a discrepancy between the treatment of PTA sensitivity curves here and the higher frequency detectors discussed in sections \ref{sec:ground} and \ref{sec:space}.
-When observing a long lived source, such as an inspiral, with a high frequency detector the convention was to define a \emph{characteristic} strain to satisfy (\ref{eq:hc}).
+When observing a long-lived source, such as an inspiral, with a high frequency detector the convention was to define a \emph{characteristic} strain to satisfy (\ref{eq:hc}).
 Here the convention is to leave the strain untouched and instead adjust the PTA sensitivity curve with observation time, again to satisfy (\ref{eq:hc}).
-This discrepancy is an unfortunate result of the conventions in use by the different GW communities; however it is also natural given the sources under observation.
-When observing a transient source, such as a burst or inspiral, which changes within the lifetime of the detector it is natural to consider the detector as performing constantly while the signal changes.
-However when observing a monochromatic source or a stochastic background which is unchanging over the detector lifetime it is more natural to consider the source as being fixed and the sensitivity of the detector gradually improving.
-All that is required by the definition in (\ref{eq:hc}) is that the ratio $h_{c}(f)/h_{n}(f)$ is constant.
+This discrepancy is an unfortunate result of the conventions in use by the different GW communities; however, it is also natural given the sources under observation.
+When observing a transient source, such as a burst or inspiral, which changes within the lifetime of the detector, it is natural to consider the detector as performing constantly while the signal changes.
+However, when observing a monochromatic source or a stochastic background, which is unchanging over the detector lifetime, it is more natural to consider the source as being fixed and the sensitivity of the detector gradually improving.
+All that is required by the definition in (\ref{eq:hc}) is that the ratio $h_\mathrm{c}(f)/h_{n}(f)$ is constant.
 

### 2014-07-05 Detectors.tex
End of PTAs

@@
 Let $F_{a}$ be the averaged response of the $a$-th interferometer, the average response of a network of $k$ detectors is obtained by adding in quadrature,
-\begin{equation} F_{\mathrm{Total}}^{2}=\frac{1}{k)}\sum_{a\,=\,1}^{k}F_{a}^{2} \; .\end{equation}
+\begin{equation} F_{\mathrm{Total}}^{2}=\frac{1}{k}\sum_{a\,=\,1}^{k}F_{a}^{2} \; .\end{equation}
 
@@
 The canonical design for a space-based detector is the Laser Interferometer Space Antenna (LISA), which is sensitive to millihertz GWs.
-LISA would consist of three satellites flying in a triangular constellation with arms of length $5\times 10^{9}~\mathrm{m}$ in a $1~\mathrm{AU}$ orbit around the sun, trailing the Earth by $20\degree$.
-The laser arms in a LISA like detector are not a cavity, the light only travels once along each arm.
+LISA would consist of three satellites flying in a triangular constellation with arms of length $5\times 10^{9}~\mathrm{m}$ in a $1~\mathrm{AU}$ orbit around the Sun, trailing the Earth by $20\degree$.
+The laser arms in a LISA-like detector are not a cavity, the light only travels once along each arm.
 eLISA is a reduced version of LISA designed to probe the same frequency range, while proposals such as the Advanced Laser Interferometer Antenna (ALIA), Big Bang Observer (BBO) and Deci-hertz Interferometer GW Observatory (DECIGO) are designed to probe decihertz GWs.
@@
 
-\subsubsection{EPTA/PPTA/NANOGrav}
-The PTAs currently in operation are the EPTA (European Pulsar Timing Array), PPTA (Parkes Pulsar Timing Array, based in  Australia) and NANOGrav (North American Nanohertz Observatory for GWs).
-There are published limits on the amplitude of the stochastic background from all three detectors: currently the lowest is from the EPTA, \cite{Haasteren}.
-The EPTA curves in the figures show the  published limit based on an analysis of 5 pulsars over approximately 10 years.
+\subsubsection{Current PTAs}
 
-\subsubsection{IPTA}
-Combining the existing arrays would yield a single PTA using approximately 4 times as many pulsars pulsars.
+The PTAs currently in operation are the European Pulsar Timing Array \citep[EPTA;][]{Ferdman2010}, the Parkes Pulsar Timing Array \citep[PPTA;][]{Manchester2013} based in Australia, and the North American Nanohertz Observatory for Gravitational waves \citep[NANOGrav;][]{Jenet2009}.
+There are published limits on the amplitude of the stochastic background from all three detectors: the most recent from EPTA is \citet{Haasteren}, from PPTA is \citet{Shannon2013} and from NANOGrav is \citet{Arzoumanian2014}.
+These are all comparable and we use the EPTA limits in \ref{app:a}, the figures show the published limit based on an analysis of five pulsars over approximately ten years.
+
+\subsubsection{The International Pulsar Timing Array}
+
+Combining the existing arrays would yield a single PTA using approximately 4 times as many pulsars; this consortium of consortia is known as the International Pulsar Timing Array \citep[IPTA;][]{Hobbs2010}.
 The IPTA curves plotted in the figures are based on 20 pulsars timed for twice as long as the current EPTA.
@@
 \subsubsection{SKA}
-The sensitivity curves plotted in \ref{app:a} for the SKA (Square Kilometer Array) assume values of $\delta t^{2}_{\mathrm{r.m.s.}}$ a factor of 10 better than current PTAs and 20 pulsars timed for twice as long as the current EPTA.
+
+The next great advancement in radio astronomy shall come with the completion of the Square Kilometre Array \citep[SKA;][]{Dewdney2009}.
+This shall greatly increasing the sensitivity of pulsar timing \citep{Kramer2004}.
+The sensitivity curves plotted in \ref{app:a} for the SKA assume values of $\delta t^{2}_{\mathrm{r.m.s.}}$ \cplb{[This is not defined above]} a factor of 10 better than current PTAs, and 20 pulsars timed for twice as long as the current EPTA.

### 2014-07-07 Detectors.tex
addressed some of Christopher's comments

@@
 Let us choose the origin of our coordinate system to be the beam-splitter of the interferometer, and $l_{1}^{i}$ and $l_{2}^{i}$ to be unit 3-vectors pointing along the two arms.
-In the absence of noise the output of the detector is the difference in strain between the two arms \citep{Thorne1987} \cplb{I think you mean this rather than MTW}
+In the absence of noise the output of the detector is the difference in strain between the two arms \citep{Thorne1987}
 \begin{equation}\label{eq:hoftxx}
@@
  \includegraphics[trim=0cm 0cm 0cm 0cm, width=0.99\textwidth]{LIGO_detectorframe.pdf}
- \caption{The angular response function of an interferometric detector shown both as a surface plot and in an Aitoff--Hammer projection \cplb{[citation needed]}.
-The two detector arms lie in the $x$--$y$ plane either side of one of the zeros in the response.
-\cplb{Can you label the colour bar/explain what the surface indicates?}}
+ \caption{The angular response function of an interferometric detector in (\ref{eq:skyav}) shown both as a surface plot and in an Aitoff--Hammer projection.
+The response is a function of two sky angles, $\theta$ and $\phi$, and varies between $0$ and $1$.
+The two detector arms lie in the $x$--$y$ plane either side of one of the zeros in the response.}
  \label{fig:LIGO}
@@
 This is the case for a stochastic background; however, for a non-inspiralling circular binary, the polarisation is a function of the two spherical polar angles ($\iota,\xi$) specifying the orientation of the binary's orbital angular momentum.
-Here, $\iota$ is the inclination angle, the polar angle between the orbital angular momentum and the line joining the source to the detector \cplb{($-\hat{r}$)} and $\xi$ is the azimuthal angle around the same line.
+Here, $\iota$ is the inclination angle, the polar angle between the orbital angular momentum and the line joining the source to the detector ($-\hat{r}$) and $\xi$ is the azimuthal angle around the same line.
 In this case, to characterise the detector sensitivity we average over all four angles ($\theta,\phi,\iota,\xi$).
 If the binary is inspiralling, then the polarisation depends on still more parameters which need to be averaged over.
-These more complicated averages all have the property that they depend on both the detector and the source, hence they are unhelpful for our present purpose \cplb{of \ldots}.
+These more complicated averages all have the property that they depend on both the detector and the source, hence they are unhelpful for our present purpose seperating the source amplitude from the detector sensitivity.
 Additionally, the different averages do not work out to be so different from each other: \citet{1993PhRvD..47.2198F} calculated the sensitivity for a detector with the LIGO geometry averaged over the four angles ($\theta,\phi,\iota,\xi$) as $\sqrt{4/25}=0.4$ times peak sensitivity, which should be compared with the value $\sqrt{1/5}\approx 0.447$ above.
@@
 They are now being upgraded to second-generation detectors, advanced LIGO (aLIGO) and advanced Virgo (AdV).
-\cplb{I think AdV is the official acronym, although it does break the pattern.} LIGO has two observatories, one at Hanford, Washington, which has two detectors, and another at Livingston, Louisiana.
+LIGO has two observatories, one at Hanford, Washington, which has two detectors, and another at Livingston, Louisiana.
 There is an agreement to move one of the upgraded Hanford detector systems to a location in India \citep{LIGO-India,Unnikrishnan2013}.
@@
 
-\subsection{Pulsar timing arrays}\label{sec:PTAgeneralproperties}
+\subsection{Pulsar timing arrays (PTAs)}\label{sec:PTAgeneralproperties}
 
@@
 \begin{equation}
-h_{ij}(t,\hat{x}) = \sum_{A\,=\,+,\,\times}\int\mathrm{d}f\;\iint_{\mathbb{S}_{2}}\mathrm{d}\hat{\Omega}\; \tilde{h}_{A}(f,\hat{\Omega})e_{ij}^{A}(\hat{\Omega}) \exp\left[2\pi i f (t-\hat{\Omega}\cdot\vec{x}\right] \; ,
+h_{ij}(t,\vec{x}) = \sum_{A\,=\,+,\,\times}\int\mathrm{d}f\;\iint_{\mathbb{S}_{2}}\mathrm{d}\hat{\Omega}\; \tilde{h}_{A}(f,\hat{\Omega})e_{ij}^{A}(\hat{\Omega}) \exp\left[2\pi i f (t-\hat{\Omega}\cdot\vec{x}\right] \; ,
 \end{equation}
-where $e^{A}_{ij}(\hat{\Omega})$ is the $A$ polarisation basis tensor for direction $\hat{\Omega}$ \cplb{What is $x$?}.
+where $e^{A}_{ij}(\hat{\Omega})$ is the $A$ polarisation basis tensor for direction $\hat{\Omega}$  and $\vec{x}$ is the spatial position.
 Choosing the SSB as the origin of our coordinate system, so the pulsar is at position $L\hat{p}$, gives
@@
 The function $F^{A}(\hat{\Omega})$ may be regarded as the PTA equivalent of the detector response functions in (\ref{eq:responsefuncs}).
-The stochastic background of GWs is fully characterised by the one-sided power spectral density via the expectation value
+The stochastic background of GWs is fully characterised by the one-sided PSD via the expectation value
 \begin{equation}\label{eq:stochback}
@@
 This shall greatly increasing the sensitivity of pulsar timing \citep{Kramer2004}.
-The sensitivity curves plotted in \ref{app:a} for the SKA assume values of $\delta t^{2}_{\mathrm{r.m.s.}}$ \cplb{[This is not defined above]} a factor of 10 better than current PTAs, and 20 pulsars timed for twice as long as the current EPTA.
+The sensitivity curves plotted in \ref{app:a} for the SKA assumes an rms error on the timing residuals a factor of 10 better than current PTAs, and 20 pulsars timed for twice as long as the current EPTA.

### 2014-07-07 Introduction.tex
Attempt at adding stuff to astrophysical sources
Ending
addressed some of Christopher's comments

@@
 Given a detector output, the challenge is to extract the signal.
-There is a well known solution to this problem which involves constructing a Wiener optimal filter \cplb{[citation needed]}.
+There is a well known solution to this problem which involves constructing a Wiener optimal filter \citep{Wiener49}.
 Let $K(t)$ be the filter function with Fourier transform $\tilde{K}(f)$.
@@
 \begin{equation}\label{eq:conv}
-\left(s*K\right)(\tau) = \int_{-\infty}^{\infty}\mathrm{d}t\;(n(t)+h(t))K(t-\tau) = \mathcal{S} + \mathcal{N} \; .
+\left(s*K\right)(\tau) = \int_{-\infty}^{\infty}\mathrm{d}t\;(n(t)+h(t))K(t-\tau) \approx \mathcal{S} + \mathcal{N} \; .
 \end{equation}
@@
  \end{eqnarray}
-using the definition of $S_{n}(f)$ from (\ref{eq:psd}). Hence the optimum signal-to-noise ratio (SNR) $\varrho$ is given by \cplb{Why is this optimal?}
+using the definition of $S_{n}(f)$ from (\ref{eq:psd}). Hence the signal-to-noise ratio (SNR) $\varrho$ is given by
 \begin{equation}\label{eq:SNRinnerprod} 
@@
 \end{equation}
-\cplb{Inner product should be between signals: you don't specify the Fourier transform} where we have introduced the inner product between signal $A$ and $B$ as \citep{Finn1992}
-\begin{equation}\label{eq:snrinnerprod} \left( A | B \right) = 4\Re\left\{\int_{0}^{\infty}\mathrm{d}f\;\frac{\tilde{A}^{*}(f)\tilde{B}(f)}{S_{n}(f)}\right\} \; .\end{equation}
+where we have introduced the inner product between signal $A$ and $B$ as \citep{Finn1992}
+\begin{equation}\label{eq:snrinnerprod} \left( \tilde{A}(f) | \tilde{B}(f) \right) = 4\Re\left\{\int_{0}^{\infty}\mathrm{d}f\;\frac{\tilde{A}^{*}(f)\tilde{B}(f)}{S_{n}(f)}\right\} \; .\end{equation}
 The optimum filter maximises the SNR in (\ref{eq:SNRinnerprod}). From the Cauchy--Schwarz inequality, it follows that the optimum filter is
@@
 
-\subsection{Characteristic strain}
+\subsection{Characteristic strain}\label{sec:character-strain}
 
@@
 
-An additional advantage of this convention is that the values on the strain axis for the detector curve $h_n(F)$ have a simple physical interpretation: they correspond to the root-mean-square noise in a bandwidth $f$.
+An additional advantage of this convention is that the values on the strain axis for the detector curve $h_n(f)$ have a simple physical interpretation: they correspond to the root-mean-square noise in a bandwidth $f$.
 One downside to plotting characteristic strain is that the values on the strain axis $h_\mathrm{c}(f)$ do not directly relate to the amplitude of the waves from the source.
@@
 
-\subsection{Power spectral density}\label{sec:psd}
+\subsection{Power spectral density (PSD)}\label{sec:psd}
 
-\cplb{Can we abbreviate to PSD?}
-
-A second commonly used quantity for sensitivity curves is the square root of the power spectral density (see figure \ref{fig:S}). When discussing a detector this is
+A second commonly used quantity for sensitivity curves is the square root of the PSD (see figure \ref{fig:S}). When discussing a detector this is
 \begin{equation}\label{eq:temp1}
@@
 
-The power spectral density has the nice property, demonstrated in (\ref{eq:meansquare1}), that integrated over all positive frequencies it gives the mean square amplitude of the signal in the detector.
+The PSD has the nice property, demonstrated in (\ref{eq:meansquare1}), that integrated over all positive frequencies it gives the mean square amplitude of the signal in the detector.
 However, in one important regard it is less appealing than characteristic strain: the height of the source above the detector curve is no longer is directly related to the SNR.
 
-\subsection{Energy density}
+\subsection{Energy density}\label{sec:energy-density}
 
@@
 
-The energy in GWs is described by the Isaacson stress tensor \citep[section 35.15]{MTW}
+The energy in GWs is described by the Isaacson stress--energy tensor \citep[section 35.15]{MTW}
 \begin{equation}
@@
 \label{eq:omega}
-\Omega_{\mathrm{GW}}(f) = \frac{fS_{\mathrm{E}}(f)}{\rho_{\mathrm{c}}c^{2}} \; .
+\Omega_\mathrm{GW}(f) = \frac{fS_{\mathrm{E}}(f)}{\rho_{\mathrm{c}}c^{2}} \; .
 \end{equation}
@@
 \subsection{Relating the different descriptions}
-The dimensionless energy density in GWs $\Omega_{\mathrm{GW}}$, spectral energy density $S_{\mathrm{E}}$, one-sided power spectral density $S_{h}$, characteristic strain $h_\mathrm{c}$ and frequency-domain strain $\tilde{h}(f)$ are related via
-\begin{equation}
-H_0^2\Omega_{\mathrm{GW}}(f)= \frac{8 \pi G}{3 c^{2}} fS_{\mathrm{E}}(f) = \frac{2\pi^2}{3} f^3 S_h(f) = \frac{2\pi^2}{3} f^2 h_\mathrm{c}^2(f) = \frac{8\pi^2}{3} f^4 \left|\tilde{h}(f)\right|^2\; ,
+The dimensionless energy density in GWs $\Omega_{\mathrm{GW}}$, spectral energy density $S_{\mathrm{E}}$, one-sided PSD, $S_{h}$, characteristic strain $h_\mathrm{c}$ and frequency-domain strain $\tilde{h}(f)$ are related via
+\begin{equation}\label{eq:differentdescriptions}
+H_0^2\Omega_\mathrm{GW}(f)= \frac{8 \pi G}{3 c^{2}} fS_{\mathrm{E}}(f) = \frac{2\pi^2}{3} f^3 S_h(f) = \frac{2\pi^2}{3} f^2 h_\mathrm{c}^2(f) = \frac{8\pi^2}{3} f^4 \left|\tilde{h}(f)\right|^2\; ,
 \end{equation}

### 2014-07-07 Sources.tex
addressed some of Christopher's comments

@@
 Descriptions of each source are given in the following sections.
-Sources with short durations (i.e., burst sources) and sources which evolve in time over much longer timescales than our observations are drawn with flat-topped boxes \cplb{for $h_\mathrm{c}(f)$}.
-Inspiralling binaries, or stochastic backgrounds of binaries, are drawn with a sloping top proportional to $f^{-2/3}$ \cplb{for $h_\mathrm{c}(f)$}, which is result derived in section \ref{sec:stoch}.
+Sources with short durations (i.e., burst sources) and sources which evolve in time over much longer timescales than our observations are drawn with flat-topped boxes for $h_\mathrm{c}(f)$.
+Inspiralling binaries, or stochastic backgrounds of binaries, are drawn with a sloping top proportional to $f^{-2/3}$ for $h_\mathrm{c}(f)$, which is result derived in section \ref{sec:stoch}.
 The width of the box gives the range of frequencies sources of a given type can have a while remaining at a detectable amplitude.
@@
 Space-based detectors shall be sensitive to equal-mass mergers in the range $(10^{4}$--$10^{7})\,\Msun$.
-Predictions of the event rate for these mergers range from ${\mathcal{O}}(3$--$300)~\mathrm{yr}^{-1}$ for LISA with SNRs of up to $10^4$ \cplb{[citation needed]}.
+Predictions of the event rate for these mergers range from ${\mathcal{O}}(10$--$100)~\mathrm{yr}^{-1}$ for eLISA with SNRs of up to $10^3$ \citep{TheGravitationalUniverse}.
 The large range in the rate reflects our uncertainty in the growth mechanisms of the supermassive black hole population \citep{Volonteri2010}.
@@
 The unresolvable binaries overlap to form a stochastic background as discussed in section \ref{sec:stoch}.
-\cplb{The distinction between resolvable and unresolvable is detector specific; here we choose LISA.
-This boundary will not be too different for eLISA but would move substantially for either of the decihertz detectors.} Plotted in \ref{app:a} is the estimate of this background due to \citet{Nelemans} where an observation time of one year has been assumed,
+The distinction between resolvable and unresolvable is detector specific; here we choose LISA.
+This boundary will not be too different for eLISA but would move substantially for either of the decihertz detectors.
+Plotted in \ref{app:a} is the estimate of this background due to \citet{Nelemans} where an observation time of one year has been assumed,
 \begin{equation}
@@
 EMRIs occur when a compact stellar mass object inspirals into a supermassive black hole.
-There is extreme uncertainty in the event rate for EMRIs due to the poorly constrained astrophysics in galactic centres \citep[e.g.,][]{Merritt2011}; the best guess estimate is ${\mathcal{O}}(10)$ events per year with eLISA with SNR $\ge 20$ \cplb{[citation needed]}.
+There is extreme uncertainty in the event rate for EMRIs due to the poorly constrained astrophysics in galactic centres \citep[e.g.,][]{Merritt2011}; the best guess estimate is around $25$ events per year with eLISA with SNR $\ge 20$ \citep{TheGravitationalUniverse}.
 The boxes plotted in \ref{app:a} have a characteristic strain of $h_{c}=3\times 10^{-20}$ at $10^{-2}~\mathrm{Hz}$, which corresponds to a $10\Msun$ BH inspiralling into a $10^{6}\Msun$ black hole at a luminosity distance of $1~\mathrm{Gpc}$.
-The frequency width of the box is somewhat unknown; EMRI events can occur into a black hole of any mass, and hence EMRIs can occur at any frequency.
-However, there is substantial uncertainty in the black hole mass function \cplb{[citation needed]}, so the boxes in \ref{app:a} are drawn with a width comparable to that of the LISA sensitivity curve.
+The frequency width of the box is somewhat unknown; EMRI events can occur into a black hole of any mass, and hence EMRIs can, in principal, occur at any frequency, so the boxes in \ref{app:a} are drawn with a width comparable to that of the LISA sensitivity curve.
 

### 2014-07-07 Typesofsource.tex
addressed some of Christopher's comments

@@
 \end{eqnarray}
-From (\ref{eq:strains}) and (\ref{eq:FTofEMRI}) the characteristic strain and the Fourier transform can be related for inspiralling sources \citep{FinnThorne}
+From (\ref{eq:differentdescriptions}) and (\ref{eq:FTofEMRI}) the characteristic strain and the Fourier transform can be related for inspiralling sources \citep{FinnThorne}
 \begin{equation}\label{eq:insphc}h_{c}(f) = \sqrt{\frac{2f^{2}}{\dot{f}}}h\;.\end{equation}

### 2014-07-07 Introduction.tex
redone derivation of equation 7, now very long and detailed

@@
 We will work with the Fourier transform of the signal, using the conventions that
-\begin{eqnarray} 
+\begin{eqnarray} \label{eq:defFourier}
 \tilde{x}(f) &= {\mathcal{F}}\left\{x(t)\right\} &= \int_{-\infty}^{\infty}\mathrm{d}t\;x(t)\exp (-2\pi \rmi f t ) \; , \\
@@
 
-When integrated over all positive frequencies the power spectral density gives the mean square noise amplitude. Starting with taking the average of the square of the detector noise:
+When integrated over all positive frequencies the power spectral density gives the mean square noise amplitude. Starting with taking the \emph{time} average of the square of the detector noise:
 \begin{eqnarray}\label{eq:meansquare}
-\overline{n(t)^{2}} &= \lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T}\mathrm{d}t\;\left| n(t) \right|^{2} \nonumber \\
- &= \lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T}\mathrm{d}t\; \int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\tilde{n}(f)\tilde{n}^{*}(f')\exp\left(2\pi \rmi ft \right)\exp\left(-2\pi \rmi f't \right)
+\overline{n(t)^{2}} &= \lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T}\mathrm{d}t\; n(t) ^{2} \nonumber \\
+ &= \lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T}\mathrm{d}t\; n(t)n^{*}(t) \nonumber \\
+ &= \lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T}\mathrm{d}t\; \int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\tilde{n}(f)\tilde{n}^{*}(f')e^{2\pi \rmi ft }e^{-2\pi \rmi f't } \nonumber\\
+ &= \lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T}\mathrm{d}t\; \int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;{\cal{F}}\left\{ n(\tau) \right\}(f){\cal{F}}\left\{ n(\tau) \right\}(f')^{*}e^{2\pi \rmi ft }e^{-2\pi \rmi f't }
 \end{eqnarray}
-Using the Fourier property ${\mathcal{F}}\left\{ n(t+\tau) \right\} = \tilde{n}(f)\exp (-2\pi \rmi f\tau )$, the exponential factors may be absorbed as a time-translation of the noise by an amount $t$.
-The noise is stationary and is invariant under this time translation.
-We are thus taking the time average...
-\cplb{I give up here. This derivation makes no sense in it's current state.}
+Where in the penultimate line we have used the definition of the Fourier transform, (\ref{eq:defFourier}), and in the last line we have just changed notation.
+A property of Fourier transforms is that in the time domain a translation by amount $t$ translates to a rotation in phase by $2\pi ft$ in the frequency domain; if ${\mathcal{F}}\left\{ n(\tau) \right\} = \tilde{n}(f)$ then ${\mathcal{F}}\left\{ n(\tau-t) \right\} = \tilde{n}(f)\exp (2\pi \rmi ft )$.
+Therefore the exponential factors in (\ref{eq:meansquare}) may be absorbed as a time-translation of the noise by an amount $t$
+\begin{eqnarray}
+\overline{n(t)^{2}} & = \lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T}\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;{\cal{F}}\left\{ n(\tau-t) \right\}(f){\cal{F}}\left\{ n(\tau-t) \right\}(f')^{*} \, .
+\end{eqnarray}
+Using the ergodic property of the system described above the time average (denoted by $\overline{(\ldots)}$) may now be changed into an ensemble average (denoted by $\left<\ldots\right>$).
+The noise is stationary so it's expectation value is unchanged by the time translation performed above. Therefore using (\ref{eq:psd}) gives
 \begin{eqnarray}\label{eq:meansquare1}
-\overline{n(t)^{2}} & = \lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T}\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;{\cal{F}}\left\{ n(\tau-t) \right\}(f){\cal{F}}\left\{ n(\tau-t) \right\}(f')^{*} \nonumber \\
 &=\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\left<\tilde{n}(f)\tilde{n}^{*}(f')\right>\nonumber \\
+&=\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\frac{1}{2}S_{n}(f)\delta(f-f') \nonumber \\
 & = \int_{0}^{\infty}\mathrm{d}f\; S_{n}(f)
 \end{eqnarray}
-Using the property of Fourier transforms, ${\cal{F}}\left\{ n(t+\tau) \right\}(f)=\tilde{n}(f)\exp (-2\pi \rmi f\tau )$, the time average can be turned into an ensemble average and hence the definition of power spectral density in (\ref{eq:psd}) may be used.
-\begin{eqnarray}\label{eq:meansquare1}
-%\left<n(t)^{2}\right>&=\lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\int_{-T}^{T}\mathrm{d}t\; {\cal{F}}\left\{n(\tau+t)\right\}\left[{\cal{F}}\left\{n(\tau+t)\right\}\right]^{*}\nonumber \\
-\overline{n(t)^{2}}& = \lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\int_{-T}^{T}\mathrm{d}t\; \tilde{n}(f)\tilde{n}^{*}(f') \nonumber \\
-&=\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\left<\tilde{n}(f)\tilde{n}^{*}(f')\right>\nonumber \\
-%&= \int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\frac{1}{2}S_{n}(f)\delta (f-f')\nonumber\\
-& = \int_{0}^{\infty}\mathrm{d}f\; S_{n}(f)
-\end{eqnarray}
-\cplb{Try someting based on Wiener--Khinchin theorem?}
 

### 2014-07-07 Introduction.tex
Premerge

--- 2014-07-07 Introduction.tex
@@
 
-When discussing the loudness of sources and the sensitivity of detectors there are three commonly used parametrizations based upon the strain: the characteristic strain, the power spectral density and the spectral energy density.
+When discussing the loudness of sources and the sensitivity of detectors there are three commonly used parametrizations based upon the strain: the characteristic strain, the PSD and the spectral energy density.
 We aim to disambiguate these three and give a concrete comparison of different detectors.
@@
 
-For simplicity, it is assumed that the noise in the GW detector is stationary and Gaussian (with zero mean); under these assumptions the noise is fully characterised via the one-sided noise power spectral density $S_{n}(f)$,
+For simplicity, it is assumed that the noise in the GW detector is stationary and Gaussian (with zero mean); under these assumptions the noise is fully characterised via the one-sided noise power spectral density (PSD) $S_{n}(f)$,
 \begin{equation}\label{eq:psd}
@@
 The procedure is to measure the noise over a sufficiently long duration $T$ and then compute the Fourier transform $\tilde{n}(f)$ with a frequency resolution $\Delta f = T^{-1}$; this is repeated many times to give an average.
-The noise power spectral density $S_{n}(f)$ has units of inverse frequency.
+The noise PSD $S_{n}(f)$ has units of inverse frequency.
 
 Since the GW signal and detector output are both real it follows that $\tilde{h}(-f)=\tilde{h}^{*}(f)$ and $\tilde{n}(-f)=\tilde{n}^{*}(f)$; therefore, $S_{n}(f)=S_{n}(-f)$.
-The fact that $S_{n}(f)$ is an even function means that Fourier integrals over all frequencies can instead be written as integrals over positive frequencies only, e.g., (\ref{eq:meansquare1}) and (\ref{eq:snrinnerprod}); it is for this reason that $S_{n}(f)$ is called the \emph{one-sided} power spectral density.\footnote{An alternative convention is to use the two-sided power spectral density $S^{(2)}_{n}(f) = S_{n}(f)/2$.}
+The fact that $S_{n}(f)$ is an even function means that Fourier integrals over all frequencies can instead be written as integrals over positive frequencies only, e.g., (\ref{eq:meansquare1}) and (\ref{eq:snrinnerprod}); it is for this reason that $S_{n}(f)$ is called the \emph{one-sided} PSD.\footnote{An alternative convention is to use the two-sided PSD $S^{(2)}_{n}(f) = S_{n}(f)/2$.}
 
-When integrated over all positive frequencies the power spectral density gives the mean square noise amplitude. Starting with taking the average of the square of the detector noise:
+When integrated over all positive frequencies the PSD gives the mean square noise amplitude. Starting with taking the average of the square of the detector noise:
 \begin{eqnarray}\label{eq:meansquare}
@@
 \end{eqnarray}
-Using the property of Fourier transforms, ${\cal{F}}\left\{ n(t+\tau) \right\}(f)=\tilde{n}(f)\exp (-2\pi \rmi f\tau )$, the time average can be turned into an ensemble average and hence the definition of power spectral density in (\ref{eq:psd}) may be used.
+Using the property of Fourier transforms, ${\cal{F}}\left\{ n(t+\tau) \right\}(f)=\tilde{n}(f)\exp (-2\pi \rmi f\tau )$, the time average can be turned into an ensemble average and hence the definition of PSD in (\ref{eq:psd}) may be used.
 \begin{eqnarray}\label{eq:meansquare1}
@@
 
-The characteristic strain $h_c$ is designed to include the effect of integrating an inspiralling signal. Its counterpart for describing noise is the noise amplitude $h_n$. These are defined as
+The characteristic strain $h_\mathrm{c}$ is designed to include the effect of integrating an inspiralling signal.
+Its counterpart for describing noise is the noise amplitude $h_n$.
+These are defined as
 \begin{eqnarray}\label{eq:strain-hc} 
@@
 \end{equation}
-The strain amplitudes $h_\mathrm{c}(f)$ and $h_{c}(f)$ are dimensionless.
+The strain amplitudes $h_\mathrm{c}(f)$ and $h_{n}(f)$ are dimensionless.
 Using this convention, when plotting on a log--log scale, the area between the source and detector curves is related to the SNR via (\ref{eq:hc}).
@@
 
-\subsection{Power spectral density (PSD)}\label{sec:psd}
+\subsection{Power spectral density }\label{sec:psd}
 
@@
 \end{equation}
-where we have used (\ref{eq:strain-hc}).
-Both $\sqrt{S_{n}(f)}$ and $\sqrt{S_{h}(f)}$ have units of $\mathrm{Hz^{-1/2}}$.
-The root power spectral density is the most frequently plotted quantity in the literature.
+where we have used (\ref{eq:strain-hc}). Both $\sqrt{S_{n}(f)}$ and $\sqrt{S_{h}(f)}$ have units of $\mathrm{Hz^{-1/2}}$. The root PSD is the most frequently plotted quantity in the literature.
 

### 2014-07-07 Sources.tex
Premerge

@@
 There is extreme uncertainty in the event rate for EMRIs due to the poorly constrained astrophysics in galactic centres \citep[e.g.,][]{Merritt2011}; the best guess estimate is around $25$ events per year with eLISA with SNR $\ge 20$ \citep{TheGravitationalUniverse}.
-The boxes plotted in \ref{app:a} have a characteristic strain of $h_{c}=3\times 10^{-20}$ at $10^{-2}~\mathrm{Hz}$, which corresponds to a $10\Msun$ BH inspiralling into a $10^{6}\Msun$ black hole at a luminosity distance of $1~\mathrm{Gpc}$.
+The boxes plotted in \ref{app:a} have a characteristic strain of $h_\mathrm{c}=3\times 10^{-20}$ at $10^{-2}~\mathrm{Hz}$, which corresponds to a $10\Msun$ BH inspiralling into a $10^{6}\Msun$ black hole at a luminosity distance of $1~\mathrm{Gpc}$.
 The frequency width of the box is somewhat unknown; EMRI events can occur into a black hole of any mass, and hence EMRIs can, in principal, occur at any frequency, so the boxes in \ref{app:a} are drawn with a width comparable to that of the LISA sensitivity curve.

### 2014-07-07 noise_curve.tex
Detectors to PTAs
Premerge

@@
 We have carefully defined these quantities and derived the relationships between them.
-The characteristic strain (section \ref{sec:character-strain}) is most directly related to the SNR, the power spectral density is mostly closely related to the mean square amplitude in the detector, and the energy density has a clear physical interpretation.
+The characteristic strain (section \ref{sec:character-strain}) is most directly related to the SNR, the PSD is mostly closely related to the mean square amplitude in the detector, and the energy density has a clear physical interpretation.
 We have produced example plots using each of these quantities for a wide range of detectors and sources.
@@
 
+\input{Plots}
+
 \bibliographystyle{apj}
@@
 
-\input{Plots}
-
 \end{document}

### 2014-07-07 Introduction.tex

--- 2014-07-07 Introduction.tex
@@
 The procedure is to measure the noise over a sufficiently long duration $T$ and then compute the Fourier transform $\tilde{n}(f)$ with a frequency resolution $\Delta f = T^{-1}$; this is repeated many times to give an average.
-The noise power spectral density $S_{n}(f)$ has units of inverse frequency.
+The noise PSD $S_{n}(f)$ has units of inverse frequency.
 
 Since the GW signal and detector output are both real it follows that $\tilde{h}(-f)=\tilde{h}^{*}(f)$ and $\tilde{n}(-f)=\tilde{n}^{*}(f)$; therefore, $S_{n}(f)=S_{n}(-f)$.
-The fact that $S_{n}(f)$ is an even function means that Fourier integrals over all frequencies can instead be written as integrals over positive frequencies only, e.g., (\ref{eq:meansquare1}) and (\ref{eq:snrinnerprod}); it is for this reason that $S_{n}(f)$ is called the \emph{one-sided} power spectral density.\footnote{An alternative convention is to use the two-sided power spectral density $S^{(2)}_{n}(f) = S_{n}(f)/2$.}
+The fact that $S_{n}(f)$ is an even function means that Fourier integrals over all frequencies can instead be written as integrals over positive frequencies only, e.g., (\ref{eq:meansquare1}) and (\ref{eq:snrinnerprod}); it is for this reason that $S_{n}(f)$ is called the \emph{one-sided} PSD.\footnote{An alternative convention is to use the two-sided PSD $S^{(2)}_{n}(f) = S_{n}(f)/2$.}
 
-When integrated over all positive frequencies the power spectral density gives the mean square noise amplitude. Starting with taking the \emph{time} average of the square of the detector noise:
+When integrated over all positive frequencies, the PSD gives the mean square noise amplitude. Starting with taking the \emph{time} average of the square of the detector noise:
 \begin{eqnarray}\label{eq:meansquare}

### 2014-07-07 Typesofsource.tex

--- 2014-06-30 Typesofsource.tex
@@
 
-Sources of GWs can be broadly split into three categories: well modelled sources, for which we have a description of the expected waveform; stochastic backgrounds, for which we can describe the statistical behaviour; and unmodelled (or poorly-modelled) transient sources.
-The classic example of a well modelled source is the inspiral of two compact objects, this is discussed in section \ref{sec:insp}.
-Stochastic backgrounds can either be formed from many overlapping sources, which are each individually susceptible to being modelled, or from some intrinsically random process, these are discussed in section \ref{sec:stoch}.
-An example of an unmodelled (or poorly-modelled) transient source is a supernova; searches of this type are often called burst searches and are discussed in section \ref{sec:bursts}.
+GW signals can be broadly split into three categories: those from well-modelled sources, for which we have a description of the expected waveform; stochastic backgrounds, for which we can describe the statistical behaviour; and unmodelled (or poorly-modelled) transient sources.
+The classic example of a well-modelled source is the inspiral of two compact objects, this is discussed in section \ref{sec:insp}.
+Stochastic backgrounds can either be formed from many overlapping sources, which individually could be modelable, or from some intrinsically random process, these are discussed in section \ref{sec:stoch}.
+An example of an unmodelled (or poorly-modelled) transient source is a supernova; searches for signals of this type are often called burst searches and are discussed in section \ref{sec:bursts}.
 
 \subsection{Inspirals}\label{sec:insp}
-Inspiralling binaries spend a variable amount of time in each frequency band. If $\phi$ is the orbital phase then the number of cycles generated at frequency $f$ can be estimated as
-\begin{equation}\label{eq:inspiral} {\cal{N}}_{\textrm{cycles}} = \frac{f}{2\pi}\frac{\textrm{d}\phi}{\textrm{d}f}=\frac{f^{2}}{\dot{f}} \; ,\end{equation}
-where an overdot represents the time derivative and $f=\dot{\phi}$.
-The squared SNR will scale with ${\cal{N}}_{\textrm{cycles}}$, so it would be expected that $h_{c}(f)\approx \sqrt{{\cal{N}}_{\textrm{cycles}}}\tilde{h}(f)$.
-The exact conversion can be derived by considering the Fourier transform in the stationary phase approximation.
-Consider a source signal with approximately constant amplitude, $h$, and central frequency $f'$.
-\begin{eqnarray}&h(t)= h\exp\left(\rmi \phi(t)\right) \\
-\Rightarrow &\tilde{h}(f)=\int_{-\infty}^{\infty}d\textrm{t}\;h\exp \left(2\pi \rmi \left(\frac{\phi(t)}{2\pi t}-f\right)t\right)\end{eqnarray}
-Without loss of generality assume an initial phase of zero and expand the exponent in powers of $t$,
-\begin{eqnarray} \label{eq:FTofEMRI} \tilde{h}(f) &\approx \int_{-\infty}^{\infty}\rmd t\; h \exp\left[2\pi\rmi \dot{f'} \left(t^{2}+\frac{f'-f}{\dot{f'}}t\right)\right]\nonumber \\
-&\approx h\exp\left[-\pi\rmi\frac{\left(f'-f\right)^{2}}{2\dot{f'}}\right]\int_{-\infty}^{\infty}\rmd t\;\exp\left[2\pi\rmi\dot{f'}\left(t+\frac{f'-f}{2\dot{f'}}\right)^{2}\right] \nonumber \\
-&\approx\frac{h}{\sqrt{2i\dot{f'}}}\exp\left[-\pi\rmi\frac{\left(f'-f\right)^{2}}{2\dot{f'}}\right]\; .
+
+Inspiralling binaries may the most important GW source.
+They spend a variable amount of time in each frequency band.
+If $\phi$ is the orbital phase, then the number of cycles generated at frequency $f$ can be estimated as
+\begin{equation}\label{eq:inspiral}
+{N}_{\mathrm{cycles}} = \frac{f}{2\pi} \frac{\mathrm{d}\phi}{\mathrm{d}f} = \frac{f^{2}}{\dot{f}} \; ,
+\end{equation}
+where an overdot represents the time derivative and $\dot{\phi} = 2\pi f$.
+The squared SNR scales with ${N}_{\mathrm{cycles}}$, so it would be expected that $h_\mathrm{c}(f)\approx \sqrt{{N}_{\mathrm{cycles}}}\tilde{h}(f)$.
+
+The form for $h_\mathrm{c}$ can be derived from the Fourier transform in the stationary-phase approximation.
+Consider a source signal with approximately constant amplitude, $h_0$, and central frequency $f'$.
+In this case,
+\begin{eqnarray}
+h(t) &=& h_0\exp\left[\rmi \phi(t)\right] \; ; \\
+\tilde{h}(f) &=& \int_{-\infty}^{\infty}\rmd t\; h_0\exp \left\{2\pi \rmi \left[\frac{\phi(t)}{2\pi t}-f\right]t\right\} \; .
 \end{eqnarray}
-From (\ref{eq:strains}) and (\ref{eq:FTofEMRI}) the characteristic strain and the Fourier transform can be related for inspiralling sources \citep{FinnThorne}
-\begin{equation}\label{eq:insphc}h_{c}(f) = \sqrt{\frac{2f^{2}}{\dot{f}}}h\;.\end{equation}
-Equation (\ref{eq:insphc}) is the relation between $h_{c}(f)$ and the instantaneous amplitude $h$ for an inspiralling source; for other types of source a new relation which satisfies (\ref{eq:hc}) will have to be found.
+Without loss of generality, we can assume an initial phase of zero, such that $\phi(t) \approx 2\pi f't$.
+The largest contribution to the integral comes from where the argument of the exponential is approximately zero.
+Expanding the exponent in powers of $t$ about $f = f'$,
+\begin{eqnarray} \label{eq:FTofEMRI}
+\tilde{h}(f) &\simeq& \int_{-\infty}^{\infty}\rmd t\; h_0 \exp\left[2\pi\rmi \dot{f'} \left(t^{2}+\frac{f'-f}{\dot{f'}}t\right)\right]\nonumber \\
+&\simeq& h_0\exp\left[-\pi\rmi\frac{\left(f'-f\right)^{2}}{2\dot{f'}}\right]\int_{-\infty}^{\infty}\rmd t\;\exp\left[2\pi\rmi\dot{f'}\left(t+\frac{f'-f}{2\dot{f'}}\right)^{2}\right] \nonumber \\
+&\simeq& \frac{h_0}{\sqrt{2\rmi\dot{f'}}}\exp\left[-\pi\rmi\frac{\left(f'-f\right)^{2}}{2\dot{f'}}\right]\; .
+\end{eqnarray}
+From (\ref{eq:strain-hc}) and (\ref{eq:FTofEMRI}), the characteristic strain for inspiralling sources is given by \citep{FinnThorne}
+\begin{equation}\label{eq:insphc}
+h_\mathrm{c}(f) = \sqrt{\frac{2f^{2}}{\dot{f}}}h_0 \;.
+\end{equation}
 Equation (\ref{eq:hc}) should be considered as the definition of characteristic strain and (\ref{eq:insphc}) a consequence of it for inspirals.
+Equation (\ref{eq:insphc}) is the relation between $h_\mathrm{c}(f)$ and the instantaneous amplitude $h_0$ for an inspiralling source; for other types of source a new relation satisfying (\ref{eq:hc}) has to be found.
 
 \subsection{Stochastic backgrounds}\label{sec:stoch}
-Another important source of GWs arises from stochastic backgrounds due to large populations of individually unresolvable sources.
-The population of sources will in general be at cosmological distances and it is necessary to distinguish the frequency in the source rest frame, $f_{\textrm{r}}$, from the measured frequency, $f$; these are related through the redshift, $z$, via $f_{\textrm{r}}=(1+z)f$.
-The comoving number density of sources, $n$, producing the background will also be a function of redshift; if the sources producing the stochastic background are all in the local universe then simply set $n(z)=\delta (z)$ and replace $d_{\textrm{L}}(z)$ with $d$ in all that follows (where $d_{\textrm{L}}(z)$ and $d$ are respectively the luminosity and comoving distances to the source, $d_{\textrm{L}}(z)=(1+z)d_{m}$).
+
+Another important source of GWs are stochastic backgrounds, which can be produced from a large population of unresolvable sources.
+These can be at cosmological distances, then it is necessary to distinguish the frequency in the source rest frame $f_{\mathrm{r}}$ from the measured frequency $f$; the two are related through the redshift $z$ via $f_{\mathrm{r}}=(1+z)f$.
+
+The comoving number density of sources $\nu$ producing the background is also a function of redshift; if the sources producing the background are all in the local universe, then simply set $\nu(z) = \nu_0\delta (z)$ and replace $d_{\mathrm{L}}(z)$ with $d$ in all that follows, where $d_{\mathrm{L}}(z)$ and $d$ are respectively the luminosity and comoving distances to the source, $d_{\mathrm{L}}(z)=(1+z)d$.
+
+We shall assume the the individual sources are binaries, in which case the number density of sources is also be a function of component mass.
+It is convenient to work in terms of the chirp mass, defined as ${\mathcal{M}}=\mu^{3/5}M^{2/5}$, where $\mu$ is the reduced mass and $M$ is the total mass of the binary.
+The comoving number density of sources shall be represented by $\nu(z, \mathcal{M})$.
+
 Equation (\ref{eq:omega}) gives an expression for the energy density in GWs per logarithmic frequency interval,
-\begin{equation}\label{eq:stoch} fS_{\textrm{E}}(f)=\frac{\pi c^{2}}{4G}f^{2}h_{c}(f)^{2} \; . \end{equation}
-The total energy emitted in the logarithmic frequency interval $\textrm{d}\left(\log (f_{\textrm{r}})\right)$ by a single binary in the population is $\left(\textrm{d}E_{\textrm{GW}}/\textrm{d}(\log f_{\textrm{r}})\right)\textrm{d}(\log f_{\textrm{r}})$; the energy density may be written as
-\begin{equation}\label{eq:Phinney} fS_{\textrm{E}}(f)=\int_{0}^{\infty}dz\; \frac{\textrm{d}n}{\textrm{d}z}\frac{1}{(1+z)}\frac{1}{d_{\textrm{L}}(z)^{2}}\frac{\textrm{d}E_{\textrm{GW}}}{\textrm{d}\left(\log f_{\textrm{r}} \right)} \; , \end{equation}
+\begin{equation}\label{eq:stoch}
+fS_{\mathrm{E}}(f)=\frac{\pi c^{2}}{4G}f^{2}h_\mathrm{c}^{2}(f) \; .
+\end{equation}
+The total energy emitted in the logarithmic frequency interval $\mathrm{d}\left(\log f_{\mathrm{r}}\right)$ by a single binary in the population is $\left[\mathrm{d}E_{\mathrm{GW}}/\mathrm{d}(\log f_{\mathrm{r}})\right]\mathrm{d}(\log f_{\mathrm{r}})$; the energy density may be written as
+\begin{equation}\label{eq:Phinney}
+fS_{\mathrm{E}}(f) = \int_{0}^{\infty} \rmd z\; \frac{\mathrm{d}\nu}{\mathrm{d}z}\frac{1}{(1+z)}\frac{1}{d_{\mathrm{L}}^{2}(z)}\frac{\mathrm{d}E_{\mathrm{GW}}}{\mathrm{d}\left(\log f_{\mathrm{r}} \right)} \; , \end{equation}
 where the factor of $\left( 1+z \right)^{-1}$ accounts for the redshift of the energy.
 
-For simplicity consider all the binaries comprising the background to be in circular orbits with frequencies $\nu=f_{\textrm{r}}/2$, and to be far from their last stable orbit; the energy radiated may then be calculated using the quadrupole approximation, \cite{petersmathews1963}.
-The chirp mass is defined as ${\cal{M}}=\mu^{3/5}M^{2/5}$, where $\mu$ is the reduced mass and $M$ is the total mass of the binary.
-The number density of sources will also be a function of chirp mass, $n(z,{\cal{M}})$.
+For simplicity, consider the background to comprise of binaries in circular orbits, with frequencies $f_\mathrm{GW} =f_{\mathrm{r}}/2$, which are far from their last stable orbit.
+The energy radiated may then be calculated using the quadrupole approximation \citep{petersmathews1963}.
 The energy in GWs from a single binary per logarithmic frequency interval is,
-\begin{equation}\label{eq:Thorne} \frac{\textrm{d}E_{\textrm{GW}}}{\textrm{d}\left(\log f_{\textrm{r}} \right)}=\frac{G^{2/3}\pi^{2/3}}{3}{\cal{M}}^{5/3}f_{\textrm{r}}^{2/3} \; , \end{equation}
+\begin{equation}\label{eq:Thorne}
+\frac{\mathrm{d}E_{\mathrm{GW}}}{\mathrm{d}\left(\log f_{\mathrm{r}} \right)} = \frac{G^{2/3}\pi^{2/3}}{3}{\cal{M}}^{5/3}f_{\mathrm{r}}^{2/3} \; ,
+\end{equation}
 between an minimum and maximum frequency set by the initial and final radius of the binary orbit.
-Here we assume that the maximum and minimum frequencies are outside of the range of frequencies probed by our detector and hence may be neglected.
-Using (\ref{eq:stoch}), (\ref{eq:Phinney}) and (\ref{eq:Thorne}) an expression for characteristic strain can now be found, \cite{SesanaVecchioColancino}
+Here, we assume that the maximum and minimum frequencies are outside of the range of our detector and hence can be neglected.
+Using (\ref{eq:stoch}), (\ref{eq:Phinney}) and (\ref{eq:Thorne}), an expression for characteristic strain can now be found \citep{SesanaVecchioColancino}
 \begin{equation}\label{eq:bigint}
-h_{c}(f)^{2}=\frac{4G^{5/3}}{3\pi^{1/3}c^{2}}f^{-4/3}\int_{0}^{\infty}\textrm{d}z\;\int_{0}^{\infty}\textrm{d}{\cal{M}}\;\frac{\textrm{d}^{2}n}{\textrm{d}z\,\textrm{d}{\cal{M}}}\frac{1}{d_{\textrm{L}}(z)^{2}}\left( \frac{{\cal{M}}^{5}}{1+z} \right)^{1/3}\; .
+h_\mathrm{c}^{2}(f) = \frac{4G^{5/3}}{3\pi^{1/3}c^{2}}f^{-4/3}\int_{0}^{\infty}\mathrm{d}\nu\;\int_{0}^{\infty}\mathrm{d}{\cal{M}}\;\frac{\mathrm{d}^{2}\nu}{\mathrm{d}z\,\mathrm{d}{\mathcal{M}}}\frac{1}{d_{\mathrm{L}}^{2}(z)}\left( \frac{{\mathcal{M}}^{5}}{1+z} \right)^{1/3}\; .
 \end{equation}
+
 From (\ref{eq:bigint}) it can be seen that the characteristic strain due to a stochastic background of binaries is a power law in frequency with spectral index $\alpha=-2/3$.
-The amplitude of the background depends on the population statistics of the binaries under consideration via $n(z,{\cal{M}})$.
+The amplitude of the background depends on the population statistics of the binaries under consideration via $\nu(z,{\mathcal{M}})$.
 The power law is often parametrised as
-\begin{equation}\label{eq:power} h_{c}(f)= A\left(\frac{f}{f_{0}}\right)^{\alpha}\; , \end{equation}
+\begin{equation}\label{eq:power} 
+h_\mathrm{c}(f) = A\left(\frac{f}{f_{0}}\right)^{\alpha}\; , 
+\end{equation}
 and constraints are then placed on $A$.
-In practice this power law will also have upper and lower frequency cut-offs related to the population of objects creating the GW spectrum.
-The stochastic background due to other sources, such as cosmic strings or relic GWs from the early Universe, can also be written in the same form as (\ref{eq:power}) but with different spectral indices: $\alpha=-7/6$ for cosmic strings or $\alpha$ in the range $-1$ to $-0.8$ for relic GWs, \cite{Jenet}.
+In practice, this power law also has upper and lower frequency cut-offs related to the population of source objects.
+A stochastic background from other sources, such as cosmic strings or relic GWs from the early Universe, can also be written in the same form as (\ref{eq:power}), but with different spectral indices: $\alpha=-7/6$ for cosmic strings or $\alpha$ in the range $-1$ to $-0.8$ for relic GWs \citep{Jenet}.
 
@@
 Some sources of GWs can produce signals with large amplitudes, greater than the detector noise.
-The duration of such a signal is typically very short, of the order of a few wave periods, and so there is not time to accumulate SNR in each frequency band in the way inspirals do.
-As a consequence, waveform models are not required to produce a detection; we simply rely on the "excess power" produced by these so-called burst sources.
-Typically, we may be looking for signals from core collapse supernovae \citep{Ott2009}, the late stages of merging compact binaries, or more generally, signals from any unexpected or poorly modelled sources.
+The typical duration of such a signal is short, of the order of a few wave periods, and so there is not time to accumulate SNR in each frequency band as for inspirals.
+As a consequence, waveform models are not required for detection; we simply rely on identifying the excess power produced by these burst sources.
+Typically, we may be looking for signals from core-collapse supernovae \citep{Ott2009}, the late stages of merging compact binaries, or more generally, signals from any unexpected or poorly modelled sources.
 
-Burst searches are often carried out using time-frequency techniques.
+Burst searches are often carried out using time--frequency techniques.
 The data stream from a detector is temporally split into segments, the length of which can be tuned to give greater sensitivity to particular sources.
-Each segment is then transformed into the frequency domain, whitened and normalised to the noise spectrum to produce a time-frequency plot.
-Potential GW signals are identified by searching for clusters of pixels that contain an excess of power (see e.g.
-\cite{Bursts}).
+Each segment is then transformed into the frequency domain, whitened and normalised to the noise spectrum to produce a time--frequency plot.
+Potential GW signals are identified by searching for clusters of pixels that contain an excess of power \citep[e.g.,][]{Bursts}.
 
@@
 An important aspect of burst search algorithms is to accurately estimate the noise properties within each time segment.
-To this extent, null data streams can be constructed that should be insensitive to real GW signals.
+To this extent, null data streams can be constructed that are insensitive to real GW signals.
 In order to estimate the false alarm rate, the data from different detectors can be shifted in time to remove any genuine coincident GW signals.
@@
 
-As discussed in \ref{sec:insp}, the expected relation between $h_{c}(f)$ and a typical waveform $\tilde{h}(f)$ is simply
-\begin{equation}\label{eq:simple} h_{c}(f)=\tilde{h}(f)\sqrt{{\cal{N}}_{\textrm{cycles}}} \; , \end{equation}
-where ${\cal{N}}_{\textrm{cycles}}$ is the number of cycles of radiation generated by the source.
-Here, ${\cal{N}}_{\textrm{cycles}}$ is of order unity and $\tilde{h}(f)$ is strongly peaked around some central frequency.
+As discussed in \ref{sec:insp}, the expected relation between $h_\mathrm{c}(f)$ and a typical waveform $\tilde{h}(f)$ is
+\begin{equation}\label{eq:simple} 
+h_\mathrm{c}(f) = \tilde{h}(f)\sqrt{{N}_{\mathrm{cycles}}} \; , 
+\end{equation}
+where ${N}_{\mathrm{cycles}}$ is the number of cycles of radiation generated by the source.
+For bursts, $N_{\mathrm{cycles}}$ is of order unity and $\tilde{h}(f)$ is strongly peaked around some central frequency.

### 2014-07-07 Detectors.tex
Premerge
PSD derivation

@@
 \begin{equation}\label{eq:pulsarredshift}
-z(t,\hat{\Omega})= \frac{1}{2} \frac{\hat{p}^{j}\hat{p}^{i}} {1+\hat{\Omega}\cdot\hat{p}}\left[h_{ij}^{\mathrm{Pulsar}}\left(t-\frac{L}/c,\hat{\Omega} \right)-h_{ij}^{\mathrm{Earth}}(t,\hat{\Omega} )\right] = \frac{1}{2}\frac{\hat{p}^{j}\hat{p}^{i}}{1+\hat{\Omega}\cdot\hat{p}} \Delta h_{ij}(t,\hat{\Omega})\; .
+z(t,\hat{\Omega})= \frac{1}{2} \frac{\hat{p}^{j}\hat{p}^{i}} {1+\hat{\Omega}\cdot\hat{p}}\left[h_{ij}^{\mathrm{Pulsar}}\left(t-\frac{L}{c},\hat{\Omega} \right)-h_{ij}^{\mathrm{Earth}}(t,\hat{\Omega} )\right] = \frac{1}{2}\frac{\hat{p}^{j}\hat{p}^{i}}{1+\hat{\Omega}\cdot\hat{p}} \Delta h_{ij}(t,\hat{\Omega})\; .
 \end{equation}
@@
 \begin{equation}
-h_{ij}(t,\vec{x}) = \sum_{A\,=\,+,\,\times}\int\mathrm{d}f\;\iint_{\mathbb{S}_{2}}\mathrm{d}\hat{\Omega}\; \tilde{h}_{A}(f,\hat{\Omega})e_{ij}^{A}(\hat{\Omega}) \exp\left[2\pi i f (t-\hat{\Omega}\cdot\vec{x}\right] \; ,
+h_{ij}(t,\vec{r}) = \sum_{A\,=\,+,\,\times}\int\mathrm{d}f\;\iint_{\mathbb{S}_{2}}\mathrm{d}\hat{\Omega}\; \tilde{h}_{A}(f,\hat{\Omega})e_{ij}^{A}(\hat{\Omega}) \exp\left[2\pi i f \left(t-\frac{\hat{\Omega}\cdot\vec{x}}{c}\right)\right] \; ,
 \end{equation}
-where $e^{A}_{ij}(\hat{\Omega})$ is the $A$ polarisation basis tensor for direction $\hat{\Omega}$  and $\vec{x}$ is the spatial position.
+where $e^{A}_{ij}(\hat{\Omega})$ is the $A$ polarisation basis tensor for direction $\hat{\Omega}$  and $\vec{r}$ is the spatial position.
 Choosing the SSB as the origin of our coordinate system, so the pulsar is at position $L\hat{p}$, gives
@@
 \begin{equation}\label{eq:stochback}
-\left<h_{A}^{*}(f,\hat{\Omega})h_{A'}(f',\hat{\Omega}')\right> = \frac{1}{2}S_{h}(f)\delta^{(2)}(\hat{\Omega}, \hat{\Omega}')\delta_{AA'}\delta(f-f') \; ,
+\left<\tilde{h}_{A}^{*}(f,\hat{\Omega})\tilde{h}_{A'}(f',\hat{\Omega}')\right> = \frac{1}{2}S_{h}(f)\delta^{(2)}(\hat{\Omega}, \hat{\Omega}')\delta_{AA'}\delta(f-f') \; ,
 \end{equation}
 where $\delta^{(2)}(\hat{\Omega},\hat{\Omega}')$ is the delta-function on the sphere.
-From (\ref{eq:45}) and (\ref{eq:stochback}), the expectation of the product of signals from two different pulsars in directions $\hat{p}_{1}$ and $\hat{p}_{2}$ may be evaluated AS
+From (\ref{eq:45}) and (\ref{eq:stochback}), the expectation of the product of signals from two different pulsars in directions $\hat{p}_{1}$ and $\hat{p}_{2}$ may be evaluated as
 \begin{equation} 
-\left<z_{1}(f)z_{2}^{*}(f')\right> = \frac{1}{2}S_{h}(f)\delta (f-f')\Gamma(f)
+\left<\tilde{z}_{1}(f)\tilde{z}_{2}^{*}(f')\right> = \frac{1}{2}S_{h}(f)\delta (f-f')\Gamma(f)
 \end{equation}

### 2014-07-07 Introduction.tex
PSD derivation

@@
 \end{equation}
-We will work with the Fourier transform of the signal, using the conventions that
+We shall work with the Fourier transform of the signal, using the conventions that
 \begin{eqnarray} \label{eq:defFourier}
-\tilde{x}(f) &= {\mathcal{F}}\left\{x(t)\right\} &= \int_{-\infty}^{\infty}\mathrm{d}t\;x(t)\exp (-2\pi \rmi f t ) \; , \\
-x(t) &= {\mathcal{F}}^{-1}\left\{\tilde{x}(f)\right\} &= \int_{-\infty}^{\infty}\mathrm{d}f\;\tilde{x}(f)\exp (2\pi \rmi f t ) \; .
+\tilde{x}(f) &= {\mathcal{F}}\left\{x(t)\right\}(f) &= \int_{-\infty}^{\infty}\mathrm{d}t\;x(t)\exp (-2\pi \rmi f t ) \; , \\
+x(t) &= {\mathcal{F}}^{-1}\left\{\tilde{x}(f)\right\}(t) &= \int_{-\infty}^{\infty}\mathrm{d}f\;\tilde{x}(f)\exp (2\pi \rmi f t ) \; .
 \end{eqnarray}
@@
 
-When integrated over all positive frequencies, the PSD gives the mean square noise amplitude. Starting with taking the \emph{time} average of the square of the detector noise:
+When integrated over all positive frequencies, the PSD gives the mean square noise amplitude. Starting by taking the time average of the square of the detector noise:
 \begin{eqnarray}\label{eq:meansquare}
-\overline{n(t)^{2}} &= \lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T}\mathrm{d}t\; n(t) ^{2} \nonumber \\
- &= \lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T}\mathrm{d}t\; n(t)n^{*}(t) \nonumber \\
- &= \lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T}\mathrm{d}t\; \int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\tilde{n}(f)\tilde{n}^{*}(f')e^{2\pi \rmi ft }e^{-2\pi \rmi f't } \nonumber\\
- &= \lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T}\mathrm{d}t\; \int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;{\cal{F}}\left\{ n(\tau) \right\}(f){\cal{F}}\left\{ n(\tau) \right\}(f')^{*}e^{2\pi \rmi ft }e^{-2\pi \rmi f't }
+\overline{\left|n(t)\right|^{2}} &= \lim_{T\rightarrow\infty}\frac{1}{2T} & \int_{-T}^{T}\mathrm{d}t\; n(t)n^{*}(t) \\
+ &= \lim_{T\rightarrow\infty}\frac{1}{2T} & \int_{-T}^{T}\mathrm{d}t\; \int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\tilde{n}(f)\tilde{n}^{*}(f')\exp\left(2\pi \rmi ft\right)\exp\left(-2\pi \rmi f't\right) \nonumber\\
+ &= \lim_{T\rightarrow\infty}\frac{1}{2T} & \int_{-T}^{T}\mathrm{d}t\; \int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;{\mathcal{F}}\left\{ n(\tau) \right\}(f)\left[{\mathcal{F}}\left\{ n(\tau) \right\}(f')\right]^{*} \nonumber \\*
+ & & \times \exp\left(2\pi \rmi ft\right)\exp\left(-2\pi \rmi f't\right)
 \end{eqnarray}
-Where in the penultimate line we have used the definition of the Fourier transform, (\ref{eq:defFourier}), and in the last line we have just changed notation.
-A property of Fourier transforms is that in the time domain a translation by amount $t$ translates to a rotation in phase by $2\pi ft$ in the frequency domain; if ${\mathcal{F}}\left\{ n(\tau) \right\} = \tilde{n}(f)$ then ${\mathcal{F}}\left\{ n(\tau-t) \right\} = \tilde{n}(f)\exp (2\pi \rmi ft )$.
-Therefore the exponential factors in (\ref{eq:meansquare}) may be absorbed as a time-translation of the noise by an amount $t$
+Where we have substituted in using the definitions of the Fourier transform and its inverse.
+A property of Fourier transforms is that in the a time-domain translation by amount $t$ translates to a frequency-domain phase change $2\pi ft$; if ${\mathcal{F}}\left\{ n(\tau) \right\}(f) = \tilde{n}(f)$, then ${\mathcal{F}}\left\{ n(\tau-t) \right\}(f) = \tilde{n}(f)\exp (2\pi \rmi ft )$.
+Therefore, the exponential factors in (\ref{eq:meansquare}) may be absorbed as
 \begin{eqnarray}
-\overline{n(t)^{2}} & = \lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T}\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;{\cal{F}}\left\{ n(\tau-t) \right\}(f){\cal{F}}\left\{ n(\tau-t) \right\}(f')^{*} \, .
+\overline{\left|n(t)\right|^{2}} &=& \lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T} \mathrm{d}t\:\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;{\mathcal{F}}\left\{ n(\tau-t) \right\}(f)\left[{\mathcal{F}}\left\{ n(\tau-t) \right\}(f')\right]^{*} \, .
 \end{eqnarray}
-Using the ergodic property of the system described above the time average (denoted by $\overline{(\ldots)}$) may now be changed into an ensemble average (denoted by $\left<\ldots\right>$).
-The noise is stationary so it's expectation value is unchanged by the time translation performed above. Therefore using (\ref{eq:psd}) gives
+Since the noise is a randomly varying signal, we can use the ergodic principle to equate a time average, denoted by $\overline{(\ldots)}$, with an ensemble average, denoted by $\left<\ldots\right>$.
+The noise is stationary, consequently, its expectation value is unchanged by time-translation performed above.
+Therefore,
 \begin{eqnarray}\label{eq:meansquare1}
-&=\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\left<\tilde{n}(f)\tilde{n}^{*}(f')\right>\nonumber \\
-&=\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\frac{1}{2}S_{n}(f)\delta(f-f') \nonumber \\
-& = \int_{0}^{\infty}\mathrm{d}f\; S_{n}(f)
+\overline{\left|n(t)\right|^{2}} &=& \int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\left<\tilde{n}(f)\tilde{n}^{*}(f')\right> \\
+&=& \int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\frac{1}{2}S_{n}(f)\delta(f-f') \nonumber \\
+&=& \int_{0}^{\infty}\mathrm{d}f\; S_{n}(f),
 \end{eqnarray}
+using (\ref{eq:psd}).
 
@@
 \begin{equation}\label{eq:conv}
-\left(s*K\right)(\tau) = \int_{-\infty}^{\infty}\mathrm{d}t\;(n(t)+h(t))K(t-\tau) \approx \mathcal{S} + \mathcal{N} \; .
+\left(s*K\right)(\tau) = \int_{-\infty}^{\infty}\mathrm{d}t\;\left[h(t)+n(t)\right]K(t-\tau) \approx \mathcal{S} + \mathcal{N} \; .
 \end{equation}
@@
 \begin{equation}\label{eq:SNRinnerprod} 
-\varrho^{2} = \frac{\mathcal{S}^{2}}{\mathcal{N}^{2}}= \frac{\innerprod{\frac{1}{2}S_{n}(f)\tilde{K}(f)}{\tilde{h}}^{2}}{\innerprod{\frac{1}{2}S_{n}(f)\tilde{K}(f)}{\frac{1}{2}S_{n}(f)\tilde{K}(f)}} \;,
+\varrho^{2} = \frac{\mathcal{S}^{2}}{\mathcal{N}^{2}}= \frac{\innerprod{\frac{1}{2}S_{n}(f)\tilde{K}(f)}{\tilde{h}(f)}^{2}}{\innerprod{\frac{1}{2}S_{n}(f)\tilde{K}(f)}{\frac{1}{2}S_{n}(f)\tilde{K}(f)}} \;,
 \end{equation}
-where we have introduced the inner product between signal $A$ and $B$ as \citep{Finn1992}
-\begin{equation}\label{eq:snrinnerprod} \left( \tilde{A}(f) | \tilde{B}(f) \right) = 4\Re\left\{\int_{0}^{\infty}\mathrm{d}f\;\frac{\tilde{A}^{*}(f)\tilde{B}(f)}{S_{n}(f)}\right\} \; .\end{equation}
+where we have introduced the inner product between signal $\tilde{A}$ and $\tilde{B}$ as \citep{Finn1992}
+\begin{equation}\label{eq:snrinnerprod} \innerprod{\tilde{A}(f)}{\tilde{B}(f)} = 4\Re\left\{\int_{0}^{\infty}\mathrm{d}f\;\frac{\tilde{A}^{*}(f)\tilde{B}(f)}{S_{n}(f)}\right\} \; .\end{equation}
 The optimum filter maximises the SNR in (\ref{eq:SNRinnerprod}). From the Cauchy--Schwarz inequality, it follows that the optimum filter is
@@
 \begin{equation}
-\varrho^2 = \int_0^\infty\mathrm{d}f \frac{4| \tilde{h}(f)|^{2}}{S_n(f)} = \innerprod{h}{h}.
+\varrho^2 = \int_0^\infty\mathrm{d}f \frac{4| \tilde{h}(f)|^{2}}{S_n(f)} = \innerprod{\tilde{h}(f)}{\tilde{h}(f)}.
 \label{eq:traditionalSNR} 
@@
 \begin{equation}\label{eq:crit-density}
-\rho_{\mathrm{c}}=3H_{0}^{2}/8\pi G \;,
+\rho_{\mathrm{c}}=\frac{3H_{0}^{2}}{8\pi G} \;,
 \end{equation}

### 2014-07-08 Detectors.tex
PSD derivation
typos

@@
 \end{equation}
-where the response functions inherit their angular dependence from the choice of coordinates,
+where the response functions inherit their angular dependence from the choice of coordinates
 \begin{eqnarray}\label{eq:responsefuncs}
@@
 \end{eqnarray}
-The response function of a two-arm interferometric detector is quadrupolar, an example is plotted in figure \ref{fig:LIGO}.
 
@@
 
+The response function of a two-arm interferometric detector is quadrupolar, an example is plotted in figure \ref{fig:LIGO}.
 Throughout this paper detector sensitivity refers to the polarisation and sky averaged sensitivity $F$, where
@@
 \begin{equation}
-\Gamma_{0} = \frac{1}{2}+\frac{3x}{2}\left(\ln x -\frac{1}{6}\right)
-\end{equation}
-where
-\begin{equation}
-x = \frac{1-\cos\theta}{2}\; .
+\Gamma_{0} = \frac{1}{2}+\frac{3x}{2}\left(\ln x -\frac{1}{6}\right)\quad \textrm{where}\quad x = \frac{1-\cos\theta}{2}\; .
 \end{equation}
@@
  \includegraphics[trim=0cm 0cm 0cm 0cm, width=0.6\textwidth]{HnDcurve.pdf}
- \caption{The Hellings and Downs curve, the correlation between two pulsars separated on the sky by an angle $\theta$ \citep{HellingsDowns}.}
+ \caption{The \citet{HellingsDowns} curve, the correlation between two pulsars separated on the sky by an angle $\theta$.}
  \label{fig:HnD}

### 2014-07-08 Introduction.tex
typos

@@
  &= \lim_{T\rightarrow\infty}\frac{1}{2T} & \int_{-T}^{T}\mathrm{d}t\; \int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;{\mathcal{F}}\left\{ n(\tau) \right\}(f)\left[{\mathcal{F}}\left\{ n(\tau) \right\}(f')\right]^{*} \nonumber \\*
- & & \times \exp\left(2\pi \rmi ft\right)\exp\left(-2\pi \rmi f't\right)
+ & & \times \exp\left(2\pi \rmi ft\right)\exp\left(-2\pi \rmi f't\right)\;.
 \end{eqnarray}
@@
 \end{eqnarray}
-Since the noise is a randomly varying signal, we can use the ergodic principle to equate a time average, denoted by $\overline{(\ldots)}$, with an ensemble average, denoted by $\left<\ldots\right>$.
+Since the noise is a randomly varying signal, we can use the ergodic principle to equate a time average, denoted by $\overline{\left(\ldots\right)}$, with an ensemble average, denoted by $\left<\ldots\right>$.
 The noise is stationary, consequently, its expectation value is unchanged by time-translation performed above.
@@
 \end{equation}
-a corresponding expression for the noise can be formulated by replacing $S_h(F)$ with $S_{n}(f)$.
+a corresponding expression for the noise can be formulated by replacing $S_h(f)$ with $S_{n}(f)$.
 

### 2014-07-08 Detectors.tex
Updated the reference to different PTAs and added in urls to their respective websites

@@
 
-The PTAs currently in operation are the European Pulsar Timing Array \citep[EPTA;][]{Ferdman2010}, the Parkes Pulsar Timing Array \citep[PPTA;][]{Manchester2013} based in Australia, and the North American Nanohertz Observatory for Gravitational waves \citep[NANOGrav;][]{Jenet2009}.
+The PTAs currently in operation are the European Pulsar Timing Array (EPTA\footnote{\url{http://www.epta.eu.org/}}, \cite{eptareview2013}), the Parkes Pulsar Timing Array (PPTA\footnote{\url{http://www.atnf.csiro.au/research/pulsar/ppta/}} \cite{parkesreview2013}) based in Australia, and the North American Nanohertz Observatory for Gravitational waves (NANOGrav\footnote{\url{http://nanograv.org/}} \cite{nanogravreview2013}).
 There are published limits on the amplitude of the stochastic background from all three detectors: the most recent from EPTA is \citet{Haasteren}, from PPTA is \citet{Shannon2013} and from NANOGrav is \citet{Arzoumanian2014}.
@@
 
-Combining the existing arrays would yield a single PTA using approximately 4 times as many pulsars; this consortium of consortia is known as the International Pulsar Timing Array \citep[IPTA;][]{Hobbs2010}.
+Combining the existing arrays would yield a single PTA using approximately 4 times as many pulsars; this consortium of consortia is known as the International Pulsar Timing Array (IPTA\footnote{\url{http://www.ipta4gw.org/}} \cite{iptareview2013}).
 The IPTA curves plotted in the figures are based on 20 pulsars timed for twice as long as the current EPTA.

### 2014-07-08 Introduction.tex
Updated the reference to different PTAs and added in urls to their respective websites

@@
 \end{equation}
-The signal contribution $\mathcal{S}$ is defined as the expectation of the convolution in (\ref{eq:conv}) when a signal is present, maximised by varying the offset $\tau$.
-\cplb{I think this needs expanding.} Since the expectation of pure noise is zero it, follows that
+The signal contribution $\mathcal{S}$ is defined as the expectation of the convolution in (\ref{eq:conv}) when a signal is present, maximised by varying the offset to achieve the best overlap with the data.
+Since the expectation of pure noise is zero it, follows that
 \begin{equation}

### 2014-07-08 Detectors.tex
Odds and Ends

--- 2014-07-07 Detectors.tex
@@
 They all simulate free-floating test masses by suspending a mass from a pendulum system with natural frequency much greater than that of the GW.
-Their sensitivity curves include narrow lines, which arise from noise sources in the instrument including resonances in the suspension system and electrical noise at multiples of $60~\mathrm{Hz}$: these have been removed in the \ref{app:a} figures for clarity.
-The detectors fall broadly into three categories: first-generation detectors, which have already operated; second-generation detectors currently under construction, and third-generation detectors at the planning stage.
+Their sensitivity curves include narrow lines that arise from noise sources in the instrument, including resonances in the suspension system and electrical noise at multiples of $60~\mathrm{Hz}$: these have been removed in the \ref{app:a} figures for clarity.
+The detectors fall broadly into three categories: first-generation detectors, which have already operated; second-generation detectors currently under construction; and third-generation detectors at the planning stage.
 
-The most notable ground-based detectors are LIGO and Virgo \citep{Abramovici1992,Acernese2008a}.
-These work in collaboration, supported by GEO600 \citep{Willke2002}.
+The most notable ground-based detectors are LIGO and Virgo, which work in collaboration, supported by GEO600.
 LIGO and Virgo have completed science runs as first-generation detectors, now referred to as Initial LIGO (iLIGO) and Initial Virgo (iVirgo).
-They are now being upgraded to second-generation detectors, advanced LIGO (aLIGO) and advanced Virgo (AdV).
-LIGO has two observatories, one at Hanford, Washington, which has two detectors, and another at Livingston, Louisiana.
+They are currently being upgraded to second-generation detectors, advanced LIGO (aLIGO) and advanced Virgo (AdV).
+LIGO has two observatories: one at Hanford, Washington, which has two detectors; and another at Livingston, Louisiana.
 There is an agreement to move one of the upgraded Hanford detector systems to a location in India \citep{LIGO-India,Unnikrishnan2013}.
@@
 Its successor is the Kamioka Gravitational Wave Detector (KAGRA), formerly the Large-scale Cryogenic Gravitational wave Telescope (LCGT), which is located underground in the Kamioka mine.
-It employs more sophisticated noise-reduction techniques than LIGO or Virgo, such as cryogenic cooling \citep{2012CQGra..29l4007S}.
+It employs more sophisticated noise-reduction techniques than LIGO or Virgo, such as cryogenic cooling.
 
-The Einstein Telescope (ET) is an ambitious proposal to construct an underground third-generation detector \citep{Punturo2010,Sathyaprakash2012}.
+The Einstein Telescope (ET) is an ambitious proposal to construct an underground third-generation detector.
 Its location would provide shielding from seismic noise, allowing it to observe frequencies of $10$--$10^4~\mathrm{Hz}$.
@@
 
-We use an interpolation to the data published on \url{https://wwwcascina.virgo.infn.it/advirgo/} (2013) for the AdV sensitivity curve, an interpolation to the data for version D of the detector published on \url{http://gwcenter.icrr.u-tokyo.ac.jp/en/researcher/parameter} (2013) for KAGRA and analytic fits to the sensitivity curves from \citet{Sathyaprakash} for the remaining detectors.
+We use an interpolation to the data published on \url{https://wwwcascina.virgo.infn.it/advirgo/} (2013) for the AdV sensitivity curve, an interpolation to the data for version D of the KAGRA detector published on \url{http://gwcenter.icrr.u-tokyo.ac.jp/en/researcher/parameter} (2013) and analytic fits to the sensitivity curves from \citet{Sathyaprakash} for the remaining detectors.
 
@@
 These missions are designed to probe the decihertz region of the GW spectrum; they are considerably more ambitious than the LISA or eLISA mission and their launches will be further into the future.
-We use a simple analytic fit to the sensitivity curve for ALIA \cplb{[citation needed]}, while for DECIGO and BBO, fits to the sensitivity curves given by \citet{2011PhRvD..83d4011Y} are used.
+We use a simple analytic fit to the sensitivity curve for ALIA~\citep{bender_possible_2013}, while for DECIGO and BBO, fits to the sensitivity curves given by \citet{2011PhRvD..83d4011Y} are used.
 

### 2014-07-08 Introduction.tex
Odds and Ends

--- 2014-07-07 Introduction.tex
@@
 where angle brackets $\left<\ldots\right>$ denote an ensemble average over many noise realisations \citep{Cutler1994}.
-In reality, we have only a realisation to work with, however the ensemble average can replaced by a time average for stationary stochastic noise.
+In reality, we have only a single realisation to work with, however the ensemble average can replaced by a time average for stationary stochastic noise.
 The procedure is to measure the noise over a sufficiently long duration $T$ and then compute the Fourier transform $\tilde{n}(f)$ with a frequency resolution $\Delta f = T^{-1}$; this is repeated many times to give an average.
@@
  &= \lim_{T\rightarrow\infty}\frac{1}{2T} & \int_{-T}^{T}\mathrm{d}t\; \int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;{\mathcal{F}}\left\{ n(\tau) \right\}(f)\left[{\mathcal{F}}\left\{ n(\tau) \right\}(f')\right]^{*} \nonumber \\*
- & & \times \exp\left(2\pi \rmi ft\right)\exp\left(-2\pi \rmi f't\right)
+ & & \times \exp\left(2\pi \rmi ft\right)\exp\left(-2\pi \rmi f't\right),
 \end{eqnarray}
-Where we have substituted in using the definitions of the Fourier transform and its inverse.
-A property of Fourier transforms is that in the a time-domain translation by amount $t$ translates to a frequency-domain phase change $2\pi ft$; if ${\mathcal{F}}\left\{ n(\tau) \right\}(f) = \tilde{n}(f)$, then ${\mathcal{F}}\left\{ n(\tau-t) \right\}(f) = \tilde{n}(f)\exp (2\pi \rmi ft )$.
+where we have substituted in using the definitions of the Fourier transform and its inverse.
+A property of Fourier transforms is that a time-domain translation by amount $t$ is equivalent to a frequency-domain phase change $2\pi ft$; if ${\mathcal{F}}\left\{ n(\tau) \right\}(f) = \tilde{n}(f)$, then ${\mathcal{F}}\left\{ n(\tau-t) \right\}(f) = \tilde{n}(f)\exp (2\pi \rmi ft )$.
 Therefore, the exponential factors in (\ref{eq:meansquare}) may be absorbed as

### 2014-07-08 Plots.tex
Ending
Odds and Ends

@@
 \section{Sensitivity curves}\label{app:a}
-The plots in this section show all of the detectors and sources described in the text.
-Interactive versions of these plots may be viewed on-line, \url{http://www.ast.cam.ac.uk/~rhc26/sources/}.
+The plots in this section show all of the detectors and sources described in the main text.
+Interactive versions of these plots, allowing for removal of any of the curves, may be created and downloaded on-line, \url{http://www.ast.cam.ac.uk/~rhc26/sources/}.
 The detector noise curves all have their resonance spikes removed for clarity.
-\begin{figure}[h!]
+\begin{figure}
  \centering
@@
 \end{figure}
-\begin{figure}[h!]
+\begin{figure}
  \centering

### 2014-07-08 Sources.tex
Odds and Ends

@@
 The event rate for supernovae is approximately $\rho_{\mathrm{SN}} = 5\times10^{-4}~\mathrm{Mpc^{-3}\,yr^{-1}}$.
-The boxes plotted in \ref{app:a} correspond to a distance $r=3~\mathrm{Mpc}$ with the frequency range quoted above.
+The boxes plotted in \ref{app:a} correspond to a distance $r=300~\mathrm{kpc}$ with the frequency range quoted above.
 The LIGO and Virgo detectors have already placed bounds on the event rate for these sources \citep{Bursts}.
@@
 This is currently uncertain \citep{Lattimer2012}.
-Several known pulsars could be sources for the advanced detectors and upper limits from the initial detectors help to constrain the deformations \citep{Aasi2014a}.
-The boxes plotted in \ref{app:a} is the same as for supernovae \cplb{for reasons I do not entirely understand}.
+Several known pulsars could be sources for the advanced detectors and upper limits from the initial detectors help to constrain the deformations.
+The boxes plotted in \ref{app:a} correspond to the upper limits placed on a GW signal from the Crab pulsar \citep{Aasi2014a}, extrapolated across a frequency range between $(20$--$10^{3})~\mathrm{Hz}$.
 

### 2014-07-08 Introduction.tex

--- 2014-07-08 Introduction.tex
@@
  &= \lim_{T\rightarrow\infty}\frac{1}{2T} & \int_{-T}^{T}\mathrm{d}t\; \int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;{\mathcal{F}}\left\{ n(\tau) \right\}(f)\left[{\mathcal{F}}\left\{ n(\tau) \right\}(f')\right]^{*} \nonumber \\*
- & & \times \exp\left(2\pi \rmi ft\right)\exp\left(-2\pi \rmi f't\right)\;.
+ & & \times \exp\left(2\pi \rmi ft\right)\exp\left(-2\pi \rmi f't\right)\;,
 \end{eqnarray}
-Where we have substituted in using the definitions of the Fourier transform and its inverse.
-A property of Fourier transforms is that in the a time-domain translation by amount $t$ translates to a frequency-domain phase change $2\pi ft$; if ${\mathcal{F}}\left\{ n(\tau) \right\}(f) = \tilde{n}(f)$, then ${\mathcal{F}}\left\{ n(\tau-t) \right\}(f) = \tilde{n}(f)\exp (2\pi \rmi ft )$.
+where we have substituted in using the definitions of the Fourier transform and its inverse.
+A property of Fourier transforms is that a time-domain translation by amount $t$ is equivalent to a frequency-domain phase change $2\pi ft$; if ${\mathcal{F}}\left\{ n(\tau) \right\}(f) = \tilde{n}(f)$, then ${\mathcal{F}}\left\{ n(\tau-t) \right\}(f) = \tilde{n}(f)\exp (2\pi \rmi ft )$.
 Therefore, the exponential factors in (\ref{eq:meansquare}) may be absorbed as
@@
 \begin{equation}\label{eq:SNRinnerprod} 
-\varrho^{2} = \frac{\mathcal{S}^{2}}{\mathcal{N}^{2}}= \frac{\innerprod{\frac{1}{2}S_{n}(f)\tilde{K}(f)}{\tilde{h}(f)}^{2}}{\innerprod{\frac{1}{2}S_{n}(f)\tilde{K}(f)}{\frac{1}{2}S_{n}(f)\tilde{K}(f)}} \;,
+\varrho^{2} = \frac{\mathcal{S}^{2}}{\mathcal{N}^{2}}= \frac{\innerprod{\frac{1}{2}S_{n}(f)\tilde{K}(f)}{\tilde{h}(f)}^{2}}{\innerprod{\frac{1}{2}S_{n}(f)\tilde{K}(f)}{\frac{1}{2}S_{n}(f)\tilde{K}(f)}},
 \end{equation}

### 2014-07-08 Plots.tex
redone the figures in appendix

@@
 The plots in this section show all of the detectors and sources described in the main text.
-Interactive versions of these plots, allowing for removal of any of the curves, may be created and downloaded on-line, \url{http://www.ast.cam.ac.uk/~rhc26/sources/}.
+Clearer, interactive versions of these plots, allowing for removal of any of the curves, may be created and downloaded on-line, \url{http://www.ast.cam.ac.uk/~rhc26/sources/}.
 The detector noise curves all have their resonance spikes removed for clarity.
@@
  \centering
- \includegraphics[trim=0cm 0cm 0cm 0cm, width=0.85\textwidth]{figure1.pdf}
+ \includegraphics[trim=0cm 0cm 0cm 0cm, width=0.8\textwidth]{figure1.pdf}
  \caption{A plot of characteristic strain against frequency for a variety of detectors and sources.}
@@
  \centering
- \includegraphics[trim=0cm 0cm 0cm 0cm, width=0.85\textwidth]{figure2.pdf}
+ \includegraphics[trim=0cm 0cm 0cm 0cm, width=0.8\textwidth]{figure2.pdf}
  \caption{A plot of the square root of power spectral density against frequency for a variety of detectors and sources.}
@@
  \centering
- \includegraphics[trim=0cm 0cm 0cm 0cm, width=0.85\textwidth]{figure3.pdf}
+ \includegraphics[trim=0cm 0cm 0cm 0cm, width=0.8\textwidth]{figure3.pdf}
  \caption{A plot of the dimensionless energy density in GWs against frequency for a variety of detectors and sources.}

### 2014-07-09 Sources.tex
Changing rho to gamma for event rate ans rho is already energy density

@@
 The inspiral and merger of a pair of neutron stars is the primary target for ground based detectors.
-The expected event rate for this type of source is uncertain, but estimates centre around $\rho_{\mathrm{NS-NS}}=1.3\times 10^{-4}~\mathrm{Mpc^{-3}\,yr^{-1}}$ \citep{CBC}.
+The expected event rate for this type of source is uncertain, but estimates centre around $\gamma_{\mathrm{NS-NS}}=1.3\times 10^{-4}~\mathrm{Mpc^{-3}\,yr^{-1}}$ \citep{CBC}.
 Plotted in \ref{app:a} is a box with an amplitude such that it produces a ratio $h_\mathrm{c}/h_{n}=16$ for Advanced LIGO at peak sensitivity and a width between $(3$--$300)~\mathrm{Hz}$.
@@
 \begin{equation}
-h=8.9\times 10^{-21}\left( \frac{10~\mathrm{kpc}}{r} \right) \; .
+h_\mathrm{max} = 8.9\times 10^{-21}\left( \frac{10~\mathrm{kpc}}{r} \right) \; .
 \end{equation}
-The event rate for supernovae is approximately $\rho_{\mathrm{SN}} = 5\times10^{-4}~\mathrm{Mpc^{-3}\,yr^{-1}}$.
+The event rate for supernovae is approximately $\gamma_{\mathrm{SN}} = 5\times10^{-4}~\mathrm{Mpc^{-3}\,yr^{-1}}$.
 The boxes plotted in \ref{app:a} correspond to a distance $r=300~\mathrm{kpc}$ with the frequency range quoted above.

### 2014-07-10 Detectors.tex
myriad typos fixed

@@
 In this section we introduce the detector noise curves used in \ref{app:a}.
-We begin with with a description of the basic operation of detectors.
-We then discuss ground-based detectors, space-based detectors and PTAs.
-The latter function somewhat different than conventional interferometers, so we include a brief introduction to PTA analysis.
-References for the noise curve used for individual detectors can be found in the relevant subsections, further information about the detectors can be in discovered these.
+We begin with a description of the basic operation of detectors.
+We then discuss ground-based detectors, space-based detectors and PTAs in turn.
+The latter function somewhat differently than conventional interferometers, so we include a brief introduction to PTA analysis.
+References for the noise curves used for individual detectors can be found in the relevant subsections, further information about the detectors can be in discovered these.
 
@@
 
-All of the man-made detectors discussed in this section work on the principle of interferometry.
+All of the man-made detectors discussed in this section utilise the principle of interferometry.
 Such detectors work by taking a beam of monochromatic light and splitting it into two beams travelling at some angle to each other.
@@
 The ends of the cavity are, in the ideal case, freely floating test masses which move in response to a passing GW, this effect is measured by observing the changing interference pattern.
-% LIGO isn't Fabry-P\'{e}rot
 
@@
 \begin{eqnarray}
-H^{+}_{ij}&=p_{i}p_{j}-q_{i}q_{j} \nonumber \\
+H^{+}_{ij}&=p_{i}p_{j}-q_{i}q_{j} \; , \\
 H^{\times}_{ij}&=p_{i}q_{j}+q_{i}p_{j} \; .
@@
 There remains a freedom in the coordinates described, a rotation of $p^{i}$ and $q^{i}$ through an angle $\psi$ about $\hat{r}^{i}$ known as the polarization angle.
-For a single frequency component, the strain induced by GW may be written as
+For a single frequency component, the strain induced by a GW may be written as
 \begin{equation}\label{eq:hijxx}
@@
 If the binary is inspiralling, then the polarisation depends on still more parameters which need to be averaged over.
-These more complicated averages all have the property that they depend on both the detector and the source, hence they are unhelpful for our present purpose seperating the source amplitude from the detector sensitivity.
+These more complicated averages all have the property that they depend on both the detector and the source, hence they are unhelpful for our present purpose separating the source amplitude from the detector sensitivity.
 Additionally, the different averages do not work out to be so different from each other: \citet{1993PhRvD..47.2198F} calculated the sensitivity for a detector with the LIGO geometry averaged over the four angles ($\theta,\phi,\iota,\xi$) as $\sqrt{4/25}=0.4$ times peak sensitivity, which should be compared with the value $\sqrt{1/5}\approx 0.447$ above.
@@
 The laser arms in a LISA-like detector are not a cavity, the light only travels once along each arm.
-eLISA is a reduced version of LISA designed to probe the same frequency range, while proposals such as the Advanced Laser Interferometer Antenna (ALIA), Big Bang Observer (BBO) and Deci-hertz Interferometer GW Observatory (DECIGO) are designed to probe decihertz GWs.
+eLISA is a rescoped version of LISA designed to probe the same frequency range, while proposals such as the Advanced Laser Interferometer Antenna (ALIA), Big Bang Observer (BBO) and Deci-hertz Interferometer GW Observatory (DECIGO) are designed to probe decihertz GWs.
 
@@
 
-The instrumental noise curves for LISA by the analytic fit given by \citet{Sathyaprakash}, which we use for the plots in \ref{app:a}.
+The instrumental noise curves for LISA are approximated by the analytic fit given by \citet{Sathyaprakash}, which we use for the plots in \ref{app:a}.
 When observing individual sources with LISA there is an additional contribution to the noise from a background of unresolvable binaries.
 This is not included here as we consider the background as a source of GWs (see section \ref{sec:GB}).
-eLISA is a re-scoped version of the classic LISA mission, the main differences are shorter arms ($10^{9}~\mathrm{m}$ instead of $5\times 10^{9}~\mathrm{m}$), two laser arms instead of three, and a different orbit (drifting away from Earth instead of $20\degree$ Earth trailing).
+eLISA is a rescoped version of the classic LISA mission, the main differences are shorter arms ($10^{9}~\mathrm{m}$ instead of $5\times 10^{9}~\mathrm{m}$), two laser arms instead of three, and a different orbit (drifting away from Earth instead of $20\degree$ Earth trailing).
 The effect of these changes is a slightly reduced peak sensitivity and a shift to higher frequencies.
@@
 
-\subsection{Pulsar timing arrays (PTAs)}\label{sec:PTAgeneralproperties}
+\subsection{Pulsar timing arrays}\label{sec:PTAgeneralproperties}
 
@@
 \begin{equation} 
-\left<\tilde{z}_{1}(f)\tilde{z}_{2}^{*}(f')\right> = \frac{1}{2}S_{h}(f)\delta (f-f')\Gamma(f)
+\left<\tilde{z}_{1}(f)\tilde{z}_{2}^{*}(f')\right> = \frac{1}{2}S_{h}(f)\delta (f-f')\Gamma(f) \; ,
 \end{equation}
-where,
+where
 \begin{eqnarray}
@@
 \begin{equation}
-\Gamma_{0} = \frac{1}{2}+\frac{3x}{2}\left(\ln x -\frac{1}{6}\right)\quad \textrm{where}\quad x = \frac{1-\cos\theta}{2}\; .
+\Gamma_{0} = \frac{1}{2}+\frac{3x}{2}\left(\ln x -\frac{1}{6}\right) \; ,
 \end{equation}
+where $x = (1-\cos\theta)/{2}.$
 \begin{figure}
@@
 The next great advancement in radio astronomy shall come with the completion of the Square Kilometre Array \citep[SKA;][]{Dewdney2009}.
-This shall greatly increasing the sensitivity of pulsar timing \citep{Kramer2004}.
-The sensitivity curves plotted in \ref{app:a} for the SKA assumes an rms error on the timing residuals a factor of 10 better than current PTAs, and 20 pulsars timed for twice as long as the current EPTA.
+This shall greatly increase the sensitivity of pulsar timing \citep{Kramer2004}.
+The sensitivity curves plotted in \ref{app:a} for the SKA assume a RMS error on the timing residuals a factor of 10 better than current PTAs, and include 20 pulsars timed for twice as long as the current EPTA.

### 2014-07-10 Introduction.tex
myriad typos fixed

@@
 This will most likely be achieved by the advanced Laser Interferometer Gravitational-wave Observatory \citep[LIGO;][]{2010CQGra..27h4006H} and Virgo \citep{Accadia2011} detectors operating in the frequency range $(10$--$10^{3})~\mathrm{Hz}$.
-By the end of the decade, it is expected that pulsar timing arrays \citep[PTAs;][]{1990ApJ361300F} shall detect very low frequency GWs around $10^{-8}~\mathrm{Hz}$.
+By the end of the decade, it is expected that pulsar timing arrays \citep[PTAs;][]{1990ApJ361300F} shall also detect very low frequency GWs around $10^{-8}~\mathrm{Hz}$.
 Further into the future, space-based detectors, such as the evolved Laser Interferometer Space Antenna \citep[eLISA;][]{2012CQGra..29l4016A}, will probe GWs in the millihertz regime.
@@
 It is common practice to summarise this information on a sensitivity-curve plot.
-When producing these plots, it is desirable to have a consistent convention between detectors and sources which allows information about both to be plotted on the same graph.
+When producing these plots, it is desirable to have a consistent convention between detectors and sources that allows information about both to be plotted on the same graph.
 Ideally, the detectors and sources are represented in such a way that the relative detectability of the signals is immediately apparent from the plot.
@@
 
-When discussing the loudness of sources and the sensitivity of detectors there are three commonly used parametrizations based upon the strain: the characteristic strain, the PSD and the spectral energy density.
+When discussing the loudness of sources and the sensitivity of detectors there are three commonly used parametrizations based upon the strain: the characteristic strain, the power spectral density (PSD) and the spectral energy density.
 We aim to disambiguate these three and give a concrete comparison of different detectors.
@@
 The sensitivity of a detector to these depends upon the relative orientations of the source and detector.
-The output of a gravitational wave detector $s(t)$ contains a superposition of noise $n(t)$ and (possibly) the signal $h(t)$,
+The output of a gravitational wave detector $s(t)$ contains a superposition of noise $n(t)$ and (possibly) a signal $h(t)$,
 \begin{equation}
@@
 
-For simplicity, it is assumed that the noise in the GW detector is stationary and Gaussian (with zero mean); under these assumptions the noise is fully characterised via the one-sided noise power spectral density (PSD) $S_{n}(f)$,
+For simplicity, it is assumed that the noise in the GW detector is stationary and Gaussian (with zero mean); under these assumptions the noise is fully characterised via the one-sided noise PSD $S_{n}(f)$,
 \begin{equation}\label{eq:psd}
@@
 When integrated over all positive frequencies, the PSD gives the mean square noise amplitude. Starting by taking the time average of the square of the detector noise:
-\begin{eqnarray}\label{eq:meansquare}
+\begin{eqnarray}
 \overline{\left|n(t)\right|^{2}} &= \lim_{T\rightarrow\infty}\frac{1}{2T} & \int_{-T}^{T}\mathrm{d}t\; n(t)n^{*}(t) \\
@@
  &= \lim_{T\rightarrow\infty}\frac{1}{2T} & \int_{-T}^{T}\mathrm{d}t\; \int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;{\mathcal{F}}\left\{ n(\tau) \right\}(f)\left[{\mathcal{F}}\left\{ n(\tau) \right\}(f')\right]^{*} \nonumber \\*
- & & \times \exp\left(2\pi \rmi ft\right)\exp\left(-2\pi \rmi f't\right)\;,
+\label{eq:meansquare} & & \times \exp\left(2\pi \rmi ft\right)\exp\left(-2\pi \rmi f't\right)\;,
 \end{eqnarray}
@@
 Since the noise is a randomly varying signal, we can use the ergodic principle to equate a time average, denoted by $\overline{\left(\ldots\right)}$, with an ensemble average, denoted by $\left<\ldots\right>$.
-The noise is stationary, consequently, its expectation value is unchanged by time-translation performed above.
+The noise is stationary, consequently, its expectation value is unchanged by the time-translation performed above.
 Therefore,
@@
 Given a detector output, the challenge is to extract the signal.
-There is a well known solution to this problem which involves constructing a Wiener optimal filter \citep{Wiener49}.
-Let $K(t)$ be the filter function with Fourier transform $\tilde{K}(f)$.
+There is a well known solution to this problem that involves constructing a Wiener optimal filter \citep{Wiener49}.
+Let $K(t)$ be a real filter function with Fourier transform $\tilde{K}(f)$.
 Convolving this with the detector output gives a contribution from the signal and a contribution from the noise
@@
 \end{equation}
-This is the Wiener filter, it may be multiplied by an arbitrary constant since this does not change the SNR. Using this form for $\tilde{K}(f)$, the squared SNR is
+This is the Wiener filter, which may be multiplied by an arbitrary constant since this does not change the SNR. Using this form for $\tilde{K}(f)$, the squared SNR is
 \begin{equation}
@@
 Whilst the magnitude of the Fourier transform of the signal $|\tilde{h}(f)|$ provides a simple quantification of the GW amplitude as a function of frequency, it has one main deficiency.
-For an inspiralling source the instantaneous amplitude can be orders of magnitude below the noise level in the detector; however, as the signal continues over many orbits, the SNR can be integrated up to a detectable level.
+For an inspiralling source the instantaneous amplitude can be orders of magnitude below the noise level in a detector; however, as the signal continues over many orbits, the SNR can be integrated up to a detectable level.
 It is useful to have a quantification of the GW amplitude that accounts for this effect; three such methods are described in the following subsections.
@@
 \end{equation}
-where we have used (\ref{eq:strain-hn}); by analogy, we can define an equivalent for for source amplitudes
+where we have used (\ref{eq:strain-hn}); by analogy, we can define an equivalent for source amplitudes
 \begin{equation}
@@
 The PSD has the nice property, demonstrated in (\ref{eq:meansquare1}), that integrated over all positive frequencies it gives the mean square amplitude of the signal in the detector.
-However, in one important regard it is less appealing than characteristic strain: the height of the source above the detector curve is no longer is directly related to the SNR.
+However, in one important regard it is less appealing than characteristic strain: the height of the source above the detector curve is no longer directly related to the SNR.
 
@@
 \end{equation}
-where the anle brackets denote averaging over several wavelengths or periods. The energy density $\rho c^{2}$ is given by the $T_{00}$ component of this tensor. Consequently
+where the angle brackets denote averaging over several wavelengths or periods. The energy density $\rho c^{2}$ is given by the $T_{00}$ component of this tensor. Consequently
 \begin{eqnarray}

### 2014-07-10 Sources.tex
myriad typos fixed

@@
 Descriptions of each source are given in the following sections.
-Sources with short durations (i.e., burst sources) and sources which evolve in time over much longer timescales than our observations are drawn with flat-topped boxes for $h_\mathrm{c}(f)$.
+Sources with short durations (i.e., burst sources) and sources that evolve in time over much longer timescales than our observations are drawn with flat-topped boxes for $h_\mathrm{c}(f)$.
 Inspiralling binaries, or stochastic backgrounds of binaries, are drawn with a sloping top proportional to $f^{-2/3}$ for $h_\mathrm{c}(f)$, which is result derived in section \ref{sec:stoch}.
-The width of the box gives the range of frequencies sources of a given type can have a while remaining at a detectable amplitude.
+The width of the box gives the range of frequencies sources of a given type can have while remaining at a detectable amplitude.
 
-\subsection{Sources for ground based detectors}
+\subsection{Sources for ground-based detectors}
 
@@
 
-The inspiral and merger of a pair of neutron stars is the primary target for ground based detectors.
+The inspiral and merger of a pair of neutron stars is the primary target for ground-based detectors.
 The expected event rate for this type of source is uncertain, but estimates centre around $\gamma_{\mathrm{NS-NS}}=1.3\times 10^{-4}~\mathrm{Mpc^{-3}\,yr^{-1}}$ \citep{CBC}.
-Plotted in \ref{app:a} is a box with an amplitude such that it produces a ratio $h_\mathrm{c}/h_{n}=16$ for Advanced LIGO at peak sensitivity and a width between $(3$--$300)~\mathrm{Hz}$.
+Plotted in \ref{app:a} are boxes with amplitudes such that a ratio $h_\mathrm{c}/h_{n}=16$ is produced for Advanced LIGO at peak sensitivity and a width between $(3$--$300)~\mathrm{Hz}$, corresponding to expected observable frequencies.
 
@@
 
-\subsection{Sources for space based detectors}
+\subsection{Sources for space-based detectors}
 
-For a review of the GW sources for space based missions see, for example, \citet{Amaro-Seoane-et-al}, \citet{Gairetal} or \citet{eLISAyellowbook}.
+For a review of the GW sources for space-based missions see, for example, \citet{Amaro-Seoane-et-al}, \citet{Gairetal} or \citet{eLISAyellowbook}.
 
@@
 Plotted in \ref{app:a} are boxes with a ratio $h_\mathrm{c}/h_{n}=100$ for eLISA at its peak sensitivity.
-The range of frequencies plotted is $(3\times 10^{-4}$--$3\times 10^{-1})~\mathrm{Hz}$, this corresponds to circular binaries in the mass range quoted above.
+The range of frequencies plotted is $(3\times 10^{-4}$--$3\times 10^{-1})~\mathrm{Hz}$; this corresponds to circular binaries in the mass range quoted above.
 
@@
 There is extreme uncertainty in the event rate for EMRIs due to the poorly constrained astrophysics in galactic centres \citep[e.g.,][]{Merritt2011}; the best guess estimate is around $25$ events per year with eLISA with SNR $\ge 20$ \citep{TheGravitationalUniverse}.
-The boxes plotted in \ref{app:a} have a characteristic strain of $h_\mathrm{c}=3\times 10^{-20}$ at $10^{-2}~\mathrm{Hz}$, which corresponds to a $10\Msun$ BH inspiralling into a $10^{6}\Msun$ black hole at a luminosity distance of $1~\mathrm{Gpc}$.
+The boxes plotted in \ref{app:a} have a characteristic strain of $h_\mathrm{c}=3\times 10^{-20}$ at $10^{-2}~\mathrm{Hz}$, which corresponds to a $10\Msun$ black hole inspiralling into a $10^{6}\Msun$ black hole at a luminosity distance of $1~\mathrm{Gpc}$.
 The frequency width of the box is somewhat unknown; EMRI events can occur into a black hole of any mass, and hence EMRIs can, in principal, occur at any frequency, so the boxes in \ref{app:a} are drawn with a width comparable to that of the LISA sensitivity curve.

### 2014-07-10 Typesofsource.tex
myriad typos fixed

@@
 The classic example of a well-modelled source is the inspiral of two compact objects, this is discussed in section \ref{sec:insp}.
-Stochastic backgrounds can either be formed from many overlapping sources, which individually could be modelable, or from some intrinsically random process, these are discussed in section \ref{sec:stoch}.
+Stochastic backgrounds can either be formed from many overlapping sources, which could be modelled individually, or from some intrinsically random process, these are discussed in section \ref{sec:stoch}.
 An example of an unmodelled (or poorly-modelled) transient source is a supernova; searches for signals of this type are often called burst searches and are discussed in section \ref{sec:bursts}.
@@
 
-Another important source of GWs are stochastic backgrounds, which can be produced from a large population of unresolvable sources.
-These can be at cosmological distances, then it is necessary to distinguish the frequency in the source rest frame $f_{\mathrm{r}}$ from the measured frequency $f$; the two are related through the redshift $z$ via $f_{\mathrm{r}}=(1+z)f$.
+Another important source of GWs is that of stochastic backgrounds, which can be produced from a large population of unresolvable sources.
+These can be at cosmological distances, where it is necessary to distinguish the frequency in the source rest frame $f_{\mathrm{r}}$ from the measured frequency $f$; the two are related through the redshift $z$ via $f_{\mathrm{r}}=(1+z)f$.
 
-The comoving number density of sources $\nu$ producing the background is also a function of redshift; if the sources producing the background are all in the local universe, then simply set $\nu(z) = \nu_0\delta (z)$ and replace $d_{\mathrm{L}}(z)$ with $d$ in all that follows, where $d_{\mathrm{L}}(z)$ and $d$ are respectively the luminosity and comoving distances to the source, $d_{\mathrm{L}}(z)=(1+z)d$.
+The comoving number density of sources $\nu$ producing the background is also a function of redshift; if the sources producing the background are all in the local Universe, then simply set $\nu(z) = \nu_0\delta (z)$ and replace $d_{\mathrm{L}}(z)$ with $d$ in all that follows, where $d_{\mathrm{L}}(z)$ and $d$ are respectively the luminosity and comoving distances to the source, $d_{\mathrm{L}}(z)=(1+z)d$.
 
-We shall assume the the individual sources are binaries, in which case the number density of sources is also be a function of component mass.
+We shall assume that the individual sources are binaries, in which case the number density of sources is also a function of the component masses.
 It is convenient to work in terms of the chirp mass, defined as ${\mathcal{M}}=\mu^{3/5}M^{2/5}$, where $\mu$ is the reduced mass and $M$ is the total mass of the binary.
@@
 
-Equation (\ref{eq:omega}) gives an expression for the energy density in GWs per logarithmic frequency interval,
+Equation (\ref{eq:differentdescriptions}) gives an expression for the energy density in GWs per logarithmic frequency interval,
 \begin{equation}\label{eq:stoch}
@@
 \begin{equation}\label{eq:bigint}
-h_\mathrm{c}^{2}(f) = \frac{4G^{5/3}}{3\pi^{1/3}c^{2}}f^{-4/3}\int_{0}^{\infty}\mathrm{d}\nu\;\int_{0}^{\infty}\mathrm{d}{\cal{M}}\;\frac{\mathrm{d}^{2}\nu}{\mathrm{d}z\,\mathrm{d}{\mathcal{M}}}\frac{1}{d_{\mathrm{L}}^{2}(z)}\left( \frac{{\mathcal{M}}^{5}}{1+z} \right)^{1/3}\; .
+h_\mathrm{c}^{2}(f) = \frac{4G^{5/3}}{3\pi^{1/3}c^{2}}f^{-4/3}\int_{0}^{\infty}\mathrm{d}z\;\int_{0}^{\infty}\mathrm{d}{\cal{M}}\;\frac{\mathrm{d}^{2}\nu}{\mathrm{d}z\,\mathrm{d}{\mathcal{M}}}\frac{1}{d_{\mathrm{L}}^{2}(z)}\left( \frac{{\mathcal{M}}^{5}}{1+z} \right)^{1/3}\; .
 \end{equation}

### 2014-07-10 noise_curve.tex
myriad typos fixed

@@
 
-\title{Gravitational-wave sensitivity curves}
+\title{Gravitational wave sensitivity curves}
 \author{C J Moore$^1$, R H Cole$^1$ \& C P L Berry$^{1,\,2}$}
@@
 \begin{abstract}
-There are several common conventions use by the gravitational-wave community to describe the amplitude of sources and the sensitivity of detectors.
+There are several common conventions in use by the gravitational wave community to describe the amplitude of sources and the sensitivity of detectors.
 These are frequently confused.
-We outline the merits of and differences between the various quantities used for parameterizing noise curves and characterizing gravitational-wave amplitudes.
+We outline the merits of and differences between the various quantities used for parameterizing noise curves and characterizing gravitational wave amplitudes.
 We conclude by producing plots that consistently compare different detectors.

### 2014-07-11 Introduction.tex
Fixing up factors of c in (22) [see Hobson, Efstathious and Lasenby] and adding gratuitous self-citation
typos and added some text to secs 5.1 5.2 and 5.3

@@
 
-The next few years promises to deliver the first direct detection of gravitational waves (GWs).
-This will most likely be achieved by the advanced Laser Interferometer Gravitational-wave Observatory \citep[LIGO;][]{2010CQGra..27h4006H} and Virgo \citep{Accadia2011} detectors operating in the frequency range $(10$--$10^{3})~\mathrm{Hz}$.
-By the end of the decade, it is expected that pulsar timing arrays \citep[PTAs;][]{1990ApJ361300F} shall also detect very low frequency GWs around $10^{-8}~\mathrm{Hz}$.
-Further into the future, space-based detectors, such as the evolved Laser Interferometer Space Antenna \citep[eLISA;][]{2012CQGra..29l4016A}, will probe GWs in the millihertz regime.
+The next few years  promise to deliver the first direct detection of gravitational waves (GWs).
+This will most likely be achieved by the advanced Laser Interferometer Gravitational-wave Observatory \citep[LIGO;][]{2010CQGra..27h4006H} and advanced Virgo \citep{Acernese2009} detectors operating in the frequency range $(10$--$10^{3})~\mathrm{Hz}$.
+By the end of the decade, it is expected that pulsar timing arrays \citep[PTAs;][]{1990ApJ361300F} will also detect very low frequency GWs around $10^{-8}~\mathrm{Hz}$.
+Further into the future, space-based detectors, such as the evolved Laser Interferometer Space Antenna \citep[eLISA;][]{TheGravitationalUniverse}, will probe GWs in the millihertz regime.
 These advances shall herald the beginning of multi-wavelength GW astronomy as a means of observing the Universe.
@@
 This gives a fractional change in length, or equivalently light travel time, across a detector.
-The strain is small, making it a challenge to measure: we are yet to obtain a direct detection
-of a GW. To calibrate our expectations for detection, it is necessary to quantify the sensitivity of our instruments and the strength of their target signals.
-
+The strain is small, making it a challenge to measure: we are yet to obtain a direct detection of a GW.
+To calibrate our expectations for future detections it is necessary to quantify the sensitivity of our instruments and the strength of their target signals.
 When discussing the loudness of sources and the sensitivity of detectors there are three commonly used parametrizations based upon the strain: the characteristic strain, the power spectral density (PSD) and the spectral energy density.
@@
 \end{equation}
-We shall work with the Fourier transform of the signal, using the conventions that
+We shall have recourse to work with the Fourier transform of the signal, using the conventions that
 \begin{eqnarray} \label{eq:defFourier}
@@
 where angle brackets $\left<\ldots\right>$ denote an ensemble average over many noise realisations \citep{Cutler1994}.
-In reality, we have only a single realisation to work with, however the ensemble average can replaced by a time average for stationary stochastic noise.
+In reality, we have only a single realisation to work with, however the ensemble average can be replaced by a time average for stationary stochastic noise.
 The procedure is to measure the noise over a sufficiently long duration $T$ and then compute the Fourier transform $\tilde{n}(f)$ with a frequency resolution $\Delta f = T^{-1}$; this is repeated many times to give an average.
@@
 The noise is stationary, consequently, its expectation value is unchanged by the time-translation performed above.
-Therefore,
+Therefore, using (\ref{eq:psd}), the mean square noise amplitude is given by
 \begin{eqnarray}\label{eq:meansquare1}
@@
 &=& \int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\frac{1}{2}S_{n}(f)\delta(f-f') \nonumber \\
-&=& \int_{0}^{\infty}\mathrm{d}f\; S_{n}(f),
+&=& \int_{0}^{\infty}\mathrm{d}f\; S_{n}(f).
 \end{eqnarray}
-using (\ref{eq:psd}).
 
@@
 Let $K(t)$ be a real filter function with Fourier transform $\tilde{K}(f)$.
-Convolving this with the detector output gives a contribution from the signal and a contribution from the noise
+Convolving this with the detector output gives a contribution from the signal and a contribution from the noise,
 \begin{equation}\label{eq:conv}
@@
 \begin{equation}\label{eq:snrinnerprod} \innerprod{\tilde{A}(f)}{\tilde{B}(f)} = 4\Re\left\{\int_{0}^{\infty}\mathrm{d}f\;\frac{\tilde{A}^{*}(f)\tilde{B}(f)}{S_{n}(f)}\right\} \; .\end{equation}
-The optimum filter maximises the SNR in (\ref{eq:SNRinnerprod}). From the Cauchy--Schwarz inequality, it follows that the optimum filter is
+The optimum filter is that function $\tilde{K}(f)$ which maximises the SNR in (\ref{eq:SNRinnerprod}). From the Cauchy--Schwarz inequality, it follows that the optimum filter is
 \begin{equation}
@@
 \end{equation}
-In order to construct the Wiener filter, it is necessary to know \emph{a priori} the form of the signal, $\tilde{h}(f)$.
+In order to construct the Wiener filter, it is necessary to know \emph{a priori} the form of the signal, $\tilde{h}(f)$, for this reason the Wiener filter is sometimes called the \emph{matched} filter.
 
@@
 \begin{equation}
-T_{\mu\nu}=\frac{c^{2}}{32\pi G}\left<\partial_{\mu}\bar{h}_{\alpha\beta}\partial_{\nu}\bar{h}^{\alpha\beta}\right> \;,
+T_{\mu\nu}=\frac{c^{4}}{32\pi G}\left<\partial_{\mu}\bar{h}_{\alpha\beta}\partial_{\nu}\bar{h}^{\alpha\beta}\right> \;,
 \end{equation}
-where the angle brackets denote averaging over several wavelengths or periods. The energy density $\rho c^{2}$ is given by the $T_{00}$ component of this tensor. Consequently
+where the angle brackets denote averaging over several wavelengths or periods and $\bar{h}_{\alpha\beta}$ is the transverse-traceless metric perturbation.
+The energy density $\rho c^{2}$ is given by the $T_{00}$ component of this tensor.
+Consequently \citep[cf.][]{Berry2013},
 \begin{eqnarray}
@@
 
-Cosmological studies often work in terms of the dimensionless quantity $\Omega_{\mathrm{GW}}$, the energy density per logarithmic frequency interval normalised to the critical density of the universe:
+Cosmological studies often work in terms of the dimensionless quantity $\Omega_{\mathrm{GW}}$, the energy density per logarithmic frequency interval normalised to the critical density of the universe, $\rho_{\mathrm{c}}$:
 \begin{equation}
 \label{eq:omega}
-\Omega_\mathrm{GW}(f) = \frac{fS_{\mathrm{E}}(f)}{\rho_{\mathrm{c}}c^{2}} \; .
-\end{equation}
-The critical density of the universe is
-\begin{equation}\label{eq:crit-density}
-\rho_{\mathrm{c}}=\frac{3H_{0}^{2}}{8\pi G} \;,
+\Omega_\mathrm{GW}(f) = \frac{fS_{\mathrm{E}}(f)}{\rho_{\mathrm{c}}c^{2}} \; , \quad \textrm{where}\quad \rho_{\mathrm{c}}=\frac{3H_{0}^{2}}{8\pi G} \;.
 \end{equation}
@@
 \end{equation}
-The reduced Hubble parameter $h_{100}$ has nothing to do with strain.
-The most common quantity related to energy density to be plotted on sensitivity curves is $\Omega_{\mathrm{GW}}h_{100}^{2}$ (figure \ref{fig:omega}) as this removes sensitivity to the measured value of the Hubble constant.
+The reduced Hubble parameter $h_{100}$ has nothing to do with strain!
+The most common quantity related to energy density to be plotted on sensitivity curves is $\Omega_{\mathrm{GW}}h_{100}^{2}$ (figure \ref{fig:omega}) as this removes sensitivity to the historically very uncertain measured value of the Hubble constant.
 

### 2014-07-11 Sources.tex
typos and added some text to secs 5.1 5.2 and 5.3

@@
 The expected event rate for this type of source is uncertain, but estimates centre around $\gamma_{\mathrm{NS-NS}}=1.3\times 10^{-4}~\mathrm{Mpc^{-3}\,yr^{-1}}$ \citep{CBC}.
-Plotted in \ref{app:a} are boxes with amplitudes such that a ratio $h_\mathrm{c}/h_{n}=16$ is produced for Advanced LIGO at peak sensitivity and a width between $(3$--$300)~\mathrm{Hz}$, corresponding to expected observable frequencies.
+Plotted in \ref{app:a} are boxes labelled ``\emph{compact binary inspirals}'' with amplitudes such that a ratio $h_\mathrm{c}/h_{n}=16$ is produced for Advanced LIGO at peak sensitivity and a width between $(3$--$300)~\mathrm{Hz}$, corresponding to expected observable frequencies.
 
@@
 The event rate for supernovae is approximately $\gamma_{\mathrm{SN}} = 5\times10^{-4}~\mathrm{Mpc^{-3}\,yr^{-1}}$.
-The boxes plotted in \ref{app:a} correspond to a distance $r=300~\mathrm{kpc}$ with the frequency range quoted above.
+The boxes labelled ``\emph{supernova}'' plotted in \ref{app:a} correspond to a distance $r=300~\mathrm{kpc}$ with the frequency range quoted above.
 The LIGO and Virgo detectors have already placed bounds on the event rate for these sources \citep{Bursts}.
@@
 Several known pulsars could be sources for the advanced detectors and upper limits from the initial detectors help to constrain the deformations.
-The boxes plotted in \ref{app:a} correspond to the upper limits placed on a GW signal from the Crab pulsar \citep{Aasi2014a}, extrapolated across a frequency range between $(20$--$10^{3})~\mathrm{Hz}$.
+The boxes labelled ``\emph{pulsars}'' plotted in \ref{app:a} correspond to the upper limits placed on a GW signal from the Crab pulsar \citep{Aasi2014a}, extrapolated across a frequency range between $(20$--$10^{3})~\mathrm{Hz}$.
 
@@
 The large range in the rate reflects our uncertainty in the growth mechanisms of the supermassive black hole population \citep{Volonteri2010}.
-Plotted in \ref{app:a} are boxes with a ratio $h_\mathrm{c}/h_{n}=100$ for eLISA at its peak sensitivity.
+Plotted in \ref{app:a} are boxes labelled ``\emph{$\approx 10^{6}$ solar mass binaries}'' with a ratio $h_\mathrm{c}/h_{n}=100$ for eLISA at its peak sensitivity.
 The range of frequencies plotted is $(3\times 10^{-4}$--$3\times 10^{-1})~\mathrm{Hz}$; this corresponds to circular binaries in the mass range quoted above.
@@
 This boundary will not be too different for eLISA but would move substantially for either of the decihertz detectors.
-Plotted in \ref{app:a} is the estimate of this background due to \citet{Nelemans} where an observation time of one year has been assumed,
+Plotted in \ref{app:a} with the label ``\emph{unresolvable galactic binaries}'' is the estimate of this background due to \citet{Nelemans} where an observation time of one year has been assumed,
 \begin{equation}
@@
 Estimates for the event rate of resolvable binaries centre around ${\mathcal{O}}(10^{3})$ events for eLISA.
-The boxes plotted in \ref{app:a} have a ratio $h_\mathrm{c}/h_{n}=50$ for eLISA at its peak sensitivity.
+The boxes plotted in \ref{app:a} with the label ``\emph{resolvable galactic binaries}'' have a ratio $h_\mathrm{c}/h_{n}=50$ for eLISA at its peak sensitivity.
 The frequency range of the box is $\left(3\times10^{-4}\right.$--$\left.10^{-2}\right)~\mathrm{Hz}$, estimated from Monte Carlo population simulation results presented in \citet{Amaro-Seoane-et-al}.
@@
 There is extreme uncertainty in the event rate for EMRIs due to the poorly constrained astrophysics in galactic centres \citep[e.g.,][]{Merritt2011}; the best guess estimate is around $25$ events per year with eLISA with SNR $\ge 20$ \citep{TheGravitationalUniverse}.
-The boxes plotted in \ref{app:a} have a characteristic strain of $h_\mathrm{c}=3\times 10^{-20}$ at $10^{-2}~\mathrm{Hz}$, which corresponds to a $10\Msun$ black hole inspiralling into a $10^{6}\Msun$ black hole at a luminosity distance of $1~\mathrm{Gpc}$.
+The boxes labelled ``\emph{extreme mass ratio inspirals}'' plotted in \ref{app:a} have a characteristic strain of $h_\mathrm{c}=3\times 10^{-20}$ at $10^{-2}~\mathrm{Hz}$, which corresponds to a $10\Msun$ black hole inspiralling into a $10^{6}\Msun$ black hole at a luminosity distance of $1~\mathrm{Gpc}$.
 The frequency width of the box is somewhat unknown; EMRI events can occur into a black hole of any mass, and hence EMRIs can, in principal, occur at any frequency, so the boxes in \ref{app:a} are drawn with a width comparable to that of the LISA sensitivity curve.
@@
 It is currently unclear whether PTAs will detect an individual binary or a stochastic background first.
-Plotted in \ref{app:a} for the unresolvable background is a third of the current limit with a cut off frequency of $f=1~\mathrm{yr}^{-1}$ which is suggested by Monte Carlo population studies \citep{SesanaVecchioColancino}.
-For the resolvable sources the amplitude of the current limit is plotted between $\left(3\times 10^{-9}\right.$--$\left.3\times 10^{-7}\right)~\mathrm{Hz}$.
+Plotted in \ref{app:a} with the label ``\emph{stochastic background}'' is a third of the current limit with a cut off frequency of $f=1~\mathrm{yr}^{-1}$ which is suggested by Monte Carlo population studies \citep{SesanaVecchioColancino}.
+For the resolvable sources, labelled ``\emph{$\approx 10^{9}$ solar mass binaries}'', the amplitude of the current limit is plotted between $\left(3\times 10^{-9}\right.$--$\left.3\times 10^{-7}\right)~\mathrm{Hz}$.
 
@@
 Cosmological stochastic backgrounds have been predicted across a range of frequencies with considerable variation in amplitude.
-As a consequence of this uncertainty, although we could learn much from measuring these signals, we have not included them amongst the sources shown in \ref{app:a}.
+As a consequence of this uncertainty, although we could learn much from measuring these signals, we have not included them amongst the sources in \ref{app:a}.

### 2014-07-14 Introduction.tex
On PSD

@@
 
-\subsection{Power spectral density }\label{sec:psd}
+\subsection{Power spectral density}\label{sec:psd}
 
-A second commonly used quantity for sensitivity curves is the square root of the PSD (see figure \ref{fig:S}). When discussing a detector this is
+A second commonly used quantity for sensitivity curves is the square root of the PSD or the amplitude spectral density (see figure \ref{fig:S}).
+When discussing a detector, rearranging (\ref{eq:strain-hn}) gives
 \begin{equation}\label{eq:temp1}
-\sqrt{S_{n}(f)} = h_{n}(f)f^{-1/2} \; ,
+\sqrt{S_{n}(f)} = h_{n}(f)f^{-1/2} \; ;
 \end{equation}
-where we have used (\ref{eq:strain-hn}); by analogy, we can define an equivalent for source amplitudes
+by analogy, we can define an equivalent for source amplitudes
 \begin{equation}
@@
 
-The PSD has the nice property, demonstrated in (\ref{eq:meansquare1}), that integrated over all positive frequencies it gives the mean square amplitude of the signal in the detector.
+The PSD, as sefined by (\ref{eq:psd}), has the nice property, demonstrated in (\ref{eq:meansquare1}), that integrated over all positive frequencies it gives the mean square amplitude of the signal in the detector.
 However, in one important regard it is less appealing than characteristic strain: the height of the source above the detector curve is no longer directly related to the SNR.
@@
 
-Cosmological studies often work in terms of the dimensionless quantity $\Omega_{\mathrm{GW}}$, the energy density per logarithmic frequency interval normalised to the critical density of the universe, $\rho_{\mathrm{c}}$:
+Cosmological studies often work in terms of the dimensionless quantity $\Omega_{\mathrm{GW}}$, the energy density per logarithmic frequency interval normalised to the critical density of the universe $\rho_{\mathrm{c}}$,
 \begin{equation}
 \label{eq:omega}
-\Omega_\mathrm{GW}(f) = \frac{fS_{\mathrm{E}}(f)}{\rho_{\mathrm{c}}c^{2}} \; , \quad \textrm{where}\quad \rho_{\mathrm{c}}=\frac{3H_{0}^{2}}{8\pi G} \;.
+\Omega_\mathrm{GW}(f) = \frac{fS_{\mathrm{E}}(f)}{\rho_{\mathrm{c}}c^{2}} \; .
 \end{equation}
-where $H_{0}$ is the Hubble constant which is commonly parametrized as
+The critical density is
 \begin{equation}
-H_0 =h_{100}\times 100~\mathrm{km\,s^{-1}\,Mpc^{-1}}.
+\label{eq:crit-density}
+\rho_{\mathrm{c}}=\frac{3H_{0}^{2}}{8\pi G} \;,
+\end{equation}
+where $H_{0}$ is the Hubble constant, commonly parametrized as
+\begin{equation}
+H_0 = h_{100}\times 100~\mathrm{km\,s^{-1}\,Mpc^{-1}}.
 \end{equation}

### 2014-07-29 Detectors.tex
small edits from P&P process

@@
  \includegraphics[trim=0cm 0cm 0cm 0cm, width=0.99\textwidth]{LIGO_detectorframe.pdf}
- \caption{The angular response function of an interferometric detector in (\ref{eq:skyav}) shown both as a surface plot and in an Aitoff--Hammer projection.
+ \caption{The angular response function of an interferometric detector shown both as a surface plot and in an Aitoff--Hammer projection.
+The quentity that is plotted is is the polarisation average, $\int\textrm{d}\psi \;((F^{+})^{2}+(F^{\times})^{2})/2$.
 The response is a function of two sky angles, $\theta$ and $\phi$, and varies between $0$ and $1$.
@@
 Additionally, the different averages do not work out to be so different from each other: \citet{1993PhRvD..47.2198F} calculated the sensitivity for a detector with the LIGO geometry averaged over the four angles ($\theta,\phi,\iota,\xi$) as $\sqrt{4/25}=0.4$ times peak sensitivity, which should be compared with the value $\sqrt{1/5}\approx 0.447$ above.
+The effect of replacing the true sensitivity with the sky-averaged sensitivity for the LISA detector was considered in detail by \cite{2012CQGra..29l4015V}.
 For the remainder of this paper the three-angle average defined in (\ref{eq:skyav}) is used.
@@
 The most notable ground-based detectors are LIGO and Virgo, which work in collaboration, supported by GEO600.
-LIGO and Virgo have completed science runs as first-generation detectors, now referred to as Initial LIGO (iLIGO) and Initial Virgo (iVirgo).
+LIGO, Virgo and GEO600 have completed science runs as first-generation detectors, now referred to as Initial LIGO (iLIGO) and Initial Virgo (iVirgo).
 They are currently being upgraded to second-generation detectors, advanced LIGO (aLIGO) and advanced Virgo (AdV).
@@
 TAMA300 is a Japanese first-generation detector.
-Its successor is the Kamioka Gravitational Wave Detector (KAGRA), formerly the Large-scale Cryogenic Gravitational wave Telescope (LCGT), which is located underground in the Kamioka mine.
+Its successor, currently under construction, is the Kamioka Gravitational Wave Detector (KAGRA), formerly the Large-scale Cryogenic Gravitational wave Telescope (LCGT), which is located underground in the Kamioka mine.
 It employs more sophisticated noise-reduction techniques than LIGO or Virgo, such as cryogenic cooling.

### 2014-07-29 Introduction.tex
small edits from P&P process

@@
 The signal contribution $\mathcal{S}$ is defined as the expectation of the convolution in (\ref{eq:conv}) when a signal is present, maximised by varying the offset to achieve the best overlap with the data.
-Since the expectation of pure noise is zero it, follows that
+Since the expectation of pure noise is zero it follows that
 \begin{equation}
@@
 \begin{equation}\label{eq:hc} 
-\varrho^{2} = \int_{0}^{\infty} \mathrm{d}\left(\log f\right)\; \left[\frac{h_\mathrm{c}(f)}{h_{n}(f)}\right]^{2} \;.
+\varrho^{2} = \int_{-\infty}^{\infty} \mathrm{d}\left(\log f\right)\; \left[\frac{h_\mathrm{c}(f)}{h_{n}(f)}\right]^{2} \;.
 \end{equation}
@@
 \begin{eqnarray}
-\label{eq:specNRGdensity}
 \rho c^{2} &=& \frac{c^{2}}{16\pi G}\int_{-\infty}^{\infty}\mathrm{d}f\;\left(2\pi f\right)^{2}\tilde{h}(f)\tilde{h}^{*}(f) \\*
- &=& \int_{0}^{\infty}\mathrm{d}f\;\frac{\pi c^{2}}{4G}f^{2}S_{h}(f)\; ,
+ &=& \int_{0}^{\infty}\mathrm{d}f\;\frac{\pi c^{2}}{4G}f^{2}S_{h}(f)\; , \label{eq:specNRGdensity}
 \end{eqnarray} 
 where the definition (\ref{eq:ShforSources}) have been used.
-Equation (\ref{eq:specNRGdensity}) gives the definition of the spectral energy density, the energy per unit volume of space, per unit frequency \citep{HellingsDowns}
+The integrand in Equation (\ref{eq:specNRGdensity}) is defined as the spectral energy density, the energy per unit volume of space, per unit frequency \citep{HellingsDowns}
 \begin{equation}\label{eq:spectralenergydensity}

### 2014-07-29 Typesofsource.tex
typos and added some text to secs 5.1 5.2 and 5.3
small edits from P&P process

@@
 where an overdot represents the time derivative and $\dot{\phi} = 2\pi f$.
-The squared SNR scales with ${N}_{\mathrm{cycles}}$, so it would be expected that $h_\mathrm{c}(f)\approx \sqrt{{N}_{\mathrm{cycles}}}\tilde{h}(f)$.
+The squared SNR scales with ${N}_{\mathrm{cycles}}$, so it would be expected that $h_\mathrm{c}(f)\approx \sqrt{{N}_{\mathrm{cycles}}}\left|\tilde{h}(f)\right|$.
 
@@
 These can be at cosmological distances, where it is necessary to distinguish the frequency in the source rest frame $f_{\mathrm{r}}$ from the measured frequency $f$; the two are related through the redshift $z$ via $f_{\mathrm{r}}=(1+z)f$.
-
 The comoving number density of sources $\nu$ producing the background is also a function of redshift; if the sources producing the background are all in the local Universe, then simply set $\nu(z) = \nu_0\delta (z)$ and replace $d_{\mathrm{L}}(z)$ with $d$ in all that follows, where $d_{\mathrm{L}}(z)$ and $d$ are respectively the luminosity and comoving distances to the source, $d_{\mathrm{L}}(z)=(1+z)d$.
@@
 
+An alternative method for graphically representing the sensitivity of a GW detector to stochastic backgrounds, called the \emph{power-law-integrated sensitivity curve}, was suggested by \cite{2013PhRvD..88l4032T}.
+This method accounts for the fact that there is power across all frequencies in the sensitivity band by integrating the noise weighted signaal over frequency.
+As our aim here is to present stochastic backgrounds alongside other types of sources for comparison we do not use this approach.
+
 \subsection{Burst sources}\label{sec:bursts}
@@
 The data stream from a detector is temporally split into segments, the length of which can be tuned to give greater sensitivity to particular sources.
-Each segment is then transformed into the frequency domain, whitened and normalised to the noise spectrum to produce a time--frequency plot.
+Each segment is then transformed into the frequency domain, whitened and normalised to the noise spectrum of the detector to produce a time--frequency plot.
 Potential GW signals are identified by searching for clusters of pixels that contain an excess of power \citep[e.g.,][]{Bursts}.
@@
 \begin{equation}\label{eq:simple} 
-h_\mathrm{c}(f) = \tilde{h}(f)\sqrt{{N}_{\mathrm{cycles}}} \; , 
+h_\mathrm{c}(f) = \left|\tilde{h}(f)\right|\sqrt{{N}_{\mathrm{cycles}}} \; , 
 \end{equation}

### 2014-07-29 noise_curve.tex
small edits from P&P process

@@
 We are indebted to Jonathan Gair, for useful advice and proof-reading the manuscript.
-We also thank Stephen Taylor and Justin Ellis for useful conversations about pulsar timing.}
+We also thank Stephen Taylor and Justin Ellis for useful conversations about pulsar timing.
+The authors thanks a number of members of the LIGO--Virgo Scientific Collaboration for useful comments and suggestions.
+This paper has been assigned LIGO document LIGO-P1400129.}
 

### 2014-07-29 Typesofsource.tex
h_rss

@@
 An alternative method for graphically representing the sensitivity of a GW detector to stochastic backgrounds, called the \emph{power-law-integrated sensitivity curve}, was suggested by \cite{2013PhRvD..88l4032T}.
-This method accounts for the fact that there is power across all frequencies in the sensitivity band by integrating the noise weighted signaal over frequency.
+This method accounts for the fact that there is power across all frequencies in the sensitivity band by integrating the noise weighted signal over frequency.
 As our aim here is to present stochastic backgrounds alongside other types of sources for comparison we do not use this approach.
@@
 \end{equation}
-where ${N}_{\mathrm{cycles}}$ is the number of cycles of radiation generated by the source.
-For bursts, $N_{\mathrm{cycles}}$ is of order unity and $\tilde{h}(f)$ is strongly peaked around some central frequency.
+where ${N}_{\mathrm{cycles}}$ is the number of cycles of radiation generated by the source, which is of order unity for bursts.
+Due to the uncertainty of burst models, the waveform is commonly indicated by a constant amplitude across some finite bandwidth.
+This is analogous to computing the root-sum-square of the waveform polarisations
+\begin{equation}
+h_\mathrm{rss} = \sqrt{\int{|h_+(t)|}^2 + {|h_\times(t)|}^2\,\rmd t},
+\end{equation}
+which is a quantity that is used to characterise the sensitivity of detectors to bursts \citep{Bursts}.
+In this work, we favour a constant $h_\mathrm{c}(f)$ over $h_\mathrm{rss}$ for consistency with the other types of source.

### 2014-07-29 Typesofsource.tex
h_rss again

@@
 where ${N}_{\mathrm{cycles}}$ is the number of cycles of radiation generated by the source, which is of order unity for bursts.
-Due to the uncertainty of burst models, the waveform is commonly indicated by a constant amplitude across some finite bandwidth.
-This is analogous to computing the root-sum-square of the waveform polarisations
+
+An alternative characterisation of the signal amplitude commonly used for burst sources is the root-sum-square of the waveform polarisations
 \begin{equation}
-h_\mathrm{rss} = \sqrt{\int{|h_+(t)|}^2 + {|h_\times(t)|}^2\,\rmd t},
+h_\mathrm{rss} = \sqrt{\int{|h_+(t)|}^2 + {|h_\times(t)|}^2\,\rmd t}.
 \end{equation}
-which is a quantity that is used to characterise the sensitivity of detectors to bursts \citep{Bursts}.
-In this work, we favour a constant $h_\mathrm{c}(f)$ over $h_\mathrm{rss}$ for consistency with the other types of source.
+If $\tilde{h}(f) = \tilde{h}$ is constant across the bandwidth $\Delta f$ then they are related by
+\begin{equation}
+h_\mathrm{rss} \simeq |\tilde{h}|\sqrt{\Delta f}.
+\end{equation}
+In this work, we favour a constant $h_\mathrm{c}(f)$ rather than $h_\mathrm{rss}$ for consistency with the other types of source.

### 2014-07-29 Detectors.tex
added GEO-HF ref and updated GEO600 ref, and a few tweaks

@@
 There is an agreement to move one of the upgraded Hanford detector systems to a location in India \citep{LIGO-India,Unnikrishnan2013}.
+The GEO600 detector can also operate in a high frequency mode \citep[GEO-HF]{2006CQGra..23S.207W}, this is not included in Table \ref{table:t}.
 The advanced detectors should start operation in the next couple of years, with LIGO-India following further in the future.
@@
 \br
-\multicolumn{5}{l}{$^{a}$\citet{2008CQGra..25k4043G}, $^{b}$\citet{2002CQGra..19.1409A}, $^{c}$\citet{2009RPPh...72g6901A}, $^{d}$\citet{accadia_virgo:2012}, }\\
-\multicolumn{5}{l}{$^{e}$\citet{2010CQGra..27h4006H}, $^{f}$\citet{Acernese2009}, $^{g}$\citet{2012CQGra..29l4007S}, $^{h}$\citet{2011CQGra..28i4013H}.}\\
+\multicolumn{5}{l}{$^{a}$\citet{2010CQGra..27h4003G}, $^{b}$\citet{2002CQGra..19.1409A}, $^{c}$\citet{2009RPPh...72g6901A},}\\
+\multicolumn{5}{l}{$^{d}$\citet{accadia_virgo:2012}, $^{e}$\citet{2010CQGra..27h4006H}, $^{f}$\citet{Acernese2009}, $^{g}$\citet{2012CQGra..29l4007S}, $^{h}$\citet{2011CQGra..28i4013H}.}\\
 \br
@@
 There are published limits on the amplitude of the stochastic background from all three detectors: the most recent from EPTA is \citet{Haasteren}, from PPTA is \citet{Shannon2013} and from NANOGrav is \citet{Arzoumanian2014}.
-These are all comparable and we use the EPTA limits in \ref{app:a}, the figures show the published limit based on an analysis of five pulsars over approximately ten years.
+These are all comparable and we use the EPTA limits in \ref{app:a}, the figures show the published limit on $A$ from equation \ref{eq:power} at a frequency of $f_{0}=1\,\textrm{year}^{-1}$, based on an analysis of five pulsars over approximately ten years.
 

### 2014-07-29 Typesofsource.tex
added GEO-HF ref and updated GEO600 ref, and a few tweaks

--- 2014-07-29 Typesofsource.tex
@@
 where ${N}_{\mathrm{cycles}}$ is the number of cycles of radiation generated by the source, which is of order unity for bursts.
-Due to the uncertainty of burst models, the waveform is commonly indicated by a constant amplitude across some finite bandwidth.
-This is analogous to computing the root-sum-square of the waveform polarisations
-\begin{equation}
-h_\mathrm{rss} = \sqrt{\int{|h_+(t)|}^2 + {|h_\times(t)|}^2\,\rmd t},
-\end{equation}
-which is a quantity that is used to characterise the sensitivity of detectors to bursts \citep{Bursts}.
-In this work, we favour a constant $h_\mathrm{c}(f)$ over $h_\mathrm{rss}$ for consistency with the other types of source.
+
+An alternative characterisation of the signal amplitude commonly used for burst sources is the root-sum-square of the waveform polarisations, $h_\mathrm{rss}$.
+If $\tilde{h}(f)$ is constant across the bandwidth $\Delta f$ this is related to the characteristic strain via Equation \ref{eq:rela}.
+In this work, we favour a constant $h_\mathrm{c}(f)$ rather than $h_\mathrm{rss}$ for consistency with the other types of source.
+\begin{eqnarray}
+&h_\mathrm{rss} = \sqrt{\int{|h_+(t)|}^2 + {|h_\times(t)|}^2\,\rmd t} \\
+&h_\mathrm{rss} \simeq |\tilde{h}(f)|\sqrt{\Delta f}\,\label{eq:rela}.
+\end{eqnarray}

### 2014-07-29 Sources.tex
first para of sec 5, rewritten to be clearer about exactly how the source boxes should be interpretted

@@
 
-All the sources described here are represented by shaded boxes in \ref{app:a}.
+%All the sources described here are represented by shaded boxes in \ref{app:a}.
 The boxes are drawn in such a way that there is a reasonable event rate for sources at a detectable SNR.
@@
 The width of the box gives the range of frequencies sources of a given type can have while remaining at a detectable amplitude.
+
+All the sources described here are represented by shaded boxes in \ref{app:a}.
+Sources with short durations (i.e., burst sources) and sources that evolve in time over much longer timescales than our observations are drawn with flat-topped boxes for $h_\mathrm{c}(f)$.
+Inspiraling binaries, or stochastic backgrounds of binaries, are drawn with a sloping top proportional to $f^{-2/3}$ for $h_\mathrm{c}(f)$, which is result derived in section \ref{sec:stoch}.
+The width of the box gives the range of frequencies sources of a given type can have while remaining at a detectable amplitude.
+The question of the height of the box is more problematic; it would be desirable to normalise each box so that there was a fixed event rate, say one event per year with an amplitude lying within the box.
+However for many of the sources considered the event rate is subject to a large degree of uncertainty (and estimates change rapidly as our understanding of the astrophysics improves).
+So instead we take the more definite, but somewhat arbitrary, approach that for each type of source a fiducial event is nominated (with parameters or amplitude detailed in the relevant section below) and the amplitude of this event is used to position the top of the box.
+The parameters for each fiducial event are chosen such that the resulting amplitude is roughly consistent with the maximum end of current predictions.
+I.e.
+the boxes indicate where we \emph{could} find GW sources; the actual event rates could turn out to be substantially lower than this upper bound.
 

### 2014-07-31 Detectors.tex
Small changes

@@
 
+The response function of a two-arm interferometric detector is quadrupolar, an example is plotted in figure \ref{fig:LIGO}.
 \begin{figure}
@@
 \end{figure}
-
-The response function of a two-arm interferometric detector is quadrupolar, an example is plotted in figure \ref{fig:LIGO}.
 Throughout this paper detector sensitivity refers to the polarisation and sky averaged sensitivity $F$, where
@@
 Additionally, the different averages do not work out to be so different from each other: \citet{1993PhRvD..47.2198F} calculated the sensitivity for a detector with the LIGO geometry averaged over the four angles ($\theta,\phi,\iota,\xi$) as $\sqrt{4/25}=0.4$ times peak sensitivity, which should be compared with the value $\sqrt{1/5}\approx 0.447$ above.
-The effect of replacing the true sensitivity with the sky-averaged sensitivity for the LISA detector was considered in detail by \cite{2012CQGra..29l4015V}.
+The effect of replacing the true sensitivity with the sky-averaged sensitivity for the LISA detector was considered in detail by \cite{2012CQGra..29l4015V}; they also found there is a small difference when considering an entire population of sources.
 For the remainder of this paper the three-angle average defined in (\ref{eq:skyav}) is used.
@@
 There is an agreement to move one of the upgraded Hanford detector systems to a location in India \citep{LIGO-India,Unnikrishnan2013}.
-The GEO600 detector can also operate in a high frequency mode \citep[GEO-HF]{2006CQGra..23S.207W}, this is not included in Table \ref{table:t}.
+The GEO600 detector can also operate in a high frequency mode \citep[GEO-HF]{2006CQGra..23S.207W}, this is not included in Table \ref{table:t} \cplb{Why not?}.
 The advanced detectors should start operation in the next couple of years, with LIGO-India following further in the future.
@@
 \end{equation}
-All of the pulsars, and the Earth, are subject to the same metric perturbation field. This may be expressed in terms of its Fourier transform
+All of the pulsars, and the Earth, are subject to the same metric-perturbation field. This may be expressed in terms of its Fourier transform
 \begin{equation}
@@
 The characteristic strain that the PTA is sensitive to scales linearly with $f$ in this range.
-This gives the wedge shaped curves plotted in \ref{app:a}.
+This gives the wedge-shaped curves plotted in \ref{app:a}.
 The absolute value of the sensitivity is fixed by normalising to a calculated limit at a given frequency for each PTA.
@@
 When observing a long-lived source, such as an inspiral, with a high frequency detector the convention was to define a \emph{characteristic} strain to satisfy (\ref{eq:hc}).
-Here the convention is to leave the strain untouched and instead adjust the PTA sensitivity curve with observation time, again to satisfy (\ref{eq:hc}).
+Here, the convention is to leave the strain untouched and instead adjust the PTA sensitivity curve with observation time, again to satisfy (\ref{eq:hc}).
 This discrepancy is an unfortunate result of the conventions in use by the different GW communities; however, it is also natural given the sources under observation.
@@
 
-The PTAs currently in operation are the European Pulsar Timing Array (EPTA\footnote{\url{http://www.epta.eu.org/}}, \cite{eptareview2013}), the Parkes Pulsar Timing Array (PPTA\footnote{\url{http://www.atnf.csiro.au/research/pulsar/ppta/}} \cite{parkesreview2013}) based in Australia, and the North American Nanohertz Observatory for Gravitational waves (NANOGrav\footnote{\url{http://nanograv.org/}} \cite{nanogravreview2013}).
+The PTAs currently in operation are the European Pulsar Timing Array \citep[EPTA\footnote{\url{http://www.epta.eu.org/}};][]{eptareview2013}, the Parkes Pulsar Timing Array \citep[PPTA\footnote{\url{http://www.atnf.csiro.au/research/pulsar/ppta/}};][]{parkesreview2013} based in Australia, and the North American Nanohertz Observatory for Gravitational waves \citep[NANOGrav\footnote{\url{http://nanograv.org/}};][]{nanogravreview2013}).
 There are published limits on the amplitude of the stochastic background from all three detectors: the most recent from EPTA is \citet{Haasteren}, from PPTA is \citet{Shannon2013} and from NANOGrav is \citet{Arzoumanian2014}.
@@
 
-Combining the existing arrays would yield a single PTA using approximately 4 times as many pulsars; this consortium of consortia is known as the International Pulsar Timing Array (IPTA\footnote{\url{http://www.ipta4gw.org/}} \cite{iptareview2013}).
+Combining the existing arrays would yield a single PTA using approximately 4 times as many pulsars; this consortium of consortia is known as the International Pulsar Timing Array \citep[IPTA\footnote{\url{http://www.ipta4gw.org/}};][]{iptareview2013}.
 The IPTA curves plotted in the figures are based on 20 pulsars timed for twice as long as the current EPTA.
@@
 This shall greatly increase the sensitivity of pulsar timing \citep{Kramer2004}.
-The sensitivity curves plotted in \ref{app:a} for the SKA assume a RMS error on the timing residuals a factor of 10 better than current PTAs, and include 20 pulsars timed for twice as long as the current EPTA.
+The sensitivity curves plotted in \ref{app:a} for the SKA assume a root-mean-square error on the timing residuals a factor of 10 better than current PTAs, and include 20 pulsars timed for twice as long as the current EPTA.

### 2014-07-31 Introduction.tex
Small changes

@@
 When producing these plots, it is desirable to have a consistent convention between detectors and sources that allows information about both to be plotted on the same graph.
-Ideally, the detectors and sources are represented in such a way that the relative detectability of the signals is immediately apparent from the plot.
+Ideally, the detectors and sources are represented in such a way that the relative detectability of the signals is immediately apparent.
 
@@
 A website where similar figures can be generated is available at \url{www.ast.cam.ac.uk/~rhc26/sources/}.
-Here the user may select which sources and detectors to include to tailor the figure to their specific requirements.
+Here, the user may select which sources and detectors to include to tailor the figure to their specific requirements.
 
@@
 \end{equation}
+
 We shall have recourse to work with the Fourier transform of the signal, using the conventions that
@@
 where angle brackets $\left<\ldots\right>$ denote an ensemble average over many noise realisations \citep{Cutler1994}.
-In reality, we have only a single realisation to work with, however the ensemble average can be replaced by a time average for stationary stochastic noise.
+In reality, we have only a single realisation to work with, but the ensemble average can be replaced by a time average for stationary stochastic noise.
 The procedure is to measure the noise over a sufficiently long duration $T$ and then compute the Fourier transform $\tilde{n}(f)$ with a frequency resolution $\Delta f = T^{-1}$; this is repeated many times to give an average.
@@
  &= \lim_{T\rightarrow\infty}\frac{1}{2T} & \int_{-T}^{T}\mathrm{d}t\; \int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;{\mathcal{F}}\left\{ n(\tau) \right\}(f)\left[{\mathcal{F}}\left\{ n(\tau) \right\}(f')\right]^{*} \nonumber \\*
-\label{eq:meansquare} & & \times \exp\left(2\pi \rmi ft\right)\exp\left(-2\pi \rmi f't\right)\;,
+\label{eq:meansquare} & & \times \hphantom{\int_{-T}^{T}} \exp\left(2\pi \rmi ft\right)\exp\left(-2\pi \rmi f't\right)\;,
 \end{eqnarray}
@@
 &=& \int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\frac{1}{2}S_{n}(f)\delta(f-f') \nonumber \\
-&=& \int_{0}^{\infty}\mathrm{d}f\; S_{n}(f).
+&=& \int_{0}^{\infty}\mathrm{d}f\; S_{n}(f) \;.
 \end{eqnarray}
@@
 Whilst the magnitude of the Fourier transform of the signal $|\tilde{h}(f)|$ provides a simple quantification of the GW amplitude as a function of frequency, it has one main deficiency.
-For an inspiralling source the instantaneous amplitude can be orders of magnitude below the noise level in a detector; however, as the signal continues over many orbits, the SNR can be integrated up to a detectable level.
+For an inspiralling source ,the instantaneous amplitude can be orders of magnitude below the noise level in a detector; however, as the signal continues over many orbits, the SNR can be integrated up to a detectable level.
 It is useful to have a quantification of the GW amplitude that accounts for this effect; three such methods are described in the following subsections.
@@
 \begin{eqnarray}\label{eq:strain-hc} 
-h_\mathrm{c}(f)^{2} &= 4f^{2}\left| \tilde{h}(f) \right|^{2} \; ,\\
-h_{n}(f)^{2} &= fS_{n}(f),
+\left[h_\mathrm{c}(f)\right]^{2} &= 4f^{2}\left| \tilde{h}(f) \right|^{2} \; ,\\
+\left[h_{n}(f)\right]^{2} &= fS_{n}(f),
 \label{eq:strain-hn}
@@
 
-The PSD, as sefined by (\ref{eq:psd}), has the nice property, demonstrated in (\ref{eq:meansquare1}), that integrated over all positive frequencies it gives the mean square amplitude of the signal in the detector.
+The PSD, as defined by (\ref{eq:psd}), has the nice property, demonstrated in (\ref{eq:meansquare1}), that integrated over all positive frequencies it gives the mean square amplitude of the signal in the detector.
 However, in one important regard it is less appealing than characteristic strain: the height of the source above the detector curve is no longer directly related to the SNR.
@@
 \end{equation}
-where the angle brackets denote averaging over several wavelengths or periods and $\bar{h}_{\alpha\beta}$ is the transverse-traceless metric perturbation.
+where the angle brackets denote averaging over several wavelengths or periods, and $\bar{h}_{\alpha\beta}$ is the transverse-traceless metric perturbation.
 The energy density $\rho c^{2}$ is given by the $T_{00}$ component of this tensor.
@@
 \end{equation}
-The reduced Hubble parameter $h_{100}$ has nothing to do with strain!
-The most common quantity related to energy density to be plotted on sensitivity curves is $\Omega_{\mathrm{GW}}h_{100}^{2}$ (figure \ref{fig:omega}) as this removes sensitivity to the historically very uncertain measured value of the Hubble constant.
+The reduced Hubble parameter $h_{100}$ has nothing to do with strain.
+The most common quantity related to energy density to be plotted on sensitivity curves is $\Omega_{\mathrm{GW}}h_{100}^{2}$ (figure \ref{fig:omega}) as this removes sensitivity to the (historically uncertain) measured value of the Hubble constant.
 

### 2014-07-31 Sources.tex
Small changes

@@
 So instead we take the more definite, but somewhat arbitrary, approach that for each type of source a fiducial event is nominated (with parameters or amplitude detailed in the relevant section below) and the amplitude of this event is used to position the top of the box.
-The parameters for each fiducial event are chosen such that the resulting amplitude is roughly consistent with the maximum end of current predictions.
-I.e.
-the boxes indicate where we \emph{could} find GW sources; the actual event rates could turn out to be substantially lower than this upper bound.
+The parameters for each fiducial event are chosen such that the resulting amplitude is roughly consistent with the maximum end of current predictions: the boxes indicate where we \emph{could} plausibly find GW sources; the actual event rates could turn out to be substantially lower than this upper bound.
 
@@
 Supermassive black holes are known to lie at the centres of most galaxies and the black hole mergers are associated with the mergers of the host galaxies \citep{Volonteri2003,Ferrarese2005}.
-The current best published limit for the amplitude of the stochastic background is $h_\mathrm{c} = 6\times 10^{-15}$ at a frequency of $f_{0}=1~\mathrm{yr}^{-1}$ \citep{Haasteren}.
+The current best published limit for the amplitude of the stochastic background is $h_\mathrm{c} = 6\times 10^{-15}$ at a frequency of $f_{0}=1~\mathrm{yr}^{-1}$ \citep{Haasteren} \cplb{Is this still the best?}.
 There is strong theoretical evidence that the actual background lies close to the current limit \citep{Sesana-2012}.

### 2014-07-31 Typesofsource.tex
Small changes

--- 2014-07-29 Typesofsource.tex
@@
 where an overdot represents the time derivative and $\dot{\phi} = 2\pi f$.
-The squared SNR scales with ${N}_{\mathrm{cycles}}$, so it would be expected that $h_\mathrm{c}(f)\approx \sqrt{{N}_{\mathrm{cycles}}}\left|\tilde{h}(f)\right|$.
+The squared SNR scales with ${N}_{\mathrm{cycles}}$, so it would be expected that $h_\mathrm{c}(f)\approx \sqrt{{N}_{\mathrm{cycles}}}|\tilde{h}(f)|$.
 
 The form for $h_\mathrm{c}$ can be derived from the Fourier transform in the stationary-phase approximation.
-Consider a source signal with approximately constant amplitude, $h_0$, and central frequency $f'$.
+Consider a source signal with approximately constant amplitude $h_0$ and central frequency $f'$.
 In this case,
@@
 fS_{\mathrm{E}}(f) = \int_{0}^{\infty} \rmd z\; \frac{\mathrm{d}\nu}{\mathrm{d}z}\frac{1}{(1+z)}\frac{1}{d_{\mathrm{L}}^{2}(z)}\frac{\mathrm{d}E_{\mathrm{GW}}}{\mathrm{d}\left(\log f_{\mathrm{r}} \right)} \; , \end{equation}
-where the factor of $\left( 1+z \right)^{-1}$ accounts for the redshift of the energy.
+where the factor of $\left( 1+z \right)^{-1}$ accounts for the redshifting of the energy.
 
@@
 The energy radiated may then be calculated using the quadrupole approximation \citep{petersmathews1963}.
-The energy in GWs from a single binary per logarithmic frequency interval is,
+The energy in GWs from a single binary per logarithmic frequency interval is
 \begin{equation}\label{eq:Thorne}
-\frac{\mathrm{d}E_{\mathrm{GW}}}{\mathrm{d}\left(\log f_{\mathrm{r}} \right)} = \frac{G^{2/3}\pi^{2/3}}{3}{\cal{M}}^{5/3}f_{\mathrm{r}}^{2/3} \; ,
+\frac{\mathrm{d}E_{\mathrm{GW}}}{\mathrm{d}\left(\log f_{\mathrm{r}} \right)} = \frac{G^{2/3}\pi^{2/3}}{3}{\cal{M}}^{5/3}f_{\mathrm{r}}^{2/3}
 \end{equation}
@@
 An alternative method for graphically representing the sensitivity of a GW detector to stochastic backgrounds, called the \emph{power-law-integrated sensitivity curve}, was suggested by \cite{2013PhRvD..88l4032T}.
-This method accounts for the fact that there is power across all frequencies in the sensitivity band by integrating the noise weighted signal over frequency.
-As our aim here is to present stochastic backgrounds alongside other types of sources for comparison we do not use this approach.
+This method accounts for the there being power across all frequencies in the sensitivity band by integrating the noise-weighted signal over frequency.
+As our aim here is to present stochastic backgrounds alongside other types of sources for comparison, we do not use this approach.
 
@@
 where ${N}_{\mathrm{cycles}}$ is the number of cycles of radiation generated by the source, which is of order unity for bursts.
-Due to the uncertainty of burst models, the waveform is commonly indicated by a constant amplitude across some finite bandwidth.
-This is analogous to computing the root-sum-square of the waveform polarisations
+
+An alternative characterisation of the signal amplitude commonly used for burst sources is the root-sum-square of the waveform polarisations:
 \begin{equation}
-h_\mathrm{rss} = \sqrt{\int{|h_+(t)|}^2 + {|h_\times(t)|}^2\,\rmd t},
+h_\mathrm{rss} = \left[\int \rmd t {|h_+(t)|}^2 + {|h_\times(t)|}^2\right]^{1/2}.
 \end{equation}
-which is a quantity that is used to characterise the sensitivity of detectors to bursts \citep{Bursts}.
-In this work, we favour a constant $h_\mathrm{c}(f)$ over $h_\mathrm{rss}$ for consistency with the other types of source.
+If $\tilde{h}(f)$ is constant across the bandwidth $\Delta f$, this is related to the characteristic strain via \cplb{What about the detector response?}.
+In this work, we favour a constant $h_\mathrm{c}(f)$ rather than $h_\mathrm{rss}$ for consistency with the other types of source where the bandwidth is detector specific.

### 2014-07-31 Detectors.tex
Christopher's to-do list

@@
 References for the noise curves used for individual detectors can be found in the relevant subsections, further information about the detectors can be in discovered these.
+Detectors are frequently upgraded and redesigned, hence, while these curves are believed to be correct at the time of writing, it is best to check for updates from the appropriate science teams before relying on the details given here, although we hope that they shall remain accurate enough for illustrative purposes.
 

### 2014-07-31 Sources.tex
Christopher's to-do list

@@
 The signals are near monochromatic with a frequency twice the rotation frequency of the neutron star, and are a potential source for ground-based detectors.
-The amplitude of the GWs depends upon the deformation of the neutron star, which itself depends upon the neutron star equation of state.
-This is currently uncertain \citep{Lattimer2012}.
+The amplitude of the GWs depends upon the deformation of the neutron star.
+The magnitude of the distortion depends upon the neutron star equation of state and the stiffness of the crust, which are currently uncertain \citep{Chamel2008,Lattimer2012}; there is a maximum deformation that be supported by a given set of material properties for the neutron-star structure.
+Deformations can also be supported by internal magnetic fields \citep{Haskell2008}; since the magnetic field can be arbitrarily small, there is no lower limit on the size of the deformation.
 Several known pulsars could be sources for the advanced detectors and upper limits from the initial detectors help to constrain the deformations.

### 2014-07-31 Typesofsource.tex
Christopher's to-do list

@@
 The form for $h_\mathrm{c}$ can be derived from the Fourier transform in the stationary-phase approximation.
-Consider a source signal with approximately constant amplitude $h_0$ and central frequency $f'$.
+Consider a source signal with approximately constant (root-mean-square) amplitude $h_0$ and central frequency $f'$.
 In this case,
 \begin{eqnarray}
-h(t) &=& h_0\exp\left[\rmi \phi(t)\right] \; ; \\
-\tilde{h}(f) &=& \int_{-\infty}^{\infty}\rmd t\; h_0\exp \left\{2\pi \rmi \left[\frac{\phi(t)}{2\pi t}-f\right]t\right\} \; .
+h(t) &=& \sqrt{2} h_0 \cos\left[\rmi \phi(t)\right] \; ; \\
+\tilde{h}(f) &=& \frac{h_0}{\sqrt{2}} \int_{-\infty}^{\infty}\rmd t\; \exp \left\{2\pi \rmi \left[\frac{\phi(t)}{2\pi t}-f\right]t\right\} + \exp \left\{-2\pi \rmi \left[\frac{\phi(t)}{2\pi t}+f\right]t\right\} \; .
+\label{eq:fourier-h}
 \end{eqnarray}
-Without loss of generality, we can assume an initial phase of zero, such that $\phi(t) \approx 2\pi f't$.
-The largest contribution to the integral comes from where the argument of the exponential is approximately zero.
-Expanding the exponent in powers of $t$ about $f = f'$,
+Without loss of generality, we can assume an initial phase of zero, such that $\phi(0) = 0$.
+The largest contribution to the integral comes from where the argument of the exponentials is approximately zero.
+For the first term, this occurs when $f = f'$, then the term in brackets is
+\begin{equation}
+\left[\frac{\phi(t)}{2\pi t}-f\right]_{f\,=\,f'} = f' + \dot{f'}t + \mathcal{O}\left(t^2\right) - f' = \dot{f'}t + \mathcal{O}\left(t^2\right).
+\end{equation}
+The higher-order terms cause the expoential to oscillate rapidly such that these terms integrate to zero and may be neglected.
+Performing a similar expansion about $f = -f'$ for the second term in (\ref{eq:fourier-h}) gives
 \begin{eqnarray} \label{eq:FTofEMRI}
-\tilde{h}(f) &\simeq& \int_{-\infty}^{\infty}\rmd t\; h_0 \exp\left[2\pi\rmi \dot{f'} \left(t^{2}+\frac{f'-f}{\dot{f'}}t\right)\right]\nonumber \\
-&\simeq& h_0\exp\left[-\pi\rmi\frac{\left(f'-f\right)^{2}}{2\dot{f'}}\right]\int_{-\infty}^{\infty}\rmd t\;\exp\left[2\pi\rmi\dot{f'}\left(t+\frac{f'-f}{2\dot{f'}}\right)^{2}\right] \nonumber \\
-&\simeq& \frac{h_0}{\sqrt{2\rmi\dot{f'}}}\exp\left[-\pi\rmi\frac{\left(f'-f\right)^{2}}{2\dot{f'}}\right]\; .
+\tilde{h}(f) &\simeq& \frac{h_0}{\sqrt{2}} \int_{-\infty}^{\infty}\rmd t\; \exp\left(2\pi\rmi \dot{f'} t^{2}\right) + \exp\left(-2\pi\rmi \dot{f'} t^{2}\right) \nonumber \\
+&\simeq& \frac{h_0}{\sqrt{2\dot{f'}}}.
 \end{eqnarray}
@@
 Equation (\ref{eq:hc}) should be considered as the definition of characteristic strain and (\ref{eq:insphc}) a consequence of it for inspirals.
-Equation (\ref{eq:insphc}) is the relation between $h_\mathrm{c}(f)$ and the instantaneous amplitude $h_0$ for an inspiralling source; for other types of source a new relation satisfying (\ref{eq:hc}) has to be found.
+Equation (\ref{eq:insphc}) is the relation between $h_\mathrm{c}(f)$ and the instantaneous root-mean-square amplitude $h_0$ for an inspiralling source; for other types of source a new relation satisfying (\ref{eq:hc}) has to be found.
 
@@
 \begin{equation}\label{eq:stoch}
-fS_{\mathrm{E}}(f)=\frac{\pi c^{2}}{4G}f^{2}h_\mathrm{c}^{2}(f) \; .
+fS_{\mathrm{E}}(f)=\frac{\pi c^{2}}{4G}f^{2}\left[h_\mathrm{c}(f)\right]^2 \; .
 \end{equation}
@@
 \begin{equation}\label{eq:Phinney}
-fS_{\mathrm{E}}(f) = \int_{0}^{\infty} \rmd z\; \frac{\mathrm{d}\nu}{\mathrm{d}z}\frac{1}{(1+z)}\frac{1}{d_{\mathrm{L}}^{2}(z)}\frac{\mathrm{d}E_{\mathrm{GW}}}{\mathrm{d}\left(\log f_{\mathrm{r}} \right)} \; , \end{equation}
+fS_{\mathrm{E}}(f) = \int_{0}^{\infty} \rmd z\; \frac{\mathrm{d}\nu}{\mathrm{d}z}\frac{1}{(1+z)}\frac{1}{\left[d_{\mathrm{L}}(z)\right]^2}\frac{\mathrm{d}E_{\mathrm{GW}}}{\mathrm{d}\left(\log f_{\mathrm{r}} \right)} \; , \end{equation}
 where the factor of $\left( 1+z \right)^{-1}$ accounts for the redshifting of the energy.
@@
 \begin{equation}\label{eq:bigint}
-h_\mathrm{c}^{2}(f) = \frac{4G^{5/3}}{3\pi^{1/3}c^{2}}f^{-4/3}\int_{0}^{\infty}\mathrm{d}z\;\int_{0}^{\infty}\mathrm{d}{\cal{M}}\;\frac{\mathrm{d}^{2}\nu}{\mathrm{d}z\,\mathrm{d}{\mathcal{M}}}\frac{1}{d_{\mathrm{L}}^{2}(z)}\left( \frac{{\mathcal{M}}^{5}}{1+z} \right)^{1/3}\; .
+\left[h_\mathrm{c}(f)\right]^2 = \frac{4G^{5/3}}{3\pi^{1/3}c^{2}}f^{-4/3}\int_{0}^{\infty}\mathrm{d}z\;\int_{0}^{\infty}\mathrm{d}{\cal{M}}\;\frac{\mathrm{d}^{2}\nu}{\mathrm{d}z\,\mathrm{d}{\mathcal{M}}}\frac{1}{\left[d_{\mathrm{L}}(z)\right]^2}\left( \frac{{\mathcal{M}}^{5}}{1+z} \right)^{1/3}\; .
 \end{equation}
@@
 \begin{equation}\label{eq:simple} 
-h_\mathrm{c}(f) = \left|\tilde{h}(f)\right|\sqrt{{N}_{\mathrm{cycles}}} \; , 
+h_\mathrm{c}(f) = \sqrt{{N}_{\mathrm{cycles}}}\left|\tilde{h}(f)\right| \; , 
 \end{equation}

### 2014-08-01 Detectors.tex
RHC and CJM alterations

@@
 The most notable ground-based detectors are LIGO and Virgo, which work in collaboration, supported by GEO600.
-LIGO, Virgo and GEO600 have completed science runs as first-generation detectors, now referred to as Initial LIGO (iLIGO) and Initial Virgo (iVirgo).
-They are currently being upgraded to second-generation detectors, advanced LIGO (aLIGO) and advanced Virgo (AdV).
+LIGO, Virgo and GEO600 have completed science runs as first-generation detectors.
+As LIGO and Virgo are currently being upgraded their initial configurations are now referred to as Initial LIGO (iLIGO) and Initial Virgo (iVirgo) respectively.
+The upgraded, second-generation versions are referred to as advanced LIGO (aLIGO) and advanced Virgo (AdV) respectively.
 LIGO has two observatories: one at Hanford, Washington, which has two detectors; and another at Livingston, Louisiana.
 There is an agreement to move one of the upgraded Hanford detector systems to a location in India \citep{LIGO-India,Unnikrishnan2013}.
-The GEO600 detector can also operate in a high frequency mode \citep[GEO-HF]{2006CQGra..23S.207W}, this is not included in Table \ref{table:t} \cplb{Why not?}.
+The GEO600 detector can also operate in a high frequency mode \citep[GEO-HF]{2006CQGra..23S.207W}.
 The advanced detectors should start operation in the next couple of years, with LIGO-India following further in the future.

### 2014-08-01 Typesofsource.tex
RHC and CJM alterations

@@
 \begin{eqnarray}
-h(t) &=& \sqrt{2} h_0 \cos\left[\rmi \phi(t)\right] \; ; \\
+h(t) &=& \sqrt{2} h_0 \cos\left[\phi(t)\right] \; , \\
 \tilde{h}(f) &=& \frac{h_0}{\sqrt{2}} \int_{-\infty}^{\infty}\rmd t\; \exp \left\{2\pi \rmi \left[\frac{\phi(t)}{2\pi t}-f\right]t\right\} + \exp \left\{-2\pi \rmi \left[\frac{\phi(t)}{2\pi t}+f\right]t\right\} \; .
@@
 The higher-order terms cause the expoential to oscillate rapidly such that these terms integrate to zero and may be neglected.
-Performing a similar expansion about $f = -f'$ for the second term in (\ref{eq:fourier-h}) gives
+Performing a similar expansion about $f = -f'$ for the second term in (\ref{eq:fourier-h}) and then evaluating the Gaussian integrals gives
 \begin{eqnarray} \label{eq:FTofEMRI}
@@
 \end{equation}
-If $\tilde{h}(f)$ is constant across the bandwidth $\Delta f$, this is related to the characteristic strain via \cplb{What about the detector response?}.
+For a linearly polarised GW, with $\tilde{h}(f)$ constant across the bandwidth $\Delta f$, this is approximately related to the characteristic strain via 
+\begin{equation}
+h_\mathrm{rss} \simeq |\tilde{h}(f)|\sqrt{\Delta f}\,,\label{eq:rela}
+\end{equation}
+where we have neglected the detector response functions (see section \ref{sec:principles}), which are of order unity.
 In this work, we favour a constant $h_\mathrm{c}(f)$ rather than $h_\mathrm{rss}$ for consistency with the other types of source where the bandwidth is detector specific.

### 2014-08-01 Detectors.tex
updated PTA refs

@@
 The PTAs currently in operation are the European Pulsar Timing Array \citep[EPTA\footnote{\url{http://www.epta.eu.org/}};][]{eptareview2013}, the Parkes Pulsar Timing Array \citep[PPTA\footnote{\url{http://www.atnf.csiro.au/research/pulsar/ppta/}};][]{parkesreview2013} based in Australia, and the North American Nanohertz Observatory for Gravitational waves \citep[NANOGrav\footnote{\url{http://nanograv.org/}};][]{nanogravreview2013}).
-There are published limits on the amplitude of the stochastic background from all three detectors: the most recent from EPTA is \citet{Haasteren}, from PPTA is \citet{Shannon2013} and from NANOGrav is \citet{Arzoumanian2014}.
+There are published limits on the amplitude of the stochastic background from all three detectors: the most recent from EPTA is \citet{Haasteren}, from PPTA is \citet{Shannon2013} and from NANOGrav is \citet{2013ApJ...762...94D}.
 These are all comparable and we use the EPTA limits in \ref{app:a}, the figures show the published limit on $A$ from equation \ref{eq:power} at a frequency of $f_{0}=1\,\textrm{year}^{-1}$, based on an analysis of five pulsars over approximately ten years.

### 2014-08-01 Sources.tex
updated PTA refs

--- 2014-07-31 Sources.tex
@@
 \section{Astrophysical sources}\label{sec:sources}
-
-%All the sources described here are represented by shaded boxes in \ref{app:a}.
-The boxes are drawn in such a way that there is a reasonable event rate for sources at a detectable SNR.
-Descriptions of each source are given in the following sections.
-Sources with short durations (i.e., burst sources) and sources that evolve in time over much longer timescales than our observations are drawn with flat-topped boxes for $h_\mathrm{c}(f)$.
-Inspiralling binaries, or stochastic backgrounds of binaries, are drawn with a sloping top proportional to $f^{-2/3}$ for $h_\mathrm{c}(f)$, which is result derived in section \ref{sec:stoch}.
-The width of the box gives the range of frequencies sources of a given type can have while remaining at a detectable amplitude.
 
@@
 The signals are near monochromatic with a frequency twice the rotation frequency of the neutron star, and are a potential source for ground-based detectors.
-The amplitude of the GWs depends upon the deformation of the neutron star, which itself depends upon the neutron star equation of state.
-This is currently uncertain \citep{Lattimer2012}.
+The amplitude of the GWs depends upon the deformation of the neutron star.
+The magnitude of the distortion depends upon the neutron star equation of state and the stiffness of the crust, which are currently uncertain \citep{Chamel2008,Lattimer2012}; deformations can also be supported by internal magnetic fields \citep{Haskell2008}.
 Several known pulsars could be sources for the advanced detectors and upper limits from the initial detectors help to constrain the deformations.
@@
 Supermassive black holes are known to lie at the centres of most galaxies and the black hole mergers are associated with the mergers of the host galaxies \citep{Volonteri2003,Ferrarese2005}.
-The current best published limit for the amplitude of the stochastic background is $h_\mathrm{c} = 6\times 10^{-15}$ at a frequency of $f_{0}=1~\mathrm{yr}^{-1}$ \citep{Haasteren} \cplb{Is this still the best?}.
+The current best 95\% confidence limit for the amplitude of the stochastic background is $h_\mathrm{c} = 2.7\times 10^{-15}$ at a frequency of $f_{0}=1~\mathrm{yr}^{-1}$ \citep{Shannon2013}.
 There is strong theoretical evidence that the actual background lies close to the current limit \citep{Sesana-2012}.

### 2014-08-02 noise_curve.tex
Small changes
updated PTA refs
Bibliography tidying

@@
 We also thank Stephen Taylor and Justin Ellis for useful conversations about pulsar timing.
-The authors thanks a number of members of the LIGO--Virgo Scientific Collaboration for useful comments and suggestions.
+The authors thanks a number of members of the LIGO--Virgo Scientific Collaboration, especially Chris Messenger, for useful comments and suggestions.
 This paper has been assigned LIGO document LIGO-P1400129.}
@@
 \input{Plots}
+\clearpage
 
-\bibliographystyle{apj}
-\bibliography{bibliography,apj-jour}    %% includes the journal abbrevs
-%\bibliography{bibliography}
+\bibliographystyle{jphysicsB}
+\bibliography{bibliography}
 

### 2014-08-04 arXiv_version/noise_curve.tex
arXivversion

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    INSTITUTE OF PHYSICS PUBLISHING                                   %
%                                                                      %
%   `Preparing an article for publication in an Institute of Physics   %
%    Publishing journal using LaTeX'                                   %
%                                                                      %
%    LaTeX source code `ioplau2e.tex' used to generate `author         %
%    guidelines', the documentation explaining and demonstrating use   %
%    of the Institute of Physics Publishing LaTeX preprint files       %
%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
%                                                                      %
%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[fleqn,12pt]{iopart}
\newcommand{\gguide}{{\it Preparing graphics for IOP journals}}
%Uncomment next line if AMS fonts required
\usepackage{iopams}  
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{url}
\usepackage{natbib}
\usepackage{appendix}
\usepackage[margin=1cm]{caption}
\usepackage{braket}
\usepackage{fullpage}
\usepackage{hyperref}
\usepackage{color}

\def\newblock{\hskip .11em plus .33em minus .07em}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% 	definition of a solar mass symbol for use in equations.      %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\Msun}{\ensuremath{M_{ \odot }}}
\newcommand{\degree}{\ensuremath{^\circ}}
\newcommand{\innerprod}[2]{\ensuremath{\left({#1}\middle|{#2}\right)}}
\newcommand{\cplb}[1]{{\color{red} #1}}
\newcommand{\rhc}[1]{{\color{green} #1}}
\newcommand{\cjm}[1]{{\color{blue} #1}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\title{Gravitational wave sensitivity curves}
\author{C J Moore$^1$, R H Cole$^1$ \& C P L Berry$^{1,\,2}$}
\address{$^1$ Institute of Astronomy, Madingley Road, Cambridge, CB3 0HA, United Kingdom}
\address{$^2$ School of Physics and Astronomy, University of Birmingham, Edgbaston, Birmingham B15 2TT, United Kingdom}
\ead{cjm96@ast.cam.ac.uk}

\begin{abstract}
There are several common conventions in use by the gravitational wave community to describe the amplitude of sources and the sensitivity of detectors. These are frequently confused. We outline the merits of and differences between the various quantities used for parameterizing noise curves and characterizing gravitational wave amplitudes. We conclude by producing plots that consistently compare different detectors. Similar figures can be generated on-line for general use at \url{http://www.ast.cam.ac.uk/~rhc26/sources/}.
\end{abstract}

\pacs{04.30.--w, 04.30.Db, 04.80.Nn, 95.55.Ym}

\section{Introduction}

The next few years  promise to deliver the first direct detection of gravitational waves (GWs). This will most likely be achieved by the advanced Laser Interferometer Gravitational-wave Observatory \citep[LIGO;][]{2010CQGra..27h4006H} and advanced Virgo \citep{Acernese2009} detectors operating in the frequency range $(10$--$10^{3})~\mathrm{Hz}$. By the end of the decade, it is expected that pulsar timing arrays \citep[PTAs;][]{1990ApJ361300F} will also detect very low frequency GWs around $10^{-8}~\mathrm{Hz}$. Further into the future, space-based detectors, such as the evolved Laser Interferometer Space Antenna \citep[eLISA;][]{TheGravitationalUniverse}, will probe GWs in the millihertz regime. These advances shall herald the beginning of multi-wavelength GW astronomy as a means of observing the Universe.

There already exists an extensive literature assessing the potential of all of these detectors to probe the astrophysics of various sources. There are several different methods commonly used to describe the sensitivity of a GW detector and the strength of a GW source. It is common practice to summarise this information on a sensitivity-curve plot. When producing these plots, it is desirable to have a consistent convention between detectors and sources that allows information about both to be plotted on the same graph. Ideally, the detectors and sources are represented in such a way that the relative detectability of the signals is immediately apparent.

In this work, we discuss the differing conventions commonly used in GW astronomy. The amplitude of a GW is a strain, a dimensionless quantity $h$. This gives a fractional change in length, or equivalently light travel time, across a detector. The strain is small, making it a challenge to measure: we are yet to obtain a direct detection of a GW. To calibrate our expectations for future detections it is necessary to quantify the sensitivity of our instruments and the strength of their target signals. When discussing the loudness of sources and the sensitivity of detectors there are three commonly used parametrizations based upon the strain: the characteristic strain, the power spectral density (PSD) and the spectral energy density. We aim to disambiguate these three and give a concrete comparison of different detectors. It is hoped that this will provide a useful reference for new and experienced researchers in this field alike.

We begin by expounding the various conventions and the relationships between the conventions in sections \ref{sec:conventions} and \ref{sec:voc}. A review of GW detectors (both current and proposed) is given in section \ref{sec:detectors} and a review of GW sources is given in section \ref{sec:sources}. In \ref{app:a} several example sensitivity curves are presented. A website where similar figures can be generated is available at \url{www.ast.cam.ac.uk/~rhc26/sources/}. Here, the user may select which sources and detectors to include to tailor the figure to their specific requirements.

\section{Signal parametrization}\label{sec:conventions}

\subsection{Signal analysis preliminaries}

Gravitational radiation has two independent polarization states denoted $+$ and $\times$; a general signal can be described as a linear combination of the two polarization states, $h \nobreak=\nobreak A_{+} h_{+}\nobreak +\nobreak A_{\times} h_{\times}$. The sensitivity of a detector to these depends upon the relative orientations of the source and detector. The output of a gravitational wave detector $s(t)$ contains a superposition of noise $n(t)$ and (possibly) a signal $h(t)$,
\begin{equation}
s(t) = n(t)+h(t) \; .
\end{equation}

We shall have recourse to work with the Fourier transform of the signal, using the conventions that
\begin{eqnarray} \label{eq:defFourier}
\tilde{x}(f) &= {\mathcal{F}}\left\{x(t)\right\}(f) &= \int_{-\infty}^{\infty}\mathrm{d}t\;x(t)\exp (-2\pi \rmi f t ) \; , \\
x(t) &= {\mathcal{F}}^{-1}\left\{\tilde{x}(f)\right\}(t) &= \int_{-\infty}^{\infty}\mathrm{d}f\;\tilde{x}(f)\exp (2\pi \rmi f t ) \; .
\end{eqnarray}

For simplicity, it is assumed that the noise in the GW detector is stationary and Gaussian (with zero mean); under these assumptions the noise is fully characterised via the one-sided noise PSD $S_{n}(f)$, 
\begin{equation}\label{eq:psd}
\left<\tilde{n}(f)\tilde{n}^{*}(f')\right>=\frac{1}{2}\delta (f-f')S_{n}(f) \; ,
\end{equation}
where angle brackets $\left<\ldots\right>$ denote an ensemble average over many noise realisations \citep{Cutler1994}. In reality, we have only a single realisation to work with, but the ensemble average can be replaced by a time average for stationary stochastic noise. The procedure is to measure the noise over a sufficiently long duration $T$ and then compute the Fourier transform $\tilde{n}(f)$ with a frequency resolution $\Delta f = T^{-1}$; this is repeated many times to give an average. The noise PSD $S_{n}(f)$ has units of inverse frequency.

Since the GW signal and detector output are both real it follows that $\tilde{h}(-f)=\tilde{h}^{*}(f)$ and $\tilde{n}(-f)=\tilde{n}^{*}(f)$; therefore, $S_{n}(f)=S_{n}(-f)$. The fact that $S_{n}(f)$ is an even function means that Fourier integrals over all frequencies can instead be written as integrals over positive frequencies only, e.g., (\ref{eq:meansquare1}) and (\ref{eq:snrinnerprod}); it is for this reason that $S_{n}(f)$ is called the \emph{one-sided} PSD.\footnote{An alternative convention is to use the two-sided PSD $S^{(2)}_{n}(f) = S_{n}(f)/2$.}

When integrated over all positive frequencies, the PSD gives the mean square noise amplitude. Starting by taking the time average of the square of the detector noise:
\begin{eqnarray}
\overline{\left|n(t)\right|^{2}} &= \lim_{T\rightarrow\infty}\frac{1}{2T} & \int_{-T}^{T}\mathrm{d}t\; n(t)n^{*}(t) \\
 &= \lim_{T\rightarrow\infty}\frac{1}{2T} & \int_{-T}^{T}\mathrm{d}t\; \int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\tilde{n}(f)\tilde{n}^{*}(f')\exp\left(2\pi \rmi ft\right)\exp\left(-2\pi \rmi f't\right) \nonumber\\
 &= \lim_{T\rightarrow\infty}\frac{1}{2T} & \int_{-T}^{T}\mathrm{d}t\; \int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;{\mathcal{F}}\left\{ n(\tau) \right\}(f)\left[{\mathcal{F}}\left\{ n(\tau) \right\}(f')\right]^{*} \nonumber \\*
\label{eq:meansquare} & & \times \hphantom{\int_{-T}^{T}} \exp\left(2\pi \rmi ft\right)\exp\left(-2\pi \rmi f't\right)\;,
\end{eqnarray}
where we have substituted in using the definitions of the Fourier transform and its inverse. A property of Fourier transforms is that a time-domain translation by amount $t$ is equivalent to a frequency-domain phase change $2\pi ft$; if ${\mathcal{F}}\left\{ n(\tau) \right\}(f) = \tilde{n}(f)$, then ${\mathcal{F}}\left\{ n(\tau-t) \right\}(f) = \tilde{n}(f)\exp (2\pi \rmi ft )$. Therefore, the exponential factors in (\ref{eq:meansquare}) may be absorbed as
\begin{eqnarray}
\overline{\left|n(t)\right|^{2}} &=& \lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T} \mathrm{d}t\:\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;{\mathcal{F}}\left\{ n(\tau-t) \right\}(f)\left[{\mathcal{F}}\left\{ n(\tau-t) \right\}(f')\right]^{*} \, .
\end{eqnarray}
Since the noise is a randomly varying signal, we can use the ergodic principle to equate a time average, denoted by $\overline{\left(\ldots\right)}$, with an ensemble average, denoted by $\left<\ldots\right>$. The noise is stationary, consequently, its expectation value is unchanged by the time-translation performed above. Therefore, using (\ref{eq:psd}), the mean square noise amplitude is given by
\begin{eqnarray}\label{eq:meansquare1}
\overline{\left|n(t)\right|^{2}} &=& \int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\left<\tilde{n}(f)\tilde{n}^{*}(f')\right> \\
&=& \int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\frac{1}{2}S_{n}(f)\delta(f-f') \nonumber \\
&=& \int_{0}^{\infty}\mathrm{d}f\; S_{n}(f) \;.
\end{eqnarray}

Given a detector output, the challenge is to extract the signal. There is a well known solution to this problem that involves constructing a Wiener optimal filter \citep{Wiener49}. Let $K(t)$ be a real filter function with Fourier transform $\tilde{K}(f)$. Convolving this with the detector output gives a contribution from the signal and a contribution from the noise,
\begin{equation}\label{eq:conv}
\left(s*K\right)(\tau) = \int_{-\infty}^{\infty}\mathrm{d}t\;\left[h(t)+n(t)\right]K(t-\tau) \approx \mathcal{S} + \mathcal{N} \; .
\end{equation}
The signal contribution $\mathcal{S}$ is defined as the expectation of the convolution in (\ref{eq:conv}) when a signal is present, maximised by varying the offset to achieve the best overlap with the data. Since the expectation of pure noise is zero it follows that
\begin{equation}
\mathcal{S} = \int_{-\infty}^{\infty}\mathrm{d}t\;h(t)K(t)=\int_{-\infty}^{\infty}\mathrm{d}t\;h(t)K^{*}(t)=\int_{-\infty}^{\infty}\mathrm{d}f\; \tilde{h}(f)\tilde{K}^{*}(f) \; .
\end{equation}
The squared contribution from noise $\mathcal{N}^{2}$ is defined as the mean square of the convolution in (\ref{eq:conv}) when no signal is present,
\begin{eqnarray} 
\mathcal{N}^{2} &= \int_{-\infty}^{\infty}\mathrm{d}t\;\int_{-\infty}^{\infty}\mathrm{d}t'\;K(t)K(t')\left<n(t)n(t')\right> \nonumber \\
 &= \int_{-\infty}^{\infty}\mathrm{d}t\;\int_{-\infty}^{\infty}\mathrm{d}t'\;K(t)K^{*}(t')\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\left<\tilde{n}(f)\tilde{n}^{*}(f')\right>\exp\left[2\pi\rmi(ft-f't')\right]\nonumber \\
 &= \int_{-\infty}^{\infty}\mathrm{d}f \; \frac{1}{2}S_{n}(f)\tilde{K}(f)\tilde{K}^{*}(f) \; ,
 \end{eqnarray}
using the definition of $S_{n}(f)$ from (\ref{eq:psd}). Hence the signal-to-noise ratio (SNR) $\varrho$ is given by
\begin{equation}\label{eq:SNRinnerprod} 
\varrho^{2} = \frac{\mathcal{S}^{2}}{\mathcal{N}^{2}}= \frac{\innerprod{\frac{1}{2}S_{n}(f)\tilde{K}(f)}{\tilde{h}(f)}^{2}}{\innerprod{\frac{1}{2}S_{n}(f)\tilde{K}(f)}{\frac{1}{2}S_{n}(f)\tilde{K}(f)}},
\end{equation}
where we have introduced the inner product between signal $\tilde{A}$ and $\tilde{B}$ as \citep{Finn1992}
\begin{equation}\label{eq:snrinnerprod} \innerprod{\tilde{A}(f)}{\tilde{B}(f)} = 4\Re\left\{\int_{0}^{\infty}\mathrm{d}f\;\frac{\tilde{A}^{*}(f)\tilde{B}(f)}{S_{n}(f)}\right\} \; .\end{equation}
The optimum filter is that function $\tilde{K}(f)$ which maximises the SNR in (\ref{eq:SNRinnerprod}). From the Cauchy--Schwarz inequality, it follows that the optimum filter is
\begin{equation}
\tilde{K}(f)=\frac{\tilde{h}(f)}{S_{n}(f)} \; .
\end{equation}
This is the Wiener filter, which may be multiplied by an arbitrary constant since this does not change the SNR. Using this form for $\tilde{K}(f)$, the squared SNR is
\begin{equation}
\varrho^2 = \int_0^\infty\mathrm{d}f \frac{4| \tilde{h}(f)|^{2}}{S_n(f)} = \innerprod{\tilde{h}(f)}{\tilde{h}(f)}.
\label{eq:traditionalSNR} 
\end{equation}
In order to construct the Wiener filter, it is necessary to know \emph{a priori} the form of the signal, $\tilde{h}(f)$, for this reason the Wiener filter is sometimes called the \emph{matched} filter.

Whilst the magnitude of the Fourier transform of the signal $|\tilde{h}(f)|$ provides a simple quantification of the GW amplitude as a function of frequency, it has one main deficiency. For an inspiralling source ,the instantaneous amplitude can be orders of magnitude below the noise level in a detector; however, as the signal continues over many orbits, the SNR can be integrated up to a detectable level. It is useful to have a quantification of the GW amplitude that accounts for this effect; three such methods are described in the following subsections.

\subsection{Characteristic strain}\label{sec:character-strain}

The characteristic strain $h_\mathrm{c}$ is designed to include the effect of integrating an inspiralling signal. Its counterpart for describing noise is the noise amplitude $h_n$. These are defined as
\begin{eqnarray}\label{eq:strain-hc} 
\left[h_\mathrm{c}(f)\right]^{2} &= 4f^{2}\left| \tilde{h}(f) \right|^{2} \; ,\\
\left[h_{n}(f)\right]^{2} &= fS_{n}(f),
\label{eq:strain-hn}
\end{eqnarray}
such that the SNR in (\ref{eq:traditionalSNR}) may be written
\begin{equation}\label{eq:hc} 
\varrho^{2} = \int_{-\infty}^{\infty} \mathrm{d}\left(\log f\right)\; \left[\frac{h_\mathrm{c}(f)}{h_{n}(f)}\right]^{2} \;.
\end{equation}
The strain amplitudes $h_\mathrm{c}(f)$ and $h_{n}(f)$ are dimensionless. Using this convention, when plotting on a log--log scale, the area between the source and detector curves is related to the SNR via (\ref{eq:hc}). This convention allows the reader to integrate by eye to assess the detectability of a given source (see figure \ref{fig:hc}).

An additional advantage of this convention is that the values on the strain axis for the detector curve $h_n(f)$ have a simple physical interpretation: they correspond to the root-mean-square noise in a bandwidth $f$. One downside to plotting characteristic strain is that the values on the strain axis $h_\mathrm{c}(f)$ do not directly relate to the amplitude of the waves from the source. Another disadvantage is that the characteristic strain for a monochromatic source is infinite.

\subsection{Power spectral density}\label{sec:psd}

A second commonly used quantity for sensitivity curves is the square root of the PSD or the amplitude spectral density (see figure \ref{fig:S}). When discussing a detector, rearranging (\ref{eq:strain-hn}) gives
\begin{equation}\label{eq:temp1}
\sqrt{S_{n}(f)} = h_{n}(f)f^{-1/2} \; ;
\end{equation}
by analogy, we can define an equivalent for source amplitudes
\begin{equation}
\sqrt{S_{h}(f)} = h_\mathrm{c}(f)f^{-1/2} = 2 f^{1/2} \left| \tilde{h}(f) \right| \; ,
\label{eq:ShforSources}
\end{equation}
where we have used (\ref{eq:strain-hc}). Both $\sqrt{S_{n}(f)}$ and $\sqrt{S_{h}(f)}$ have units of $\mathrm{Hz^{-1/2}}$. The root PSD is the most frequently plotted quantity in the literature.

The PSD, as defined by (\ref{eq:psd}), has the nice property, demonstrated in (\ref{eq:meansquare1}), that integrated over all positive frequencies it gives the mean square amplitude of the signal in the detector. However, in one important regard it is less appealing than characteristic strain: the height of the source above the detector curve is no longer directly related to the SNR.

\subsection{Energy density}\label{sec:energy-density}

A third way of describing the amplitude of a GW is through the energy carried by the waves. This has the advantage of having a clear physical significance. The energy density is most commonly used in sensitivity curves showing stochastic backgrounds of GWs (see section \ref{sec:stoch}).

The energy in GWs is described by the Isaacson stress--energy tensor \citep[section 35.15]{MTW}
\begin{equation}
T_{\mu\nu}=\frac{c^{4}}{32\pi G}\left<\partial_{\mu}\bar{h}_{\alpha\beta}\partial_{\nu}\bar{h}^{\alpha\beta}\right> \;,
\end{equation}
where the angle brackets denote averaging over several wavelengths or periods, and $\bar{h}_{\alpha\beta}$ is the transverse-traceless metric perturbation. The energy density $\rho c^{2}$ is given by the $T_{00}$ component of this tensor. Consequently \citep[cf.][]{Berry2013},
\begin{eqnarray}
\rho c^{2} &=& \frac{c^{2}}{16\pi G}\int_{-\infty}^{\infty}\mathrm{d}f\;\left(2\pi f\right)^{2}\tilde{h}(f)\tilde{h}^{*}(f) \\*
 &=& \int_{0}^{\infty}\mathrm{d}f\;\frac{\pi c^{2}}{4G}f^{2}S_{h}(f)\; , \label{eq:specNRGdensity}
\end{eqnarray} 
where the definition (\ref{eq:ShforSources}) have been used. The integrand in Equation (\ref{eq:specNRGdensity}) is defined as the spectral energy density, the energy per unit volume of space, per unit frequency \citep{HellingsDowns}
\begin{equation}\label{eq:spectralenergydensity}
S_{\mathrm{E}}(f)=\frac{\pi c^{2}}{4G} f^{2}S_{h}(f) \; ;
\end{equation}
a corresponding expression for the noise can be formulated by replacing $S_h(f)$ with $S_{n}(f)$.

Cosmological studies often work in terms of the dimensionless quantity $\Omega_{\mathrm{GW}}$, the energy density per logarithmic frequency interval normalised to the critical density of the universe $\rho_{\mathrm{c}}$,
\begin{equation}
\label{eq:omega}
\Omega_\mathrm{GW}(f) = \frac{fS_{\mathrm{E}}(f)}{\rho_{\mathrm{c}}c^{2}} \; .
\end{equation}
The critical density is
\begin{equation}
\label{eq:crit-density}
\rho_{\mathrm{c}}=\frac{3H_{0}^{2}}{8\pi G} \;,
\end{equation}
where $H_{0}$ is the Hubble constant, commonly parametrized as
\begin{equation}
H_0 = h_{100}\times 100~\mathrm{km\,s^{-1}\,Mpc^{-1}}.
\end{equation}
The reduced Hubble parameter $h_{100}$ has nothing to do with strain. The most common quantity related to energy density to be plotted on sensitivity curves is $\Omega_{\mathrm{GW}}h_{100}^{2}$ (figure \ref{fig:omega}) as this removes sensitivity to the (historically uncertain) measured value of the Hubble constant.

This quantity has one aesthetic advantage over the others: it automatically accounts for the fact that there is less energy in low frequency waves of the same amplitude. However, unlike characteristic strain, the area between the source and detector curves is no longer simply related to the SNR.

\subsection{Relating the different descriptions}
The dimensionless energy density in GWs $\Omega_{\mathrm{GW}}$, spectral energy density $S_{\mathrm{E}}$, one-sided PSD, $S_{h}$, characteristic strain $h_\mathrm{c}$ and frequency-domain strain $\tilde{h}(f)$ are related via
\begin{equation}\label{eq:differentdescriptions}
H_0^2\Omega_\mathrm{GW}(f)= \frac{8 \pi G}{3 c^{2}} fS_{\mathrm{E}}(f) = \frac{2\pi^2}{3} f^3 S_h(f) = \frac{2\pi^2}{3} f^2 \left[h_\mathrm{c}(f)\right]^2 = \frac{8\pi^2}{3} f^4 \left|\tilde{h}(f)\right|^2\; ,
\end{equation}
using (\ref{eq:strain-hc}), (\ref{eq:ShforSources}), (\ref{eq:spectralenergydensity}), (\ref{eq:omega}) and (\ref{eq:crit-density}).
Corresponding expressions for the noise are obtained by substituting $S_{n}(f)$ for $S_h(f)$, $h_{n}(f)$ for $h_\mathrm{c}(f)$ and $\tilde{n}(f)$ for $\tilde{h}(f)$.

\section{Types of source}\label{sec:voc}

GW signals can be broadly split into three categories: those from well-modelled sources, for which we have a description of the expected waveform; stochastic backgrounds, for which we can describe the statistical behaviour; and unmodelled (or poorly-modelled) transient sources. The classic example of a well-modelled source is the inspiral of two compact objects, this is discussed in section \ref{sec:insp}. Stochastic backgrounds can either be formed from many overlapping sources, which could be modelled individually, or from some intrinsically random process, these are discussed in section \ref{sec:stoch}. An example of an unmodelled (or poorly-modelled) transient source is a supernova; searches for signals of this type are often called burst searches and are discussed in section \ref{sec:bursts}.

\subsection{Inspirals}\label{sec:insp}

Inspiralling binaries may the most important GW source. They spend a variable amount of time in each frequency band. If $\phi$ is the orbital phase, then the number of cycles generated at frequency $f$ can be estimated as
\begin{equation}\label{eq:inspiral}
{N}_{\mathrm{cycles}} = \frac{f}{2\pi} \frac{\mathrm{d}\phi}{\mathrm{d}f} = \frac{f^{2}}{\dot{f}} \; ,
\end{equation}
where an overdot represents the time derivative and $\dot{\phi} = 2\pi f$. The squared SNR scales with ${N}_{\mathrm{cycles}}$, so it would be expected that $h_\mathrm{c}(f)\approx \sqrt{{N}_{\mathrm{cycles}}}|\tilde{h}(f)|$.

The form for $h_\mathrm{c}$ can be derived from the Fourier transform in the stationary-phase approximation. Consider a source signal with approximately constant (root-mean-square) amplitude $h_0$ and central frequency $f'$. In this case,
\begin{eqnarray}
h(t) &=& \sqrt{2} h_0 \cos\left[\phi(t)\right] \; , \\
\tilde{h}(f) &=& \frac{h_0}{\sqrt{2}} \int_{-\infty}^{\infty}\rmd t\; \exp \left\{2\pi \rmi \left[\frac{\phi(t)}{2\pi t}-f\right]t\right\} + \exp \left\{-2\pi \rmi \left[\frac{\phi(t)}{2\pi t}+f\right]t\right\} \; .
\label{eq:fourier-h}
\end{eqnarray}
Without loss of generality, we can assume an initial phase of zero, such that $\phi(0) = 0$. The largest contribution to the integral comes from where the argument of the exponentials is approximately zero. For the first term, this occurs when $f = f'$, then the term in brackets is
\begin{equation}
\left[\frac{\phi(t)}{2\pi t}-f\right]_{f\,=\,f'} = f' + \dot{f'}t + \mathcal{O}\left(t^2\right) - f' = \dot{f'}t + \mathcal{O}\left(t^2\right).
\end{equation}
The higher-order terms cause the expoential to oscillate rapidly such that these terms integrate to zero and may be neglected. Performing a similar expansion about $f = -f'$ for the second term in (\ref{eq:fourier-h}) and then evaluating the Gaussian integrals gives
\begin{eqnarray} \label{eq:FTofEMRI}
\tilde{h}(f) &\simeq& \frac{h_0}{\sqrt{2}} \int_{-\infty}^{\infty}\rmd t\; \exp\left(2\pi\rmi \dot{f'} t^{2}\right) + \exp\left(-2\pi\rmi \dot{f'} t^{2}\right) \nonumber \\
&\simeq& \frac{h_0}{\sqrt{2\dot{f'}}}.
\end{eqnarray}
From (\ref{eq:strain-hc}) and (\ref{eq:FTofEMRI}), the characteristic strain for inspiralling sources is given by \citep{FinnThorne}
\begin{equation}\label{eq:insphc}
h_\mathrm{c}(f) = \sqrt{\frac{2f^{2}}{\dot{f}}}h_0 \;.
\end{equation}
Equation (\ref{eq:hc}) should be considered as the definition of characteristic strain and (\ref{eq:insphc}) a consequence of it for inspirals. Equation (\ref{eq:insphc}) is the relation between $h_\mathrm{c}(f)$ and the instantaneous root-mean-square amplitude $h_0$ for an inspiralling source; for other types of source a new relation satisfying (\ref{eq:hc}) has to be found.

\subsection{Stochastic backgrounds}\label{sec:stoch}

Another important source of GWs is that of stochastic backgrounds, which can be produced from a large population of unresolvable sources. These can be at cosmological distances, where it is necessary to distinguish the frequency in the source rest frame $f_{\mathrm{r}}$ from the measured frequency $f$; the two are related through the redshift $z$ via $f_{\mathrm{r}}=(1+z)f$. The comoving number density of sources $\nu$ producing the background is also a function of redshift; if the sources producing the background are all in the local Universe, then simply set $\nu(z) = \nu_0\delta (z)$ and replace $d_{\mathrm{L}}(z)$ with $d$ in all that follows, where $d_{\mathrm{L}}(z)$ and $d$ are respectively the luminosity and comoving distances to the source, $d_{\mathrm{L}}(z)=(1+z)d$.

We shall assume that the individual sources are binaries, in which case the number density of sources is also a function of the component masses. It is convenient to work in terms of the chirp mass, defined as ${\mathcal{M}}=\mu^{3/5}M^{2/5}$, where $\mu$ is the reduced mass and $M$ is the total mass of the binary. The comoving number density of sources shall be represented by $\nu(z, \mathcal{M})$.

Equation (\ref{eq:differentdescriptions}) gives an expression for the energy density in GWs per logarithmic frequency interval,
\begin{equation}\label{eq:stoch}
fS_{\mathrm{E}}(f)=\frac{\pi c^{2}}{4G}f^{2}\left[h_\mathrm{c}(f)\right]^2 \; .
\end{equation}
The total energy emitted in the logarithmic frequency interval $\mathrm{d}\left(\log f_{\mathrm{r}}\right)$ by a single binary in the population is $\left[\mathrm{d}E_{\mathrm{GW}}/\mathrm{d}(\log f_{\mathrm{r}})\right]\mathrm{d}(\log f_{\mathrm{r}})$; the energy density may be written as
\begin{equation}\label{eq:Phinney}
fS_{\mathrm{E}}(f) = \int_{0}^{\infty} \rmd z\; \frac{\mathrm{d}\nu}{\mathrm{d}z}\frac{1}{(1+z)}\frac{1}{\left[d_{\mathrm{L}}(z)\right]^2}\frac{\mathrm{d}E_{\mathrm{GW}}}{\mathrm{d}\left(\log f_{\mathrm{r}} \right)} \; , \end{equation}
where the factor of $\left( 1+z \right)^{-1}$ accounts for the redshifting of the energy.

For simplicity, consider the background to comprise of binaries in circular orbits, with frequencies $f_\mathrm{GW} =f_{\mathrm{r}}/2$, which are far from their last stable orbit. The energy radiated may then be calculated using the quadrupole approximation \citep{petersmathews1963}. The energy in GWs from a single binary per logarithmic frequency interval is
\begin{equation}\label{eq:Thorne}
\frac{\mathrm{d}E_{\mathrm{GW}}}{\mathrm{d}\left(\log f_{\mathrm{r}} \right)} = \frac{G^{2/3}\pi^{2/3}}{3}{\cal{M}}^{5/3}f_{\mathrm{r}}^{2/3}
\end{equation}
between an minimum and maximum frequency set by the initial and final radius of the binary orbit. Here, we assume that the maximum and minimum frequencies are outside of the range of our detector and hence can be neglected. Using (\ref{eq:stoch}), (\ref{eq:Phinney}) and (\ref{eq:Thorne}), an expression for characteristic strain can now be found \citep{SesanaVecchioColancino}
\begin{equation}\label{eq:bigint}
\left[h_\mathrm{c}(f)\right]^2 = \frac{4G^{5/3}}{3\pi^{1/3}c^{2}}f^{-4/3}\int_{0}^{\infty}\mathrm{d}z\;\int_{0}^{\infty}\mathrm{d}{\cal{M}}\;\frac{\mathrm{d}^{2}\nu}{\mathrm{d}z\,\mathrm{d}{\mathcal{M}}}\frac{1}{\left[d_{\mathrm{L}}(z)\right]^2}\left( \frac{{\mathcal{M}}^{5}}{1+z} \right)^{1/3}\; .
\end{equation}

From (\ref{eq:bigint}) it can be seen that the characteristic strain due to a stochastic background of binaries is a power law in frequency with spectral index $\alpha=-2/3$. The amplitude of the background depends on the population statistics of the binaries under consideration via $\nu(z,{\mathcal{M}})$. The power law is often parametrised as
\begin{equation}\label{eq:power} 
h_\mathrm{c}(f) = A\left(\frac{f}{f_{0}}\right)^{\alpha}\; , 
\end{equation}
and constraints are then placed on $A$. In practice, this power law also has upper and lower frequency cut-offs related to the population of source objects. A stochastic background from other sources, such as cosmic strings or relic GWs from the early Universe, can also be written in the same form as (\ref{eq:power}), but with different spectral indices: $\alpha=-7/6$ for cosmic strings or $\alpha$ in the range $-1$ to $-0.8$ for relic GWs \citep{Jenet}.

An alternative method for graphically representing the sensitivity of a GW detector to stochastic backgrounds, called the \emph{power-law-integrated sensitivity curve}, was suggested by \cite{2013PhRvD..88l4032T}. This method accounts for the there being power across all frequencies in the sensitivity band by integrating the noise-weighted signal over frequency. As our aim here is to present stochastic backgrounds alongside other types of sources for comparison, we do not use this approach.

\subsection{Burst sources}\label{sec:bursts}

Some sources of GWs can produce signals with large amplitudes, greater than the detector noise. The typical duration of such a signal is short, of the order of a few wave periods, and so there is not time to accumulate SNR in each frequency band as for inspirals. As a consequence, waveform models are not required for detection; we simply rely on identifying the excess power produced by these burst sources. Typically, we may be looking for signals from core-collapse supernovae \citep{Ott2009}, the late stages of merging compact binaries, or more generally, signals from any unexpected or poorly modelled sources.

Burst searches are often carried out using time--frequency techniques. The data stream from a detector is temporally split into segments, the length of which can be tuned to give greater sensitivity to particular sources. Each segment is then transformed into the frequency domain, whitened and normalised to the noise spectrum of the detector to produce a time--frequency plot. Potential GW signals are identified by searching for clusters of pixels that contain an excess of power \citep[e.g.,][]{Bursts}.

The presence of excess power across a number of pixels eliminates modelled noise sources, but such a cluster may also be caused by atypical noise within a detector. We can improve our confidence of a GW signal by making use of information obtained from other GW detectors. Signals across a network of detectors should have compatible arrival times (given the sky direction) as well as consistent amplitudes, frequencies and shapes of the waveform. Different pipelines are currently in use that analyse the signal consistency in different ways: both coincidence searches \citep{Chatterji2004} and fully coherent methods \citep{Klimenko2008} are used.

An important aspect of burst search algorithms is to accurately estimate the noise properties within each time segment. To this extent, null data streams can be constructed that are insensitive to real GW signals. In order to estimate the false alarm rate, the data from different detectors can be shifted in time to remove any genuine coincident GW signals. These time-shifts are then analysed to simulate the potential occurence of coincident noise events. The algorithms are tuned using time-shifted data to ensure there is no bias in the final search.

As discussed in \ref{sec:insp}, the expected relation between $h_\mathrm{c}(f)$ and a typical waveform $\tilde{h}(f)$ is
\begin{equation}\label{eq:simple} 
h_\mathrm{c}(f) = \sqrt{{N}_{\mathrm{cycles}}}\left|\tilde{h}(f)\right| \; , 
\end{equation}
where ${N}_{\mathrm{cycles}}$ is the number of cycles of radiation generated by the source, which is of order unity for bursts.

An alternative characterisation of the signal amplitude commonly used for burst sources is the root-sum-square of the waveform polarisations:
\begin{equation}
h_\mathrm{rss} = \left[\int \rmd t {|h_+(t)|}^2 + {|h_\times(t)|}^2\right]^{1/2}.
\end{equation}
For a linearly polarised GW, with $\tilde{h}(f)$ constant across the bandwidth $\Delta f$, this is approximately related to the characteristic strain via 
\begin{equation}
h_\mathrm{rss} \simeq |\tilde{h}(f)|\sqrt{\Delta f}\,,\label{eq:rela}
\end{equation}
where we have neglected the detector response functions (see section \ref{sec:principles}), which are of order unity. In this work, we favour a constant $h_\mathrm{c}(f)$ rather than $h_\mathrm{rss}$ for consistency with the other types of source where the bandwidth is detector specific.

\section{Detectors}\label{sec:detectors}

In this section we introduce the detector noise curves used in \ref{app:a}. We begin with a description of the basic operation of detectors. We then discuss ground-based detectors, space-based detectors and PTAs in turn. The latter function somewhat differently than conventional interferometers, so we include a brief introduction to PTA analysis. References for the noise curves used for individual detectors can be found in the relevant subsections, further information about the detectors can be in discovered these. Detectors are frequently upgraded and redesigned, hence, while these curves are believed to be correct at the time of writing, it is best to check for updates from the appropriate science teams before relying on the details given here, although we hope that they shall remain accurate enough for illustrative purposes.

\subsection{Operating principle of an interferometric detector}\label{sec:principles}

All of the man-made detectors discussed in this section utilise the principle of interferometry. Such detectors work by taking a beam of monochromatic light and splitting it into two beams travelling at some angle to each other. Each beam is passed in to an optical cavity where it undergoes a number of round trips before being recombined to form an interference pattern. The ends of the cavity are, in the ideal case, freely floating test masses which move in response to a passing GW, this effect is measured by observing the changing interference pattern.

The response of a detector to an incident plane-fronted GW depends upon the relative orientations of the detector and the incoming wave. Let us choose the origin of our coordinate system to be the beam-splitter of the interferometer, and $l_{1}^{i}$ and $l_{2}^{i}$ to be unit 3-vectors pointing along the two arms. In the absence of noise the output of the detector is the difference in strain between the two arms \citep{Thorne1987}
\begin{equation}\label{eq:hoftxx}
h(t)=\frac{1}{2}h_{ij}\left( l_{1}^{i}l_{1}^{j}-l_{2}^{i}l_{2}^{j} \right)\; ,
\end{equation}
where $h_{ij}$ are the spatial components of the GW metric perturbation. Let $\hat{r}^{i}$ be the unit 3-vector pointing towards the source of the GWs, with spherical polar angles $(\theta,\phi)$ relative to some axes fixed to the detector, and let $p^{i}$ and $q^{i}$ be unit vectors orthogonal to $\hat{r}^{i}$. We can now define the basis tensors
\begin{eqnarray}
H^{+}_{ij}&=p_{i}p_{j}-q_{i}q_{j} \; , \\
H^{\times}_{ij}&=p_{i}q_{j}+q_{i}p_{j} \; .
\end{eqnarray}
There remains a freedom in the coordinates described, a rotation of $p^{i}$ and $q^{i}$ through an angle $\psi$ about $\hat{r}^{i}$ known as the polarization angle. For a single frequency component, the strain induced by a GW may be written as
\begin{equation}\label{eq:hijxx}
h_{ij}=A_{+}H^{+}_{ij}\cos\left(2\pi ft\right)+A_{\times}H^{\times}_{ij}\cos\left(2\pi ft+\Delta \phi\right) \; ,
\end{equation}
where $A_{+}$ and $A_{\times}$ are the amplitudes of the two polarisation states. Combining (\ref{eq:hoftxx}) and (\ref{eq:hijxx}) allows the detector output to be written as
\begin{equation}
h(t)=F^{+}(\theta,\phi,\psi)A_{+}\cos\left(2\pi ft\right)+F^{\times}(\theta,\phi,\psi)A_{\times}\cos\left(2\pi f t + \Delta\phi \right)\; ,
\end{equation}
where the response functions inherit their angular dependence from the choice of coordinates
\begin{eqnarray}\label{eq:responsefuncs}
F^{+}(\theta,\phi,\psi)&=\frac{1}{2}H^{+}_{ij}\left(l_{1}^{i}l_{1}^{j}-l_{2}^{i}l_{2}^{j}\right) \; , \\
F^{\times}(\theta,\phi,\psi)&=\frac{1}{2}H^{\times}_{ij}\left(l_{1}^{i}l_{1}^{j}-l_{2}^{i}l_{2}^{j}\right) \; . 
\end{eqnarray}

The response function of a two-arm interferometric detector is quadrupolar, an example is plotted in figure \ref{fig:LIGO}.
\begin{figure}
 \centering
 \includegraphics[trim=0cm 0cm 0cm 0cm, width=0.99\textwidth]{LIGO_detectorframe.pdf}
 \caption{The angular response function of an interferometric detector shown both as a surface plot and in an Aitoff--Hammer projection. The quentity that is plotted is is the polarisation average, $\int\textrm{d}\psi \;((F^{+})^{2}+(F^{\times})^{2})/2$. The response is a function of two sky angles, $\theta$ and $\phi$, and varies between $0$ and $1$. The two detector arms lie in the $x$--$y$ plane either side of one of the zeros in the response.}
 \label{fig:LIGO}
\end{figure}
Throughout this paper detector sensitivity refers to the polarisation and sky averaged sensitivity $F$, where
\begin{equation}\label{eq:skyav}
F^{2}=\int_{0}^{2\pi}\frac{\mathrm{d}\psi}{2\pi}\; \int_{0}^{2\pi} \frac{\mathrm{d}\phi}{2\pi}\; \int_{0}^{\pi}\frac{\sin\theta\,\mathrm{d}\theta}{2}\;\left[\frac{F^{+}\left(\theta,\phi,\psi\right)^{2}+F^{\times}\left(\theta,\phi,\psi\right)^{2}}{2}\right]\; .
\end{equation}
For a single $90\degree$-interferometer, such as LIGO, the sky and polarisation averaged response is $F=\sqrt{1/5}\approx 0.447$.

A detector may consist of several interferometers. Let $F_{a}$ be the averaged response of the $a$-th interferometer, the average response of a network of $k$ detectors is obtained by adding in quadrature,
\begin{equation} F_{\mathrm{Total}}^{2}=\frac{1}{k}\sum_{a\,=\,1}^{k}F_{a}^{2} \; .\end{equation}

The averaging in (\ref{eq:skyav}) assumes a uniform distribution of polarisation angles $\psi$. This is the case for a stochastic background; however, for a non-inspiralling circular binary, the polarisation is a function of the two spherical polar angles ($\iota,\xi$) specifying the orientation of the binary's orbital angular momentum. Here, $\iota$ is the inclination angle, the polar angle between the orbital angular momentum and the line joining the source to the detector ($-\hat{r}$) and $\xi$ is the azimuthal angle around the same line. In this case, to characterise the detector sensitivity we average over all four angles ($\theta,\phi,\iota,\xi$). If the binary is inspiralling, then the polarisation depends on still more parameters which need to be averaged over. These more complicated averages all have the property that they depend on both the detector and the source, hence they are unhelpful for our present purpose separating the source amplitude from the detector sensitivity. Additionally, the different averages do not work out to be so different from each other: \citet{1993PhRvD..47.2198F} calculated the sensitivity for a detector with the LIGO geometry averaged over the four angles ($\theta,\phi,\iota,\xi$) as $\sqrt{4/25}=0.4$ times peak sensitivity, which should be compared with the value $\sqrt{1/5}\approx 0.447$ above. The effect of replacing the true sensitivity with the sky-averaged sensitivity for the LISA detector was considered in detail by \cite{2012CQGra..29l4015V}; they also found there is a small difference when considering an entire population of sources. For the remainder of this paper the three-angle average defined in (\ref{eq:skyav}) is used.

\subsection{Ground-based detectors}\label{sec:ground}

Ground-based detectors are the most numerous. A collection of interferometric detectors are listed in table \ref{table:t}, these are sensitive to GWs in the frequency range ${\mathcal{O}}(10$--$10^{3})~\mathrm{Hz}$. They all simulate free-floating test masses by suspending a mass from a pendulum system with natural frequency much greater than that of the GW. Their sensitivity curves include narrow lines that arise from noise sources in the instrument, including resonances in the suspension system and electrical noise at multiples of $60~\mathrm{Hz}$: these have been removed in the \ref{app:a} figures for clarity. The detectors fall broadly into three categories: first-generation detectors, which have already operated; second-generation detectors currently under construction; and third-generation detectors at the planning stage.

The most notable ground-based detectors are LIGO and Virgo, which work in collaboration, supported by GEO600. LIGO, Virgo and GEO600 have completed science runs as first-generation detectors. As LIGO and Virgo are currently being upgraded their initial configurations are now referred to as Initial LIGO (iLIGO) and Initial Virgo (iVirgo) respectively. The upgraded, second-generation versions are referred to as advanced LIGO (aLIGO) and advanced Virgo (AdV) respectively. LIGO has two observatories: one at Hanford, Washington, which has two detectors; and another at Livingston, Louisiana. There is an agreement to move one of the upgraded Hanford detector systems to a location in India \citep{LIGO-India,Unnikrishnan2013}. The GEO600 detector can also operate in a high frequency mode \citep[GEO-HF]{2006CQGra..23S.207W}. The advanced detectors should start operation in the next couple of years, with LIGO-India following further in the future.

TAMA300 is a Japanese first-generation detector. Its successor, currently under construction, is the Kamioka Gravitational Wave Detector (KAGRA), formerly the Large-scale Cryogenic Gravitational wave Telescope (LCGT), which is located underground in the Kamioka mine. It employs more sophisticated noise-reduction techniques than LIGO or Virgo, such as cryogenic cooling.

The Einstein Telescope (ET) is an ambitious proposal to construct an underground third-generation detector. Its location would provide shielding from seismic noise, allowing it to observe frequencies of $10$--$10^4~\mathrm{Hz}$.

\begin{table}
\caption{\label{table:t} Summary of ground-based laser interferometers.}
\begin{indented}
\item[]\begin{tabular}{ l l l l l }
\br
{\bf Detector} & {\bf Country} & {\bf Arm length} & {\bf  Approximate date} & {\bf Generation} \\
\mr
  GEO600$^{a}$	&	Germany 	& $600~\mathrm{m}$ 	& 2001--present	   & First \\
  TAMA300$^{b}$ & 	Japan		& $300~\mathrm{m}$ 	& 1995--present    & First \\
  iLIGO$^{c}$	&	USA		& $4~\mathrm{km}$ 	& 2004--2010 	   & First \\
  Virgo$^{d}$	& 	Italy		& $3~\mathrm{km}$ 	& 2007--2010 	   & First \\
  aLIGO$^{e}$	&	USA		& $4~\mathrm{km}$ 	& \emph{est.} 2015 & Second \\
  AdV$^{f}$	&	Italy	 	& $3~\mathrm{km}$ 	& \emph{est.} 2016 & Second \\
  KAGRA$^{g}$	&	Japan		& $3~\mathrm{km}$ 	& \emph{est.} 2018 & Second \\
  ET$^{h}$      & 	---		& $10~\mathrm{km}$ 	& \emph{est.} 2025 & Third \\
\br
\multicolumn{5}{l}{$^{a}$\citet{2010CQGra..27h4003G}, $^{b}$\citet{2002CQGra..19.1409A}, $^{c}$\citet{2009RPPh...72g6901A},}\\
\multicolumn{5}{l}{$^{d}$\citet{accadia_virgo:2012}, $^{e}$\citet{2010CQGra..27h4006H}, $^{f}$\citet{Acernese2009}, $^{g}$\citet{2012CQGra..29l4007S}, $^{h}$\citet{2011CQGra..28i4013H}.}\\
\br
\end{tabular}
\end{indented}
\end{table}

We use an interpolation to the data published on \url{https://wwwcascina.virgo.infn.it/advirgo/} (2013) for the AdV sensitivity curve, an interpolation to the data for version D of the KAGRA detector published on \url{http://gwcenter.icrr.u-tokyo.ac.jp/en/researcher/parameter} (2013) and analytic fits to the sensitivity curves from \citet{Sathyaprakash} for the remaining detectors.

\subsection{Space-based detectors}\label{sec:space}

Space-based detectors work on similar principles to ground based detectors, but with the test masses residing inside of independent, widely separated satellites. Space-based detectors are sensitive to lower frequency GWs than their ground-based counterparts; this is partly because space based detectors can have much longer arms, and partly because they are unaffected by seismic noise which limits the low frequency performance of ground-based detectors.

The canonical design for a space-based detector is the Laser Interferometer Space Antenna (LISA), which is sensitive to millihertz GWs. LISA would consist of three satellites flying in a triangular constellation with arms of length $5\times 10^{9}~\mathrm{m}$ in a $1~\mathrm{AU}$ orbit around the Sun, trailing the Earth by $20\degree$. The laser arms in a LISA-like detector are not a cavity, the light only travels once along each arm. eLISA is a rescoped version of LISA designed to probe the same frequency range, while proposals such as the Advanced Laser Interferometer Antenna (ALIA), Big Bang Observer (BBO) and Deci-hertz Interferometer GW Observatory (DECIGO) are designed to probe decihertz GWs.

\subsubsection{LISA and eLISA}

The instrumental noise curves for LISA are approximated by the analytic fit given by \citet{Sathyaprakash}, which we use for the plots in \ref{app:a}. When observing individual sources with LISA there is an additional contribution to the noise from a background of unresolvable binaries. This is not included here as we consider the background as a source of GWs (see section \ref{sec:GB}). eLISA is a rescoped version of the classic LISA mission, the main differences are shorter arms ($10^{9}~\mathrm{m}$ instead of $5\times 10^{9}~\mathrm{m}$), two laser arms instead of three, and a different orbit (drifting away from Earth instead of $20\degree$ Earth trailing). The effect of these changes is a slightly reduced peak sensitivity and a shift to higher frequencies. We use an analytic fit to the instrumental noise curve given by \citet{Amaro-Seoane-et-al}.

\subsubsection{DECIGO, ALIA and BBO}

These missions are designed to probe the decihertz region of the GW spectrum; they are considerably more ambitious than the LISA or eLISA mission and their launches will be further into the future. We use a simple analytic fit to the sensitivity curve for ALIA~\citep{bender_possible_2013}, while for DECIGO and BBO, fits to the sensitivity curves given by \citet{2011PhRvD..83d4011Y} are used.

\subsection{Pulsar timing arrays}\label{sec:PTAgeneralproperties}

PTAs can be thought of as naturally occurring interferometers with galactic-scale arm lengths. Accordingly, they are sensitive to much lower frequencies than the detectors previously discussed. Each pulsar is a regular clock and the measured pulse arrival time can be compared against a prediction, leaving a residual which includes the effects of passing GWs. Using an array of these pulsars spread across the sky allows us to correlate residuals between different pulsars, to exploit the fact that GWs influence all pulsars whereas intrinsic pulsar noise does not. The correlation between different pulsars depends only on their angular separation on the sky, and has a distinctive shape, known as the Hellings and Downs curve \citep{HellingsDowns}.

The redshift of the rate of arrival of pulses for a pulsar at a distance $L$ from the Solar-System barycentre (SSB), in the direction of the unit spatial vector $\hat{p}$ induced by a GW travelling in direction of the unit vector $\hat{\Omega}$ is \citep{anholm-2009}
\begin{equation}\label{eq:pulsarredshift}
z(t,\hat{\Omega})= \frac{1}{2} \frac{\hat{p}^{j}\hat{p}^{i}} {1+\hat{\Omega}\cdot\hat{p}}\left[h_{ij}^{\mathrm{Pulsar}}\left(t-\frac{L}{c},\hat{\Omega} \right)-h_{ij}^{\mathrm{Earth}}(t,\hat{\Omega} )\right] = \frac{1}{2}\frac{\hat{p}^{j}\hat{p}^{i}}{1+\hat{\Omega}\cdot\hat{p}} \Delta h_{ij}(t,\hat{\Omega})\; .
\end{equation}
The redshift includes two terms: the pulsar term and the Earth term. The pulsar term is often neglected in PTA analysis as it can be considered as an extra noise term which averages to zero across the array. The experimentally measured quantity is not the redshift but the timing residual, the two are related via
\begin{equation}\label{eq:restored}
R(t,\hat{\Omega}) = \int_{0}^{t}\mathrm{d}t'\;z(t',\hat{\Omega}) \; .
\end{equation}
All of the pulsars, and the Earth, are subject to the same metric-perturbation field. This may be expressed in terms of its Fourier transform
\begin{equation}
h_{ij}(t,\vec{r}) = \sum_{A\,=\,+,\,\times}\int\mathrm{d}f\;\iint_{\mathbb{S}_{2}}\mathrm{d}\hat{\Omega}\; \tilde{h}_{A}(f,\hat{\Omega})e_{ij}^{A}(\hat{\Omega}) \exp\left[2\pi i f \left(t-\frac{\hat{\Omega}\cdot\vec{x}}{c}\right)\right] \; ,
\end{equation}
where $e^{A}_{ij}(\hat{\Omega})$ is the $A$ polarisation basis tensor for direction $\hat{\Omega}$  and $\vec{r}$ is the spatial position. Choosing the SSB as the origin of our coordinate system, so the pulsar is at position $L\hat{p}$, gives
\begin{equation}\label{eq:deltahij}
\Delta h_{ij}(t,\hat{\Omega}) = \sum_{A\,=\,+,\,\times}\int\mathrm{d}f\; \tilde{h}_{A}(f,\hat{\Omega})e_{ij}^{A}(\hat{\Omega})\exp\left(2\pi \rmi f t\right)\left\{\exp\left[-2\pi\rmi fL \left(1+\hat{p}\cdot\hat{\Omega}\right)\right]-1\right\} \; .
\end{equation}
From (\ref{eq:pulsarredshift}) and (\ref{eq:deltahij}), the Fourier transform of the redshift $\tilde{z}(f,\hat{\Omega})$ can be identified as
\begin{equation}\label{eq:45}
\tilde{z}(f,\hat{\Omega}) = \left\{\exp\left[-2\pi\rmi fL \left(1+\hat{p}\cdot\hat{\Omega}\right)\right]-1\right\} \sum_{A\,=\,+,\,\times} \tilde{h}_{A}(f,\hat{\Omega})F^{A}(\hat{\Omega}) \; ,
\end{equation}
where
\begin{equation}
F^{A}(\hat{\Omega}) = \frac{e_{ij}^{A}(\hat{\Omega})\hat{p}^{j}\hat{p}^{i}} {2\left(1+\hat{\Omega}\cdot\hat{p}\right)} \, .
\end{equation}
The function $F^{A}(\hat{\Omega})$ may be regarded as the PTA equivalent of the detector response functions in (\ref{eq:responsefuncs}). The stochastic background of GWs is fully characterised by the one-sided PSD via the expectation value
\begin{equation}\label{eq:stochback}
\left<\tilde{h}_{A}^{*}(f,\hat{\Omega})\tilde{h}_{A'}(f',\hat{\Omega}')\right> = \frac{1}{2}S_{h}(f)\delta^{(2)}(\hat{\Omega}, \hat{\Omega}')\delta_{AA'}\delta(f-f') \; ,
\end{equation}
where $\delta^{(2)}(\hat{\Omega},\hat{\Omega}')$ is the delta-function on the sphere. From (\ref{eq:45}) and (\ref{eq:stochback}), the expectation of the product of signals from two different pulsars in directions $\hat{p}_{1}$ and $\hat{p}_{2}$ may be evaluated as
\begin{equation} 
\left<\tilde{z}_{1}(f)\tilde{z}_{2}^{*}(f')\right> = \frac{1}{2}S_{h}(f)\delta (f-f')\Gamma(f) \; ,
\end{equation}
where
\begin{eqnarray}
\Gamma(f) &= \sum_{A\,=\,+,\,\times} & \iint_{{\mathbb{S}}^{2}}\mathrm{d}\hat{\Omega}\,\left\{\exp\left[2\pi\rmi fL_{1}\left(1+\hat{\Omega}\cdot\hat{p}_{1}\right)\right]-1\right\} \nonumber \\*
 & & \times \left\{\exp\left[-2\pi\rmi fL_{2}\left(1+\hat{\Omega}\cdot\hat{p}_{2}\right)\right]-1\right\} F_{1}^{A}(\hat{\Omega})F_{2}^{A}(\hat{\Omega}) \; .
\end{eqnarray}
The overlap function $\Gamma(f)$ tends to a constant value in the limit that the distances to the pulsars are large compared to the wavelength of GWs; PTAs operate in this limit, so the overlap may be approximated as a constant, 
\begin{equation}
\Gamma(f) \approx \Gamma_{0} = \sum_{A\,=\,+,\,\times}\iint_{{\mathbb{S}}^{2}}\mathrm{d}\hat{\Omega}\;F_{1}^{A}(\hat{\Omega})F_{2}^{A}(\hat{\Omega})\;.
\end{equation}
Neglecting the exponential terms in the overlap is the frequency-domain equivalent of neglecting the pulsar term in (\ref{eq:pulsarredshift}). The integral may be evaluated to give an expression depending only on the angle $\theta$ between the two pulsars; this is the famous Hellings and Downs curve, shown in figure \ref{fig:HnD},
\begin{equation}
\Gamma_{0} = \frac{1}{2}+\frac{3x}{2}\left(\ln x -\frac{1}{6}\right) \; ,
\end{equation}
where $x = (1-\cos\theta)/{2}.$
\begin{figure}
 \centering
 \includegraphics[trim=0cm 0cm 0cm 0cm, width=0.6\textwidth]{HnDcurve.pdf}
 \caption{The \citet{HellingsDowns} curve, the correlation between two pulsars separated on the sky by an angle $\theta$.}
 \label{fig:HnD}
\end{figure}

The sensitivity bandwidth of a PTA is set by the sampling properties of the data set. If measurements are spaced in time by $\delta t$ and taken for a total length of time $T$, then the PTA is sensitive to frequencies in the range $(1/T) < f < (1/\delta t)$. The characteristic strain that the PTA is sensitive to scales linearly with $f$ in this range. This gives the wedge-shaped curves plotted in \ref{app:a}. The absolute value of the sensitivity is fixed by normalising to a calculated limit at a given frequency for each PTA. For a discussion of the sensitivities of PTA to both individual sources and stochastic background, see \citet{MooreTaylorGair}.

There is a discrepancy between the treatment of PTA sensitivity curves here and the higher frequency detectors discussed in sections \ref{sec:ground} and \ref{sec:space}. When observing a long-lived source, such as an inspiral, with a high frequency detector the convention was to define a \emph{characteristic} strain to satisfy (\ref{eq:hc}). Here, the convention is to leave the strain untouched and instead adjust the PTA sensitivity curve with observation time, again to satisfy (\ref{eq:hc}). This discrepancy is an unfortunate result of the conventions in use by the different GW communities; however, it is also natural given the sources under observation. When observing a transient source, such as a burst or inspiral, which changes within the lifetime of the detector, it is natural to consider the detector as performing constantly while the signal changes. However, when observing a monochromatic source or a stochastic background, which is unchanging over the detector lifetime, it is more natural to consider the source as being fixed and the sensitivity of the detector gradually improving. All that is required by the definition in (\ref{eq:hc}) is that the ratio $h_\mathrm{c}(f)/h_{n}(f)$ is constant.

\subsubsection{Current PTAs}

The PTAs currently in operation are the European Pulsar Timing Array \citep[EPTA\footnote{\url{http://www.epta.eu.org/}};][]{eptareview2013}, the Parkes Pulsar Timing Array \citep[PPTA\footnote{\url{http://www.atnf.csiro.au/research/pulsar/ppta/}};][]{parkesreview2013} based in Australia, and the North American Nanohertz Observatory for Gravitational waves \citep[NANOGrav\footnote{\url{http://nanograv.org/}};][]{nanogravreview2013}). There are published limits on the amplitude of the stochastic background from all three detectors: the most recent from EPTA is \citet{Haasteren}, from PPTA is \citet{Shannon2013} and from NANOGrav is \citet{2013ApJ...762...94D}. These are all comparable and we use the EPTA limits in \ref{app:a}, the figures show the published limit on $A$ from equation \ref{eq:power} at a frequency of $f_{0}=1\,\textrm{year}^{-1}$, based on an analysis of five pulsars over approximately ten years.

\subsubsection{The International Pulsar Timing Array}

Combining the existing arrays would yield a single PTA using approximately 4 times as many pulsars; this consortium of consortia is known as the International Pulsar Timing Array \citep[IPTA\footnote{\url{http://www.ipta4gw.org/}};][]{iptareview2013}. The IPTA curves plotted in the figures are based on 20 pulsars timed for twice as long as the current EPTA.

\subsubsection{SKA}

The next great advancement in radio astronomy shall come with the completion of the Square Kilometre Array \citep[SKA;][]{Dewdney2009}. This shall greatly increase the sensitivity of pulsar timing \citep{Kramer2004}. The sensitivity curves plotted in \ref{app:a} for the SKA assume a root-mean-square error on the timing residuals a factor of 10 better than current PTAs, and include 20 pulsars timed for twice as long as the current EPTA.

\section{Astrophysical sources}\label{sec:sources}

All the sources described here are represented by shaded boxes in \ref{app:a}. Sources with short durations (i.e., burst sources) and sources that evolve in time over much longer timescales than our observations are drawn with flat-topped boxes for $h_\mathrm{c}(f)$. Inspiraling binaries, or stochastic backgrounds of binaries, are drawn with a sloping top proportional to $f^{-2/3}$ for $h_\mathrm{c}(f)$, which is result derived in section \ref{sec:stoch}. The width of the box gives the range of frequencies sources of a given type can have while remaining at a detectable amplitude. The question of the height of the box is more problematic; it would be desirable to normalise each box so that there was a fixed event rate, say one event per year with an amplitude lying within the box. However for many of the sources considered the event rate is subject to a large degree of uncertainty (and estimates change rapidly as our understanding of the astrophysics improves). So instead we take the more definite, but somewhat arbitrary, approach that for each type of source a fiducial event is nominated (with parameters or amplitude detailed in the relevant section below) and the amplitude of this event is used to position the top of the box. The parameters for each fiducial event are chosen such that the resulting amplitude is roughly consistent with the maximum end of current predictions: the boxes indicate where we \emph{could} plausibly find GW sources; the actual event rates could turn out to be substantially lower than this upper bound.

\subsection{Sources for ground-based detectors}

\subsubsection{Neutron star binaries}

The inspiral and merger of a pair of neutron stars is the primary target for ground-based detectors. The expected event rate for this type of source is uncertain, but estimates centre around $\gamma_{\mathrm{NS-NS}}=1.3\times 10^{-4}~\mathrm{Mpc^{-3}\,yr^{-1}}$ \citep{CBC}. Plotted in \ref{app:a} are boxes labelled ``\emph{compact binary inspirals}'' with amplitudes such that a ratio $h_\mathrm{c}/h_{n}=16$ is produced for Advanced LIGO at peak sensitivity and a width between $(3$--$300)~\mathrm{Hz}$, corresponding to expected observable frequencies.

\subsubsection{Supernovae}

Simulations of core-collapse supernovae show that GWs between $(10^{2}$--$10^{3})~\mathrm{Hz}$ can be produced \citep{Kotake2006}. The GW signal undergoes ${\mathcal{O}}(1)$ oscillation and is hence burst-like. \citet{2002A&A...393..523D} calculate the average maximum amplitude of GWs for a supernova at distance $r$ as
\begin{equation}
h_\mathrm{max} = 8.9\times 10^{-21}\left( \frac{10~\mathrm{kpc}}{r} \right) \; .
\end{equation}
The event rate for supernovae is approximately $\gamma_{\mathrm{SN}} = 5\times10^{-4}~\mathrm{Mpc^{-3}\,yr^{-1}}$. The boxes labelled ``\emph{supernova}'' plotted in \ref{app:a} correspond to a distance $r=300~\mathrm{kpc}$ with the frequency range quoted above. The LIGO and Virgo detectors have already placed bounds on the event rate for these sources \citep{Bursts}.

\subsubsection{Continuous waves from rotating neutron stars}

Rotating neutron stars are a source of continuous GWs if they possess some degree of axial asymmetry \citep{Abbott2007, Prix2009, Einstein@Home}. The signals are near monochromatic with a frequency twice the rotation frequency of the neutron star, and are a potential source for ground-based detectors. The amplitude of the GWs depends upon the deformation of the neutron star. The magnitude of the distortion depends upon the neutron star equation of state and the stiffness of the crust, which are currently uncertain \citep{Chamel2008,Lattimer2012}; deformations can also be supported by internal magnetic fields \citep{Haskell2008}. Several known pulsars could be sources for the advanced detectors and upper limits from the initial detectors help to constrain the deformations. The boxes labelled ``\emph{pulsars}'' plotted in \ref{app:a} correspond to the upper limits placed on a GW signal from the Crab pulsar \citep{Aasi2014a}, extrapolated across a frequency range between $(20$--$10^{3})~\mathrm{Hz}$.

\subsection{Sources for space-based detectors}

For a review of the GW sources for space-based missions see, for example, \citet{Amaro-Seoane-et-al}, \citet{Gairetal} or \citet{eLISAyellowbook}.

\subsubsection{Massive black hole binaries}

Space-based detectors shall be sensitive to equal-mass mergers in the range $(10^{4}$--$10^{7})\,\Msun$. Predictions of the event rate for these mergers range from ${\mathcal{O}}(10$--$100)~\mathrm{yr}^{-1}$ for eLISA with SNRs of up to $10^3$ \citep{TheGravitationalUniverse}. The large range in the rate reflects our uncertainty in the growth mechanisms of the supermassive black hole population \citep{Volonteri2010}. Plotted in \ref{app:a} are boxes labelled ``\emph{$\approx 10^{6}$ solar mass binaries}'' with a ratio $h_\mathrm{c}/h_{n}=100$ for eLISA at its peak sensitivity. The range of frequencies plotted is $(3\times 10^{-4}$--$3\times 10^{-1})~\mathrm{Hz}$; this corresponds to circular binaries in the mass range quoted above.

\subsubsection{Galactic white dwarf binaries} \label{sec:GB}

For space-based detectors, these are the most numerous GW sources; they are also the only guaranteed source since several detectable systems (known as verification binaries) have already been identified by electromagnetic observations \citep{2006CQGra..23S.809S}.

Galactic binaries divide into two classes: the unresolvable and the resolvable galactic binaries. The unresolvable binaries overlap to form a stochastic background as discussed in section \ref{sec:stoch}. The distinction between resolvable and unresolvable is detector specific; here we choose LISA. This boundary will not be too different for eLISA but would move substantially for either of the decihertz detectors. Plotted in \ref{app:a} with the label ``\emph{unresolvable galactic binaries}'' is the estimate of this background due to \citet{Nelemans} where an observation time of one year has been assumed,
\begin{equation}
h_\mathrm{c}(f)= 5\times 10^{-21} \left(\frac{f}{10^{-3}~\mathrm{Hz}}\right)^{-2/3} \; .
 \end{equation}
Estimates for the event rate of resolvable binaries centre around ${\mathcal{O}}(10^{3})$ events for eLISA. The boxes plotted in \ref{app:a} with the label ``\emph{resolvable galactic binaries}'' have a ratio $h_\mathrm{c}/h_{n}=50$ for eLISA at its peak sensitivity. The frequency range of the box is $\left(3\times10^{-4}\right.$--$\left.10^{-2}\right)~\mathrm{Hz}$, estimated from Monte Carlo population simulation results presented in \citet{Amaro-Seoane-et-al}.

\subsubsection{Extreme mass-ratio inspirals}

EMRIs occur when a compact stellar mass object inspirals into a supermassive black hole. There is extreme uncertainty in the event rate for EMRIs due to the poorly constrained astrophysics in galactic centres \citep[e.g.,][]{Merritt2011}; the best guess estimate is around $25$ events per year with eLISA with SNR $\ge 20$ \citep{TheGravitationalUniverse}. The boxes labelled ``\emph{extreme mass ratio inspirals}'' plotted in \ref{app:a} have a characteristic strain of $h_\mathrm{c}=3\times 10^{-20}$ at $10^{-2}~\mathrm{Hz}$, which corresponds to a $10\Msun$ black hole inspiralling into a $10^{6}\Msun$ black hole at a luminosity distance of $1~\mathrm{Gpc}$. The frequency width of the box is somewhat unknown; EMRI events can occur into a black hole of any mass, and hence EMRIs can, in principal, occur at any frequency, so the boxes in \ref{app:a} are drawn with a width comparable to that of the LISA sensitivity curve.

\subsection{Sources for PTAs}

\subsubsection{Supermassive black hole binaries}

The main target for PTAs is a stochastic background of GWs produced by a population of supermassive black hole binaries at cosmological distances \citep{SesanaVecchioColancino}. Supermassive black holes are known to lie at the centres of most galaxies and the black hole mergers are associated with the mergers of the host galaxies \citep{Volonteri2003,Ferrarese2005}. The current best 95\% confidence limit for the amplitude of the stochastic background is $h_\mathrm{c} = 2.7\times 10^{-15}$ at a frequency of $f_{0}=1~\mathrm{yr}^{-1}$ \citep{Shannon2013}. There is strong theoretical evidence that the actual background lies close to the current limit \citep{Sesana-2012}.

Supermassive black hole binaries at higher frequencies are inspiralling faster and hence there are fewer of them per frequency bin. At a certain frequency, these sources will cease to be a background and become individually resolvable. It is currently unclear whether PTAs will detect an individual binary or a stochastic background first. Plotted in \ref{app:a} with the label ``\emph{stochastic background}'' is a third of the current limit with a cut off frequency of $f=1~\mathrm{yr}^{-1}$ which is suggested by Monte Carlo population studies \citep{SesanaVecchioColancino}. For the resolvable sources, labelled ``\emph{$\approx 10^{9}$ solar mass binaries}'', the amplitude of the current limit is plotted between $\left(3\times 10^{-9}\right.$--$\left.3\times 10^{-7}\right)~\mathrm{Hz}$.

\subsection{Cosmological sources}

In addition to the sources above, early Universe processes, such as inflation \citep{Grishchuk2005} or a first-order phase transition \citep{Binetruy2012}, could have created GWs. More speculatively, it has been hypothesised that cosmic strings could also be a potential source \citep{Damour2005,Binetruy2012,Aasi2014}. These relic GWs allow us to explore energy scales far beyond those accessible by other means, providing insight into new and exotic physics. The excitement surrounding the tentative discovery by BICEP2 of the imprint of primordial GWs (generated during inflation) in the cosmic microwave background \citep{Ade2014}, and the subsequent flurry of activity, has shown the scientific potential of such cosmological GWs. These GW signals are so alluring because they probe unknown physics; this also makes them difficult to predict. Cosmological stochastic backgrounds have been predicted across a range of frequencies with considerable variation in amplitude. As a consequence of this uncertainty, although we could learn much from measuring these signals, we have not included them amongst the sources in \ref{app:a}.

\section{Concluding remarks}\label{sec:discussion}

When quantifying the sensitivity of a GW detector and the loudness of a GW source, there are three commonly used quantities: the characteristic strain, the power spectral density, and the spectral energy density. We have carefully defined these quantities and derived the relationships between them. The characteristic strain (section \ref{sec:character-strain}) is most directly related to the SNR, the PSD is mostly closely related to the mean square amplitude in the detector, and the energy density has a clear physical interpretation. We have produced example plots using each of these quantities for a wide range of detectors and sources. The predicted source amplitudes are based on astrophysical estimates of the event rates and are subject to varying degrees of uncertainty.\footnote{There is also the exciting possibility of sources yet to be considered.} Interactive versions of these plots, with user-specified detectors and sources, are available on-line at \url{http://www.ast.cam.ac.uk/~rhc26/sources/}. Trying to summarise an entire field of astronomy on one plot is an impossible task; however, we hope that the figures and analysis presented here provide useful insight.

\ack{CJM, RHC and CPLB are supported by the Science \& Technology Facilities Council. We are indebted to Jonathan Gair, for useful advice and proof-reading the manuscript. We also thank Stephen Taylor and Justin Ellis for useful conversations about pulsar timing. The authors thanks a number of members of the LIGO--Virgo Scientific Collaboration, especially Chris Messenger, for useful comments and suggestions. This paper has been assigned LIGO document LIGO-P1400129.}

\appendix
\section{Sensitivity curves}\label{app:a}
The plots in this section show all of the detectors and sources described in the main text. Clearer, interactive versions of these plots, allowing for removal of any of the curves, may be created and downloaded on-line, \url{http://www.ast.cam.ac.uk/~rhc26/sources/}. The detector noise curves all have their resonance spikes removed for clarity.

\begin{figure}
 \centering
 \includegraphics[trim=0cm 0cm 0cm 0cm, width=0.8\textwidth]{figure1.pdf}
 \caption{A plot of characteristic strain against frequency for a variety of detectors and sources.}
 \label{fig:hc}
\end{figure}

\begin{figure}
 \centering
 \includegraphics[trim=0cm 0cm 0cm 0cm, width=0.8\textwidth]{figure2.pdf}
 \caption{A plot of the square root of power spectral density against frequency for a variety of detectors and sources.}
 \label{fig:S}
\end{figure}

\begin{figure}
 \centering
 \includegraphics[trim=0cm 0cm 0cm 0cm, width=0.8\textwidth]{figure3.pdf}
 \caption{A plot of the dimensionless energy density in GWs against frequency for a variety of detectors and sources.}
 \label{fig:omega}
\end{figure}
\clearpage

\bibliographystyle{jphysicsCB}
\bibliography{bibliography}

\end{document}

### 2014-08-04 arXiv_version/noise_curve.tex

@@
 These are frequently confused.
-We outline the merits of and differences between the various quantities used for parameterizing noise curves and characterizing gravitational wave amplitudes.
+We outline the merits of and differences between the various quantities used for parameterizing noise curves and characterising gravitational wave amplitudes.
 We conclude by producing plots that consistently compare different detectors.

### 2014-08-06 Detectors.tex
Post-arXiv 1

@@
 The latter function somewhat differently than conventional interferometers, so we include a brief introduction to PTA analysis.
-References for the noise curves used for individual detectors can be found in the relevant subsections, further information about the detectors can be in discovered these.
+References for the noise curves used for individual detectors can be found in the relevant subsections and further information about the detectors can be discovered in these.
 Detectors are frequently upgraded and redesigned, hence, while these curves are believed to be correct at the time of writing, it is best to check for updates from the appropriate science teams before relying on the details given here, although we hope that they shall remain accurate enough for illustrative purposes.
@@
  \caption{The angular response function of an interferometric detector shown both as a surface plot and in an Aitoff--Hammer projection.
-The quentity that is plotted is is the polarisation average, $\int\textrm{d}\psi \;((F^{+})^{2}+(F^{\times})^{2})/2$.
+The quentity that is plotted is is the polarisation average, $\int\textrm{d}\psi \;((F^{+})^{2}+(F^{\times})^{2})/2$ \cplb{Should this be divided by $2\pi$ too?}.
 The response is a function of two sky angles, $\theta$ and $\phi$, and varies between $0$ and $1$.
@@
 The most notable ground-based detectors are LIGO and Virgo, which work in collaboration, supported by GEO600.
-LIGO, Virgo and GEO600 have completed science runs as first-generation detectors.
+LIGO, Virgo and GEO600 have completed science runs as first-generation detectors \cplb{[citation needed]}.
 As LIGO and Virgo are currently being upgraded their initial configurations are now referred to as Initial LIGO (iLIGO) and Initial Virgo (iVirgo) respectively.
-The upgraded, second-generation versions are referred to as advanced LIGO (aLIGO) and advanced Virgo (AdV) respectively.
+The upgraded, second-generation versions are referred to as Advanced LIGO (aLIGO) and Advanced Virgo (AdV) respectively.
 LIGO has two observatories: one at Hanford, Washington, which has two detectors; and another at Livingston, Louisiana.
 There is an agreement to move one of the upgraded Hanford detector systems to a location in India \citep{LIGO-India,Unnikrishnan2013}.
-The GEO600 detector can also operate in a high frequency mode \citep[GEO-HF]{2006CQGra..23S.207W}.
+The GEO600 detector can also operate in a high frequency mode \citep[GEO-HF]{2006CQGra..23S.207W} \cplb{Add to table}.
 The advanced detectors should start operation in the next couple of years, with LIGO-India following further in the future.
@@
 Space-based detectors work on similar principles to ground based detectors, but with the test masses residing inside of independent, widely separated satellites.
-Space-based detectors are sensitive to lower frequency GWs than their ground-based counterparts; this is partly because space based detectors can have much longer arms, and partly because they are unaffected by seismic noise which limits the low frequency performance of ground-based detectors.
+Space-based detectors are sensitive to lower frequency GWs than their ground-based counterparts; this is partly because space-based detectors can have much longer arms, and partly because they are unaffected by seismic noise which limits the low frequency performance of ground-based detectors.
 
@@
 \begin{equation}
-\Gamma_{0} = \frac{1}{2}+\frac{3x}{2}\left(\ln x -\frac{1}{6}\right) \; ,
+\Gamma_{0} = \frac{1}{2}+\frac{3\xi}{2}\left(\ln \xi -\frac{1}{6}\right) \; ,
 \end{equation}
-where $x = (1-\cos\theta)/{2}.$
+where $\xi = (1-\cos\theta)/{2}.$
 \begin{figure}
  \centering
- \includegraphics[trim=0cm 0cm 0cm 0cm, width=0.6\textwidth]{HnDcurve.pdf}
+ \includegraphics[trim=0cm 0cm 0cm 0cm, width=0.6\textwidth]{Fig_Hellings_Downs}
  \caption{The \citet{HellingsDowns} curve, the correlation between two pulsars separated on the sky by an angle $\theta$.}
@@
 Here, the convention is to leave the strain untouched and instead adjust the PTA sensitivity curve with observation time, again to satisfy (\ref{eq:hc}).
-This discrepancy is an unfortunate result of the conventions in use by the different GW communities; however, it is also natural given the sources under observation.
+This discrepancy is an unfortunate result of the conventions in use by the different GW communities; however, it is natural given the sources under observation.
 When observing a transient source, such as a burst or inspiral, which changes within the lifetime of the detector, it is natural to consider the detector as performing constantly while the signal changes.
@@
 
-The PTAs currently in operation are the European Pulsar Timing Array \citep[EPTA\footnote{\url{http://www.epta.eu.org/}};][]{eptareview2013}, the Parkes Pulsar Timing Array \citep[PPTA\footnote{\url{http://www.atnf.csiro.au/research/pulsar/ppta/}};][]{parkesreview2013} based in Australia, and the North American Nanohertz Observatory for Gravitational waves \citep[NANOGrav\footnote{\url{http://nanograv.org/}};][]{nanogravreview2013}).
+The PTAs currently in operation are the European Pulsar Timing Array \citep[EPTA\footnote{\url{http://www.epta.eu.org/}};][]{eptareview2013}, the Parkes Pulsar Timing Array \citep[PPTA\footnote{\url{http://www.atnf.csiro.au/research/pulsar/ppta/}};][]{parkesreview2013} based in Australia, and the North American Nanohertz Observatory for Gravitational waves \citep[NANOGrav\footnote{\url{http://nanograv.org/}};][]{nanogravreview2013}.
 There are published limits on the amplitude of the stochastic background from all three detectors: the most recent from EPTA is \citet{Haasteren}, from PPTA is \citet{Shannon2013} and from NANOGrav is \citet{2013ApJ...762...94D}.
-These are all comparable and we use the EPTA limits in \ref{app:a}, the figures show the published limit on $A$ from equation \ref{eq:power} at a frequency of $f_{0}=1\,\textrm{year}^{-1}$, based on an analysis of five pulsars over approximately ten years.
+These are all comparable and we use the EPTA limits in \ref{app:a}, the figures show the published limit on $A$ from equation \ref{eq:power} at a frequency of $f_{0}=1~\mathrm{yr^{-1}}$, based on an analysis of $5$ pulsars over approximately $10~\mathrm{yr}$.
 
@@
 
-Combining the existing arrays would yield a single PTA using approximately 4 times as many pulsars; this consortium of consortia is known as the International Pulsar Timing Array \citep[IPTA\footnote{\url{http://www.ipta4gw.org/}};][]{iptareview2013}.
-The IPTA curves plotted in the figures are based on 20 pulsars timed for twice as long as the current EPTA.
+Combining the existing arrays would yield a single PTA using approximately $4$ times as many pulsars; this consortium of consortia is known as the International Pulsar Timing Array \citep[IPTA\footnote{\url{http://www.ipta4gw.org/}};][]{iptareview2013}.
+The IPTA curves plotted in the figures are based on $20$ pulsars timed for twice as long as the current EPTA.
 
@@
 This shall greatly increase the sensitivity of pulsar timing \citep{Kramer2004}.
-The sensitivity curves plotted in \ref{app:a} for the SKA assume a root-mean-square error on the timing residuals a factor of 10 better than current PTAs, and include 20 pulsars timed for twice as long as the current EPTA.
+The sensitivity curves plotted in \ref{app:a} for the SKA assume a root-mean-square error on the timing residuals a factor of $10$ better than current PTAs, and include $20$ pulsars timed for twice as long as the current EPTA.

### 2014-08-06 Introduction.tex
Post-arXiv 1

@@
 The next few years  promise to deliver the first direct detection of gravitational waves (GWs).
-This will most likely be achieved by the advanced Laser Interferometer Gravitational-wave Observatory \citep[LIGO;][]{2010CQGra..27h4006H} and advanced Virgo \citep{Acernese2009} detectors operating in the frequency range $(10$--$10^{3})~\mathrm{Hz}$.
+This will most likely be achieved by the advanced versions of the Laser Interferometer Gravitational-wave Observatory \citep[LIGO;][]{2010CQGra..27h4006H} and Virgo \citep{Acernese2009} detectors operating in the frequency range $(10$--$10^{3})~\mathrm{Hz}$.
 By the end of the decade, it is expected that pulsar timing arrays \citep[PTAs;][]{1990ApJ361300F} will also detect very low frequency GWs around $10^{-8}~\mathrm{Hz}$.
@@
 A review of GW detectors (both current and proposed) is given in section \ref{sec:detectors} and a review of GW sources is given in section \ref{sec:sources}.
-In \ref{app:a} several example sensitivity curves are presented.
+In \ref{app:a}, several example sensitivity curves are presented.
 A website where similar figures can be generated is available at \url{www.ast.cam.ac.uk/~rhc26/sources/}.
@@
 Since the GW signal and detector output are both real it follows that $\tilde{h}(-f)=\tilde{h}^{*}(f)$ and $\tilde{n}(-f)=\tilde{n}^{*}(f)$; therefore, $S_{n}(f)=S_{n}(-f)$.
-The fact that $S_{n}(f)$ is an even function means that Fourier integrals over all frequencies can instead be written as integrals over positive frequencies only, e.g., (\ref{eq:meansquare1}) and (\ref{eq:snrinnerprod}); it is for this reason that $S_{n}(f)$ is called the \emph{one-sided} PSD.\footnote{An alternative convention is to use the two-sided PSD $S^{(2)}_{n}(f) = S_{n}(f)/2$.}
+The fact that $S_{n}(f)$ is an even function means that Fourier integrals over all frequencies can instead be written as integrals over positive frequencies only, e.g., (\ref{eq:meansquare2}) and (\ref{eq:snrinnerprod}); it is for this reason that $S_{n}(f)$ is called the \emph{one-sided} PSD.\footnote{An alternative convention is to use the two-sided PSD $S^{(2)}_{n}(f) = S_{n}(f)/2$.}
 
@@
  &= \lim_{T\rightarrow\infty}\frac{1}{2T} & \int_{-T}^{T}\mathrm{d}t\; \int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;{\mathcal{F}}\left\{ n(\tau) \right\}(f)\left[{\mathcal{F}}\left\{ n(\tau) \right\}(f')\right]^{*} \nonumber \\*
-\label{eq:meansquare} & & \times \hphantom{\int_{-T}^{T}} \exp\left(2\pi \rmi ft\right)\exp\left(-2\pi \rmi f't\right)\;,
+\label{eq:meansquare} & & \times \vphantom{\int_{-T}^{T}} \exp\left(2\pi \rmi ft\right)\exp\left(-2\pi \rmi f't\right)\;,
 \end{eqnarray}
@@
 &=& \int_{0}^{\infty}\mathrm{d}f\; S_{n}(f) \;.
+\label{eq:meansquare2}
 \end{eqnarray}
@@
 Whilst the magnitude of the Fourier transform of the signal $|\tilde{h}(f)|$ provides a simple quantification of the GW amplitude as a function of frequency, it has one main deficiency.
-For an inspiralling source ,the instantaneous amplitude can be orders of magnitude below the noise level in a detector; however, as the signal continues over many orbits, the SNR can be integrated up to a detectable level.
-It is useful to have a quantification of the GW amplitude that accounts for this effect; three such methods are described in the following subsections.
+For an inspiralling source, the instantaneous amplitude can be orders of magnitude below the noise level in a detector; however, as the signal continues over many orbits, the SNR can be integrated up to a detectable level.
+It is useful to have a quantification of the GW amplitude that accounts for this effect; we shall now describe how this can be achieved.
 
@@
 
-The PSD, as defined by (\ref{eq:psd}), has the nice property, demonstrated in (\ref{eq:meansquare1}), that integrated over all positive frequencies it gives the mean square amplitude of the signal in the detector.
+The PSD, as defined by (\ref{eq:psd}), has the nice property, demonstrated in (\ref{eq:meansquare2}), that integrated over all positive frequencies it gives the mean square amplitude of the signal in the detector.
 However, in one important regard it is less appealing than characteristic strain: the height of the source above the detector curve is no longer directly related to the SNR.
@@
 
-This quantity has one aesthetic advantage over the others: it automatically accounts for the fact that there is less energy in low frequency waves of the same amplitude.
+This quantity has one aesthetic advantage over the others: it automatically accounts for there being less energy in low frequency waves of the same amplitude.
 However, unlike characteristic strain, the area between the source and detector curves is no longer simply related to the SNR.
@@
 \begin{equation}\label{eq:differentdescriptions}
-H_0^2\Omega_\mathrm{GW}(f)= \frac{8 \pi G}{3 c^{2}} fS_{\mathrm{E}}(f) = \frac{2\pi^2}{3} f^3 S_h(f) = \frac{2\pi^2}{3} f^2 h_\mathrm{c}^2(f) = \frac{8\pi^2}{3} f^4 \left|\tilde{h}(f)\right|^2\; ,
+H_0^2\Omega_\mathrm{GW}(f)= \frac{8 \pi G}{3 c^{2}} fS_{\mathrm{E}}(f) = \frac{2\pi^2}{3} f^3 S_h(f) = \frac{2\pi^2}{3} f^2 \left[h_\mathrm{c}(f)\right]^2 = \frac{8\pi^2}{3} f^4 \left|\tilde{h}(f)\right|^2\; ,
 \end{equation}

### 2014-08-06 Sources.tex
Post-arXiv 1

@@
 
-\subsubsection{Neutron star binaries}
+\subsubsection{Neutron-star binaries}
 
@@
 The amplitude of the GWs depends upon the deformation of the neutron star.
-The magnitude of the distortion depends upon the neutron star equation of state and the stiffness of the crust, which are currently uncertain \citep{Chamel2008,Lattimer2012}; deformations can also be supported by internal magnetic fields \citep{Haskell2008}.
+The magnitude of the distortion depends upon the neutron star equation of state and the stiffness of the crust, which are currently uncertain \citep{Chamel2008,Lattimer2012}.
+Deformations can also be supported by internal magnetic fields \citep{Haskell2008}.
 Several known pulsars could be sources for the advanced detectors and upper limits from the initial detectors help to constrain the deformations.
@@
 
-\subsubsection{Massive black hole binaries}
+\subsubsection{Massive black-hole binaries}
 
@@
 The large range in the rate reflects our uncertainty in the growth mechanisms of the supermassive black hole population \citep{Volonteri2010}.
-Plotted in \ref{app:a} are boxes labelled ``\emph{$\approx 10^{6}$ solar mass binaries}'' with a ratio $h_\mathrm{c}/h_{n}=100$ for eLISA at its peak sensitivity.
+Plotted in \ref{app:a} are boxes labelled ``\emph{$\mathit{\approx 10^{6}}$ solar mass binaries}'' with a ratio $h_\mathrm{c}/h_{n}=100$ for eLISA at its peak sensitivity.
 The range of frequencies plotted is $(3\times 10^{-4}$--$3\times 10^{-1})~\mathrm{Hz}$; this corresponds to circular binaries in the mass range quoted above.
 
-\subsubsection{Galactic white dwarf binaries} \label{sec:GB}
+\subsubsection{Galactic white-dwarf binaries} \label{sec:GB}
 
@@
 
-EMRIs occur when a compact stellar mass object inspirals into a supermassive black hole.
+EMRIs occur when a compact stellar-mass object inspirals into a supermassive black hole.
 There is extreme uncertainty in the event rate for EMRIs due to the poorly constrained astrophysics in galactic centres \citep[e.g.,][]{Merritt2011}; the best guess estimate is around $25$ events per year with eLISA with SNR $\ge 20$ \citep{TheGravitationalUniverse}.
 The boxes labelled ``\emph{extreme mass ratio inspirals}'' plotted in \ref{app:a} have a characteristic strain of $h_\mathrm{c}=3\times 10^{-20}$ at $10^{-2}~\mathrm{Hz}$, which corresponds to a $10\Msun$ black hole inspiralling into a $10^{6}\Msun$ black hole at a luminosity distance of $1~\mathrm{Gpc}$.
-The frequency width of the box is somewhat unknown; EMRI events can occur into a black hole of any mass, and hence EMRIs can, in principal, occur at any frequency, so the boxes in \ref{app:a} are drawn with a width comparable to that of the LISA sensitivity curve.
+The frequency width of the box is somewhat uncertain; EMRI events can occur around a black hole of any mass, and hence EMRIs can, in principal, occur at any frequency.
+The boxes in \ref{app:a} are drawn with a width comparable to that of the LISA sensitivity curve.
 
@@
 
-The main target for PTAs is a stochastic background of GWs produced by a population of supermassive black hole binaries at cosmological distances \citep{SesanaVecchioColancino}.
-Supermassive black holes are known to lie at the centres of most galaxies and the black hole mergers are associated with the mergers of the host galaxies \citep{Volonteri2003,Ferrarese2005}.
+The main target for PTAs is a stochastic background of GWs produced by a population of supermassive black-hole binaries at cosmological distances \citep{SesanaVecchioColancino}.
+Supermassive black holes are known to lie at the centres of most galaxies and black-hole mergers are associated with the mergers of their host galaxies \citep{Volonteri2003,Ferrarese2005}.
 The current best 95\% confidence limit for the amplitude of the stochastic background is $h_\mathrm{c} = 2.7\times 10^{-15}$ at a frequency of $f_{0}=1~\mathrm{yr}^{-1}$ \citep{Shannon2013}.
@@
 
-Supermassive black hole binaries at higher frequencies are inspiralling faster and hence there are fewer of them per frequency bin.
-At a certain frequency, these sources will cease to be a background and become individually resolvable.
-It is currently unclear whether PTAs will detect an individual binary or a stochastic background first.
+Supermassive black-hole binaries at higher frequencies are inspiralling faster and, hence, there are fewer of them per frequency bin.
+At a certain frequency, these sources cease to be a background and become individually resolvable \cplb{[citation needed]}.
+It is currently unclear whether PTAs shall detect an individual binary or a stochastic background first.
 Plotted in \ref{app:a} with the label ``\emph{stochastic background}'' is a third of the current limit with a cut off frequency of $f=1~\mathrm{yr}^{-1}$ which is suggested by Monte Carlo population studies \citep{SesanaVecchioColancino}.
-For the resolvable sources, labelled ``\emph{$\approx 10^{9}$ solar mass binaries}'', the amplitude of the current limit is plotted between $\left(3\times 10^{-9}\right.$--$\left.3\times 10^{-7}\right)~\mathrm{Hz}$.
+For the resolvable sources, labelled ``\emph{$\mathit{\approx 10^{9}}$ solar mass binaries}'', the amplitude of the current limit is plotted between $\left(3\times 10^{-9}\right.$--$\left.3\times 10^{-7}\right)~\mathrm{Hz}$.
 

### 2014-08-06 Typesofsource.tex
Post-arXiv 1

@@
 \end{equation}
-between an minimum and maximum frequency set by the initial and final radius of the binary orbit.
+between minimum and maximum frequencies set by the initial and final radii of the binary orbit respectively.
 Here, we assume that the maximum and minimum frequencies are outside of the range of our detector and hence can be neglected.
@@
 \end{equation}
-and constraints are then placed on $A$.
-In practice, this power law also has upper and lower frequency cut-offs related to the population of source objects.
+and constraints are then placed on $A$. In practice, this power law also has upper and lower frequency cut-offs related to the population of source objects.
+
 A stochastic background from other sources, such as cosmic strings or relic GWs from the early Universe, can also be written in the same form as (\ref{eq:power}), but with different spectral indices: $\alpha=-7/6$ for cosmic strings or $\alpha$ in the range $-1$ to $-0.8$ for relic GWs \citep{Jenet}.
@@
 This method accounts for the there being power across all frequencies in the sensitivity band by integrating the noise-weighted signal over frequency.
-As our aim here is to present stochastic backgrounds alongside other types of sources for comparison, we do not use this approach.
+As our aim here is to present stochastic backgrounds alongside other types of sources for comparison, we will not use this approach.
 
@@
 
-An important aspect of burst search algorithms is to accurately estimate the noise properties within each time segment.
+\cplb{This paragraph needs references} An important aspect of burst search algorithms is to accurately estimate the noise properties within each time segment.
 To this extent, null data streams can be constructed that are insensitive to real GW signals.
@@
 \begin{equation}
-h_\mathrm{rss} = \left[\int \rmd t {|h_+(t)|}^2 + {|h_\times(t)|}^2\right]^{1/2}.
+h_\mathrm{rss}^2 = \int \rmd t \; {|h_+(t)|}^2 + {|h_\times(t)|}^2 \; .
 \end{equation}
@@
 \begin{equation}
-h_\mathrm{rss} \simeq |\tilde{h}(f)|\sqrt{\Delta f}\,,\label{eq:rela}
+h_\mathrm{rss} \simeq |\tilde{h}(f)|\sqrt{\Delta f}\;,\label{eq:rela}
 \end{equation}

### 2014-08-06 Detectors.tex
GEO related things

@@
 The most notable ground-based detectors are LIGO and Virgo, which work in collaboration, supported by GEO600.
-LIGO, Virgo and GEO600 have completed science runs as first-generation detectors \cplb{[citation needed]}.
+LIGO, Virgo and GEO600 have completed science runs as first-generation detectors \citep[e.g.,][]{Abadie2010}.
 As LIGO and Virgo are currently being upgraded their initial configurations are now referred to as Initial LIGO (iLIGO) and Initial Virgo (iVirgo) respectively.
@@
 There is an agreement to move one of the upgraded Hanford detector systems to a location in India \citep{LIGO-India,Unnikrishnan2013}.
-The GEO600 detector can also operate in a high frequency mode \citep[GEO-HF]{2006CQGra..23S.207W} \cplb{Add to table}.
+The GEO600 detector is not subject to a major upgrade plan, but since summer 2009 it has enhanced by a series of smaller improvements, notably improving high-frequency sensitivity \citep[GEO-HF;][]{2006CQGra..23S.207W}.
 The advanced detectors should start operation in the next couple of years, with LIGO-India following further in the future.

### 2014-08-08 Typesofsource.tex
Adding referencing details

@@
 As a consequence, waveform models are not required for detection; we simply rely on identifying the excess power produced by these burst sources.
-Typically, we may be looking for signals from core-collapse supernovae \citep{Ott2009}, the late stages of merging compact binaries, or more generally, signals from any unexpected or poorly modelled sources.
+Typically, we may be looking for signals from core-collapse supernovae \citep{Ott2009}, the late stages of merging compact binaries \citep{Abadie2012}, cosmic strings \citep{Aasi2014}, or more generally, signals from any unexpected or poorly modelled sources.
 
@@
 
-\cplb{This paragraph needs references} An important aspect of burst search algorithms is to accurately estimate the noise properties within each time segment.
+An important aspect of burst search algorithms is to accurately estimate the noise properties within each time segment.
 To this extent, null data streams can be constructed that are insensitive to real GW signals.
-In order to estimate the false alarm rate, the data from different detectors can be shifted in time to remove any genuine coincident GW signals.
+In order to estimate the false alarm rate, the data from different detectors can be shifted in time to remove any genuine coincident GW signals \citep{Cannon2008}.
 These time-shifts are then analysed to simulate the potential occurence of coincident noise events.

### 2014-08-12 Detectors.tex
removed factor of 1/2 in Eq. 51
tweak fig 1 caption and update cover letter
Cover letter read

@@
  \caption{The angular response function of an interferometric detector shown both as a surface plot and in an Aitoff--Hammer projection.
-The quentity that is plotted is is the polarisation average, $\int\textrm{d}\psi \;((F^{+})^{2}+(F^{\times})^{2})/2$ \cplb{Should this be divided by $2\pi$ too?}.
+The quantity that is plotted is is the polarisation average, $\left[(1/2\pi)\int\mathrm{d}\psi \;(F^{+})^{2}+(F^{\times})^{2}\right]^{1/2}$.
 The response is a function of two sky angles, $\theta$ and $\phi$, and varies between $0$ and $1$.
@@
 \begin{equation}\label{eq:skyav}
-F^{2}=\int_{0}^{2\pi}\frac{\mathrm{d}\psi}{2\pi}\; \int_{0}^{2\pi} \frac{\mathrm{d}\phi}{2\pi}\; \int_{0}^{\pi}\frac{\sin\theta\,\mathrm{d}\theta}{2}\;\left[\frac{F^{+}\left(\theta,\phi,\psi\right)^{2}+F^{\times}\left(\theta,\phi,\psi\right)^{2}}{2}\right]\; .
+F^{2}=\int_{0}^{2\pi}\frac{\mathrm{d}\psi}{2\pi}\; \int_{0}^{2\pi} \frac{\mathrm{d}\phi}{2\pi}\; \int_{0}^{\pi}\frac{\sin\theta\,\mathrm{d}\theta}{2}\;\left[F^{+}\left(\theta,\phi,\psi\right)^{2}+F^{\times}\left(\theta,\phi,\psi\right)^{2}\right]\; .
 \end{equation}

### 2014-09-24 V2/ReplyToReferees.txt
reply to referees in folder V2/

Dear Sir/Madam

Please find below a point by point reply to the comments of the two referees and a description of the changes we have made to the manuscript.

We have made changes to address all of the comments by the referees. These changes have removed a number of small typographical mistakes and improved the overall readability of the manuscript. All of the authors would like to thank the referees for their comments.

Yours faithfully,

Christopher Moore, Robert Cole and Christopher Berry.

Referee 1
It is very gratifying to hear that the referee enjoyed reading the manuscript. We would like to thank the referee for their time and their kind comments.

Referee 2
Major points:
1)	DISCUSS WITH ROB.

2)	We agree with this comment and think that it's well worth emphasising this point. We have added the following passage to the end of sec 5.4...

"It should be noted, however, that the sensitivity curves plotted in Appendix A would change substantially for stochastic backgrounds. This is because searches for stochastic backgrounds utilise the cross correlation of the outputs of multiple detectors for detection (Allen and Romano, 1999; Maggiore, 2000), instead of the individual outputs. Therefore the sensitivity of a network of ground based detectors is much greater than that of any individual detector. The shape of the sensitivity curve for a stochastic background would also differ from the curves plotted here, for a discussion of this point see Thrane and Romano (2013)."

Other comments (in order of appearance in the text):
1)	We thank the referee for pointing out this interesting source of confusion. We have checked this, and it seems that it was originally "evolved LISA", but the alternative "evolving LISA" has gradually become more popular and is now the prefered option. We have found numerous examples of both acronyms being used; for two conflicting examples on the same website see https://www.elisascience.org/news/top-news/gravitationaluniverseselectedasl3 and https://www.elisascience.org/articles/elisa-consortium/home . We have changed the main body of the text to read "evolving" added the following footnote...

"The “e” in eLISA originally stood for evolved, however, the preference is now for evolving."

2)	We agree with the referees comment and have amended the passage to read...

"Another disadvantage is that applying the definition 17 to a monochromatic source gives a formally undefined answer. The correct identification of characteristic strain for a monochromatic source is the amplitude of the wave times the square root of the number of periods observed."

3)	We have made this correction.

4)	We have made this correction.

5)	We have made this correction.

6)	We agree with the referee on this comment. The message we were trying to convey here is that current limits are comparable in order of magnitude. We have changed the relevant passage in sec 4.4.1 to read...

"The limits from the existing detectors are all of a similar magnitude; the PPTA limit is a factor of ∼ 2.5 lower than the other two. In Appendix A we use the EPTA limits, the figures show the published limit on A from equation 40 at a frequency of f0 = 1 yr−1 , based on an analysis of 5 pulsars over approximately 10 yr."

7)	THE EPTA CURVE DOESN'T ASSUME A DELTA_T. NEED TO CLARIFY ASSUMPTIONS FOR OTHER CURVES.

### 2014-09-24 V2/Detectors.tex
reply to referees in folder V2/

--- 2014-08-12 Detectors.tex
@@
  \centering
- \includegraphics[trim=0cm 0cm 0cm 0cm, width=0.6\textwidth]{Fig_Hellings_Downs}
+ \includegraphics[trim=0cm 0cm 0cm 0cm, width=0.55\textwidth]{Fig_Hellings_Downs}
  \caption{The \citet{HellingsDowns} curve, the correlation between two pulsars separated on the sky by an angle $\theta$.}
@@
 There are published limits on the amplitude of the stochastic background from all three detectors: the most recent from EPTA is \citet{Haasteren}, from PPTA is \citet{Shannon2013} and from NANOGrav is \citet{2013ApJ...762...94D}.
-These are all comparable and we use the EPTA limits in \ref{app:a}, the figures show the published limit on $A$ from equation \ref{eq:power} at a frequency of $f_{0}=1~\mathrm{yr^{-1}}$, based on an analysis of $5$ pulsars over approximately $10~\mathrm{yr}$.
+The limits from the existing detectors are all of a similar magnitude; the PPTA limit is a factor of $\sim 2.5$ lower than the other two.
+In \ref{app:a} we use the EPTA limits, this was based on an analysis of $5$ pulsars over approximately $10~\mathrm{yr}$.
+The curve labelled ``EPTA'' is normalised to agree with the limit published in \cite{Haasteren} at a frequency of $f_{0}=1~\mathrm{yr^{-1}}$.
 

### 2014-09-24 V2/Introduction.tex
reply to referees in folder V2/

--- 2014-08-06 Introduction.tex
@@
 By the end of the decade, it is expected that pulsar timing arrays \citep[PTAs;][]{1990ApJ361300F} will also detect very low frequency GWs around $10^{-8}~\mathrm{Hz}$.
-Further into the future, space-based detectors, such as the evolved Laser Interferometer Space Antenna \citep[eLISA;][]{TheGravitationalUniverse}, will probe GWs in the millihertz regime.
+Further into the future, space-based detectors, such as the evolving Laser Interferometer Space Antenna \citep[eLISA;][]{TheGravitationalUniverse}
+\footnote{The “e” in eLISA originally stood for evolved, however, the preference is now for evolving.}
 These advances shall herald the beginning of multi-wavelength GW astronomy as a means of observing the Universe.
@@
 One downside to plotting characteristic strain is that the values on the strain axis $h_\mathrm{c}(f)$ do not directly relate to the amplitude of the waves from the source.
-Another disadvantage is that the characteristic strain for a monochromatic source is infinite.
+Another disadvantage is that applying the definition \ref{eq:strain-hc} to a monochromatic source gives a formally undefined answer.
+The correct identification of characteristic strain for a monochromatic source is the amplitude of the wave times the square root of the number of periods observed.
+
+%Another disadvantage is that the characteristic strain for a monochromatic source is infinite.
 
@@
 
-The PSD, as defined by (\ref{eq:psd}), has the nice property, demonstrated in (\ref{eq:meansquare2}), that integrated over all positive frequencies it gives the mean square amplitude of the signal in the detector.
+The PSD, as defined by (\ref{eq:psd}), has the nice property, demonstrated in (\ref{eq:meansquare2}), that integrated over all positive frequencies it gives the mean square amplitude of the noise in the detector.
 However, in one important regard it is less appealing than characteristic strain: the height of the source above the detector curve is no longer directly related to the SNR.
@@
 \end{eqnarray} 
-where the definition (\ref{eq:ShforSources}) have been used.
+where the definition (\ref{eq:ShforSources}) has been used.
 The integrand in Equation (\ref{eq:specNRGdensity}) is defined as the spectral energy density, the energy per unit volume of space, per unit frequency \citep{HellingsDowns}

### 2014-09-24 V2/Sources.tex
reply to referees in folder V2/

--- 2014-08-06 Sources.tex
@@
 Supermassive black-hole binaries at higher frequencies are inspiralling faster and, hence, there are fewer of them per frequency bin.
-At a certain frequency, these sources cease to be a background and become individually resolvable \cplb{[citation needed]}.
+At a certain frequency, these sources cease to be a background and become individually resolvable \citep{SesanaVecchioColancino,Sesana2009}.
 It is currently unclear whether PTAs shall detect an individual binary or a stochastic background first.
@@
 As a consequence of this uncertainty, although we could learn much from measuring these signals, we have not included them amongst the sources in \ref{app:a}.
+
+It should be noted, however, that the sensitivity curves plotted in \ref{app:a} would change substantially for stochastic backgrounds.
+This is because searches for stochastic backgrounds utilise the cross correlation of the outputs of multiple detectors for detection \citep{1999PhRvD..59j2001A,2000PhR...331..283M}, instead of the individual outputs.
+Therefore the sensitivity of a network of ground based detectors is much greater than that of any individual detector.
+The shape of the sensitivity curve for a stochastic background would also differ from the curves plotted here, for a discussion of this point see \cite{2013PhRvD..88l4032T}.

### 2014-09-24 V2/Typesofsource.tex
reply to referees in folder V2/

--- 2014-08-08 Typesofsource.tex
@@
 \end{equation}
-Equation (\ref{eq:hc}) should be considered as the definition of characteristic strain and (\ref{eq:insphc}) a consequence of it for inspirals.
+Equation (\ref{eq:strain-hc}) should be considered as the definition of characteristic strain and (\ref{eq:insphc}) a consequence of it for inspirals.
 Equation (\ref{eq:insphc}) is the relation between $h_\mathrm{c}(f)$ and the instantaneous root-mean-square amplitude $h_0$ for an inspiralling source; for other types of source a new relation satisfying (\ref{eq:hc}) has to be found.

### 2014-09-24 V2/Introduction.tex
tweak

@@
 Further into the future, space-based detectors, such as the evolving Laser Interferometer Space Antenna \citep[eLISA;][]{TheGravitationalUniverse}
-\footnote{The “e” in eLISA originally stood for evolved, however, the preference is now for evolving.}
+\footnote{The e in eLISA originally stood for evolved, however, the preference is now for evolving.}
 These advances shall herald the beginning of multi-wavelength GW astronomy as a means of observing the Universe.

### 2014-09-24 V2/ReplyToReferees.txt
tweak
done point 7

@@
 
-"The “e” in eLISA originally stood for evolved, however, the preference is now for evolving."
+"The e in eLISA originally stood for evolved, however, the preference is now for evolving."
 
@@
 The message we were trying to convey here is that current limits are comparable in order of magnitude.
-We have changed the relevant passage in sec 4.4.1 to read...
+We have changed the relevant passage in sec 4.4.1 to now state that the PPTA limit is lower by a factor of ~2.5.
+Also see comment (7) for more changes to section 4.4.1.
 
-"The limits from the existing detectors are all of a similar magnitude; the PPTA limit is a factor of ∼ 2.5 lower than the other two.
-In Appendix A we use the EPTA limits, the figures show the published limit on A from equation 40 at a frequency of f0 = 1 yr−1 , based on an analysis of 5 pulsars over approximately 10 yr."
-
-7)	THE EPTA CURVE DOESN'T ASSUME A DELTA_T. NEED TO CLARIFY ASSUMPTIONS FOR OTHER CURVES.
+7)	We agree with the referee's comment that more information on the assumptions behind the PTA sensitivity curves in required.
+We have modified sections 4.4.1, 4.4.2 and 4.4.3 so that they now clearly state the total time, frequency of measurement, r.m.s.
+timing precision and number of pulsars for each PTA.

### 2014-09-24 V2/Detectors.tex
done point 7

@@
 There are published limits on the amplitude of the stochastic background from all three detectors: the most recent from EPTA is \citet{Haasteren}, from PPTA is \citet{Shannon2013} and from NANOGrav is \citet{2013ApJ...762...94D}.
-The limits from the existing detectors are all of a similar magnitude; the PPTA limit is a factor of $\sim 2.5$ lower than the other two.
-In \ref{app:a} we use the EPTA limits, this was based on an analysis of $5$ pulsars over approximately $10~\mathrm{yr}$.
-The curve labelled ``EPTA'' is normalised to agree with the limit published in \cite{Haasteren} at a frequency of $f_{0}=1~\mathrm{yr^{-1}}$.
+The limits from the existing detectors are all of a similar magnitude.
+In \ref{app:a} we use the EPTA limit based on an analysis of $5$ pulsars over approximately $10~\mathrm{years}$.
+The curve labelled ``EPTA'' assumes that all the pulsar were timed identically every 2 weeks with an r.m.s.\ error in each timing residual of $100\,\textrm{ns}$.
+The NANOGrav limit is comparable to this, and the PPTA limit is a factor $\sim 2.5$ lower.
 
@@
 Combining the existing arrays would yield a single PTA using approximately $4$ times as many pulsars; this consortium of consortia is known as the International Pulsar Timing Array \citep[IPTA\footnote{\url{http://www.ipta4gw.org/}};][]{iptareview2013}.
-The IPTA curves plotted in the figures are based on $20$ pulsars timed for twice as long as the current EPTA.
+The IPTA curve plotted in the figures assumes $20$ pulsars timed every 2 weeks for 15 years with an r.m.s.\ error in each timing residual of $100\,\textrm{ns}$.
 
@@
 This shall greatly increase the sensitivity of pulsar timing \citep{Kramer2004}.
-The sensitivity curves plotted in \ref{app:a} for the SKA assume a root-mean-square error on the timing residuals a factor of $10$ better than current PTAs, and include $20$ pulsars timed for twice as long as the current EPTA.
+The sensitivity curves plotted in \ref{app:a} for the SKA assumes $50$ pulsars timed every 2 weeks for 15 years with an r.m.s.\ error in each timing residual of $100\,\textrm{ns}$.
+Our choice of 50 pulsars might be quite conservative, it is possible that the SKA will discover many more suitable milli-second pulsars than this but (initially at least) these won't all have been timed for at least a few years.

### 2014-09-24 V2/ReplyToReferees.txt
more additions

@@
 Referee 2
+(**) The referees suggestion of a dynamic online plotting tool, allowing for user-specified values (of chirp masses for example) is an excellent idea.
+We intend to gradually improve the website over time and this is certainly on our to-do list.
+However, the current structure is not well-suited to this kind of customisation and so a radical overhaul is required beyond the current scope of this paper.
+
 Major points:
-1)	DISCUSS WITH ROB.
+1)	We agree with the referee that the sources sections needed some improvements.
+Before we discuss the individual improvements, it should be noted that it is in general impossible to treat all the sources rigorously and consistently in a way that produces a meaningful sensitivity curve.
+We discuss the difficulties in producing a unified treatment in detail in sections 2 and 3.
+
+However, we have adjusted some of our treatments, to take into account the referee's specific points.
+
+a-Sources with flat-topped boxes
+(i) SMBH binaries.
+We agree with the referee's discussion of characteristic strain for a binary with fixed chirp mass, h ~ f^(7/6).
+However, PTAs may detect a heavier (lighter) binary at an earlier (later) stage of the inspiral radiating at the same frequency.
+The desired shape of the box therefore depends not only on the scaling pointed out by the referee, but also on the distribution of SMBH chirp masses.
+This is astrophysically uncertain and so it is unclear what the best shape is.
+
+We have added the following text to sec 5.3.1 ...
+
+"The distribution of source amplitudes with frequency is dependent on the chirp mass distribution of binaries, which is astrophysically uncertain.
+Reflecting our lack of information about the true shape of this box, we draw it with a flat top in characteristic strain."
+
+An alternative method would be to plot the h ~ f^(7/6) strain for a fiducial binary, as suggested by the referee, and allow for varying chirp masses in the online curve generation tool.
+This relates to our previous point (**) about the website.
+We believe our current convention is equally valid.
+
+(ii) Pulsars. Previously we plotted the pulsar box with a flat top. We have now modified this, see sec 5.1.3.
 

### 2014-09-24 V2/Sources.tex
more additions

@@
 The signals are near monochromatic with a frequency twice the rotation frequency of the neutron star, and are a potential source for ground-based detectors.
-The amplitude of the GWs depends upon the deformation of the neutron star.
+The amplitude of the GWs depends upon the deformation of the neutron star, and its spin frequency.
 The magnitude of the distortion depends upon the neutron star equation of state and the stiffness of the crust, which are currently uncertain \citep{Chamel2008,Lattimer2012}.
@@
 Several known pulsars could be sources for the advanced detectors and upper limits from the initial detectors help to constrain the deformations.
+
 The boxes labelled ``\emph{pulsars}'' plotted in \ref{app:a} correspond to the upper limits placed on a GW signal from the Crab pulsar \citep{Aasi2014a}, extrapolated across a frequency range between $(20$--$10^{3})~\mathrm{Hz}$.
+The extrapolation was performed using the scaling $h_0 \propto \sqrt{\dot{f}/f}$~\citep{Aasi2014a} (this is the spin-down limit, which assumes that the only loss of energy from the system is due to GW emission) and (\ref{eq:insphc}).
+This gives $h_\mathrm{c} \propto f^{1/2}$; higher frequency sources are observed at a louder SNR because they undergo more cycles.
 
@@
 There is strong theoretical evidence that the actual background lies close to the current limit \citep{Sesana-2012}.
+The distribution of source amplitudes with frequency is dependent on the chirp mass distribution of binaries, which is astrophysically uncertain.
+Reflecting our lack of information about the true shape of this box, we draw it with a flat top in characteristic strain.
 
+\subsubsection {Stochastic background of supermassive binaries}
 Supermassive black-hole binaries at higher frequencies are inspiralling faster and, hence, there are fewer of them per frequency bin.

### 2014-09-24 V2/ReplyToReferees.txt
point c

@@
 
+b-The referee's suggestion for the addition of plots showing a range of amplitudes would not be feasible within the paper itself, but would be a possible future addition to the online tool, see (**).
+To address this point within the paper itself we have added the following passage to section 5.3.2...
+
+"Plotted in Appendix A with the label “stochastic background ” is a third of the current limit with a cut off frequency of f = 1 yr−1.
+This is suggested by Monte Carlo population studies (Sesana et al., 2008), which give a range of plausible amplitudes the mean of which is plotted here."
+
+c-We agree with the referee's comment that there was little astrophysical information in the statement that h_c/h_n = 100.
+We have replaced this with a fiducial source, at the suggestion of the referee.
+The relevant passage in the text now reads ...
+
+""
+
 2)	We agree with this comment and think that it's well worth emphasising this point. We have added the following passage to the end of sec 5.4...

### 2014-09-24 V2/Sources.tex
point c

@@
 The large range in the rate reflects our uncertainty in the growth mechanisms of the supermassive black hole population \citep{Volonteri2010}.
-Plotted in \ref{app:a} are boxes labelled ``\emph{$\mathit{\approx 10^{6}}$ solar mass binaries}'' with a ratio $h_\mathrm{c}/h_{n}=100$ for eLISA at its peak sensitivity.
-The range of frequencies plotted is $(3\times 10^{-4}$--$3\times 10^{-1})~\mathrm{Hz}$; this corresponds to circular binaries in the mass range quoted above.
+A $10^5 \Msun$ binary can be observed out to a redshift $z \sim 5$ with an SNR of $100$ \citep{TheGravitationalUniverse}.
+This fiducial source gives the amplitude of the boxes labelled ``\emph{$\mathit{\approx 10^{6}}$ solar mass binaries}'' in \ref{app:a}.
+The range of frequencies is $(3\times 10^{-4}$--$3\times 10^{-1})~\mathrm{Hz}$, extrapolated using the slope $h_\mathrm{c}(f) \propto f^{-2/3}$ derived in section \ref{sec:stoch}.
 
@@
 It is currently unclear whether PTAs shall detect an individual binary or a stochastic background first.
-Plotted in \ref{app:a} with the label ``\emph{stochastic background}'' is a third of the current limit with a cut off frequency of $f=1~\mathrm{yr}^{-1}$ which is suggested by Monte Carlo population studies \citep{SesanaVecchioColancino}.
+Plotted in \ref{app:a} with the label ``\emph{stochastic background}'' is a third of the current limit with a cut off frequency of $f=1~\mathrm{yr}^{-1}$.
+This is suggested by Monte Carlo population studies \citep{SesanaVecchioColancino}, which give a range of plausible amplitudes the mean of which is plotted here.
 For the resolvable sources, labelled ``\emph{$\mathit{\approx 10^{9}}$ solar mass binaries}'', the amplitude of the current limit is plotted between $\left(3\times 10^{-9}\right.$--$\left.3\times 10^{-7}\right)~\mathrm{Hz}$.

### 2014-09-24 V2/ReplyToReferees.txt
all done

@@
 
-""
+"A 105 M binary can be observed out to a redshift z ∼ 5 with an SNR of 100 (Seoane et al., 2013).
+This fiducial source gives the amplitude of the boxes labelled “≈ 10 6 solar mass binaries” in Appendix A.
+The range of frequencies is (3 × 10−4 –3 × 10−1 ) Hz, extrapolated using the slope hc (f ) ∝ f −2/3 derived in section 3.2."
 

### 2014-09-25 V2/ReplyToReferees.txt
V2 proof read

--- 2014-09-24 V2/ReplyToReferees.txt
@@
-Dear Sir/Madam
-
-Please find below a point by point reply to the comments of the two referees and a description of the changes we have made to the manuscript.
-
-We have made changes to address all of the comments by the referees.
-These changes have removed a number of small typographical mistakes and improved the overall readability of the manuscript.
-All of the authors would like to thank the referees for their comments.
+Please find below a point-by-point reply to the comments of the two referees, and a description of the changes we have made to the manuscript.
+We hope that these changes address all of the comments by the referees.
+They have corrected a number of small typographical mistakes and improved the overall readability of the manuscript.
+We would like to thank both the referees for their comments.
 
@@
 Referee 1
+
 It is very gratifying to hear that the referee enjoyed reading the manuscript. We would like to thank the referee for their time and their kind comments.
@@
 Referee 2
+
 (**) The referees suggestion of a dynamic online plotting tool, allowing for user-specified values (of chirp masses for example) is an excellent idea.
@@
 1)	We agree with the referee that the sources sections needed some improvements.
-Before we discuss the individual improvements, it should be noted that it is in general impossible to treat all the sources rigorously and consistently in a way that produces a meaningful sensitivity curve.
+Before we discuss the individual improvements, it should be noted that it is in general not possible to treat all the sources rigorously and consistently in a way that produces a meaningful sensitivity curve.
 We discuss the difficulties in producing a unified treatment in detail in sections 2 and 3.
-
 However, we have adjusted some of our treatments, to take into account the referee's specific points.
@@
 
-We have added the following text to sec 5.3.1 ...
+We have added to sec 5.3.1:
 
@@
 b-The referee's suggestion for the addition of plots showing a range of amplitudes would not be feasible within the paper itself, but would be a possible future addition to the online tool, see (**).
-To address this point within the paper itself we have added the following passage to section 5.3.2...
+To address this point within the paper itself we have added to section 5.3.2:
 
@@
 We have replaced this with a fiducial source, at the suggestion of the referee.
-The relevant passage in the text now reads ...
+The relevant passage now reads:
 
-""
+"A 10^5 M binary can be observed out to a redshift z ∼ 5 with an SNR of 100 (Seoane et al., 2013).
+This fiducial source gives the amplitude of the boxes labelled “≈ 10 6 solar mass binaries” in Appendix A.
+The range of frequencies is (3 × 10^−4 –3 × 10^−1 ) Hz, extrapolated using the slope hc (f ) ∝ f −2/3 derived in section 3.2."
 
-2)	We agree with this comment and think that it's well worth emphasising this point. We have added the following passage to the end of sec 5.4...
+2)	We agree with this comment and think that it's well worth emphasising this point. We have added to the end of sec 5.4:
 
-"It should be noted, however, that the sensitivity curves plotted in Appendix A would change substantially for stochastic backgrounds.
+"Tthe sensitivity curves plotted in Appendix A would change substantially for stochastic backgrounds.
 This is because searches for stochastic backgrounds utilise the cross correlation of the outputs of multiple detectors for detection (Allen and Romano, 1999; Maggiore, 2000), instead of the individual outputs.
-Therefore the sensitivity of a network of ground based detectors is much greater than that of any individual detector.
+Therefore, the sensitivity of a network of ground based detectors is much greater than that of any individual detector.
 The shape of the sensitivity curve for a stochastic background would also differ from the curves plotted here, for a discussion of this point see Thrane and Romano (2013)."
@@
 We have found numerous examples of both acronyms being used; for two conflicting examples on the same website see https://www.elisascience.org/news/top-news/gravitationaluniverseselectedasl3 and https://www.elisascience.org/articles/elisa-consortium/home .
-We have changed the main body of the text to read "evolving" added the following footnote...
+We have changed the main body of the text to read "evolving" added the footnote:
 
-"The e in eLISA originally stood for evolved, however, the preference is now for evolving."
+"The e in eLISA originally stood for evolved, but the preference now is for evolving."
 
-2)	We agree with the referees comment and have amended the passage to read...
+2)	We agree with the referees comment and have amended the passage to read:
 

### 2014-09-25 V2/Detectors.tex
V2 proof read

@@
 Each beam is passed in to an optical cavity where it undergoes a number of round trips before being recombined to form an interference pattern.
-The ends of the cavity are, in the ideal case, freely floating test masses which move in response to a passing GW, this effect is measured by observing the changing interference pattern.
+The ends of the cavity are, in the ideal case, freely floating test masses which move relative to each other in response to a passing GW, this effect is measured by observing the changing interference pattern.
 
@@
 \end{figure}
-Throughout this paper detector sensitivity refers to the polarisation and sky averaged sensitivity $F$, where
+Throughout this paper, detector sensitivity refers to the polarisation and sky averaged sensitivity $F$, where
 \begin{equation}\label{eq:skyav}
@@
 A collection of interferometric detectors are listed in table \ref{table:t}, these are sensitive to GWs in the frequency range ${\mathcal{O}}(10$--$10^{3})~\mathrm{Hz}$.
-They all simulate free-floating test masses by suspending a mass from a pendulum system with natural frequency much greater than that of the GW.
+They all simulate free-floating test masses by suspending a mass from a pendulum system with natural frequency far removed from than that of the GW.
 Their sensitivity curves include narrow lines that arise from noise sources in the instrument, including resonances in the suspension system and electrical noise at multiples of $60~\mathrm{Hz}$: these have been removed in the \ref{app:a} figures for clarity.
@@
 There is an agreement to move one of the upgraded Hanford detector systems to a location in India \citep{LIGO-India,Unnikrishnan2013}.
-The GEO600 detector is not subject to a major upgrade plan, but since summer 2009 it has enhanced by a series of smaller improvements, notably improving high-frequency sensitivity \citep[GEO-HF;][]{2006CQGra..23S.207W}.
+The GEO600 detector is not subject to a major upgrade plan, but since summer 2009 it has been enhanced by a series of smaller improvements, notably improving high-frequency sensitivity \citep[GEO-HF;][]{2006CQGra..23S.207W}.
 The advanced detectors should start operation in the next couple of years, with LIGO-India following further in the future.
@@
 The Einstein Telescope (ET) is an ambitious proposal to construct an underground third-generation detector.
-Its location would provide shielding from seismic noise, allowing it to observe frequencies of $10$--$10^4~\mathrm{Hz}$.
+Its location would provide shielding from seismic noise, allowing it to observe frequencies of $(10$--$10^4)~\mathrm{Hz}$.
 
@@
 \end{equation}
-where $e^{A}_{ij}(\hat{\Omega})$ is the $A$ polarisation basis tensor for direction $\hat{\Omega}$  and $\vec{r}$ is the spatial position.
+where $e^{A}_{ij}(\hat{\Omega})$ is the $A$ polarisation basis tensor for direction $\hat{\Omega}$, and $\vec{r}$ is the spatial position.
 Choosing the SSB as the origin of our coordinate system, so the pulsar is at position $L\hat{p}$, gives
@@
 \end{eqnarray}
-The overlap function $\Gamma(f)$ tends to a constant value in the limit that the distances to the pulsars are large compared to the wavelength of GWs; PTAs operate in this limit, so the overlap may be approximated as a constant,
+The overlap function $\Gamma(f)$ tends to a constant value in the limit that the distances to the pulsars are large compared to the wavelength of GWs; PTAs operate in this limit \citep{Mingarelli2014}, so the overlap may be approximated as a constant,
 \begin{equation}
@@
 There is a discrepancy between the treatment of PTA sensitivity curves here and the higher frequency detectors discussed in sections \ref{sec:ground} and \ref{sec:space}.
-When observing a long-lived source, such as an inspiral, with a high frequency detector the convention was to define a \emph{characteristic} strain to satisfy (\ref{eq:hc}).
+When observing a long-lived source, such as an inspiral, with a high frequency detector, the convention was to define a \emph{characteristic} strain to satisfy (\ref{eq:hc}).
 Here, the convention is to leave the strain untouched and instead adjust the PTA sensitivity curve with observation time, again to satisfy (\ref{eq:hc}).
@@
 The limits from the existing detectors are all of a similar magnitude.
+
 In \ref{app:a} we use the EPTA limit based on an analysis of $5$ pulsars over approximately $10~\mathrm{years}$.
-The curve labelled ``EPTA'' assumes that all the pulsar were timed identically every 2 weeks with an r.m.s.\ error in each timing residual of $100\,\textrm{ns}$.
+The curve labelled ``EPTA'' assumes that all the pulsar were timed identically every 2 weeks with an r.m.s.\ error in each timing residual of $100~\textrm{ns}$.
 The NANOGrav limit is comparable to this, and the PPTA limit is a factor $\sim 2.5$ lower.
@@
 Combining the existing arrays would yield a single PTA using approximately $4$ times as many pulsars; this consortium of consortia is known as the International Pulsar Timing Array \citep[IPTA\footnote{\url{http://www.ipta4gw.org/}};][]{iptareview2013}.
-The IPTA curve plotted in the figures assumes $20$ pulsars timed every 2 weeks for 15 years with an r.m.s.\ error in each timing residual of $100\,\textrm{ns}$.
+The IPTA curve plotted in the figures assumes $20$ pulsars timed every 2 weeks for 15 years with an r.m.s.\ error in each timing residual of $100~\textrm{ns}$.
 
@@
 The sensitivity curves plotted in \ref{app:a} for the SKA assumes $50$ pulsars timed every 2 weeks for 15 years with an r.m.s.\ error in each timing residual of $100\,\textrm{ns}$.
-Our choice of 50 pulsars might be quite conservative, it is possible that the SKA will discover many more suitable milli-second pulsars than this but (initially at least) these won't all have been timed for at least a few years.
+Our choice of 50 pulsars might be conservative, it is possible that the SKA will discover many more suitable milli-second pulsars than this; however, not all pulsars will be discovered soon enough to accumulate sufficient observations to be of use in an analysis.
+%% CPLB: rewording, please check!

### 2014-09-25 V2/Introduction.tex
V2 proof read

@@
 \section{Introduction}
-
-%CPLB: alternate openning I put together a while ago.
-%Even the best of us can be frustrated by the use of different conventions or units when describing a single quantity.
-It is a common occurrence in physics that  there is a particular parameterization that is perfect for one experiment or calculation, but this is not of universal applicability.
-When comparing results across studies it is then necessary to trawl through the literature to check that $x$ is really the same parameter.
-Astronomy and astrophysics are rife with specialized units, from the jansky (a spectral flux density of $10^{-26}$ watts per square metre per hertz) to the solar neutrino unit (a neutrino flux that produces an interaction rate of $10^{-36}$ per second per target atom).
-Inevitably, things get lost in translation.
-The mismatch of conventions can be confusing to those new to the field, especially students.
 
@@
 Further into the future, space-based detectors, such as the evolving Laser Interferometer Space Antenna \citep[eLISA;][]{TheGravitationalUniverse}
-\footnote{The e in eLISA originally stood for evolved, however, the preference is now for evolving.}
+\footnote{The e in eLISA originally stood for evolved, but the preference now is for evolving.}
 These advances shall herald the beginning of multi-wavelength GW astronomy as a means of observing the Universe.
@@
 
-Since the GW signal and detector output are both real it follows that $\tilde{h}(-f)=\tilde{h}^{*}(f)$ and $\tilde{n}(-f)=\tilde{n}^{*}(f)$; therefore, $S_{n}(f)=S_{n}(-f)$.
+Since the GW signal and detector output are both real, it follows that $\tilde{h}(-f)=\tilde{h}^{*}(f)$ and $\tilde{n}(-f)=\tilde{n}^{*}(f)$; therefore, $S_{n}(f)=S_{n}(-f)$.
 The fact that $S_{n}(f)$ is an even function means that Fourier integrals over all frequencies can instead be written as integrals over positive frequencies only, e.g., (\ref{eq:meansquare2}) and (\ref{eq:snrinnerprod}); it is for this reason that $S_{n}(f)$ is called the \emph{one-sided} PSD.\footnote{An alternative convention is to use the two-sided PSD $S^{(2)}_{n}(f) = S_{n}(f)/2$.}
@@
 \left[h_\mathrm{c}(f)\right]^{2} &= 4f^{2}\left| \tilde{h}(f) \right|^{2} \; ,\\
-\left[h_{n}(f)\right]^{2} &= fS_{n}(f),
+\left[h_{n}(f)\right]^{2} &= fS_{n}(f) \; ,
 \label{eq:strain-hn}
@@
 One downside to plotting characteristic strain is that the values on the strain axis $h_\mathrm{c}(f)$ do not directly relate to the amplitude of the waves from the source.
-Another disadvantage is that applying the definition \ref{eq:strain-hc} to a monochromatic source gives a formally undefined answer.
+Another disadvantage is that applying (\ref{eq:strain-hc}) to a monochromatic source gives a formally undefined answer.
 The correct identification of characteristic strain for a monochromatic source is the amplitude of the wave times the square root of the number of periods observed.
-
-%Another disadvantage is that the characteristic strain for a monochromatic source is infinite.
+%% CPLB: Didn't we used to have a section explaining this?
+Perhaps add ref to 3.1?
 
@@
 
-A third way of describing the amplitude of a GW is through the energy carried by the waves.
+A third way of describing the amplitude of a GW is through the energy carried by the radiation.
 This has the advantage of having a clear physical significance.
@@
 where the definition (\ref{eq:ShforSources}) has been used.
-The integrand in Equation (\ref{eq:specNRGdensity}) is defined as the spectral energy density, the energy per unit volume of space, per unit frequency \citep{HellingsDowns}
+The integrand in (\ref{eq:specNRGdensity}) is defined as the spectral energy density, the energy per unit volume of space, per unit frequency \citep{HellingsDowns}
 \begin{equation}\label{eq:spectralenergydensity}
@@
 
-Cosmological studies often work in terms of the dimensionless quantity $\Omega_{\mathrm{GW}}$, the energy density per logarithmic frequency interval normalised to the critical density of the universe $\rho_{\mathrm{c}}$,
+Cosmological studies often work in terms of the dimensionless quantity $\Omega_{\mathrm{GW}}$, the energy density per logarithmic frequency interval, normalised to the critical density of the Universe $\rho_{\mathrm{c}}$,
 \begin{equation}
@@
 
-\subsection{Relating the different descriptions}
-The dimensionless energy density in GWs $\Omega_{\mathrm{GW}}$, spectral energy density $S_{\mathrm{E}}$, one-sided PSD, $S_{h}$, characteristic strain $h_\mathrm{c}$ and frequency-domain strain $\tilde{h}(f)$ are related via
+\subsection{Relating the descriptions}
+
+The dimensionless energy density in GWs $\Omega_{\mathrm{GW}}$, spectral energy density $S_{\mathrm{E}}$, one-sided PSD $S_{h}$, characteristic strain $h_\mathrm{c}$ and frequency-domain strain $\tilde{h}(f)$ are related via
 \begin{equation}\label{eq:differentdescriptions}

### 2014-09-25 V2/Sources.tex
V2 proof read

@@
 The question of the height of the box is more problematic; it would be desirable to normalise each box so that there was a fixed event rate, say one event per year with an amplitude lying within the box.
-However for many of the sources considered the event rate is subject to a large degree of uncertainty (and estimates change rapidly as our understanding of the astrophysics improves).
-So instead we take the more definite, but somewhat arbitrary, approach that for each type of source a fiducial event is nominated (with parameters or amplitude detailed in the relevant section below) and the amplitude of this event is used to position the top of the box.
-The parameters for each fiducial event are chosen such that the resulting amplitude is roughly consistent with the maximum end of current predictions: the boxes indicate where we \emph{could} plausibly find GW sources; the actual event rates could turn out to be substantially lower than this upper bound.
+However, for many of the sources considered, the event rate is subject to a large degree of uncertainty (and estimates change rapidly as our understanding of the astrophysics improves).
+Consequently, we take the more definite, but somewhat arbitrary, approach that for each type of source a fiducial event is nominated (with parameters or amplitude detailed in the relevant section below) and the amplitude of this event is used to position the top of the box.
+The parameters for each fiducial event are chosen such that the resulting amplitude is roughly consistent with the optimistic end of current predictions: the boxes indicate where we \emph{could} plausibly find GW sources; the actual event rates could turn out to be substantially lower than this upper bound.
 
@@
 The distinction between resolvable and unresolvable is detector specific; here we choose LISA.
-This boundary will not be too different for eLISA but would move substantially for either of the decihertz detectors.
+This boundary shall not be too different for eLISA, but would move substantially for decihertz detectors.
 Plotted in \ref{app:a} with the label ``\emph{unresolvable galactic binaries}'' is the estimate of this background due to \citet{Nelemans} where an observation time of one year has been assumed,
@@
 
-It should be noted, however, that the sensitivity curves plotted in \ref{app:a} would change substantially for stochastic backgrounds.
+The sensitivity curves plotted in \ref{app:a} would change substantially for stochastic backgrounds.
 This is because searches for stochastic backgrounds utilise the cross correlation of the outputs of multiple detectors for detection \citep{1999PhRvD..59j2001A,2000PhR...331..283M}, instead of the individual outputs.
-Therefore the sensitivity of a network of ground based detectors is much greater than that of any individual detector.
+Therefore, the sensitivity of a network of ground based detectors is much greater than that of any individual detector.
 The shape of the sensitivity curve for a stochastic background would also differ from the curves plotted here, for a discussion of this point see \cite{2013PhRvD..88l4032T}.

### 2014-09-25 V2/Typesofsource.tex
V2 proof read

@@
 \end{equation}
-The higher-order terms cause the expoential to oscillate rapidly such that these terms integrate to zero and may be neglected.
-Performing a similar expansion about $f = -f'$ for the second term in (\ref{eq:fourier-h}) and then evaluating the Gaussian integrals gives
+The higher-order terms cause the expoential to oscillate rapidly, such that these terms integrate to zero and may be neglected.
+Performing a similar expansion about $f = -f'$ for the second term in (\ref{eq:fourier-h}), and then evaluating the Gaussian integrals gives
 \begin{eqnarray} \label{eq:FTofEMRI}
@@
 \begin{equation}
-h_\mathrm{rss} \simeq |\tilde{h}(f)|\sqrt{\Delta f}\;,\label{eq:rela}
+h_\mathrm{rss} \simeq \left|\tilde{h}(f)\right|\sqrt{\Delta f}\;,\label{eq:rela}
 \end{equation}

### 2014-09-25 V2/noise_curve.tex
reply to referees in folder V2/
V2 proof read

--- 2014-08-02 noise_curve.tex
@@
 \usepackage{amsmath}
+\usepackage{microtype}
 \usepackage{url}
@@
 
-\title{Gravitational wave sensitivity curves}
+\title{Gravitational-wave sensitivity curves}
 \author{C J Moore$^1$, R H Cole$^1$ \& C P L Berry$^{1,\,2}$}
@@
 \address{$^2$ School of Physics and Astronomy, University of Birmingham, Edgbaston, Birmingham B15 2TT, United Kingdom}
-\ead{cjm96@ast.cam.ac.uk}
+\ead{cjm96@cam.ac.uk}
 
 \begin{abstract}
-There are several common conventions in use by the gravitational wave community to describe the amplitude of sources and the sensitivity of detectors.
+There are several common conventions in use by the gravitational-wave community to describe the amplitude of sources and the sensitivity of detectors.
 These are frequently confused.
-We outline the merits of and differences between the various quantities used for parameterizing noise curves and characterizing gravitational wave amplitudes.
+We outline the merits of and differences between the various quantities used for parameterizing noise curves and characterizing gravitational-wave amplitudes.
 We conclude by producing plots that consistently compare different detectors.
@@
 
-\bibliographystyle{jphysicsB}
+\bibliographystyle{jphysicsCB}
 \bibliography{bibliography}

### 2014-09-26 V2/ReplyToReferees.txt
final tweaks

@@
 "Another disadvantage is that applying the definition 17 to a monochromatic source gives a formally undefined answer.
-The correct identification of characteristic strain for a monochromatic source is the amplitude of the wave times the square root of the number of periods observed."
+The correct identification of characteristic strain for a monochromatic source is the amplitude of the wave times the square root of the number of periods observed (see section \ref{sec:insp})."
 

### 2014-09-26 V2/Detectors.tex
final tweaks

@@
 The sensitivity curves plotted in \ref{app:a} for the SKA assumes $50$ pulsars timed every 2 weeks for 15 years with an r.m.s.\ error in each timing residual of $100\,\textrm{ns}$.
-Our choice of 50 pulsars might be conservative, it is possible that the SKA will discover many more suitable milli-second pulsars than this; however, not all pulsars will be discovered soon enough to accumulate sufficient observations to be of use in an analysis.
-%% CPLB: rewording, please check!
+Our choice of 50 pulsars might be conservative, it is possible that the SKA will discover many more suitable milli-second pulsars than this during its operation, in this case the corresponding sensitivity curve would be lower.

### 2014-09-26 V2/Introduction.tex
final tweaks

@@
  &= \int_{-\infty}^{\infty}\mathrm{d}t\;\int_{-\infty}^{\infty}\mathrm{d}t'\;K(t)K^{*}(t')\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\left<\tilde{n}(f)\tilde{n}^{*}(f')\right>\exp\left[2\pi\rmi(ft-f't')\right]\nonumber \\
-%&=\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}t\;\int_{-\infty}^{\infty}\mathrm{d}t'\;\frac{1}{2}S_{n}(f)K(t)K^{*}(t')\exp\left(2\pi\rmi f (t-t')\right)\nonumber \\
  &= \int_{-\infty}^{\infty}\mathrm{d}f \; \frac{1}{2}S_{n}(f)\tilde{K}(f)\tilde{K}^{*}(f) \; ,
@@
 Another disadvantage is that applying (\ref{eq:strain-hc}) to a monochromatic source gives a formally undefined answer.
-The correct identification of characteristic strain for a monochromatic source is the amplitude of the wave times the square root of the number of periods observed.
-%% CPLB: Didn't we used to have a section explaining this?
-Perhaps add ref to 3.1?
+The correct identification of characteristic strain for a monochromatic source is the amplitude of the wave times the square root of the number of periods observed (see section \ref{sec:insp}).
 

### 2014-09-26 V2/Detectors.tex
SKA thingy

@@
 The sensitivity curves plotted in \ref{app:a} for the SKA assumes $50$ pulsars timed every 2 weeks for 15 years with an r.m.s.\ error in each timing residual of $100\,\textrm{ns}$.
-Our choice of 50 pulsars might be conservative, it is possible that the SKA will discover many more suitable milli-second pulsars than this during its operation, in this case the corresponding sensitivity curve would be lower.
+Our choice of 50 pulsars might be conservative, it is possible that the SKA will discover many more suitable milli-second pulsars than this throughout its operation, in which case the corresponding sensitivity curve would become lower.

### 2014-10-02 V2/Detectors.tex
final tweaks from the second lot of referees comments

@@
 This shall greatly increase the sensitivity of pulsar timing \citep{Kramer2004}.
-The sensitivity curves plotted in \ref{app:a} for the SKA assumes $50$ pulsars timed every 2 weeks for 15 years with an r.m.s.\ error in each timing residual of $100\,\textrm{ns}$.
+The sensitivity curves plotted in \ref{app:a} for the SKA assumes $50$ pulsars timed every 2 weeks for 20 years with an r.m.s.\ error in each timing residual of $30\,\textrm{ns}$.
 Our choice of 50 pulsars might be conservative, it is possible that the SKA will discover many more suitable milli-second pulsars than this throughout its operation, in which case the corresponding sensitivity curve would become lower.

### 2014-10-02 V2/Detectors.tex
Merge

--- 2014-09-26 V2/Detectors.tex
@@
 The sensitivity curves plotted in \ref{app:a} for the SKA assumes $50$ pulsars timed every 2 weeks for 15 years with an r.m.s.\ error in each timing residual of $100\,\textrm{ns}$.
-Our choice of 50 pulsars might be conservative, it is possible that the SKA will discover many more suitable milli-second pulsars than this during its operation, in this case the corresponding sensitivity curve would be lower.
+Our choice of 50 pulsars might be conservative, it is possible that the SKA will discover many more suitable milli-second pulsars than this throughout its operation, in which case the corresponding sensitivity curve would becomew lower.

### 2014-10-02 V2/Detectors.tex

--- 2014-10-02 V2/Detectors.tex
@@
 This shall greatly increase the sensitivity of pulsar timing \citep{Kramer2004}.
-The sensitivity curves plotted in \ref{app:a} for the SKA assumes $50$ pulsars timed every 2 weeks for 20 years with an r.m.s.\ error in each timing residual of $30\,\textrm{ns}$.
+The sensitivity curves plotted in \ref{app:a} for the SKA assumes $50$ pulsars timed every 2 weeks for 20 years with an r.m.s.\ error in each timing residual of $30~\textrm{ns}$.
 Our choice of 50 pulsars might be conservative, it is possible that the SKA will discover many more suitable milli-second pulsars than this throughout its operation, in which case the corresponding sensitivity curve would become lower.

### 2014-10-30 arXiv_version/noise_curve.tex
Hellings-Downs normalisation

@@
 Neglecting the exponential terms in the overlap is the frequency-domain equivalent of neglecting the pulsar term in (\ref{eq:pulsarredshift}).
-The integral may be evaluated to give an expression depending only on the angle $\theta$ between the two pulsars; this is the famous Hellings and Downs curve, shown in figure \ref{fig:HnD},
+The integral may be evaluated to give an expression depending only on the angle $\theta$ between the two pulsars; appropriately renormalised, this is the famous Hellings and Downs curve, shown in figure \ref{fig:HnD},
 \begin{equation}
-\Gamma_{0} = \frac{1}{2}+\frac{3x}{2}\left(\ln x -\frac{1}{6}\right) \; ,
+\Gamma_{0} \propto \frac{1}{2}+\frac{3x}{2}\left(\ln x -\frac{1}{6}\right) \; ,
 \end{equation}

### 2014-11-13 V2/Plots.tex
web address update

--- 2014-08-04 arXiv_version/noise_curve.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%    INSTITUTE OF PHYSICS PUBLISHING                                   %
-%                                                                      %
-%   `Preparing an article for publication in an Institute of Physics   %
-%    Publishing journal using LaTeX'                                   %
-%                                                                      %
-%    LaTeX source code `ioplau2e.tex' used to generate `author         %
-%    guidelines', the documentation explaining and demonstrating use   %
-%    of the Institute of Physics Publishing LaTeX preprint files       %
-%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
-%                                                                      %
-%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
-%                                                                      %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\documentclass[fleqn,12pt]{iopart}
-\newcommand{\gguide}{{\it Preparing graphics for IOP journals}}
-%Uncomment next line if AMS fonts required
-\usepackage{iopams}  
-\usepackage{graphicx}
-\usepackage{multirow}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{url}
-\usepackage{natbib}
-\usepackage{appendix}
-\usepackage[margin=1cm]{caption}
-\usepackage{braket}
-\usepackage{fullpage}
-\usepackage{hyperref}
-\usepackage{color}
-
-\def\newblock{\hskip .11em plus .33em minus .07em}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% 	definition of a solar mass symbol for use in equations.      %%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newcommand{\Msun}{\ensuremath{M_{ \odot }}}
-\newcommand{\degree}{\ensuremath{^\circ}}
-\newcommand{\innerprod}[2]{\ensuremath{\left({#1}\middle|{#2}\right)}}
-\newcommand{\cplb}[1]{{\color{red} #1}}
-\newcommand{\rhc}[1]{{\color{green} #1}}
-\newcommand{\cjm}[1]{{\color{blue} #1}}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{document}
-
-\title{Gravitational wave sensitivity curves}
-\author{C J Moore$^1$, R H Cole$^1$ \& C P L Berry$^{1,\,2}$}
-\address{$^1$ Institute of Astronomy, Madingley Road, Cambridge, CB3 0HA, United Kingdom}
-\address{$^2$ School of Physics and Astronomy, University of Birmingham, Edgbaston, Birmingham B15 2TT, United Kingdom}
-\ead{cjm96@ast.cam.ac.uk}
-
-\begin{abstract}
-There are several common conventions in use by the gravitational wave community to describe the amplitude of sources and the sensitivity of detectors.
-These are frequently confused.
-We outline the merits of and differences between the various quantities used for parameterizing noise curves and characterising gravitational wave amplitudes.
-We conclude by producing plots that consistently compare different detectors.
-Similar figures can be generated on-line for general use at \url{http://www.ast.cam.ac.uk/~rhc26/sources/}.
-\end{abstract}
-
-\pacs{04.30.--w, 04.30.Db, 04.80.Nn, 95.55.Ym}
-
-\section{Introduction}
-
-The next few years  promise to deliver the first direct detection of gravitational waves (GWs).
-This will most likely be achieved by the advanced Laser Interferometer Gravitational-wave Observatory \citep[LIGO;][]{2010CQGra..27h4006H} and advanced Virgo \citep{Acernese2009} detectors operating in the frequency range $(10$--$10^{3})~\mathrm{Hz}$.
-By the end of the decade, it is expected that pulsar timing arrays \citep[PTAs;][]{1990ApJ361300F} will also detect very low frequency GWs around $10^{-8}~\mathrm{Hz}$.
-Further into the future, space-based detectors, such as the evolved Laser Interferometer Space Antenna \citep[eLISA;][]{TheGravitationalUniverse}, will probe GWs in the millihertz regime.
-These advances shall herald the beginning of multi-wavelength GW astronomy as a means of observing the Universe.
-
-There already exists an extensive literature assessing the potential of all of these detectors to probe the astrophysics of various sources.
-There are several different methods commonly used to describe the sensitivity of a GW detector and the strength of a GW source.
-It is common practice to summarise this information on a sensitivity-curve plot.
-When producing these plots, it is desirable to have a consistent convention between detectors and sources that allows information about both to be plotted on the same graph.
-Ideally, the detectors and sources are represented in such a way that the relative detectability of the signals is immediately apparent.
-
-In this work, we discuss the differing conventions commonly used in GW astronomy.
-The amplitude of a GW is a strain, a dimensionless quantity $h$.
-This gives a fractional change in length, or equivalently light travel time, across a detector.
-The strain is small, making it a challenge to measure: we are yet to obtain a direct detection of a GW.
-To calibrate our expectations for future detections it is necessary to quantify the sensitivity of our instruments and the strength of their target signals.
-When discussing the loudness of sources and the sensitivity of detectors there are three commonly used parametrizations based upon the strain: the characteristic strain, the power spectral density (PSD) and the spectral energy density.
-We aim to disambiguate these three and give a concrete comparison of different detectors.
-It is hoped that this will provide a useful reference for new and experienced researchers in this field alike.
-
-We begin by expounding the various conventions and the relationships between the conventions in sections \ref{sec:conventions} and \ref{sec:voc}.
-A review of GW detectors (both current and proposed) is given in section \ref{sec:detectors} and a review of GW sources is given in section \ref{sec:sources}.
-In \ref{app:a} several example sensitivity curves are presented.
-A website where similar figures can be generated is available at \url{www.ast.cam.ac.uk/~rhc26/sources/}.
-Here, the user may select which sources and detectors to include to tailor the figure to their specific requirements.
-
-\section{Signal parametrization}\label{sec:conventions}
-
-\subsection{Signal analysis preliminaries}
-
-Gravitational radiation has two independent polarization states denoted $+$ and $\times$; a general signal can be described as a linear combination of the two polarization states, $h \nobreak=\nobreak A_{+} h_{+}\nobreak +\nobreak A_{\times} h_{\times}$.
-The sensitivity of a detector to these depends upon the relative orientations of the source and detector.
-The output of a gravitational wave detector $s(t)$ contains a superposition of noise $n(t)$ and (possibly) a signal $h(t)$,
-\begin{equation}
-s(t) = n(t)+h(t) \; .
-\end{equation}
-
-We shall have recourse to work with the Fourier transform of the signal, using the conventions that
-\begin{eqnarray} \label{eq:defFourier}
-\tilde{x}(f) &= {\mathcal{F}}\left\{x(t)\right\}(f) &= \int_{-\infty}^{\infty}\mathrm{d}t\;x(t)\exp (-2\pi \rmi f t ) \; , \\
-x(t) &= {\mathcal{F}}^{-1}\left\{\tilde{x}(f)\right\}(t) &= \int_{-\infty}^{\infty}\mathrm{d}f\;\tilde{x}(f)\exp (2\pi \rmi f t ) \; .
-\end{eqnarray}
-
-For simplicity, it is assumed that the noise in the GW detector is stationary and Gaussian (with zero mean); under these assumptions the noise is fully characterised via the one-sided noise PSD $S_{n}(f)$,
-\begin{equation}\label{eq:psd}
-\left<\tilde{n}(f)\tilde{n}^{*}(f')\right>=\frac{1}{2}\delta (f-f')S_{n}(f) \; ,
-\end{equation}
-where angle brackets $\left<\ldots\right>$ denote an ensemble average over many noise realisations \citep{Cutler1994}.
-In reality, we have only a single realisation to work with, but the ensemble average can be replaced by a time average for stationary stochastic noise.
-The procedure is to measure the noise over a sufficiently long duration $T$ and then compute the Fourier transform $\tilde{n}(f)$ with a frequency resolution $\Delta f = T^{-1}$; this is repeated many times to give an average.
-The noise PSD $S_{n}(f)$ has units of inverse frequency.
-
-Since the GW signal and detector output are both real it follows that $\tilde{h}(-f)=\tilde{h}^{*}(f)$ and $\tilde{n}(-f)=\tilde{n}^{*}(f)$; therefore, $S_{n}(f)=S_{n}(-f)$.
-The fact that $S_{n}(f)$ is an even function means that Fourier integrals over all frequencies can instead be written as integrals over positive frequencies only, e.g., (\ref{eq:meansquare1}) and (\ref{eq:snrinnerprod}); it is for this reason that $S_{n}(f)$ is called the \emph{one-sided} PSD.\footnote{An alternative convention is to use the two-sided PSD $S^{(2)}_{n}(f) = S_{n}(f)/2$.}
-
-When integrated over all positive frequencies, the PSD gives the mean square noise amplitude. Starting by taking the time average of the square of the detector noise:
-\begin{eqnarray}
-\overline{\left|n(t)\right|^{2}} &= \lim_{T\rightarrow\infty}\frac{1}{2T} & \int_{-T}^{T}\mathrm{d}t\; n(t)n^{*}(t) \\
- &= \lim_{T\rightarrow\infty}\frac{1}{2T} & \int_{-T}^{T}\mathrm{d}t\; \int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\tilde{n}(f)\tilde{n}^{*}(f')\exp\left(2\pi \rmi ft\right)\exp\left(-2\pi \rmi f't\right) \nonumber\\
- &= \lim_{T\rightarrow\infty}\frac{1}{2T} & \int_{-T}^{T}\mathrm{d}t\; \int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;{\mathcal{F}}\left\{ n(\tau) \right\}(f)\left[{\mathcal{F}}\left\{ n(\tau) \right\}(f')\right]^{*} \nonumber \\*
-\label{eq:meansquare} & & \times \hphantom{\int_{-T}^{T}} \exp\left(2\pi \rmi ft\right)\exp\left(-2\pi \rmi f't\right)\;,
-\end{eqnarray}
-where we have substituted in using the definitions of the Fourier transform and its inverse.
-A property of Fourier transforms is that a time-domain translation by amount $t$ is equivalent to a frequency-domain phase change $2\pi ft$; if ${\mathcal{F}}\left\{ n(\tau) \right\}(f) = \tilde{n}(f)$, then ${\mathcal{F}}\left\{ n(\tau-t) \right\}(f) = \tilde{n}(f)\exp (2\pi \rmi ft )$.
-Therefore, the exponential factors in (\ref{eq:meansquare}) may be absorbed as
-\begin{eqnarray}
-\overline{\left|n(t)\right|^{2}} &=& \lim_{T\rightarrow\infty}\frac{1}{2T}\int_{-T}^{T} \mathrm{d}t\:\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;{\mathcal{F}}\left\{ n(\tau-t) \right\}(f)\left[{\mathcal{F}}\left\{ n(\tau-t) \right\}(f')\right]^{*} \, .
-\end{eqnarray}
-Since the noise is a randomly varying signal, we can use the ergodic principle to equate a time average, denoted by $\overline{\left(\ldots\right)}$, with an ensemble average, denoted by $\left<\ldots\right>$.
-The noise is stationary, consequently, its expectation value is unchanged by the time-translation performed above.
-Therefore, using (\ref{eq:psd}), the mean square noise amplitude is given by
-\begin{eqnarray}\label{eq:meansquare1}
-\overline{\left|n(t)\right|^{2}} &=& \int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\left<\tilde{n}(f)\tilde{n}^{*}(f')\right> \\
-&=& \int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\frac{1}{2}S_{n}(f)\delta(f-f') \nonumber \\
-&=& \int_{0}^{\infty}\mathrm{d}f\; S_{n}(f) \;.
-\end{eqnarray}
-
-Given a detector output, the challenge is to extract the signal.
-There is a well known solution to this problem that involves constructing a Wiener optimal filter \citep{Wiener49}.
-Let $K(t)$ be a real filter function with Fourier transform $\tilde{K}(f)$.
-Convolving this with the detector output gives a contribution from the signal and a contribution from the noise,
-\begin{equation}\label{eq:conv}
-\left(s*K\right)(\tau) = \int_{-\infty}^{\infty}\mathrm{d}t\;\left[h(t)+n(t)\right]K(t-\tau) \approx \mathcal{S} + \mathcal{N} \; .
-\end{equation}
-The signal contribution $\mathcal{S}$ is defined as the expectation of the convolution in (\ref{eq:conv}) when a signal is present, maximised by varying the offset to achieve the best overlap with the data.
-Since the expectation of pure noise is zero it follows that
-\begin{equation}
-\mathcal{S} = \int_{-\infty}^{\infty}\mathrm{d}t\;h(t)K(t)=\int_{-\infty}^{\infty}\mathrm{d}t\;h(t)K^{*}(t)=\int_{-\infty}^{\infty}\mathrm{d}f\; \tilde{h}(f)\tilde{K}^{*}(f) \; .
-\end{equation}
-The squared contribution from noise $\mathcal{N}^{2}$ is defined as the mean square of the convolution in (\ref{eq:conv}) when no signal is present,
-\begin{eqnarray} 
-\mathcal{N}^{2} &= \int_{-\infty}^{\infty}\mathrm{d}t\;\int_{-\infty}^{\infty}\mathrm{d}t'\;K(t)K(t')\left<n(t)n(t')\right> \nonumber \\
- &= \int_{-\infty}^{\infty}\mathrm{d}t\;\int_{-\infty}^{\infty}\mathrm{d}t'\;K(t)K^{*}(t')\int_{-\infty}^{\infty}\mathrm{d}f\;\int_{-\infty}^{\infty}\mathrm{d}f'\;\left<\tilde{n}(f)\tilde{n}^{*}(f')\right>\exp\left[2\pi\rmi(ft-f't')\right]\nonumber \\
- &= \int_{-\infty}^{\infty}\mathrm{d}f \; \frac{1}{2}S_{n}(f)\tilde{K}(f)\tilde{K}^{*}(f) \; ,
- \end{eqnarray}
-using the definition of $S_{n}(f)$ from (\ref{eq:psd}). Hence the signal-to-noise ratio (SNR) $\varrho$ is given by
-\begin{equation}\label{eq:SNRinnerprod} 
-\varrho^{2} = \frac{\mathcal{S}^{2}}{\mathcal{N}^{2}}= \frac{\innerprod{\frac{1}{2}S_{n}(f)\tilde{K}(f)}{\tilde{h}(f)}^{2}}{\innerprod{\frac{1}{2}S_{n}(f)\tilde{K}(f)}{\frac{1}{2}S_{n}(f)\tilde{K}(f)}},
-\end{equation}
-where we have introduced the inner product between signal $\tilde{A}$ and $\tilde{B}$ as \citep{Finn1992}
-\begin{equation}\label{eq:snrinnerprod} \innerprod{\tilde{A}(f)}{\tilde{B}(f)} = 4\Re\left\{\int_{0}^{\infty}\mathrm{d}f\;\frac{\tilde{A}^{*}(f)\tilde{B}(f)}{S_{n}(f)}\right\} \; .\end{equation}
-The optimum filter is that function $\tilde{K}(f)$ which maximises the SNR in (\ref{eq:SNRinnerprod}). From the Cauchy--Schwarz inequality, it follows that the optimum filter is
-\begin{equation}
-\tilde{K}(f)=\frac{\tilde{h}(f)}{S_{n}(f)} \; .
-\end{equation}
-This is the Wiener filter, which may be multiplied by an arbitrary constant since this does not change the SNR. Using this form for $\tilde{K}(f)$, the squared SNR is
-\begin{equation}
-\varrho^2 = \int_0^\infty\mathrm{d}f \frac{4| \tilde{h}(f)|^{2}}{S_n(f)} = \innerprod{\tilde{h}(f)}{\tilde{h}(f)}.
-\label{eq:traditionalSNR} 
-\end{equation}
-In order to construct the Wiener filter, it is necessary to know \emph{a priori} the form of the signal, $\tilde{h}(f)$, for this reason the Wiener filter is sometimes called the \emph{matched} filter.
-
-Whilst the magnitude of the Fourier transform of the signal $|\tilde{h}(f)|$ provides a simple quantification of the GW amplitude as a function of frequency, it has one main deficiency.
-For an inspiralling source ,the instantaneous amplitude can be orders of magnitude below the noise level in a detector; however, as the signal continues over many orbits, the SNR can be integrated up to a detectable level.
-It is useful to have a quantification of the GW amplitude that accounts for this effect; three such methods are described in the following subsections.
-
-\subsection{Characteristic strain}\label{sec:character-strain}
-
-The characteristic strain $h_\mathrm{c}$ is designed to include the effect of integrating an inspiralling signal.
-Its counterpart for describing noise is the noise amplitude $h_n$.
-These are defined as
-\begin{eqnarray}\label{eq:strain-hc} 
-\left[h_\mathrm{c}(f)\right]^{2} &= 4f^{2}\left| \tilde{h}(f) \right|^{2} \; ,\\
-\left[h_{n}(f)\right]^{2} &= fS_{n}(f),
-\label{eq:strain-hn}
-\end{eqnarray}
-such that the SNR in (\ref{eq:traditionalSNR}) may be written
-\begin{equation}\label{eq:hc} 
-\varrho^{2} = \int_{-\infty}^{\infty} \mathrm{d}\left(\log f\right)\; \left[\frac{h_\mathrm{c}(f)}{h_{n}(f)}\right]^{2} \;.
-\end{equation}
-The strain amplitudes $h_\mathrm{c}(f)$ and $h_{n}(f)$ are dimensionless.
-Using this convention, when plotting on a log--log scale, the area between the source and detector curves is related to the SNR via (\ref{eq:hc}).
-This convention allows the reader to integrate by eye to assess the detectability of a given source (see figure \ref{fig:hc}).
-
-An additional advantage of this convention is that the values on the strain axis for the detector curve $h_n(f)$ have a simple physical interpretation: they correspond to the root-mean-square noise in a bandwidth $f$.
-One downside to plotting characteristic strain is that the values on the strain axis $h_\mathrm{c}(f)$ do not directly relate to the amplitude of the waves from the source.
-Another disadvantage is that the characteristic strain for a monochromatic source is infinite.
-
-\subsection{Power spectral density}\label{sec:psd}
-
-A second commonly used quantity for sensitivity curves is the square root of the PSD or the amplitude spectral density (see figure \ref{fig:S}).
-When discussing a detector, rearranging (\ref{eq:strain-hn}) gives
-\begin{equation}\label{eq:temp1}
-\sqrt{S_{n}(f)} = h_{n}(f)f^{-1/2} \; ;
-\end{equation}
-by analogy, we can define an equivalent for source amplitudes
-\begin{equation}
-\sqrt{S_{h}(f)} = h_\mathrm{c}(f)f^{-1/2} = 2 f^{1/2} \left| \tilde{h}(f) \right| \; ,
-\label{eq:ShforSources}
-\end{equation}
-where we have used (\ref{eq:strain-hc}). Both $\sqrt{S_{n}(f)}$ and $\sqrt{S_{h}(f)}$ have units of $\mathrm{Hz^{-1/2}}$. The root PSD is the most frequently plotted quantity in the literature.
-
-The PSD, as defined by (\ref{eq:psd}), has the nice property, demonstrated in (\ref{eq:meansquare1}), that integrated over all positive frequencies it gives the mean square amplitude of the signal in the detector.
-However, in one important regard it is less appealing than characteristic strain: the height of the source above the detector curve is no longer directly related to the SNR.
-
-\subsection{Energy density}\label{sec:energy-density}
-
-A third way of describing the amplitude of a GW is through the energy carried by the waves.
-This has the advantage of having a clear physical significance.
-The energy density is most commonly used in sensitivity curves showing stochastic backgrounds of GWs (see section \ref{sec:stoch}).
-
-The energy in GWs is described by the Isaacson stress--energy tensor \citep[section 35.15]{MTW}
-\begin{equation}
-T_{\mu\nu}=\frac{c^{4}}{32\pi G}\left<\partial_{\mu}\bar{h}_{\alpha\beta}\partial_{\nu}\bar{h}^{\alpha\beta}\right> \;,
-\end{equation}
-where the angle brackets denote averaging over several wavelengths or periods, and $\bar{h}_{\alpha\beta}$ is the transverse-traceless metric perturbation.
-The energy density $\rho c^{2}$ is given by the $T_{00}$ component of this tensor.
-Consequently \citep[cf.][]{Berry2013},
-\begin{eqnarray}
-\rho c^{2} &=& \frac{c^{2}}{16\pi G}\int_{-\infty}^{\infty}\mathrm{d}f\;\left(2\pi f\right)^{2}\tilde{h}(f)\tilde{h}^{*}(f) \\*
- &=& \int_{0}^{\infty}\mathrm{d}f\;\frac{\pi c^{2}}{4G}f^{2}S_{h}(f)\; , \label{eq:specNRGdensity}
-\end{eqnarray} 
-where the definition (\ref{eq:ShforSources}) have been used.
-The integrand in Equation (\ref{eq:specNRGdensity}) is defined as the spectral energy density, the energy per unit volume of space, per unit frequency \citep{HellingsDowns}
-\begin{equation}\label{eq:spectralenergydensity}
-S_{\mathrm{E}}(f)=\frac{\pi c^{2}}{4G} f^{2}S_{h}(f) \; ;
-\end{equation}
-a corresponding expression for the noise can be formulated by replacing $S_h(f)$ with $S_{n}(f)$.
-
-Cosmological studies often work in terms of the dimensionless quantity $\Omega_{\mathrm{GW}}$, the energy density per logarithmic frequency interval normalised to the critical density of the universe $\rho_{\mathrm{c}}$,
-\begin{equation}
-\label{eq:omega}
-\Omega_\mathrm{GW}(f) = \frac{fS_{\mathrm{E}}(f)}{\rho_{\mathrm{c}}c^{2}} \; .
-\end{equation}
-The critical density is
-\begin{equation}
-\label{eq:crit-density}
-\rho_{\mathrm{c}}=\frac{3H_{0}^{2}}{8\pi G} \;,
-\end{equation}
-where $H_{0}$ is the Hubble constant, commonly parametrized as
-\begin{equation}
-H_0 = h_{100}\times 100~\mathrm{km\,s^{-1}\,Mpc^{-1}}.
-\end{equation}
-The reduced Hubble parameter $h_{100}$ has nothing to do with strain.
-The most common quantity related to energy density to be plotted on sensitivity curves is $\Omega_{\mathrm{GW}}h_{100}^{2}$ (figure \ref{fig:omega}) as this removes sensitivity to the (historically uncertain) measured value of the Hubble constant.
-
-This quantity has one aesthetic advantage over the others: it automatically accounts for the fact that there is less energy in low frequency waves of the same amplitude.
-However, unlike characteristic strain, the area between the source and detector curves is no longer simply related to the SNR.
-
-\subsection{Relating the different descriptions}
-The dimensionless energy density in GWs $\Omega_{\mathrm{GW}}$, spectral energy density $S_{\mathrm{E}}$, one-sided PSD, $S_{h}$, characteristic strain $h_\mathrm{c}$ and frequency-domain strain $\tilde{h}(f)$ are related via
-\begin{equation}\label{eq:differentdescriptions}
-H_0^2\Omega_\mathrm{GW}(f)= \frac{8 \pi G}{3 c^{2}} fS_{\mathrm{E}}(f) = \frac{2\pi^2}{3} f^3 S_h(f) = \frac{2\pi^2}{3} f^2 \left[h_\mathrm{c}(f)\right]^2 = \frac{8\pi^2}{3} f^4 \left|\tilde{h}(f)\right|^2\; ,
-\end{equation}
-using (\ref{eq:strain-hc}), (\ref{eq:ShforSources}), (\ref{eq:spectralenergydensity}), (\ref{eq:omega}) and (\ref{eq:crit-density}).
-Corresponding expressions for the noise are obtained by substituting $S_{n}(f)$ for $S_h(f)$, $h_{n}(f)$ for $h_\mathrm{c}(f)$ and $\tilde{n}(f)$ for $\tilde{h}(f)$.
-
-\section{Types of source}\label{sec:voc}
-
-GW signals can be broadly split into three categories: those from well-modelled sources, for which we have a description of the expected waveform; stochastic backgrounds, for which we can describe the statistical behaviour; and unmodelled (or poorly-modelled) transient sources.
-The classic example of a well-modelled source is the inspiral of two compact objects, this is discussed in section \ref{sec:insp}.
-Stochastic backgrounds can either be formed from many overlapping sources, which could be modelled individually, or from some intrinsically random process, these are discussed in section \ref{sec:stoch}.
-An example of an unmodelled (or poorly-modelled) transient source is a supernova; searches for signals of this type are often called burst searches and are discussed in section \ref{sec:bursts}.
-
-\subsection{Inspirals}\label{sec:insp}
-
-Inspiralling binaries may the most important GW source.
-They spend a variable amount of time in each frequency band.
-If $\phi$ is the orbital phase, then the number of cycles generated at frequency $f$ can be estimated as
-\begin{equation}\label{eq:inspiral}
-{N}_{\mathrm{cycles}} = \frac{f}{2\pi} \frac{\mathrm{d}\phi}{\mathrm{d}f} = \frac{f^{2}}{\dot{f}} \; ,
-\end{equation}
-where an overdot represents the time derivative and $\dot{\phi} = 2\pi f$.
-The squared SNR scales with ${N}_{\mathrm{cycles}}$, so it would be expected that $h_\mathrm{c}(f)\approx \sqrt{{N}_{\mathrm{cycles}}}|\tilde{h}(f)|$.
-
-The form for $h_\mathrm{c}$ can be derived from the Fourier transform in the stationary-phase approximation.
-Consider a source signal with approximately constant (root-mean-square) amplitude $h_0$ and central frequency $f'$.
-In this case,
-\begin{eqnarray}
-h(t) &=& \sqrt{2} h_0 \cos\left[\phi(t)\right] \; , \\
-\tilde{h}(f) &=& \frac{h_0}{\sqrt{2}} \int_{-\infty}^{\infty}\rmd t\; \exp \left\{2\pi \rmi \left[\frac{\phi(t)}{2\pi t}-f\right]t\right\} + \exp \left\{-2\pi \rmi \left[\frac{\phi(t)}{2\pi t}+f\right]t\right\} \; .
-\label{eq:fourier-h}
-\end{eqnarray}
-Without loss of generality, we can assume an initial phase of zero, such that $\phi(0) = 0$.
-The largest contribution to the integral comes from where the argument of the exponentials is approximately zero.
-For the first term, this occurs when $f = f'$, then the term in brackets is
-\begin{equation}
-\left[\frac{\phi(t)}{2\pi t}-f\right]_{f\,=\,f'} = f' + \dot{f'}t + \mathcal{O}\left(t^2\right) - f' = \dot{f'}t + \mathcal{O}\left(t^2\right).
-\end{equation}
-The higher-order terms cause the expoential to oscillate rapidly such that these terms integrate to zero and may be neglected.
-Performing a similar expansion about $f = -f'$ for the second term in (\ref{eq:fourier-h}) and then evaluating the Gaussian integrals gives
-\begin{eqnarray} \label{eq:FTofEMRI}
-\tilde{h}(f) &\simeq& \frac{h_0}{\sqrt{2}} \int_{-\infty}^{\infty}\rmd t\; \exp\left(2\pi\rmi \dot{f'} t^{2}\right) + \exp\left(-2\pi\rmi \dot{f'} t^{2}\right) \nonumber \\
-&\simeq& \frac{h_0}{\sqrt{2\dot{f'}}}.
-\end{eqnarray}
-From (\ref{eq:strain-hc}) and (\ref{eq:FTofEMRI}), the characteristic strain for inspiralling sources is given by \citep{FinnThorne}
-\begin{equation}\label{eq:insphc}
-h_\mathrm{c}(f) = \sqrt{\frac{2f^{2}}{\dot{f}}}h_0 \;.
-\end{equation}
-Equation (\ref{eq:hc}) should be considered as the definition of characteristic strain and (\ref{eq:insphc}) a consequence of it for inspirals.
-Equation (\ref{eq:insphc}) is the relation between $h_\mathrm{c}(f)$ and the instantaneous root-mean-square amplitude $h_0$ for an inspiralling source; for other types of source a new relation satisfying (\ref{eq:hc}) has to be found.
-
-\subsection{Stochastic backgrounds}\label{sec:stoch}
-
-Another important source of GWs is that of stochastic backgrounds, which can be produced from a large population of unresolvable sources.
-These can be at cosmological distances, where it is necessary to distinguish the frequency in the source rest frame $f_{\mathrm{r}}$ from the measured frequency $f$; the two are related through the redshift $z$ via $f_{\mathrm{r}}=(1+z)f$.
-The comoving number density of sources $\nu$ producing the background is also a function of redshift; if the sources producing the background are all in the local Universe, then simply set $\nu(z) = \nu_0\delta (z)$ and replace $d_{\mathrm{L}}(z)$ with $d$ in all that follows, where $d_{\mathrm{L}}(z)$ and $d$ are respectively the luminosity and comoving distances to the source, $d_{\mathrm{L}}(z)=(1+z)d$.
-
-We shall assume that the individual sources are binaries, in which case the number density of sources is also a function of the component masses.
-It is convenient to work in terms of the chirp mass, defined as ${\mathcal{M}}=\mu^{3/5}M^{2/5}$, where $\mu$ is the reduced mass and $M$ is the total mass of the binary.
-The comoving number density of sources shall be represented by $\nu(z, \mathcal{M})$.
-
-Equation (\ref{eq:differentdescriptions}) gives an expression for the energy density in GWs per logarithmic frequency interval,
-\begin{equation}\label{eq:stoch}
-fS_{\mathrm{E}}(f)=\frac{\pi c^{2}}{4G}f^{2}\left[h_\mathrm{c}(f)\right]^2 \; .
-\end{equation}
-The total energy emitted in the logarithmic frequency interval $\mathrm{d}\left(\log f_{\mathrm{r}}\right)$ by a single binary in the population is $\left[\mathrm{d}E_{\mathrm{GW}}/\mathrm{d}(\log f_{\mathrm{r}})\right]\mathrm{d}(\log f_{\mathrm{r}})$; the energy density may be written as
-\begin{equation}\label{eq:Phinney}
-fS_{\mathrm{E}}(f) = \int_{0}^{\infty} \rmd z\; \frac{\mathrm{d}\nu}{\mathrm{d}z}\frac{1}{(1+z)}\frac{1}{\left[d_{\mathrm{L}}(z)\right]^2}\frac{\mathrm{d}E_{\mathrm{GW}}}{\mathrm{d}\left(\log f_{\mathrm{r}} \right)} \; , \end{equation}
-where the factor of $\left( 1+z \right)^{-1}$ accounts for the redshifting of the energy.
-
-For simplicity, consider the background to comprise of binaries in circular orbits, with frequencies $f_\mathrm{GW} =f_{\mathrm{r}}/2$, which are far from their last stable orbit.
-The energy radiated may then be calculated using the quadrupole approximation \citep{petersmathews1963}.
-The energy in GWs from a single binary per logarithmic frequency interval is
-\begin{equation}\label{eq:Thorne}
-\frac{\mathrm{d}E_{\mathrm{GW}}}{\mathrm{d}\left(\log f_{\mathrm{r}} \right)} = \frac{G^{2/3}\pi^{2/3}}{3}{\cal{M}}^{5/3}f_{\mathrm{r}}^{2/3}
-\end{equation}
-between an minimum and maximum frequency set by the initial and final radius of the binary orbit.
-Here, we assume that the maximum and minimum frequencies are outside of the range of our detector and hence can be neglected.
-Using (\ref{eq:stoch}), (\ref{eq:Phinney}) and (\ref{eq:Thorne}), an expression for characteristic strain can now be found \citep{SesanaVecchioColancino}
-\begin{equation}\label{eq:bigint}
-\left[h_\mathrm{c}(f)\right]^2 = \frac{4G^{5/3}}{3\pi^{1/3}c^{2}}f^{-4/3}\int_{0}^{\infty}\mathrm{d}z\;\int_{0}^{\infty}\mathrm{d}{\cal{M}}\;\frac{\mathrm{d}^{2}\nu}{\mathrm{d}z\,\mathrm{d}{\mathcal{M}}}\frac{1}{\left[d_{\mathrm{L}}(z)\right]^2}\left( \frac{{\mathcal{M}}^{5}}{1+z} \right)^{1/3}\; .
-\end{equation}
-
-From (\ref{eq:bigint}) it can be seen that the characteristic strain due to a stochastic background of binaries is a power law in frequency with spectral index $\alpha=-2/3$.
-The amplitude of the background depends on the population statistics of the binaries under consideration via $\nu(z,{\mathcal{M}})$.
-The power law is often parametrised as
-\begin{equation}\label{eq:power} 
-h_\mathrm{c}(f) = A\left(\frac{f}{f_{0}}\right)^{\alpha}\; , 
-\end{equation}
-and constraints are then placed on $A$.
-In practice, this power law also has upper and lower frequency cut-offs related to the population of source objects.
-A stochastic background from other sources, such as cosmic strings or relic GWs from the early Universe, can also be written in the same form as (\ref{eq:power}), but with different spectral indices: $\alpha=-7/6$ for cosmic strings or $\alpha$ in the range $-1$ to $-0.8$ for relic GWs \citep{Jenet}.
-
-An alternative method for graphically representing the sensitivity of a GW detector to stochastic backgrounds, called the \emph{power-law-integrated sensitivity curve}, was suggested by \cite{2013PhRvD..88l4032T}.
-This method accounts for the there being power across all frequencies in the sensitivity band by integrating the noise-weighted signal over frequency.
-As our aim here is to present stochastic backgrounds alongside other types of sources for comparison, we do not use this approach.
-
-\subsection{Burst sources}\label{sec:bursts}
-
-Some sources of GWs can produce signals with large amplitudes, greater than the detector noise.
-The typical duration of such a signal is short, of the order of a few wave periods, and so there is not time to accumulate SNR in each frequency band as for inspirals.
-As a consequence, waveform models are not required for detection; we simply rely on identifying the excess power produced by these burst sources.
-Typically, we may be looking for signals from core-collapse supernovae \citep{Ott2009}, the late stages of merging compact binaries, or more generally, signals from any unexpected or poorly modelled sources.
-
-Burst searches are often carried out using time--frequency techniques.
-The data stream from a detector is temporally split into segments, the length of which can be tuned to give greater sensitivity to particular sources.
-Each segment is then transformed into the frequency domain, whitened and normalised to the noise spectrum of the detector to produce a time--frequency plot.
-Potential GW signals are identified by searching for clusters of pixels that contain an excess of power \citep[e.g.,][]{Bursts}.
-
-The presence of excess power across a number of pixels eliminates modelled noise sources, but such a cluster may also be caused by atypical noise within a detector.
-We can improve our confidence of a GW signal by making use of information obtained from other GW detectors.
-Signals across a network of detectors should have compatible arrival times (given the sky direction) as well as consistent amplitudes, frequencies and shapes of the waveform.
-Different pipelines are currently in use that analyse the signal consistency in different ways: both coincidence searches \citep{Chatterji2004} and fully coherent methods \citep{Klimenko2008} are used.
-
-An important aspect of burst search algorithms is to accurately estimate the noise properties within each time segment.
-To this extent, null data streams can be constructed that are insensitive to real GW signals.
-In order to estimate the false alarm rate, the data from different detectors can be shifted in time to remove any genuine coincident GW signals.
-These time-shifts are then analysed to simulate the potential occurence of coincident noise events.
-The algorithms are tuned using time-shifted data to ensure there is no bias in the final search.
-
-As discussed in \ref{sec:insp}, the expected relation between $h_\mathrm{c}(f)$ and a typical waveform $\tilde{h}(f)$ is
-\begin{equation}\label{eq:simple} 
-h_\mathrm{c}(f) = \sqrt{{N}_{\mathrm{cycles}}}\left|\tilde{h}(f)\right| \; , 
-\end{equation}
-where ${N}_{\mathrm{cycles}}$ is the number of cycles of radiation generated by the source, which is of order unity for bursts.
-
-An alternative characterisation of the signal amplitude commonly used for burst sources is the root-sum-square of the waveform polarisations:
-\begin{equation}
-h_\mathrm{rss} = \left[\int \rmd t {|h_+(t)|}^2 + {|h_\times(t)|}^2\right]^{1/2}.
-\end{equation}
-For a linearly polarised GW, with $\tilde{h}(f)$ constant across the bandwidth $\Delta f$, this is approximately related to the characteristic strain via 
-\begin{equation}
-h_\mathrm{rss} \simeq |\tilde{h}(f)|\sqrt{\Delta f}\,,\label{eq:rela}
-\end{equation}
-where we have neglected the detector response functions (see section \ref{sec:principles}), which are of order unity.
-In this work, we favour a constant $h_\mathrm{c}(f)$ rather than $h_\mathrm{rss}$ for consistency with the other types of source where the bandwidth is detector specific.
-
-\section{Detectors}\label{sec:detectors}
-
-In this section we introduce the detector noise curves used in \ref{app:a}.
-We begin with a description of the basic operation of detectors.
-We then discuss ground-based detectors, space-based detectors and PTAs in turn.
-The latter function somewhat differently than conventional interferometers, so we include a brief introduction to PTA analysis.
-References for the noise curves used for individual detectors can be found in the relevant subsections, further information about the detectors can be in discovered these.
-Detectors are frequently upgraded and redesigned, hence, while these curves are believed to be correct at the time of writing, it is best to check for updates from the appropriate science teams before relying on the details given here, although we hope that they shall remain accurate enough for illustrative purposes.
-
-\subsection{Operating principle of an interferometric detector}\label{sec:principles}
-
-All of the man-made detectors discussed in this section utilise the principle of interferometry.
-Such detectors work by taking a beam of monochromatic light and splitting it into two beams travelling at some angle to each other.
-Each beam is passed in to an optical cavity where it undergoes a number of round trips before being recombined to form an interference pattern.
-The ends of the cavity are, in the ideal case, freely floating test masses which move in response to a passing GW, this effect is measured by observing the changing interference pattern.
-
-The response of a detector to an incident plane-fronted GW depends upon the relative orientations of the detector and the incoming wave.
-Let us choose the origin of our coordinate system to be the beam-splitter of the interferometer, and $l_{1}^{i}$ and $l_{2}^{i}$ to be unit 3-vectors pointing along the two arms.
-In the absence of noise the output of the detector is the difference in strain between the two arms \citep{Thorne1987}
-\begin{equation}\label{eq:hoftxx}
-h(t)=\frac{1}{2}h_{ij}\left( l_{1}^{i}l_{1}^{j}-l_{2}^{i}l_{2}^{j} \right)\; ,
-\end{equation}
-where $h_{ij}$ are the spatial components of the GW metric perturbation.
-Let $\hat{r}^{i}$ be the unit 3-vector pointing towards the source of the GWs, with spherical polar angles $(\theta,\phi)$ relative to some axes fixed to the detector, and let $p^{i}$ and $q^{i}$ be unit vectors orthogonal to $\hat{r}^{i}$.
-We can now define the basis tensors
-\begin{eqnarray}
-H^{+}_{ij}&=p_{i}p_{j}-q_{i}q_{j} \; , \\
-H^{\times}_{ij}&=p_{i}q_{j}+q_{i}p_{j} \; .
-\end{eqnarray}
-There remains a freedom in the coordinates described, a rotation of $p^{i}$ and $q^{i}$ through an angle $\psi$ about $\hat{r}^{i}$ known as the polarization angle.
-For a single frequency component, the strain induced by a GW may be written as
-\begin{equation}\label{eq:hijxx}
-h_{ij}=A_{+}H^{+}_{ij}\cos\left(2\pi ft\right)+A_{\times}H^{\times}_{ij}\cos\left(2\pi ft+\Delta \phi\right) \; ,
-\end{equation}
-where $A_{+}$ and $A_{\times}$ are the amplitudes of the two polarisation states. Combining (\ref{eq:hoftxx}) and (\ref{eq:hijxx}) allows the detector output to be written as
-\begin{equation}
-h(t)=F^{+}(\theta,\phi,\psi)A_{+}\cos\left(2\pi ft\right)+F^{\times}(\theta,\phi,\psi)A_{\times}\cos\left(2\pi f t + \Delta\phi \right)\; ,
-\end{equation}
-where the response functions inherit their angular dependence from the choice of coordinates
-\begin{eqnarray}\label{eq:responsefuncs}
-F^{+}(\theta,\phi,\psi)&=\frac{1}{2}H^{+}_{ij}\left(l_{1}^{i}l_{1}^{j}-l_{2}^{i}l_{2}^{j}\right) \; , \\
-F^{\times}(\theta,\phi,\psi)&=\frac{1}{2}H^{\times}_{ij}\left(l_{1}^{i}l_{1}^{j}-l_{2}^{i}l_{2}^{j}\right) \; . 
-\end{eqnarray}
-
-The response function of a two-arm interferometric detector is quadrupolar, an example is plotted in figure \ref{fig:LIGO}.
-\begin{figure}
- \centering
- \includegraphics[trim=0cm 0cm 0cm 0cm, width=0.99\textwidth]{LIGO_detectorframe.pdf}
- \caption{The angular response function of an interferometric detector shown both as a surface plot and in an Aitoff--Hammer projection.
-The quentity that is plotted is is the polarisation average, $\int\textrm{d}\psi \;((F^{+})^{2}+(F^{\times})^{2})/2$.
-The response is a function of two sky angles, $\theta$ and $\phi$, and varies between $0$ and $1$.
-The two detector arms lie in the $x$--$y$ plane either side of one of the zeros in the response.}
- \label{fig:LIGO}
-\end{figure}
-Throughout this paper detector sensitivity refers to the polarisation and sky averaged sensitivity $F$, where
-\begin{equation}\label{eq:skyav}
-F^{2}=\int_{0}^{2\pi}\frac{\mathrm{d}\psi}{2\pi}\; \int_{0}^{2\pi} \frac{\mathrm{d}\phi}{2\pi}\; \int_{0}^{\pi}\frac{\sin\theta\,\mathrm{d}\theta}{2}\;\left[\frac{F^{+}\left(\theta,\phi,\psi\right)^{2}+F^{\times}\left(\theta,\phi,\psi\right)^{2}}{2}\right]\; .
-\end{equation}
-For a single $90\degree$-interferometer, such as LIGO, the sky and polarisation averaged response is $F=\sqrt{1/5}\approx 0.447$.
-
-A detector may consist of several interferometers.
-Let $F_{a}$ be the averaged response of the $a$-th interferometer, the average response of a network of $k$ detectors is obtained by adding in quadrature,
-\begin{equation} F_{\mathrm{Total}}^{2}=\frac{1}{k}\sum_{a\,=\,1}^{k}F_{a}^{2} \; .\end{equation}
-
-The averaging in (\ref{eq:skyav}) assumes a uniform distribution of polarisation angles $\psi$.
-This is the case for a stochastic background; however, for a non-inspiralling circular binary, the polarisation is a function of the two spherical polar angles ($\iota,\xi$) specifying the orientation of the binary's orbital angular momentum.
-Here, $\iota$ is the inclination angle, the polar angle between the orbital angular momentum and the line joining the source to the detector ($-\hat{r}$) and $\xi$ is the azimuthal angle around the same line.
-In this case, to characterise the detector sensitivity we average over all four angles ($\theta,\phi,\iota,\xi$).
-If the binary is inspiralling, then the polarisation depends on still more parameters which need to be averaged over.
-These more complicated averages all have the property that they depend on both the detector and the source, hence they are unhelpful for our present purpose separating the source amplitude from the detector sensitivity.
-Additionally, the different averages do not work out to be so different from each other: \citet{1993PhRvD..47.2198F} calculated the sensitivity for a detector with the LIGO geometry averaged over the four angles ($\theta,\phi,\iota,\xi$) as $\sqrt{4/25}=0.4$ times peak sensitivity, which should be compared with the value $\sqrt{1/5}\approx 0.447$ above.
-The effect of replacing the true sensitivity with the sky-averaged sensitivity for the LISA detector was considered in detail by \cite{2012CQGra..29l4015V}; they also found there is a small difference when considering an entire population of sources.
-For the remainder of this paper the three-angle average defined in (\ref{eq:skyav}) is used.
-
-\subsection{Ground-based detectors}\label{sec:ground}
-
-Ground-based detectors are the most numerous.
-A collection of interferometric detectors are listed in table \ref{table:t}, these are sensitive to GWs in the frequency range ${\mathcal{O}}(10$--$10^{3})~\mathrm{Hz}$.
-They all simulate free-floating test masses by suspending a mass from a pendulum system with natural frequency much greater than that of the GW.
-Their sensitivity curves include narrow lines that arise from noise sources in the instrument, including resonances in the suspension system and electrical noise at multiples of $60~\mathrm{Hz}$: these have been removed in the \ref{app:a} figures for clarity.
-The detectors fall broadly into three categories: first-generation detectors, which have already operated; second-generation detectors currently under construction; and third-generation detectors at the planning stage.
-
-The most notable ground-based detectors are LIGO and Virgo, which work in collaboration, supported by GEO600.
-LIGO, Virgo and GEO600 have completed science runs as first-generation detectors.
-As LIGO and Virgo are currently being upgraded their initial configurations are now referred to as Initial LIGO (iLIGO) and Initial Virgo (iVirgo) respectively.
-The upgraded, second-generation versions are referred to as advanced LIGO (aLIGO) and advanced Virgo (AdV) respectively.
-LIGO has two observatories: one at Hanford, Washington, which has two detectors; and another at Livingston, Louisiana.
-There is an agreement to move one of the upgraded Hanford detector systems to a location in India \citep{LIGO-India,Unnikrishnan2013}.
-The GEO600 detector can also operate in a high frequency mode \citep[GEO-HF]{2006CQGra..23S.207W}.
-The advanced detectors should start operation in the next couple of years, with LIGO-India following further in the future.
-
-TAMA300 is a Japanese first-generation detector.
-Its successor, currently under construction, is the Kamioka Gravitational Wave Detector (KAGRA), formerly the Large-scale Cryogenic Gravitational wave Telescope (LCGT), which is located underground in the Kamioka mine.
-It employs more sophisticated noise-reduction techniques than LIGO or Virgo, such as cryogenic cooling.
-
-The Einstein Telescope (ET) is an ambitious proposal to construct an underground third-generation detector.
-Its location would provide shielding from seismic noise, allowing it to observe frequencies of $10$--$10^4~\mathrm{Hz}$.
-
-\begin{table}
-\caption{\label{table:t} Summary of ground-based laser interferometers.}
-\begin{indented}
-\item[]\begin{tabular}{ l l l l l }
-\br
-{\bf Detector} & {\bf Country} & {\bf Arm length} & {\bf  Approximate date} & {\bf Generation} \\
-\mr
-  GEO600$^{a}$	&	Germany 	& $600~\mathrm{m}$ 	& 2001--present	   & First \\
-  TAMA300$^{b}$ & 	Japan		& $300~\mathrm{m}$ 	& 1995--present    & First \\
-  iLIGO$^{c}$	&	USA		& $4~\mathrm{km}$ 	& 2004--2010 	   & First \\
-  Virgo$^{d}$	& 	Italy		& $3~\mathrm{km}$ 	& 2007--2010 	   & First \\
-  aLIGO$^{e}$	&	USA		& $4~\mathrm{km}$ 	& \emph{est.} 2015 & Second \\
-  AdV$^{f}$	&	Italy	 	& $3~\mathrm{km}$ 	& \emph{est.} 2016 & Second \\
-  KAGRA$^{g}$	&	Japan		& $3~\mathrm{km}$ 	& \emph{est.} 2018 & Second \\
-  ET$^{h}$      & 	---		& $10~\mathrm{km}$ 	& \emph{est.} 2025 & Third \\
-\br
-\multicolumn{5}{l}{$^{a}$\citet{2010CQGra..27h4003G}, $^{b}$\citet{2002CQGra..19.1409A}, $^{c}$\citet{2009RPPh...72g6901A},}\\
-\multicolumn{5}{l}{$^{d}$\citet{accadia_virgo:2012}, $^{e}$\citet{2010CQGra..27h4006H}, $^{f}$\citet{Acernese2009}, $^{g}$\citet{2012CQGra..29l4007S}, $^{h}$\citet{2011CQGra..28i4013H}.}\\
-\br
-\end{tabular}
-\end{indented}
-\end{table}
-
-We use an interpolation to the data published on \url{https://wwwcascina.virgo.infn.it/advirgo/} (2013) for the AdV sensitivity curve, an interpolation to the data for version D of the KAGRA detector published on \url{http://gwcenter.icrr.u-tokyo.ac.jp/en/researcher/parameter} (2013) and analytic fits to the sensitivity curves from \citet{Sathyaprakash} for the remaining detectors.
-
-\subsection{Space-based detectors}\label{sec:space}
-
-Space-based detectors work on similar principles to ground based detectors, but with the test masses residing inside of independent, widely separated satellites.
-Space-based detectors are sensitive to lower frequency GWs than their ground-based counterparts; this is partly because space based detectors can have much longer arms, and partly because they are unaffected by seismic noise which limits the low frequency performance of ground-based detectors.
-
-The canonical design for a space-based detector is the Laser Interferometer Space Antenna (LISA), which is sensitive to millihertz GWs.
-LISA would consist of three satellites flying in a triangular constellation with arms of length $5\times 10^{9}~\mathrm{m}$ in a $1~\mathrm{AU}$ orbit around the Sun, trailing the Earth by $20\degree$.
-The laser arms in a LISA-like detector are not a cavity, the light only travels once along each arm.
-eLISA is a rescoped version of LISA designed to probe the same frequency range, while proposals such as the Advanced Laser Interferometer Antenna (ALIA), Big Bang Observer (BBO) and Deci-hertz Interferometer GW Observatory (DECIGO) are designed to probe decihertz GWs.
-
-\subsubsection{LISA and eLISA}
-
-The instrumental noise curves for LISA are approximated by the analytic fit given by \citet{Sathyaprakash}, which we use for the plots in \ref{app:a}.
-When observing individual sources with LISA there is an additional contribution to the noise from a background of unresolvable binaries.
-This is not included here as we consider the background as a source of GWs (see section \ref{sec:GB}).
-eLISA is a rescoped version of the classic LISA mission, the main differences are shorter arms ($10^{9}~\mathrm{m}$ instead of $5\times 10^{9}~\mathrm{m}$), two laser arms instead of three, and a different orbit (drifting away from Earth instead of $20\degree$ Earth trailing).
-The effect of these changes is a slightly reduced peak sensitivity and a shift to higher frequencies.
-We use an analytic fit to the instrumental noise curve given by \citet{Amaro-Seoane-et-al}.
-
-\subsubsection{DECIGO, ALIA and BBO}
-
-These missions are designed to probe the decihertz region of the GW spectrum; they are considerably more ambitious than the LISA or eLISA mission and their launches will be further into the future.
-We use a simple analytic fit to the sensitivity curve for ALIA~\citep{bender_possible_2013}, while for DECIGO and BBO, fits to the sensitivity curves given by \citet{2011PhRvD..83d4011Y} are used.
-
-\subsection{Pulsar timing arrays}\label{sec:PTAgeneralproperties}
-
-PTAs can be thought of as naturally occurring interferometers with galactic-scale arm lengths.
-Accordingly, they are sensitive to much lower frequencies than the detectors previously discussed.
-Each pulsar is a regular clock and the measured pulse arrival time can be compared against a prediction, leaving a residual which includes the effects of passing GWs.
-Using an array of these pulsars spread across the sky allows us to correlate residuals between different pulsars, to exploit the fact that GWs influence all pulsars whereas intrinsic pulsar noise does not.
-The correlation between different pulsars depends only on their angular separation on the sky, and has a distinctive shape, known as the Hellings and Downs curve \citep{HellingsDowns}.
-
-The redshift of the rate of arrival of pulses for a pulsar at a distance $L$ from the Solar-System barycentre (SSB), in the direction of the unit spatial vector $\hat{p}$ induced by a GW travelling in direction of the unit vector $\hat{\Omega}$ is \citep{anholm-2009}
-\begin{equation}\label{eq:pulsarredshift}
-z(t,\hat{\Omega})= \frac{1}{2} \frac{\hat{p}^{j}\hat{p}^{i}} {1+\hat{\Omega}\cdot\hat{p}}\left[h_{ij}^{\mathrm{Pulsar}}\left(t-\frac{L}{c},\hat{\Omega} \right)-h_{ij}^{\mathrm{Earth}}(t,\hat{\Omega} )\right] = \frac{1}{2}\frac{\hat{p}^{j}\hat{p}^{i}}{1+\hat{\Omega}\cdot\hat{p}} \Delta h_{ij}(t,\hat{\Omega})\; .
-\end{equation}
-The redshift includes two terms: the pulsar term and the Earth term.
-The pulsar term is often neglected in PTA analysis as it can be considered as an extra noise term which averages to zero across the array.
-The experimentally measured quantity is not the redshift but the timing residual, the two are related via
-\begin{equation}\label{eq:restored}
-R(t,\hat{\Omega}) = \int_{0}^{t}\mathrm{d}t'\;z(t',\hat{\Omega}) \; .
-\end{equation}
-All of the pulsars, and the Earth, are subject to the same metric-perturbation field. This may be expressed in terms of its Fourier transform
-\begin{equation}
-h_{ij}(t,\vec{r}) = \sum_{A\,=\,+,\,\times}\int\mathrm{d}f\;\iint_{\mathbb{S}_{2}}\mathrm{d}\hat{\Omega}\; \tilde{h}_{A}(f,\hat{\Omega})e_{ij}^{A}(\hat{\Omega}) \exp\left[2\pi i f \left(t-\frac{\hat{\Omega}\cdot\vec{x}}{c}\right)\right] \; ,
-\end{equation}
-where $e^{A}_{ij}(\hat{\Omega})$ is the $A$ polarisation basis tensor for direction $\hat{\Omega}$  and $\vec{r}$ is the spatial position.
-Choosing the SSB as the origin of our coordinate system, so the pulsar is at position $L\hat{p}$, gives
-\begin{equation}\label{eq:deltahij}
-\Delta h_{ij}(t,\hat{\Omega}) = \sum_{A\,=\,+,\,\times}\int\mathrm{d}f\; \tilde{h}_{A}(f,\hat{\Omega})e_{ij}^{A}(\hat{\Omega})\exp\left(2\pi \rmi f t\right)\left\{\exp\left[-2\pi\rmi fL \left(1+\hat{p}\cdot\hat{\Omega}\right)\right]-1\right\} \; .
-\end{equation}
-From (\ref{eq:pulsarredshift}) and (\ref{eq:deltahij}), the Fourier transform of the redshift $\tilde{z}(f,\hat{\Omega})$ can be identified as
-\begin{equation}\label{eq:45}
-\tilde{z}(f,\hat{\Omega}) = \left\{\exp\left[-2\pi\rmi fL \left(1+\hat{p}\cdot\hat{\Omega}\right)\right]-1\right\} \sum_{A\,=\,+,\,\times} \tilde{h}_{A}(f,\hat{\Omega})F^{A}(\hat{\Omega}) \; ,
-\end{equation}
-where
-\begin{equation}
-F^{A}(\hat{\Omega}) = \frac{e_{ij}^{A}(\hat{\Omega})\hat{p}^{j}\hat{p}^{i}} {2\left(1+\hat{\Omega}\cdot\hat{p}\right)} \, .
-\end{equation}
-The function $F^{A}(\hat{\Omega})$ may be regarded as the PTA equivalent of the detector response functions in (\ref{eq:responsefuncs}).
-The stochastic background of GWs is fully characterised by the one-sided PSD via the expectation value
-\begin{equation}\label{eq:stochback}
-\left<\tilde{h}_{A}^{*}(f,\hat{\Omega})\tilde{h}_{A'}(f',\hat{\Omega}')\right> = \frac{1}{2}S_{h}(f)\delta^{(2)}(\hat{\Omega}, \hat{\Omega}')\delta_{AA'}\delta(f-f') \; ,
-\end{equation}
-where $\delta^{(2)}(\hat{\Omega},\hat{\Omega}')$ is the delta-function on the sphere.
-From (\ref{eq:45}) and (\ref{eq:stochback}), the expectation of the product of signals from two different pulsars in directions $\hat{p}_{1}$ and $\hat{p}_{2}$ may be evaluated as
-\begin{equation} 
-\left<\tilde{z}_{1}(f)\tilde{z}_{2}^{*}(f')\right> = \frac{1}{2}S_{h}(f)\delta (f-f')\Gamma(f) \; ,
-\end{equation}
-where
-\begin{eqnarray}
-\Gamma(f) &= \sum_{A\,=\,+,\,\times} & \iint_{{\mathbb{S}}^{2}}\mathrm{d}\hat{\Omega}\,\left\{\exp\left[2\pi\rmi fL_{1}\left(1+\hat{\Omega}\cdot\hat{p}_{1}\right)\right]-1\right\} \nonumber \\*
- & & \times \left\{\exp\left[-2\pi\rmi fL_{2}\left(1+\hat{\Omega}\cdot\hat{p}_{2}\right)\right]-1\right\} F_{1}^{A}(\hat{\Omega})F_{2}^{A}(\hat{\Omega}) \; .
-\end{eqnarray}
-The overlap function $\Gamma(f)$ tends to a constant value in the limit that the distances to the pulsars are large compared to the wavelength of GWs; PTAs operate in this limit, so the overlap may be approximated as a constant,
-\begin{equation}
-\Gamma(f) \approx \Gamma_{0} = \sum_{A\,=\,+,\,\times}\iint_{{\mathbb{S}}^{2}}\mathrm{d}\hat{\Omega}\;F_{1}^{A}(\hat{\Omega})F_{2}^{A}(\hat{\Omega})\;.
-\end{equation}
-Neglecting the exponential terms in the overlap is the frequency-domain equivalent of neglecting the pulsar term in (\ref{eq:pulsarredshift}).
-The integral may be evaluated to give an expression depending only on the angle $\theta$ between the two pulsars; this is the famous Hellings and Downs curve, shown in figure \ref{fig:HnD},
-\begin{equation}
-\Gamma_{0} = \frac{1}{2}+\frac{3x}{2}\left(\ln x -\frac{1}{6}\right) \; ,
-\end{equation}
-where $x = (1-\cos\theta)/{2}.$
-\begin{figure}
- \centering
- \includegraphics[trim=0cm 0cm 0cm 0cm, width=0.6\textwidth]{HnDcurve.pdf}
- \caption{The \citet{HellingsDowns} curve, the correlation between two pulsars separated on the sky by an angle $\theta$.}
- \label{fig:HnD}
-\end{figure}
-
-The sensitivity bandwidth of a PTA is set by the sampling properties of the data set.
-If measurements are spaced in time by $\delta t$ and taken for a total length of time $T$, then the PTA is sensitive to frequencies in the range $(1/T) < f < (1/\delta t)$.
-The characteristic strain that the PTA is sensitive to scales linearly with $f$ in this range.
-This gives the wedge-shaped curves plotted in \ref{app:a}.
-The absolute value of the sensitivity is fixed by normalising to a calculated limit at a given frequency for each PTA.
-For a discussion of the sensitivities of PTA to both individual sources and stochastic background, see \citet{MooreTaylorGair}.
-
-There is a discrepancy between the treatment of PTA sensitivity curves here and the higher frequency detectors discussed in sections \ref{sec:ground} and \ref{sec:space}.
-When observing a long-lived source, such as an inspiral, with a high frequency detector the convention was to define a \emph{characteristic} strain to satisfy (\ref{eq:hc}).
-Here, the convention is to leave the strain untouched and instead adjust the PTA sensitivity curve with observation time, again to satisfy (\ref{eq:hc}).
-This discrepancy is an unfortunate result of the conventions in use by the different GW communities; however, it is also natural given the sources under observation.
-When observing a transient source, such as a burst or inspiral, which changes within the lifetime of the detector, it is natural to consider the detector as performing constantly while the signal changes.
-However, when observing a monochromatic source or a stochastic background, which is unchanging over the detector lifetime, it is more natural to consider the source as being fixed and the sensitivity of the detector gradually improving.
-All that is required by the definition in (\ref{eq:hc}) is that the ratio $h_\mathrm{c}(f)/h_{n}(f)$ is constant.
-
-\subsubsection{Current PTAs}
-
-The PTAs currently in operation are the European Pulsar Timing Array \citep[EPTA\footnote{\url{http://www.epta.eu.org/}};][]{eptareview2013}, the Parkes Pulsar Timing Array \citep[PPTA\footnote{\url{http://www.atnf.csiro.au/research/pulsar/ppta/}};][]{parkesreview2013} based in Australia, and the North American Nanohertz Observatory for Gravitational waves \citep[NANOGrav\footnote{\url{http://nanograv.org/}};][]{nanogravreview2013}).
-There are published limits on the amplitude of the stochastic background from all three detectors: the most recent from EPTA is \citet{Haasteren}, from PPTA is \citet{Shannon2013} and from NANOGrav is \citet{2013ApJ...762...94D}.
-These are all comparable and we use the EPTA limits in \ref{app:a}, the figures show the published limit on $A$ from equation \ref{eq:power} at a frequency of $f_{0}=1\,\textrm{year}^{-1}$, based on an analysis of five pulsars over approximately ten years.
-
-\subsubsection{The International Pulsar Timing Array}
-
-Combining the existing arrays would yield a single PTA using approximately 4 times as many pulsars; this consortium of consortia is known as the International Pulsar Timing Array \citep[IPTA\footnote{\url{http://www.ipta4gw.org/}};][]{iptareview2013}.
-The IPTA curves plotted in the figures are based on 20 pulsars timed for twice as long as the current EPTA.
-
-\subsubsection{SKA}
-
-The next great advancement in radio astronomy shall come with the completion of the Square Kilometre Array \citep[SKA;][]{Dewdney2009}.
-This shall greatly increase the sensitivity of pulsar timing \citep{Kramer2004}.
-The sensitivity curves plotted in \ref{app:a} for the SKA assume a root-mean-square error on the timing residuals a factor of 10 better than current PTAs, and include 20 pulsars timed for twice as long as the current EPTA.
-
-\section{Astrophysical sources}\label{sec:sources}
-
-All the sources described here are represented by shaded boxes in \ref{app:a}.
-Sources with short durations (i.e., burst sources) and sources that evolve in time over much longer timescales than our observations are drawn with flat-topped boxes for $h_\mathrm{c}(f)$.
-Inspiraling binaries, or stochastic backgrounds of binaries, are drawn with a sloping top proportional to $f^{-2/3}$ for $h_\mathrm{c}(f)$, which is result derived in section \ref{sec:stoch}.
-The width of the box gives the range of frequencies sources of a given type can have while remaining at a detectable amplitude.
-The question of the height of the box is more problematic; it would be desirable to normalise each box so that there was a fixed event rate, say one event per year with an amplitude lying within the box.
-However for many of the sources considered the event rate is subject to a large degree of uncertainty (and estimates change rapidly as our understanding of the astrophysics improves).
-So instead we take the more definite, but somewhat arbitrary, approach that for each type of source a fiducial event is nominated (with parameters or amplitude detailed in the relevant section below) and the amplitude of this event is used to position the top of the box.
-The parameters for each fiducial event are chosen such that the resulting amplitude is roughly consistent with the maximum end of current predictions: the boxes indicate where we \emph{could} plausibly find GW sources; the actual event rates could turn out to be substantially lower than this upper bound.
-
-\subsection{Sources for ground-based detectors}
-
-\subsubsection{Neutron star binaries}
-
-The inspiral and merger of a pair of neutron stars is the primary target for ground-based detectors.
-The expected event rate for this type of source is uncertain, but estimates centre around $\gamma_{\mathrm{NS-NS}}=1.3\times 10^{-4}~\mathrm{Mpc^{-3}\,yr^{-1}}$ \citep{CBC}.
-Plotted in \ref{app:a} are boxes labelled ``\emph{compact binary inspirals}'' with amplitudes such that a ratio $h_\mathrm{c}/h_{n}=16$ is produced for Advanced LIGO at peak sensitivity and a width between $(3$--$300)~\mathrm{Hz}$, corresponding to expected observable frequencies.
-
-\subsubsection{Supernovae}
-
-Simulations of core-collapse supernovae show that GWs between $(10^{2}$--$10^{3})~\mathrm{Hz}$ can be produced \citep{Kotake2006}.
-The GW signal undergoes ${\mathcal{O}}(1)$ oscillation and is hence burst-like.
-\citet{2002A&A...393..523D} calculate the average maximum amplitude of GWs for a supernova at distance $r$ as
-\begin{equation}
-h_\mathrm{max} = 8.9\times 10^{-21}\left( \frac{10~\mathrm{kpc}}{r} \right) \; .
-\end{equation}
-The event rate for supernovae is approximately $\gamma_{\mathrm{SN}} = 5\times10^{-4}~\mathrm{Mpc^{-3}\,yr^{-1}}$.
-The boxes labelled ``\emph{supernova}'' plotted in \ref{app:a} correspond to a distance $r=300~\mathrm{kpc}$ with the frequency range quoted above.
-The LIGO and Virgo detectors have already placed bounds on the event rate for these sources \citep{Bursts}.
-
-\subsubsection{Continuous waves from rotating neutron stars}
-
-Rotating neutron stars are a source of continuous GWs if they possess some degree of axial asymmetry \citep{Abbott2007, Prix2009, Einstein@Home}.
-The signals are near monochromatic with a frequency twice the rotation frequency of the neutron star, and are a potential source for ground-based detectors.
-The amplitude of the GWs depends upon the deformation of the neutron star.
-The magnitude of the distortion depends upon the neutron star equation of state and the stiffness of the crust, which are currently uncertain \citep{Chamel2008,Lattimer2012}; deformations can also be supported by internal magnetic fields \citep{Haskell2008}.
-Several known pulsars could be sources for the advanced detectors and upper limits from the initial detectors help to constrain the deformations.
-The boxes labelled ``\emph{pulsars}'' plotted in \ref{app:a} correspond to the upper limits placed on a GW signal from the Crab pulsar \citep{Aasi2014a}, extrapolated across a frequency range between $(20$--$10^{3})~\mathrm{Hz}$.
-
-\subsection{Sources for space-based detectors}
-
-For a review of the GW sources for space-based missions see, for example, \citet{Amaro-Seoane-et-al}, \citet{Gairetal} or \citet{eLISAyellowbook}.
-
-\subsubsection{Massive black hole binaries}
-
-Space-based detectors shall be sensitive to equal-mass mergers in the range $(10^{4}$--$10^{7})\,\Msun$.
-Predictions of the event rate for these mergers range from ${\mathcal{O}}(10$--$100)~\mathrm{yr}^{-1}$ for eLISA with SNRs of up to $10^3$ \citep{TheGravitationalUniverse}.
-The large range in the rate reflects our uncertainty in the growth mechanisms of the supermassive black hole population \citep{Volonteri2010}.
-Plotted in \ref{app:a} are boxes labelled ``\emph{$\approx 10^{6}$ solar mass binaries}'' with a ratio $h_\mathrm{c}/h_{n}=100$ for eLISA at its peak sensitivity.
-The range of frequencies plotted is $(3\times 10^{-4}$--$3\times 10^{-1})~\mathrm{Hz}$; this corresponds to circular binaries in the mass range quoted above.
-
-\subsubsection{Galactic white dwarf binaries} \label{sec:GB}
-
-For space-based detectors, these are the most numerous GW sources; they are also the only guaranteed source since several detectable systems (known as verification binaries) have already been identified by electromagnetic observations \citep{2006CQGra..23S.809S}.
-
-Galactic binaries divide into two classes: the unresolvable and the resolvable galactic binaries.
-The unresolvable binaries overlap to form a stochastic background as discussed in section \ref{sec:stoch}.
-The distinction between resolvable and unresolvable is detector specific; here we choose LISA.
-This boundary will not be too different for eLISA but would move substantially for either of the decihertz detectors.
-Plotted in \ref{app:a} with the label ``\emph{unresolvable galactic binaries}'' is the estimate of this background due to \citet{Nelemans} where an observation time of one year has been assumed,
-\begin{equation}
-h_\mathrm{c}(f)= 5\times 10^{-21} \left(\frac{f}{10^{-3}~\mathrm{Hz}}\right)^{-2/3} \; .
- \end{equation}
-Estimates for the event rate of resolvable binaries centre around ${\mathcal{O}}(10^{3})$ events for eLISA.
-The boxes plotted in \ref{app:a} with the label ``\emph{resolvable galactic binaries}'' have a ratio $h_\mathrm{c}/h_{n}=50$ for eLISA at its peak sensitivity.
-The frequency range of the box is $\left(3\times10^{-4}\right.$--$\left.10^{-2}\right)~\mathrm{Hz}$, estimated from Monte Carlo population simulation results presented in \citet{Amaro-Seoane-et-al}.
-
-\subsubsection{Extreme mass-ratio inspirals}
-
-EMRIs occur when a compact stellar mass object inspirals into a supermassive black hole.
-There is extreme uncertainty in the event rate for EMRIs due to the poorly constrained astrophysics in galactic centres \citep[e.g.,][]{Merritt2011}; the best guess estimate is around $25$ events per year with eLISA with SNR $\ge 20$ \citep{TheGravitationalUniverse}.
-The boxes labelled ``\emph{extreme mass ratio inspirals}'' plotted in \ref{app:a} have a characteristic strain of $h_\mathrm{c}=3\times 10^{-20}$ at $10^{-2}~\mathrm{Hz}$, which corresponds to a $10\Msun$ black hole inspiralling into a $10^{6}\Msun$ black hole at a luminosity distance of $1~\mathrm{Gpc}$.
-The frequency width of the box is somewhat unknown; EMRI events can occur into a black hole of any mass, and hence EMRIs can, in principal, occur at any frequency, so the boxes in \ref{app:a} are drawn with a width comparable to that of the LISA sensitivity curve.
-
-\subsection{Sources for PTAs}
-
-\subsubsection{Supermassive black hole binaries}
-
-The main target for PTAs is a stochastic background of GWs produced by a population of supermassive black hole binaries at cosmological distances \citep{SesanaVecchioColancino}.
-Supermassive black holes are known to lie at the centres of most galaxies and the black hole mergers are associated with the mergers of the host galaxies \citep{Volonteri2003,Ferrarese2005}.
-The current best 95\% confidence limit for the amplitude of the stochastic background is $h_\mathrm{c} = 2.7\times 10^{-15}$ at a frequency of $f_{0}=1~\mathrm{yr}^{-1}$ \citep{Shannon2013}.
-There is strong theoretical evidence that the actual background lies close to the current limit \citep{Sesana-2012}.
-
-Supermassive black hole binaries at higher frequencies are inspiralling faster and hence there are fewer of them per frequency bin.
-At a certain frequency, these sources will cease to be a background and become individually resolvable.
-It is currently unclear whether PTAs will detect an individual binary or a stochastic background first.
-Plotted in \ref{app:a} with the label ``\emph{stochastic background}'' is a third of the current limit with a cut off frequency of $f=1~\mathrm{yr}^{-1}$ which is suggested by Monte Carlo population studies \citep{SesanaVecchioColancino}.
-For the resolvable sources, labelled ``\emph{$\approx 10^{9}$ solar mass binaries}'', the amplitude of the current limit is plotted between $\left(3\times 10^{-9}\right.$--$\left.3\times 10^{-7}\right)~\mathrm{Hz}$.
-
-\subsection{Cosmological sources}
-
-In addition to the sources above, early Universe processes, such as inflation \citep{Grishchuk2005} or a first-order phase transition \citep{Binetruy2012}, could have created GWs.
-More speculatively, it has been hypothesised that cosmic strings could also be a potential source \citep{Damour2005,Binetruy2012,Aasi2014}.
-These relic GWs allow us to explore energy scales far beyond those accessible by other means, providing insight into new and exotic physics.
-The excitement surrounding the tentative discovery by BICEP2 of the imprint of primordial GWs (generated during inflation) in the cosmic microwave background \citep{Ade2014}, and the subsequent flurry of activity, has shown the scientific potential of such cosmological GWs.
-These GW signals are so alluring because they probe unknown physics; this also makes them difficult to predict.
-Cosmological stochastic backgrounds have been predicted across a range of frequencies with considerable variation in amplitude.
-As a consequence of this uncertainty, although we could learn much from measuring these signals, we have not included them amongst the sources in \ref{app:a}.
-
-\section{Concluding remarks}\label{sec:discussion}
-
-When quantifying the sensitivity of a GW detector and the loudness of a GW source, there are three commonly used quantities: the characteristic strain, the power spectral density, and the spectral energy density.
-We have carefully defined these quantities and derived the relationships between them.
-The characteristic strain (section \ref{sec:character-strain}) is most directly related to the SNR, the PSD is mostly closely related to the mean square amplitude in the detector, and the energy density has a clear physical interpretation.
-We have produced example plots using each of these quantities for a wide range of detectors and sources.
-The predicted source amplitudes are based on astrophysical estimates of the event rates and are subject to varying degrees of uncertainty.\footnote{There is also the exciting possibility of sources yet to be considered.} Interactive versions of these plots, with user-specified detectors and sources, are available on-line at \url{http://www.ast.cam.ac.uk/~rhc26/sources/}.
-Trying to summarise an entire field of astronomy on one plot is an impossible task; however, we hope that the figures and analysis presented here provide useful insight.
-
-\ack{CJM, RHC and CPLB are supported by the Science \& Technology Facilities Council.
-We are indebted to Jonathan Gair, for useful advice and proof-reading the manuscript.
-We also thank Stephen Taylor and Justin Ellis for useful conversations about pulsar timing.
-The authors thanks a number of members of the LIGO--Virgo Scientific Collaboration, especially Chris Messenger, for useful comments and suggestions.
-This paper has been assigned LIGO document LIGO-P1400129.}
-
 \appendix
@@
 The plots in this section show all of the detectors and sources described in the main text.
-Clearer, interactive versions of these plots, allowing for removal of any of the curves, may be created and downloaded on-line, \url{http://www.ast.cam.ac.uk/~rhc26/sources/}.
+Clearer, interactive versions of these plots, allowing for removal of any of the curves, may be created and downloaded on-line, \url{http://rhcole.com/apps/GWplotter}.
 The detector noise curves all have their resonance spikes removed for clarity.
@@
 \end{figure}
-\clearpage
-
-\bibliographystyle{jphysicsCB}
-\bibliography{bibliography}
-
-\end{document}

### 2014-11-13 V2/noise_curve.tex
web address update

@@
 We conclude by producing plots that consistently compare different detectors.
-Similar figures can be generated on-line for general use at \url{http://www.ast.cam.ac.uk/~rhc26/sources/}.
+Similar figures can be generated on-line for general use at \url{http://rhcole.com/apps/GWplotter}.
 \end{abstract}
@@
 We have produced example plots using each of these quantities for a wide range of detectors and sources.
-The predicted source amplitudes are based on astrophysical estimates of the event rates and are subject to varying degrees of uncertainty.\footnote{There is also the exciting possibility of sources yet to be considered.} Interactive versions of these plots, with user-specified detectors and sources, are available on-line at \url{http://www.ast.cam.ac.uk/~rhc26/sources/}.
+The predicted source amplitudes are based on astrophysical estimates of the event rates and are subject to varying degrees of uncertainty.\footnote{There is also the exciting possibility of sources yet to be considered.} Interactive versions of these plots, with user-specified detectors and sources, are available on-line at \url{http://rhcole.com/apps/GWplotter}.
 Trying to summarise an entire field of astronomy on one plot is an impossible task; however, we hope that the figures and analysis presented here provide useful insight.
